### 2017-01-27 research/paper/draft.tex
new title

%\documentclass[iop]{emulateapj}
\documentclass[preprint]{aastex}
%\documentclass[12pt, onecolumn]{emulateapj}
%\documentstyle[aas2pp4,natbib209]{article}

\usepackage{tikz}
\usepackage{natbib}
\usepackage{amsmath}

\usetikzlibrary{shapes.geometric, arrows}
\usetikzlibrary{fit}

\tikzstyle{hyper} = [circle, text centered, draw=black]
\tikzstyle{param} = [circle, text centered, draw=black]
\tikzstyle{data} = [circle, text centered, draw=black, line width=2pt]
\tikzstyle{arrow} = [thick,->,>=stealth]

\newcommand{\myemail}{aimalz@nyu.edu}
\newcommand{\textul}{\underline}

\shorttitle{How to estimate the redshift density distribution}
\shortauthors{Malz and Hogg}

\begin{document}

\title{How to estimate the redshift density distribution}

\author{Alex Malz\altaffilmark{1} \& David W. Hogg\altaffilmark{1,2,3,4}}
\email{aimalz@nyu.edu}

\altaffiltext{1}{Center for Cosmology and Particle Physics, Department of Physics,
  New York University, 4 Washington Pl., room 424, New York, NY 10003, USA}
\altaffiltext{2}{Simons Center for Data Analysis, 160 Fifth Avenue, 7th floor, New York, NY 10010, USA}
\altaffiltext{3}{Center for Data Science, New York University, 726 Broadway, 7th floor, New York, NY 10003, USA}
\altaffiltext{4}{Max-Planck-Institut f\"ur Astronomie, K\"onigstuhl 17, D-69117 Heidelberg, Germany}

\begin{abstract}
\end{abstract}

\keywords{methods: analytical "\_\_\_" methods: data analysis "\_\_\_" methods: statistical "\_\_\_" techniques: photometric "\_\_\_" galaxies: statistics}

\section{Introduction}
\label{sec:intro}

%\bibliographystyle{apj}
\%bibliography{references}

\end{document}

### 2017-01-27 research/paper/draft.tex
new abstract

@@
 
-\shorttitle{How to estimate the redshift density distribution}
+\shorttitle{How to estimate the redshift density function}
 \shortauthors{Malz and Hogg}
@@
 \begin{abstract}
+The redshift density function $n(z)$ is a crucial ingredient for weak lensing cosmology.
+Spectroscopically confirmed redshifts for the dim and numerous galaxies observed by past, present, and future weak lensing surveys are inaccessible, making photometric redshifts (photo-$z$s) the next best alternative.
+Due to their intrinsic scatter and susceptibility to catastrophic outliers, photo-$z$ point estimates are being superseded by photo-$z$ probability distribution functions (PDFs).
+However, analytic methods for utilizing these new data products in cosmological inference are still evolving.
+
+This paper presents a novel approach to estimating the posterior distribution of $n(z)$ from a survey of galaxy photo-$z$ PDFs based upon a probabilistic graphical model.
+We present the Cosmological Hierarchical Inference with Probabilistic Photometric Redshifts (CHIPPR) code implementing this technique, as well as its validation on mock data and testing on a subset of BOSS DR10.
+The code and method are easily extensible to other one-point statistics of redshift that are informative to galaxy evolution and large-scale structure.
+
+%Ongoing and upcoming galaxy surveys aiming to constrain cosmological parameters will produce photometric redshift (photo-$z$) probability distribution functions (PDFs), which are, generally, posteriors under an interim prior.
+It is desirable to use this highly informative data product to estimate redshift-dependent distribution functions such as the redshift density function, stellar mass function, or We present a mathematically rigorous approach to obtain the posterior distribution on a one-point statistic of redshift from a survey providing photo-$z$ PDFs and their interim prior by way of a probabilistic graphical model corresponding to a hierarchical Bayesian relation of photo-$z$ PDFs to a posterior distribution over the one-point statistic.
+This fully consistent method is applied to the redshift density function $n(z)$ necessary for the calculation of cosmological parameters from weak lensing surveys, validated on synthetic data and tested on a small subset of BOSS DR10 data.
+It is demonstrated that this method improves accuracy, both qualitatively and as quantified by the Kullback-Leibler Divergence, compared to popular alternatives (such as stacking of photo-$z$ PDFs and conversion of photo-$z$ PDFs to point estimators of redshift).
+The method can be straightforwardly generalized to other one-point statistics of redshift including distribution functions over luminosity, mass, star formation rate, etc.
+We publicly release the Cosmological Hierarchical Inference with Probabilistic Photometric Redshifts (CHIPPR) code that provides a general framework for estimating the posterior distribution of a general one-point statistic of redshift.
 \end{abstract}

### 2017-01-27 research/paper/draft.tex
updating text to be ready for references

@@
 
-This paper presents a novel approach to estimating the posterior distribution of $n(z)$ from a survey of galaxy photo-$z$ PDFs based upon a probabilistic graphical model.
+This paper presents a novel approach to estimating the posterior distribution over $n(z)$ from a survey of galaxy photo-$z$ PDFs based upon a probabilistic graphical model.
 We present the Cosmological Hierarchical Inference with Probabilistic Photometric Redshifts (CHIPPR) code implementing this technique, as well as its validation on mock data and testing on a subset of BOSS DR10.
@@
 We publicly release the Cosmological Hierarchical Inference with Probabilistic Photometric Redshifts (CHIPPR) code that provides a general framework for estimating the posterior distribution of a general one-point statistic of redshift.
+
 \end{abstract}
@@
 \section{Introduction}
+
 \label{sec:intro}
 
+%\appendix{}
+
+\begin{acknowledgements}
+AIM thanks Mohammadjavad Vakili for insightful input on statistic, Geoffrey Ryan for assistance in debugging code, and Boris Leistedt for helpful comments provided in the preparation of this paper.
+\end{acknowledgements}
+
 %\bibliographystyle{apj}
-\%bibliography{references}
+%\bibliography{references}
 

### 2017-01-27 research/paper/draft.tex
small changes to abstract

@@
 We present the Cosmological Hierarchical Inference with Probabilistic Photometric Redshifts (CHIPPR) code implementing this technique, as well as its validation on mock data and testing on a subset of BOSS DR10.
-The code and method are easily extensible to other one-point statistics of redshift that are informative to galaxy evolution and large-scale structure.
+CHIPPR yields a more comprehensive and accurate characterization of $n(z)$ than traditional procedures.
+The code is easily extensible to other one-point statistics of redshift that are informative to galaxy evolution and large-scale structure.
 

### 2017-01-30 research/paper/draft.tex
added placeholder reference and bibliography
periodic backup

@@
 
-%Ongoing and upcoming galaxy surveys aiming to constrain cosmological parameters will produce photometric redshift (photo-$z$) probability distribution functions (PDFs), which are, generally, posteriors under an interim prior.
-It is desirable to use this highly informative data product to estimate redshift-dependent distribution functions such as the redshift density function, stellar mass function, or We present a mathematically rigorous approach to obtain the posterior distribution on a one-point statistic of redshift from a survey providing photo-$z$ PDFs and their interim prior by way of a probabilistic graphical model corresponding to a hierarchical Bayesian relation of photo-$z$ PDFs to a posterior distribution over the one-point statistic.
-This fully consistent method is applied to the redshift density function $n(z)$ necessary for the calculation of cosmological parameters from weak lensing surveys, validated on synthetic data and tested on a small subset of BOSS DR10 data.
-It is demonstrated that this method improves accuracy, both qualitatively and as quantified by the Kullback-Leibler Divergence, compared to popular alternatives (such as stacking of photo-$z$ PDFs and conversion of photo-$z$ PDFs to point estimators of redshift).
-The method can be straightforwardly generalized to other one-point statistics of redshift including distribution functions over luminosity, mass, star formation rate, etc.
-We publicly release the Cosmological Hierarchical Inference with Probabilistic Photometric Redshifts (CHIPPR) code that provides a general framework for estimating the posterior distribution of a general one-point statistic of redshift.
-
 \end{abstract}
@@
 \section{Introduction}
+\label{sec:intro}
 
-\label{sec:intro}
+The redshift density function $n(z)$ is necessary for calculating two-point statistics of galaxy properties used to determine the cosmological parameter values that inform our understanding of the evolution of large-scale structure in the universe \citep{masters_mapping_2015}.
+Inaccurate estimates of $n(z)$ can significantly impact the constraining power of a galaxy survey \citep{hildebrandt_kids-450:_2016}, biasing recovery of the cosmological parameters.
+
+Though the redshift density function has traditionally been determined from spectroscopically observed redshifts, modern galaxy surveys seek to obtain two-point statistics of redshift from samples of galaxies for which spectroscopic redshifts are unavailable, either due to their large numbers or their low brightnesses.
+For decades \citep{baum_photoelectric_1962}, photometrically estimated redshifts (photo-$z$s) have been the leading alternative to spectroscopically observed redshifts, though they suffer from issues of precision in the form of an intrinsic scatter and accuracy in the form of catastrophic outliers.
+
+These weaknesses are illuminated by a probabilistic interpretation of photo-$zs$; if these complicated uncertainties were expressed as a probability distribution function (PDF) over redshift, photo-$z$s could be replaced by photo-$z$ PDFs containing more information than a simple point estimate.
+
+A more recent
+
+how surveys get $n(z)$
+
+objective: given photo-$z$ PDFs, seek $n(z)$ PDF
 
@@
 \begin{acknowledgements}
-AIM thanks Mohammadjavad Vakili for insightful input on statistic, Geoffrey Ryan for assistance in debugging code, and Boris Leistedt for helpful comments provided in the preparation of this paper.
+AIM thanks Mohammadjavad Vakili for insightful input on statistics, Geoffrey Ryan for assistance in debugging code, and Boris Leistedt for helpful comments provided in the preparation of this paper.
 \end{acknowledgements}
 
-%\bibliographystyle{apj}
-%\bibliography{references}
+\bibliographystyle{apj}
+\bibliography{references}
 

### 2017-01-31 research/paper/draft.tex
wrapping up super-brief introduction

@@
 \section{Introduction}
-\label{sec:intro}
+\label{sec:introduction}
 
@@
 Though the redshift density function has traditionally been determined from spectroscopically observed redshifts, modern galaxy surveys seek to obtain two-point statistics of redshift from samples of galaxies for which spectroscopic redshifts are unavailable, either due to their large numbers or their low brightnesses.
-For decades \citep{baum_photoelectric_1962}, photometrically estimated redshifts (photo-$z$s) have been the leading alternative to spectroscopically observed redshifts, though they suffer from issues of precision in the form of an intrinsic scatter and accuracy in the form of catastrophic outliers.
+For decades \citep{baum_photoelectric_1962}, photometrically estimated redshifts (photo-$z$s) have been the leading alternative to spectroscopically observed redshifts, though they suffer from issues of precision in the form of an intrinsic scatter and accuracy in the form of catastrophic outliers, as well as other systematics imparted by the properties of the survey, data reduction pipeline(s), and assumptions underlying the analysis.
 
-These weaknesses are illuminated by a probabilistic interpretation of photo-$zs$; if these complicated uncertainties were expressed as a probability distribution function (PDF) over redshift, photo-$z$s could be replaced by photo-$z$ PDFs containing more information than a simple point estimate.
+These weaknesses are illuminated by a probabilistic interpretation of photo-$zs$; if these nontrivial uncertainties were expressed as a probability distribution function (PDF) over redshift, photo-$z$s could be replaced by photo-$z$ PDFs containing more information than a simple point estimate \citep{koo_overview_1999}.
+However, methods for using photo-$z$ PDFs in cosmological inference remain underdeveloped, with many survey pipelines reducing them to familiar point estimators that are compatible with existing technology or engaging with them heuristically in a manner inconsistent with their probabilistic nature.
 
-A more recent
+It is desirable to create robust methods for using photo-$z$ PDFs in inference, beginning with the simplest one-point statistic of redshift, $n(z)$.
+In the spirit of \citet{hogg_inferring_2010, foreman-mackey_exoplanet_2014}, this paper derives a mathematically rigorous approach to inferring $n(z)$ in Sec.
+\ref{sec:method}.
+The presentation of a public code implementing this novel technique is given in Sec.
+\ref{sec:model_specifics}.
+The code is validated on mock data and tested on a subset of BOSS DR10 data in Sec.
+\ref{sec:experiments}.
+We conclude and discuss future directions in Sec.
+\ref{sec:discussion}.
 
-how surveys get $n(z)$
+\section{Method}
+\label{sec:method}
 
-objective: given photo-$z$ PDFs, seek $n(z)$ PDF
+\section{Model specifics}
+\label{sec:model_specifics}
+
+\section{Experiments}
+\label{sec:experiments}
+
+\section{Discussion}
+\label{sec:discussion}
 

### 2017-01-31 chippr/catalog_plots.py
make plots slightly prettier

@@ -27,7 +27,7 @@ def plot_true_histogram(true_samps, n_bins=50, plot_loc='', plot_name='true_hist
     pu.set_up_plot()
     f = plt.figure(figsize=(5, 5))
     sps = f.add_subplot(1, 1, 1)
-    sps.hist(true_samps, bins=n_bins, normed=1)
+    sps.hist(true_samps, bins=n_bins, normed=1, color='k')
     sps.set_xlabel(r'$z_{true}$')
     sps.set_ylabel(r'$n(z_{true})$')
     f.savefig(os.path.join(plot_loc, plot_name))
@@ -50,7 +50,7 @@ def plot_obs_scatter(true_samps, obs_samps, plot_loc='', plot_name='obs_scatter.
     pu.set_up_plot()
     f = plt.figure(figsize=(5, 5))
     sps = f.add_subplot(1, 1, 1)
-    sps.scatter(true_samps, obs_samps)
+    sps.scatter(true_samps, obs_samps, 'k.')
     sps.set_xlabel(r'$z_{true}$')
     sps.set_ylabel(r'$z_{obs}$')
     f.savefig(os.path.join(plot_loc, plot_name))

### 2017-01-31 chippr/catalog_plots.py
added in some plots for the sampler

@@ -32,6 +32,8 @@ def plot_true_histogram(true_samps, n_bins=50, plot_loc='', plot_name='true_hist
     sps.set_ylabel(r'$n(z_{true})$')
     f.savefig(os.path.join(plot_loc, plot_name))
 
+    return
+
 def plot_obs_scatter(true_samps, obs_samps, plot_loc='', plot_name='obs_scatter.png'):
     """
     Plots a scatterplot of true and observed redshift values
@@ -50,7 +52,15 @@ def plot_obs_scatter(true_samps, obs_samps, plot_loc='', plot_name='obs_scatter.
     pu.set_up_plot()
     f = plt.figure(figsize=(5, 5))
     sps = f.add_subplot(1, 1, 1)
-    sps.scatter(true_samps, obs_samps, 'k.')
+    sps.scatter(true_samps, obs_samps, c='k', linewidth=0., alpha=0.1)
     sps.set_xlabel(r'$z_{true}$')
     sps.set_ylabel(r'$z_{obs}$')
     f.savefig(os.path.join(plot_loc, plot_name))
+
+    return
+
+def plot_pdfs():
+    """
+    Plots a small number of randomly chosen likelihood functions and parametrized posteriors
+    """
+    return

### 2017-01-31 chippr/log_z_dens.py

@@ -344,21 +344,23 @@ class log_z_dens(object):
             self.n_walkers = len(ivals)
             self.sampler = emcee.EnsembleSampler(self.n_walkers, self.n_bins, self.evaluate_log_hyper_posterior)
             self.burn_ins = 0
-            self.tot_runs = 0
             self.burning_in = True
             vals = ivals
+            canvas = plots.set_up_burn_in_plots()
             while self.burning_in:
-                self.burn_ins += 1
                 if vb:
                     print('beginning sampling '+str(self.burn_ins))
                 burn_in_mcmc_outputs = self.sample(vals, n_burn_test)
+                canvas = plots.plot_sampler_progress(canvas, burn_in_mcmc_outputs, n_burn_test, self.burn_ins, self.n_walkers, self.n_bins, self.plot_dir)
                 self.burning_in = s.gr_test(burn_in_mcmc_outputs['chains'])
                 if save:
-                    with open('mcmc'+str(self.burn_ins)+'.p', 'wb') as file_location:
+                    with open(os.path.join(self.res_dir, 'mcmc'+str(self.burn_ins)+'.p'), 'wb') as file_location:
                         cpkl.dump(burn_in_mcmc_outputs, file_location)
                 vals = np.array([item[-1] for item in burn_in_mcmc_outputs['chains']])
+                self.burn_ins += 1
 
             mcmc_products = self.sample(vals, n_accepted)
+
             self.log_smp_nz = mcmc_products['chains']
             self.smp_nz = np.exp(self.log_smp_nz)
             self.info['estimators']['log_sampled_nz'] = self.log_smp_nz
@@ -414,6 +416,7 @@ class log_z_dens(object):
         """
         f = plots.plot_estimators(self.info)
         f.savefig(os.path.join(self.plot_dir, plot_loc))
+        return
 
     def read(self, read_loc, style='pickle', vb=True):
         """

### 2017-01-31 chippr/log_z_dens_plots.py

@@ -1,4 +1,7 @@
 import numpy as np
+import os
+import scipy as sp
+from scipy import stats
 
 import matplotlib as mpl
 mpl.use('PS')
@@ -18,6 +21,111 @@ s_exp, w_exp, a_exp, c_exp, d_exp, l_exp = '--', 1., 1., 'k', [(0, (2.5, 2.5))],
 s_mle, w_mle, a_mle, c_mle, d_mle, l_mle = '--', 2., 1., 'k', [(0, (2.5, 2.5))], 'MMLE '
 # s_smp, w_smp, a_smp, c_smp, d_smp, l_smp = '--', 1., 1., 'k', [(0, (1, 1))], 'Sampled '
 s_bfe, w_bfe, a_bfe, c_bfe, d_bfe, l_bfe = '--', 2., 1., 'k', [(0, (1, 1))], 'Mean of\n Samples '
+
+def plot_ivals(info):
+    """
+    Plots the initial values given to the sampler
+
+    Parameters
+    ----------
+    info: dict
+        dictionary of stored information from log_z_dens object
+
+    Returns
+    -------
+    f: matplotlib figure
+        figure object
+    """
+    return f
+
+def set_up_burn_in_plots():
+    """
+    Creates plotting objects for sampler progress
+    """
+    f_autocorrelation_times = plt.figure(figsize=(5, 5))
+    sps_autocorrelation_times = f_autocorrelation_times.add_subplot(1, 1, 1)
+    f_autocorrelation_times.subplots_adjust(hspace=0, wspace=0)
+    sps_autocorrelation_times.set_ylabel(r'autocorrelation time')
+    sps_autocorrelation_times.set_xlabel(r'accepted sample number')
+    sps_autocorrelation_times.set_ylim(0, 100)
+    autocorrelation_times_plot = [f_autocorrelation_times, sps_autocorrelation_times]
+
+    f_acceptance_fractions = plt.figure(figsize=(5, 5))
+    sps_acceptance_fractions = f_acceptance_fractions.add_subplot(1, 1, 1)
+    f_acceptance_fractions.subplots_adjust(hspace=0, wspace=0)
+    sps_acceptance_fractions.set_ylim(0, 1)
+    sps_acceptance_fractions.set_ylabel('acceptance fraction per bin')
+    sps_acceptance_fractions.set_xlabel('number of iterations')
+    acceptance_fractions_plot = [f_acceptance_fractions, sps_acceptance_fractions]
+
[…]

### 2017-01-31 chippr/defaults.py
plotting samples does not quite work yet

@@ -17,6 +17,8 @@ gr_threshold = 1.1
 n_accepted = 10**3
 n_burned = 10**2
 
+plot_colors = 5
+
 def check_sim_params(params={}):
     """
     Checks simulation parameter dictionary for various keywords and sets to default values if not present

### 2017-01-31 chippr/log_z_dens.py

@@ -302,8 +302,8 @@ class log_z_dens(object):
 
         Returns
         -------
-        mcmc_outputs: ndarray
-            array of sampled redshift density function bin values
+        mcmc_outputs: dict
+            dictionary containing array of sampled redshift density function bin values as well as posterior probabilities, acceptance fractions, and autocorrelation times
         """
         self.sampler.reset()
         pos, prob, state = self.sampler.run_mcmc(ivals, n_samps)
@@ -346,12 +346,14 @@ class log_z_dens(object):
             self.burn_ins = 0
             self.burning_in = True
             vals = ivals
-            canvas = plots.set_up_burn_in_plots()
+            if vb:
+                canvas = plots.set_up_burn_in_plots()
             while self.burning_in:
                 if vb:
                     print('beginning sampling '+str(self.burn_ins))
                 burn_in_mcmc_outputs = self.sample(vals, n_burn_test)
-                canvas = plots.plot_sampler_progress(canvas, burn_in_mcmc_outputs, n_burn_test, self.burn_ins, self.n_walkers, self.n_bins, self.plot_dir)
+                if vb:
+                    canvas = plots.plot_sampler_progress(canvas, burn_in_mcmc_outputs, n_burn_test, self.burn_ins, self.n_walkers, self.n_bins, self.plot_dir)
                 self.burning_in = s.gr_test(burn_in_mcmc_outputs['chains'])
                 if save:
                     with open(os.path.join(self.res_dir, 'mcmc'+str(self.burn_ins)+'.p'), 'wb') as file_location:
@@ -373,6 +375,9 @@ class log_z_dens(object):
             self.smp_nz = np.exp(self.log_smp_nz)
             self.log_bfe_nz = self.info['estimators']['log_mean_sampled_nz']
             self.bfe_nz = np.exp(self.log_smp_nz)
+
+        if vb:
+            plots.plot_samples(self.info, self.plot_dir)
 
         return self.log_smp_nz
 
@@ -405,17 +410,11 @@ class log_z_dens(object):
             print(self.info['stats'])
         return self.info['stats']
 
-    def plot_estimators(self, plot_loc):
+    def plot_estimators(self):
         """
         Plots all available estimators of the redshift density function.
-
-        Parameters
-        ----------
-        plot_loc: string
-            destination where plot should be stored
         """
-        f = plots.plot_estimators(self.info)
-        f.savefig(os.path.join(self.plot_dir, plot_loc))
[…]

### 2017-01-31 chippr/log_z_dens_plots.py

@@ -6,8 +6,10 @@ from scipy import stats
 import matplotlib as mpl
 mpl.use('PS')
 import matplotlib.pyplot as plt
+import matplotlib.cm as cm
 
 import chippr
+from chippr import defaults as d
 from chippr import utils as u
 from chippr import plot_utils as pu
 
@@ -19,8 +21,11 @@ s_stk, w_stk, a_stk, c_stk, d_stk, l_stk = '--', 1.5, 1., 'k', [(0, (7.5, 7.5))]
 s_map, w_map, a_map, c_map, d_map, l_map = '--', 1., 1., 'k', [(0, (7.5, 7.5))], 'MMAP '
 s_exp, w_exp, a_exp, c_exp, d_exp, l_exp = '--', 1., 1., 'k', [(0, (2.5, 2.5))], 'MExp '
 s_mle, w_mle, a_mle, c_mle, d_mle, l_mle = '--', 2., 1., 'k', [(0, (2.5, 2.5))], 'MMLE '
-# s_smp, w_smp, a_smp, c_smp, d_smp, l_smp = '--', 1., 1., 'k', [(0, (1, 1))], 'Sampled '
+s_smp, w_smp, a_smp, c_smp, d_smp, l_smp = '--', 1., 1., 'k', [(0, (1, 1))], 'Sampled '
 s_bfe, w_bfe, a_bfe, c_bfe, d_bfe, l_bfe = '--', 2., 1., 'k', [(0, (1, 1))], 'Mean of\n Samples '
+
+cmap = np.linspace(0., 1., d.plot_colors)
+colors = [cm.viridis(i) for i in cmap]
 
 def plot_ivals(info):
     """
@@ -41,7 +46,14 @@ def plot_ivals(info):
 def set_up_burn_in_plots():
     """
     Creates plotting objects for sampler progress
+
+    Returns
+    -------
+    plot_information: tuple
+        contains figure and subplot objects for autocorrelation times, acceptance fractions, and posterior probabilities
     """
+    pu.set_up_plot()
+
     f_autocorrelation_times = plt.figure(figsize=(5, 5))
     sps_autocorrelation_times = f_autocorrelation_times.add_subplot(1, 1, 1)
     f_autocorrelation_times.subplots_adjust(hspace=0, wspace=0)
@@ -71,6 +83,33 @@ def set_up_burn_in_plots():
[…]

### 2017-02-21 research/paper/draft.tex
unclear why the paper is altered in this branch but it will be overwritten by a merge from issue/41/reboot_paper eventually anyway

--- 2017-01-30 research/paper/draft.tex
@@
 
-\shorttitle{How to estimate the redshift density function}
+\shorttitle{One-point galaxy population statistics from photo-$z$ PDFs}
 \shortauthors{Malz and Hogg}
@@
 
-\title{How to estimate the redshift density distribution}
+\title{Using a Bayesian hierarchical model to infer the redshift density function from photometric redshift probability distribution functions}
 
-\author{Alex Malz\altaffilmark{1} \& David W. Hogg\altaffilmark{1,2,3,4}}
+\author{A.I. Malz\altaffilmark{1} \& David W. Hogg\altaffilmark{1,2,3,4} %\& Boris Leistedt\altaffilmark{1,5}
+}
 \email{aimalz@nyu.edu}
@@
 \altaffiltext{4}{Max-Planck-Institut f\"ur Astronomie, K\"onigstuhl 17, D-69117 Heidelberg, Germany}
+%\altaffiltext{5}{Department of Physics \& Astronomy, University College London, Gower Street, London, WC1E 6BT, UK}
 
 \begin{abstract}
-The redshift density function $n(z)$ is a crucial ingredient for weak lensing cosmology.
-Spectroscopically confirmed redshifts for the dim and numerous galaxies observed by past, present, and future weak lensing surveys are inaccessible, making photometric redshifts (photo-$z$s) the next best alternative.
-Due to their intrinsic scatter and susceptibility to catastrophic outliers, photo-$z$ point estimates are being superseded by photo-$z$ probability distribution functions (PDFs).
-However, analytic methods for utilizing these new data products in cosmological inference are still evolving.
-
-This paper presents a novel approach to estimating the posterior distribution over $n(z)$ from a survey of galaxy photo-$z$ PDFs based upon a probabilistic graphical model.
-We present the Cosmological Hierarchical Inference with Probabilistic Photometric Redshifts (CHIPPR) code implementing this technique, as well as its validation on mock data and testing on a subset of BOSS DR10.
-CHIPPR yields a more comprehensive and accurate characterization of $n(z)$ than traditional procedures.
-The code is easily extensible to other one-point statistics of redshift that are informative to galaxy evolution and large-scale structure.
-
+Ongoing and upcoming galaxy surveys aiming to constrain cosmological parameters will produce photometric redshift (photo-$z$) probability distribution functions (PDFs), which are, generally, posteriors under an interim prior.
+It is desirable to use this highly informative data product to estimate redshift-dependent distribution functions such as the redshift density function, stellar mass function, or We present a mathematically rigorous approach to obtain the posterior distribution on a one-point statistic of redshift from a survey providing photo-$z$ PDFs and their interim prior by way of a probabilistic graphical model corresponding to a hierarchical Bayesian relation of photo-$z$ PDFs to a posterior distribution over the one-point statistic.
+This fully consistent method is applied to the redshift density function $n(z)$ necessary for the calculation of cosmological parameters from weak lensing surveys, validated on synthetic data and tested on a small subset of BOSS DR10 data.
+It is demonstrated that this method improves accuracy, both qualitatively and as quantified by the Kullback-Leibler Divergence, compared to popular alternatives (such as stacking of photo-$z$ PDFs and conversion of photo-$z$ PDFs to point estimators of redshift).
+The method can be straightforwardly generalized to other one-point statistics of redshift including distribution functions over luminosity, mass, star formation rate, etc.
+We publicly release the Cosmological Hierarchical Inference with Probabilistic Photometric Redshifts (CHIPPR) code that provides a general framework for estimating the posterior distribution of a general one-point statistic of redshift.
 \end{abstract}
@@
 
+%\clearpage
 \section{Introduction}
 \label{sec:intro}
-
-The redshift density function $n(z)$ is necessary for calculating two-point statistics of galaxy properties used to determine the cosmological parameter values that inform our understanding of the evolution of large-scale structure in the universe \citep{masters_mapping_2015}.
-Inaccurate estimates of $n(z)$ can significantly impact the constraining power of a galaxy survey \citep{hildebrandt_kids-450:_2016}, biasing recovery of the cosmological parameters.
-
-Though the redshift density function has traditionally been determined from spectroscopically observed redshifts, modern galaxy surveys seek to obtain two-point statistics of redshift from samples of galaxies for which spectroscopic redshifts are unavailable, either due to their large numbers or their low brightnesses.
-For decades \citep{baum_photoelectric_1962}, photometrically estimated redshifts (photo-$z$s) have been the leading alternative to spectroscopically observed redshifts, though they suffer from issues of precision in the form of an intrinsic scatter and accuracy in the form of catastrophic outliers.
-
-These weaknesses are illuminated by a probabilistic interpretation of photo-$zs$; if these complicated uncertainties were expressed as a probability distribution function (PDF) over redshift, photo-$z$s could be replaced by photo-$z$ PDFs containing more information than a simple point estimate.
-
-A more recent
-
-how surveys get $n(z)$
-
-objective: given photo-$z$ PDFs, seek $n(z)$ PDF
-
-%\appendix{}
-
-\begin{acknowledgements}
-AIM thanks Mohammadjavad Vakili for insightful input on statistics, Geoffrey Ryan for assistance in debugging code, and Boris Leistedt for helpful comments provided in the preparation of this paper.
-\end{acknowledgements}
-
-\bibliographystyle{apj}
-\bibliography{references}
-
-\end{document}

### 2017-02-21 research/paper/draft.tex
updated address

--- 2017-01-31 research/paper/draft.tex
@@
 
-\title{How to estimate the redshift density distribution}
+\title{How to estimate the redshift density function}
 
@@
 \altaffiltext{1}{Center for Cosmology and Particle Physics, Department of Physics,
-  New York University, 4 Washington Pl., room 424, New York, NY 10003, USA}
+  New York University, 726 Broadway, 9th floor, New York, NY 10003, USA}
 \altaffiltext{2}{Simons Center for Data Analysis, 160 Fifth Avenue, 7th floor, New York, NY 10010, USA}

### 2017-02-27 chippr/stat_utils.py
trying different ways to calculate mean and standard deviation

@@ -43,7 +43,7 @@ def norm_fit(population):
     flat = population.reshape(np.prod(shape[:-1]), shape[-1]).T
     locs, scales = [], []
     for k in range(shape[-1]):
-        loc, scale = sp.stats.norm.fit_loc_scale(flat[k])
+        (loc, scale) = sp.stats.norm.fit_loc_scale(flat[k])
         locs.append(loc)
         scales.append(scale)
     locs = np.array(locs)

### 2017-02-27 research/paper/draft.tex
revised title, abstract, introduction

@@
 
-\shorttitle{How to estimate the redshift density function}
+\shorttitle{How to obtain the redshift distribution from probabilistic redshift estimates}
 \shortauthors{Malz and Hogg}
@@
 
-\title{How to estimate the redshift density function}
+\title{How to obtain the redshift distribution from probabilistic redshift estimates}
 
@@
   New York University, 726 Broadway, 9th floor, New York, NY 10003, USA}
-\altaffiltext{2}{Simons Center for Data Analysis, 160 Fifth Avenue, 7th floor, New York, NY 10010, USA}
-\altaffiltext{3}{Center for Data Science, New York University, 726 Broadway, 7th floor, New York, NY 10003, USA}
+\altaffiltext{2}{Simons Center for Computational Astrophysics, 162 Fifth Avenue, 7th floor, New York, NY 10010, USA}
+\altaffiltext{3}{Center for Data Science, New York University, 60 Fifth Avenue, 7th floor, New York, NY 10003, USA}
 \altaffiltext{4}{Max-Planck-Institut f\"ur Astronomie, K\"onigstuhl 17, D-69117 Heidelberg, Germany}
@@
 \begin{abstract}
-The redshift density function $n(z)$ is a crucial ingredient for weak lensing cosmology.
-Spectroscopically confirmed redshifts for the dim and numerous galaxies observed by past, present, and future weak lensing surveys are inaccessible, making photometric redshifts (photo-$z$s) the next best alternative.
-Due to their intrinsic scatter and susceptibility to catastrophic outliers, photo-$z$ point estimates are being superseded by photo-$z$ probability distribution functions (PDFs).
+The redshift distribution $n(z)$ is a crucial ingredient for weak lensing cosmology.
+Spectroscopically confirmed redshifts for the dim and numerous galaxies observed by weak lensing surveys are expected to be inaccessible, making photometric redshifts (photo-$z$s) the next best alternative.
+Because of the nontrivial inference involved in their determination, photo-$z$ point estimates are being superseded by photo-$z$ probability distribution functions (PDFs).
 However, analytic methods for utilizing these new data products in cosmological inference are still evolving.
 
-This paper presents a novel approach to estimating the posterior distribution over $n(z)$ from a survey of galaxy photo-$z$ PDFs based upon a probabilistic graphical model.
+This paper presents a novel approach to estimating the posterior distribution over $n(z)$ from a survey of galaxy photo-$z$ PDFs based upon a probabilistic graphical model of hierarchical inference.
 We present the Cosmological Hierarchical Inference with Probabilistic Photometric Redshifts (CHIPPR) code implementing this technique, as well as its validation on mock data and testing on a subset of BOSS DR10.
-CHIPPR yields a more comprehensive and accurate characterization of $n(z)$ than traditional procedures.
-The code is easily extensible to other one-point statistics of redshift that are informative to galaxy evolution and large-scale structure.
+CHIPPR yields an accurate characterization of $n(z)$ containing information beyond the best-fit estimator produced by traditional procedures.
+The publicly available code is easily extensible to other one-point statistics that depend on redshift.
 
@@
 
-\keywords{methods: analytical "\_\_\_" methods: data analysis "\_\_\_" methods: statistical "\_\_\_" techniques: photometric "\_\_\_" galaxies: statistics}
+\keywords{catalogs --- cosmology: cosmological parameters --- galaxies: statistics --- gravitational lensing: weak --- methods: analytical --- methods: data analysis --- methods: statistical --- techniques: photometric}
 
@@
 
-The redshift density function $n(z)$ is necessary for calculating two-point statistics of galaxy properties used to determine the cosmological parameter values that inform our understanding of the evolution of large-scale structure in the universe \citep{masters_mapping_2015}.
-Inaccurate estimates of $n(z)$ can significantly impact the constraining power of a galaxy survey \citep{hildebrandt_kids-450:_2016}, biasing recovery of the cosmological parameters.
+The redshift distribution $n(z)$ is necessary for calculating two-point statistics of galaxy properties used to determine the cosmological parameter values that inform our understanding of the evolution of large-scale structure in the universe \citep{masters_mapping_2015}.
+Inaccurate estimates of $n(z)$ can significantly impact the constraining power of a galaxy survey, biasing recovery of the cosmological parameters.
+For example, if the $n(z)$ used in analysis is offset from the true $n(z)$ by a small, negative constant, it results in underestimating $w_{0}$ and overestimating $\sigma_{8}$ \citep{samuroff_simultaneous_2017}.
 
-Though the redshift density function has traditionally been determined from spectroscopically observed redshifts, modern galaxy surveys seek to obtain two-point statistics of redshift from samples of galaxies for which spectroscopic redshifts are unavailable, either due to their large numbers or their low brightnesses.
-For decades \citep{baum_photoelectric_1962}, photometrically estimated redshifts (photo-$z$s) have been the leading alternative to spectroscopically observed redshifts, though they suffer from issues of precision in the form of an intrinsic scatter and accuracy in the form of catastrophic outliers, as well as other systematics imparted by the properties of the survey, data reduction pipeline(s), and assumptions underlying the analysis.
+Though the redshift density function has traditionally been determined from spectroscopically observed redshifts, modern galaxy surveys including DES, LSST, Euclid, and WFIRST seek to obtain two-point statistics of redshift from samples of galaxies for which spectroscopic redshifts are unavailable, either due to their large numbers or their low brightnesses.
+For decades, photometrically estimated redshifts (photo-$z$s) have been the leading alternative to spectroscopically observed redshifts, though they suffer from issues of precision in the form of an intrinsic scatter and accuracy in the form of catastrophic outliers, as well as other systematics imparted by the properties of the survey, data reduction pipeline(s), and assumptions underlying the analysis \citep{baum_photoelectric_1962}.
+These weaknesses are illuminated by a probabilistic interpretation of photo-$zs$; if these nontrivial uncertainties were expressed as a probability distribution function (PDF) over redshift, photo-$z$s could be replaced by photo-$z$ PDFs containing more information than a simple point estimate \citep{koo_overview_1999}.
+Such data products have been commonly released by photometric galaxy surveys since SDSS DR7 \citep{abazajian_seventh_2009} using a great variety of methods.
 
-These weaknesses are illuminated by a probabilistic interpretation of photo-$zs$; if these nontrivial uncertainties were expressed as a probability distribution function (PDF) over redshift, photo-$z$s could be replaced by photo-$z$ PDFs containing more information than a simple point estimate \citep{koo_overview_1999}.
-However, methods for using photo-$z$ PDFs in cosmological inference remain underdeveloped, with many survey pipelines reducing them to familiar point estimators that are compatible with existing technology or engaging with them heuristically in a manner inconsistent with their probabilistic nature.
+Methods for using photo-$z$ PDFs in cosmological inference remain underdeveloped, with many survey pipelines reducing them to familiar point estimators that are compatible with existing technology or engaging with them heuristically in a manner inconsistent with their probabilistic nature.
+Stacking photo-$z$ PDFs, or other mathematically equivalent methods, to obtain an estimator of $n(z)$ is especially popular \citep{cunha_estimating_2009, sheldon_photometric_2012}.
+However, photo-$z$ PDFs are probabilistic data products that must be handled in a fully consistent, probabilistic manner; hierarchical inference is the only mathematically valid way to do inference with photo-$z$ PDFs and other probabilistic quantities.
 
-It is desirable to create robust methods for using photo-$z$ PDFs in inference, beginning with the simplest one-point statistic of redshift, $n(z)$.
+It is desirable to create rigorous methods for using photo-$z$ PDFs in inference, beginning with the simplest one-point statistic of redshift, $n(z)$.
 In the spirit of \citet{hogg_inferring_2010, foreman-mackey_exoplanet_2014}, this paper derives a mathematically rigorous approach to inferring $n(z)$ in Sec.

### 2017-02-28 research/paper/draft.tex
some progress on methods section

@@
 
-\section{Model specifics}
-\label{sec:model_specifics}
+The redshift distribution $n(z)$ may be understood as the probability of finding a galaxy at a redshift $z$.
+We may express $n(z)$ in terms of some functional form $f_{\vec{\theta}}(z)$ with parameters comprising $\vec{\theta}$ according to Eq.
+\ref{eq:parametrization}.
+\begin{align}
+\label{eq:parametrization}
+n(z) &\equiv f_{\vec{\theta}}(z)\ \equiv\ p(z|\vec{\theta})
+\end{align}
+In this work, we wish to characterize the posterior distribution $p(\vec{\theta}|D)$ of the parameters contained in $\vec{\theta}$ defining $n(z)$ given all available data $D$.
+The likelihood $p(z|\vec{\theta})$ represents the probability of a random variable, in this case redshift $z$, given the parameters in $\vec{\theta}$ defining the distribution $n(z)$ from which it is drawn.
+Our model for inferring $p(\vec{\theta}|D)$ will be introduced in Sec.
+\ref{sec:derivation}, and alternative methods of estimating $\vec{\theta}$ will be presented in Sec.
+\ref{sec:alternatives}.
+
+\subsection{Model Specifics}
+\label{sec:derivation}
+
+Inference of $p(\vec{\theta}|D)$ is perfectly suited to a hierarchical Bayesian model.
+In this problem, the available data $D$ is a catalog of photometry $\{\vec{d}_{i}\}$ of individual galaxies $i$ where each $\vec{d}_{i}$ is a random variable that is a function of its redshift parameter $z_{i}$; the redshift parameters $\{z_{i}\}$ are independently drawn from a function defined by hyperparameters in $\vec{\theta}$.
+These physical relationships may be illustrated in Fig.
+\ref{fig:pgm} by a directed acyclic graph representing a probabilistic graphical model (PGM).
+\begin{figure}
+\begin{center}
+\includegraphics[width=0.5\textwidth]{pgm.png}
+\caption{This directed acyclic graph corresponds to a PGM for a hierarchical inference of $p(\vec{\theta}|\{\vec{d}_{i}\})$.
+In this graph, all random variables are shown in circles, with observed variables shown in shaded circles.
+Relationships between variables are indicated by arrows.
+The box indicates that there are a number of copies of the relationships between boxed parameters, each independent of all others.
+The hyperparameters $\vec{\theta}$ representing $n(z)$ are at the top.
+Independently drawn from a function of the hyperparameters $\vec{\theta}$ are galaxy redshifts $\{z_{i}\}$ below.
+The observed galaxy photometry $\{\vec{d}_{i}\}$, shown in shaded circles, is determined by the redshifts above.}
+\label{fig:pgm}
+\end{center}
+\end{figure}
+This PGM may be used as the basis for a derivation of the desired hyperposterior $p(\vec{\theta}|\{\vec{d}_{i}\})$, assuming our physical model is complete.
+From this point on, we will work solely with log probabilities.
+
+We begin with Eq.
+\ref{eq:bayes} by applying Bayes' Rule to express the log-hyperposterior $\ln\left[p(\vec{\theta}|\{\vec{d}_{i}\})\right]$ in terms of the log-hyperlikelihood $\ln\left[p(\{\vec{d}_{i}\}|\vec{\theta})\right]$.
+\begin{align}
+\label{eq:bayes}
+\ln\left[p(\vec{\theta}|\{\vec{d}_{i}\})\right] &\propto \ln\left[p(\vec{\theta})\right] + \ln\left[p(\{\vec{d}_{i}\}|\vec{\theta})\right]
+\end{align}
+To do this, we will choose a log-hyperprior distribution $\ln\left[p(\vec{\theta})\right]$ representing our beliefs about the distribution of the hyperparameters $\vec{\theta}$ in terms of fixed variables that we will assume are known.
+
+The log-hyperlikelihood $\ln\left[p(\{\vec{d}_{i}\}|\vec{\theta})\right]$ contains no explicit reference to redshifts, therefore, we employ the PGM to write the marginalization over the redshifts in Eq.
+\ref{eq:marginalization}.
+\begin{align}
+\label{eq:marginalization}
+\ln\left[p(\{\vec{d}_{i}\}|\vec{\theta})\right] &= \ln\left[\int\ p(\{\vec{d}_{i}\}|\{z_{i}\})\ p(\{z_{i}\}|\vec{\theta})\ d\{z_{i}\}\right]
+\end{align}
+We shall handle the two terms in the integral separately.
+
+The first term is the likelihood of all galaxy photometry given all galaxy redshifts.
+If we assume independence of galaxy photometry, such that $p(\vec{d}_{i}|\{\vec{d}_{i'\neq i}\}, \{z_{i}\})=p(\vec{d}_{i}|z_{i})$, then we may write Eq.
+\ref{eq:independence}.
+\begin{align}
+\label{eq:independence}
+\ln\left[p(\{\vec{d}_{i}\}|\{z_{i}\})\right] &= \sum_{i}\ln\left[p(\vec{d}_{i}|z_{i})\right]
+\end{align}
+The second term may also be expanded by assuming that all galaxy redshifts are drawn independently from $n(z)$.
+\begin{align}
+\label{eq:poisson}
+\ln\left[p(\{z_{i}\}|\vec{\theta})\right] &= \sum_{i}\ln\left[p(z_{i}|\vec{\theta})\right]
+\end{align}
+
+Photo-$z$ PDFs, on the other hand, are posteriors, probabilities of a parameter, in this case redshift $z$, given the random variable derived from the parameter, in this case photometric data $\vec{d}$.
+Though they are commonly written as $p(z)$, a photo-$z$ PDF for galaxy $i$ in the catalog is truly a posterior $p(z_{i}|\vec{d}_{i})$ where $\vec{d}_{i}$ represents photometric data.
+
+This approach makes several assumptions that must be explicitly stated.
+\begin{enumerate}
+\item The PGM is an expression of our beliefs about the physics of the problem, and the inference will only be valid to the degree that the model is complete.
+\item We must choose a hyperprior $p(\vec{\theta})$; though we aim to choose a sufficiently general hyperprior that will not be a dominant source of information in the hyperposterior, it must still be assumed that whatever hyperprior is chosen will be accurate.
+\item We have assumed independence of galaxies in our catalog such that each galaxy's photometry $\vec{d}_{i}$ is independent from all other galaxies' photometry $\{\vec{d}_{i'\neq i}\}$ and all other galaxies' redshifts $\{z_{i'\neq i}\}$.
+\end{enumerate}
+
+\section{Alternative methods}
+\label{sec:alternatives}
 
@@
 \begin{acknowledgements}
-AIM thanks Mohammadjavad Vakili for insightful input on statistics, Geoffrey Ryan for assistance in debugging code, and Boris Leistedt for helpful comments provided in the preparation of this paper.
+AIM thanks Phil Marshall for advising on the production of usable code, Mohammadjavad Vakili for insightful input on statistics, Geoffrey Ryan for assistance in debugging code, and Boris Leistedt for helpful comments provided in the preparation of this paper.
 \end{acknowledgements}

### 2017-02-28 chippr/log_z_dens.py
modified likelihood function in accordance with paper

@@ -128,7 +128,7 @@ class log_z_dens(object):
             log prior probability associated with parameters in log_nz
         """
         log_hyper_prior = -0.5 * np.dot(np.dot(self.hyper_prior.invvar, log_nz), log_nz)
-        log_hyper_prior -= np.abs(np.dot(np.exp(log_nz), self.bin_difs) - 1.)
+        log_hyper_prior -= np.dot(np.exp(log_nz), self.bin_difs)
 
         return log_hyper_prior
 

### 2017-02-28 research/paper/draft.tex
finished derivation section

@@
 \end{align}
-In this work, we wish to characterize the posterior distribution $p(\vec{\theta}|D)$ of the parameters contained in $\vec{\theta}$ defining $n(z)$ given all available data $D$.
+In this work, we wish to characterize the posterior distribution $p(\vec{\theta}|\textul{D})$ of the parameters contained in $\vec{\theta}$ defining $n(z)$ given all available data $\textul{D}$.
 The likelihood $p(z|\vec{\theta})$ represents the probability of a random variable, in this case redshift $z$, given the parameters in $\vec{\theta}$ defining the distribution $n(z)$ from which it is drawn.
-Our model for inferring $p(\vec{\theta}|D)$ will be introduced in Sec.
+Our model for inferring $p(\vec{\theta}|\textul{D})$ will be introduced in Sec.
 \ref{sec:derivation}, and alternative methods of estimating $\vec{\theta}$ will be presented in Sec.
@@
 
-\subsection{Model Specifics}
+\subsection{Derivation}
 \label{sec:derivation}
 
-Inference of $p(\vec{\theta}|D)$ is perfectly suited to a hierarchical Bayesian model.
-In this problem, the available data $D$ is a catalog of photometry $\{\vec{d}_{i}\}$ of individual galaxies $i$ where each $\vec{d}_{i}$ is a random variable that is a function of its redshift parameter $z_{i}$; the redshift parameters $\{z_{i}\}$ are independently drawn from a function defined by hyperparameters in $\vec{\theta}$.
+Inference of $p(\vec{\theta}|\textul{D})$ is perfectly suited to a hierarchical Bayesian model.
+In this problem, the available data $\textul{D}$ is a catalog of photometry $\{\vec{d}_{i}\}$ of individual galaxies $i$ where each $\vec{d}_{i}$ is a random variable that is a function of its redshift parameter $z_{i}$; the redshift parameters $\{z_{i}\}$ are independently drawn from a function defined by hyperparameters in $\vec{\theta}$.
 These physical relationships may be illustrated in Fig.
@@
 We begin with Eq.
-\ref{eq:bayes} by applying Bayes' Rule to express the log-hyperposterior $\ln\left[p(\vec{\theta}|\{\vec{d}_{i}\})\right]$ in terms of the log-hyperlikelihood $\ln\left[p(\{\vec{d}_{i}\}|\vec{\theta})\right]$.
+\ref{eq:hyperbayes} by applying Bayes' Rule to express the log-hyperposterior $\ln\left[p(\vec{\theta}|\{\vec{d}_{i}\})\right]$ in terms of the log-hyperlikelihood $\ln\left[p(\{\vec{d}_{i}\}|\vec{\theta})\right]$.
 \begin{align}
-\label{eq:bayes}
+\label{eq:hyperbayes}
 \ln\left[p(\vec{\theta}|\{\vec{d}_{i}\})\right] &\propto \ln\left[p(\vec{\theta})\right] + \ln\left[p(\{\vec{d}_{i}\}|\vec{\theta})\right]
@@
 If we assume independence of galaxy photometry, such that $p(\vec{d}_{i}|\{\vec{d}_{i'\neq i}\}, \{z_{i}\})=p(\vec{d}_{i}|z_{i})$, then we may write Eq.
-\ref{eq:independence}.
+\ref{eq:independentdata}.
 \begin{align}
-\label{eq:independence}
+\label{eq:independentdata}
 \ln\left[p(\{\vec{d}_{i}\}|\{z_{i}\})\right] &= \sum_{i}\ln\left[p(\vec{d}_{i}|z_{i})\right]
 \end{align}
-The second term may also be expanded by assuming that all galaxy redshifts are drawn independently from $n(z)$.
+The second term may also be expanded as in Eq. \ref{eq:independentredshifts} by assuming that all galaxy redshifts are drawn independently from $n(z)$.
 \begin{align}
-\label{eq:poisson}
+\label{eq:independentredshifts}
 \ln\left[p(\{z_{i}\}|\vec{\theta})\right] &= \sum_{i}\ln\left[p(z_{i}|\vec{\theta})\right]
 \end{align}
+Since all galaxy redshifts are independent of one another, an integral over the aggregate $\int\dots d\{z_{i}\}$ is simply the product of the integrals over each one $\prod_{i}\int\dots dz_{i}$.
+Thus, the log-hyperlikelihood can be written according to Eq.
+\ref{eq:log-hyperlikelihood}.
+\begin{align}
+\label{eq:log-hyperlikelihood}
+\ln\left[p(\{\vec{d}_{i}\}|\vec{\theta})\right] &= \sum_{i}\ln\left[\int\ p(\vec{d}_{i}|z_{i})\ p(z_{i}|\vec{\theta})\ dz_{i}\right]
+\end{align}
+The expression of Eq.
+\ref{eq:log-hyperlikelihood} contains two types of quantities.
+The $\{p(z_{i}|\vec{\theta})\}$ are known likelihoods that will be equal to $\{f_{\vec{\theta}}(z_{i})\}$.
+The $\{p(\vec{d}_{i}|z_{i})\}$, however, are likelihoods that are unknown and in general unknowable, and it is worth discussing these facts.
 
-Photo-$z$ PDFs, on the other hand, are posteriors, probabilities of a parameter, in this case redshift $z$, given the random variable derived from the parameter, in this case photometric data $\vec{d}$.
-Though they are commonly written as $p(z)$, a photo-$z$ PDF for galaxy $i$ in the catalog is truly a posterior $p(z_{i}|\vec{d}_{i})$ where $\vec{d}_{i}$ represents photometric data.
+Photo-$z$ PDFs are commonly written simply as $p(z)$, but this notation oversimplifies their substance.
+Whether determined by way of template-fitting or machine learning, photo-$z$ PDFs are dependent on the data $\vec{d}$ from which they are calculated.
+This data $\vec{d}$ must be considered to be a quantity upon which the redshift $z$ is conditioned, otherwise normalizing a photo-$z$ PDF would require integrating over all possible values of $\vec{d}$.
+This means that photo-$z$ PDFs are posteriors, probabilities of parameters $z$ conditioned on data $\vec{d}$.
+However, where there is a posterior, there is always a prior.
+Photo-$z$ PDFs are in general \textit{interim} posteriors, because in addition to being conditioned on observations, they are also conditioned on an interim prior in the form of a particular value $\vec{\theta}^{*}$ of $\vec{\theta}$ necessary for computing the photo-$z$ PDF.
+In the case of template-fitting methods, the interim prior is usually specified as an input to the photo-$z$ PDF production code, and it often takes the form of an initial guess for $\vec{\theta}$ based on the results of previous galaxy surveys or simulations thereof; in the case of machine learning methods, the interim prior may be explicitly derived from the training set or implicitly produced in the process of determining the photo-$z$ PDFs, and the interim prior is not always revealed to the user.
+Hierarchical inference may be performed only when the interim prior is known, which means it will only be possible for photo-$z$ PDFs made via certain methods.
 
-This approach makes several assumptions that must be explicitly stated.
+Since the data products from which we hope to probe the log-hyperposterior are themselves posteriors, we must express the likelihoods $\{p(\vec{d}_{i}|z_{i})\}$ of Eq.
+\ref{eq:log-hyperlikelihood} in terms of the photo-$z$ PDFs $\{p(z_{i}|\vec{d}_{i}, \vec{\theta}^{*})\}$.
+We start by multiplying the likelihood by an inspired factor of unity written in terms of the photo-$z$ interim posterior we have at hand, as in Eq.
+\ref{eq:unity}.
+\begin{align}
+\label{eq:unity}
+p(\vec{d}_{i}|z_{i}) &= p(\vec{d}_{i}|z_{i})\ \frac{p(z_{i}|\vec{d}_{i}, \vec{\theta}^{*})}{p(z_{i}|\vec{d}_{i}, \vec{\theta}^{*})}
+\end{align}
+Next, we expand the denominator in terms of Bayes' Rule to obtain Eq. \ref{eq:bayes}.
+\begin{align}
+\label{eq:bayes}
+p(\vec{d}_{i}|z_{i}) &= p(\vec{d}_{i}|z_{i})\ p(z_{i}|\vec{d}_{i}, \vec{\theta}^{*})\ \frac{p(\vec{d}_{i}|\vec{\theta}^{*})}{p(z_{i}|\vec{\theta}^{*})\ p(\vec{d}_{i}|z_{i}, \vec{\theta}^{*})}
+\end{align}
+Because the redshift $z_{i}$ is independent of the interim prior $\vec{\theta}^{*}$, the interim likelihood $p(\vec{d}_{i}|z_{i}, \vec{\theta}^{*})$ may be expanded further as in Eq.
+\ref{eq:expansion}.
+\begin{align}
+\label{eq:expansion}
+p(\vec{d}_{i}|z_{i}) &= p(\vec{d}_{i}|z_{i})\ p(z_{i}|\vec{d}_{i}, \vec{\theta}^{*})\ \frac{p(\vec{d}_{i}|\vec{\theta}^{*})}{p(z_{i}|\vec{\theta}^{*})\ p(\vec{d}_{i}|z_{i})\ p(\vec{d}_{i}|\vec{\theta}^{*})}
+\end{align}
+We cancel the terms $p(\vec{d}_{i}|z_{i})$ and $p(\vec{d}_{i}|\vec{\theta}^{*})$ that appear in both the numerator and denominator of Eq. \ref{eq:expansion} to obtain Eq. \ref{eq:cancellation}.
+\begin{align}
+\label{eq:cancellation}
+p(\vec{d}_{i}|z_{i}) &= \frac{p(z_{i}|\vec{d}_{i}, \vec{\theta}^{*})}{p(z_{i}|\vec{\theta}^{*})}
+\end{align}
+
+Finally we may return to Eq.
+\ref{eq:hyperbayes} to express the log-hyperposterior $\ln\left[p(\vec{\theta}|\{\vec{d}_{i}\})\right]$ in terms of the photo-$z$ interim posteriors $\{p(z_{i}|\vec{d}_{i}, \vec{\theta}^{*})\}$, as in Eq.
+\ref{eq:final}.
+\begin{align}
+\label{eq:final}
+\ln\left[p(\vec{\theta}|\{\vec{d}_{i}\})\right] &\propto \ln\left[p(\vec{\theta})\right] + \sum_{i}\ln\left[\int\ p(z_{i}|\vec{d}_{i}, \vec{\theta}^{*})\ \frac{p(z_{i}|\vec{\theta})}{p(z_{i}|\vec{\theta}^{*})}\ dz_{i}\right]
+\end{align}
+Now we have an expression for a quantity proportional to the log-hyperposterior we desire!
+Though the constant of proportionality, $p(\{\vec{d}_{i}\})$ is in general not possible to calculate, we may still characterize the log-hyperposterior by way of MCMC sampling.
+
+Several assumptions were noted in the above derivation; for clarity, they will be enumerated below, where their limitations will be discussed.
 \begin{enumerate}
-\item The PGM is an expression of our beliefs about the physics of the problem, and the inference will only be valid to the degree that the model is complete.
-\item We must choose a hyperprior $p(\vec{\theta})$; though we aim to choose a sufficiently general hyperprior that will not be a dominant source of information in the hyperposterior, it must still be assumed that whatever hyperprior is chosen will be accurate.
-\item We have assumed independence of galaxies in our catalog such that each galaxy's photometry $\vec{d}_{i}$ is independent from all other galaxies' photometry $\{\vec{d}_{i'\neq i}\}$ and all other galaxies' redshifts $\{z_{i'\neq i}\}$.
+\item The PGM is an expression of our beliefs about the physics of the problem, and the inference will only be valid to the degree that the model is correct.
+One can easily think of ways in which the PGM of Fig.
+\ref{fig:pgm} is incomplete; for example, intrinsic galaxy SEDs will evolve with redshift.
+However, we assume that physical processes not represented in the PGM of Fig.
+\ref{fig:pgm} are subdominant and may thus be neglected.
+\item We must choose a hyperprior $p(\vec{\theta})$; we assume that our choice of a sufficiently general hyperprior that will not be a dominant source of information in the hyperposterior is successful.
+If we choose poorly, the hyperprior may dominate over the hyperlikelihood, downweighting the significance of the data in determining the hyperposterior or biasing the result based on a misunderstanding of the underlying physics.
+\item We assume independence of galaxies in our catalog such that each galaxy's photometry $\vec{d}_{i}$ is independent from all other galaxies' photometry $\{\vec{d}_{i'\neq i}\}$ and all other galaxies' redshifts $\{z_{i'\neq i}\}$.
+However, the photometry $\{\vec{d}_{i}\}$ will inherently share instrumental and systematic effects due to being observed with the same telescope as part of a single survey project.
+Furthermore, there may be blended objects or other correlations between the redshifts of different galaxies.
+We must assume that such effects are negligible and will not consider them in this treatment.
+\item We assume that the interim prior is known to us, either as input to or output of the method producing the photo-$z$ interim posteriors.
+Furthermore, it must be representable in the parametrization of $f_{\vec{\theta}}(z)$, i.e.
+our model $f_{\vec{\theta}}(z)$ for $n(z)$ must be flexible enough to encompass both the true distribution we aim to probe and the interim prior.
+\item Finally, we assume that the photo-$z$ interim posteriors $\{p(z_{i}|\vec{d}_{i}, \vec{\theta}^{*})\}$ are accurate.
+There is some evidence that photo-$z$ PDFs derived by different methods disagree with one another or are strongly dependent on the interim prior.
+However, the choice of the technique by which the photo-$z$ interim posteriors are determined is outside the scope of this paper.
 \end{enumerate}
 
-\section{Alternative methods}
+\subsection{Model Specifics}
+\label{sec:choices}
+
+\subsection{Alternative methods}
 \label{sec:alternatives}

### 2017-06-06 research/scripts/multi_inf_script.py
updating all scripts

@@ -86,9 +86,12 @@ def set_up_prior(data, params):
 
     prior_mean = log_nz_intp
     prior = mvn(prior_mean, prior_var)
-    if params['prior_mean'] is 'sample':
-        prior_mean = prior.sample_one()
-        prior = mvn(prior_mean, prior_var)
+    if params['prior_mean'] == 'sample':
+        new_mean = prior.sample_one()
+        prior = mvn(new_mean, prior_var)
+        print(params['prior_mean'], prior_mean, new_mean)
+    else:
+        print(params['prior_mean'], prior_mean)
 
     return (prior, prior_var)
 
@@ -136,16 +139,13 @@ def do_inference(given_key):
     nz.plot_estimators()
     nz.write('nz.p')
 
-    # #start_mean = mvn(nz_mmle, cov).sample_one()
-    # start = prior#mvn(data['log_interim_prior'], cov)
-    #
     # n_bins = len(nz_mmle)
-    # if params['n_walkers'] is not None:
-    #     n_ivals = params['n_walkers']
-    # else:
-    #     n_ivals = 10 * n_bins
-    # initial_values = start.sample(n_ivals)
-    #
+    if params['n_walkers'] is not None:
+        n_ivals = params['n_walkers']
+    else:
+        n_ivals = 10 * n_bins
+    initial_values = prior.sample(n_ivals)
+    log_z_dens_plots.plot_ivals(initial_values, nz.info, nz.plot_dir)
     # nz_samps = nz.calculate_samples(initial_values, no_data=params['no_data'], no_prior=params['no_prior'])
     #
     # nz_stats = nz.compare()

### 2017-06-06 research/paper/draft.tex

@@
 Now we have an expression for a quantity proportional to the log-hyperposterior we desire!
-Though the constant of proportionality, $p(\{\vec{d}_{i}\})$ is in general not possible to calculate, we may still characterize the log-hyperposterior by way of MCMC sampling.
+Though the constant of proportionality, $p(\{\vec{d}_{i}\})$ is in general not possible to calculate, we may still characterize the log-hyperposterior by way of MCMC sampling and optimization, which will be discussed in Sec.
+\ref{sec:implementation}.
 
@@
 
-\subsection{Model Specifics}
-\label{sec:choices}
-
 \subsection{Alternative methods}
 \label{sec:alternatives}
+
+This method is presented in contrast to existing methods for inferring $n(z)$ from photo-$z$ PDFs, which fall into two broad categories: stacking and point estimation.
+Stacking assumes Eq.
+\ref{eq:stacking}, where the constant of proportionality is chosen to ensure $\int f_{\vec{\theta}}(z)\ dz$ is unity.
+\begin{align}
+\label{eq:stacking}
+f_{\vec{\theta}}(z) &\propto \sum_{i}p(z_{i}|\vec{d}_{i}, \vec{\theta}^{*})
+\end{align}
+
+\section{Implementation}
+\label{sec:implementation}
+
+We introduce the Cosmological Hierarchical Inference with Probabilistic Photometric Redshifts (CHIPPR) code as a comprehensive package for estimating $n(z)$ from catalogs of photo-$z$ PDFs.
+In particular, CHIPPR implements both sampling and optimization of Eq.
+\ref{eq:final} and compares the results to those of the more traditional approaches presented in Sec.
+\ref{sec:alternatives}.
 

### 2017-06-07 research/paper/draft.tex
reviving the paper, starting with an outline

--- 2017-02-28 research/paper/draft.tex
@@
-%\documentclass[iop]{emulateapj}
 \documentclass[preprint]{aastex}
-%\documentclass[12pt, onecolumn]{emulateapj}
-%\documentstyle[aas2pp4,natbib209]{article}
 
@@
 \newcommand{\textul}{\underline}
+\newcommand{\chippr}{\texttt{CHIPPR} }
 
-\shorttitle{How to obtain the redshift distribution from probabilistic redshift estimates}
+\shorttitle{How to obtain the redshift distribution from probabilistic redshift 
+estimates}
 \shortauthors{Malz and Hogg}
@@
 
-\title{How to obtain the redshift distribution from probabilistic redshift estimates}
+\title{How to obtain the redshift distribution from probabilistic redshift 
+estimates}
 
@@
 
-\altaffiltext{1}{Center for Cosmology and Particle Physics, Department of Physics,
+\altaffiltext{1}{Center for Cosmology and Particle Physics, Department of 
+Physics,
   New York University, 726 Broadway, 9th floor, New York, NY 10003, USA}
-\altaffiltext{2}{Simons Center for Computational Astrophysics, 162 Fifth Avenue, 7th floor, New York, NY 10010, USA}
-\altaffiltext{3}{Center for Data Science, New York University, 60 Fifth Avenue, 7th floor, New York, NY 10003, USA}
-\altaffiltext{4}{Max-Planck-Institut f\"ur Astronomie, K\"onigstuhl 17, D-69117 Heidelberg, Germany}
+\altaffiltext{2}{Simons Center for Computational Astrophysics, 162 Fifth 
+Avenue, 7th floor, New York, NY 10010, USA}
+\altaffiltext{3}{Center for Data Science, New York University, 60 Fifth Avenue, 
+7th floor, New York, NY 10003, USA}
+\altaffiltext{4}{Max-Planck-Institut f\"ur Astronomie, K\"onigstuhl 17, D-69117 
+Heidelberg, Germany}
 
 \begin{abstract}
-The redshift distribution $n(z)$ is a crucial ingredient for weak lensing cosmology.
-Spectroscopically confirmed redshifts for the dim and numerous galaxies observed by weak lensing surveys are expected to be inaccessible, making photometric redshifts (photo-$z$s) the next best alternative.
-Because of the nontrivial inference involved in their determination, photo-$z$ point estimates are being superseded by photo-$z$ probability distribution functions (PDFs).
-However, analytic methods for utilizing these new data products in cosmological inference are still evolving.
+The redshift distribution $n(z)$ is a crucial ingredient for weak lensing 
+cosmology.  Spectroscopically confirmed redshifts for the dim and numerous 
+galaxies observed by weak lensing surveys are expected to be inaccessible, 
+making photometric redshifts (photo-$z$s) the next best alternative.  Because 
+of the nontrivial inference involved in their determination, photo-$z$ point 
+estimates are being superseded by photo-$z$ probability distribution functions 
+(PDFs).  However, analytic methods for utilizing these new data products in 
+cosmological inference are still evolving.
 
-This paper presents a novel approach to estimating the posterior distribution over $n(z)$ from a survey of galaxy photo-$z$ PDFs based upon a probabilistic graphical model of hierarchical inference.
-We present the Cosmological Hierarchical Inference with Probabilistic Photometric Redshifts (CHIPPR) code implementing this technique, as well as its validation on mock data and testing on a subset of BOSS DR10.
-CHIPPR yields an accurate characterization of $n(z)$ containing information beyond the best-fit estimator produced by traditional procedures.
-The publicly available code is easily extensible to other one-point statistics that depend on redshift.
+This paper presents a novel approach to estimating the posterior distribution 
+over $n(z)$ from a survey of galaxy photo-$z$ PDFs based upon a probabilistic 
+graphical model of hierarchical inference.  We present the Cosmological 
+Hierarchical Inference with Probabilistic Photometric Redshifts (\chippr) code 
+implementing this technique, as well as its validation on mock data and testing 
+on the \texttt{Buzzard} simulations.  \chippr yields a more accurate 
+characterization of $n(z)$ containing information beyond the best-fit estimator 
+produced by traditional procedures.  The publicly available code is easily 
+extensible to other one-point statistics that depend on redshift.
 
@@
 
-\keywords{catalogs --- cosmology: cosmological parameters --- galaxies: statistics --- gravitational lensing: weak --- methods: analytical --- methods: data analysis --- methods: statistical --- techniques: photometric}
+\keywords{catalogs --- cosmology: cosmological parameters --- galaxies: 
+statistics --- gravitational lensing: weak --- methods: analytical --- methods: 
+data analysis --- methods: statistical --- techniques: photometric}
 
@@
 
-The redshift distribution $n(z)$ is necessary for calculating two-point statistics of galaxy properties used to determine the cosmological parameter values that inform our understanding of the evolution of large-scale structure in the universe \citep{masters_mapping_2015}.
-Inaccurate estimates of $n(z)$ can significantly impact the constraining power of a galaxy survey, biasing recovery of the cosmological parameters.
-For example, if the $n(z)$ used in analysis is offset from the true $n(z)$ by a small, negative constant, it results in underestimating $w_{0}$ and overestimating $\sigma_{8}$ \citep{samuroff_simultaneous_2017}.
+\chippr aims to answer the following questions:
 
-Though the redshift density function has traditionally been determined from spectroscopically observed redshifts, modern galaxy surveys including DES, LSST, Euclid, and WFIRST seek to obtain two-point statistics of redshift from samples of galaxies for which spectroscopic redshifts are unavailable, either due to their large numbers or their low brightnesses.
-For decades, photometrically estimated redshifts (photo-$z$s) have been the leading alternative to spectroscopically observed redshifts, though they suffer from issues of precision in the form of an intrinsic scatter and accuracy in the form of catastrophic outliers, as well as other systematics imparted by the properties of the survey, data reduction pipeline(s), and assumptions underlying the analysis \citep{baum_photoelectric_1962}.
-These weaknesses are illuminated by a probabilistic interpretation of photo-$zs$; if these nontrivial uncertainties were expressed as a probability distribution function (PDF) over redshift, photo-$z$s could be replaced by photo-$z$ PDFs containing more information than a simple point estimate \citep{koo_overview_1999}.
-Such data products have been commonly released by photometric galaxy surveys since SDSS DR7 \citep{abazajian_seventh_2009} using a great variety of methods.
-
-Methods for using photo-$z$ PDFs in cosmological inference remain underdeveloped, with many survey pipelines reducing them to familiar point estimators that are compatible with existing technology or engaging with them heuristically in a manner inconsistent with their probabilistic nature.
-Stacking photo-$z$ PDFs, or other mathematically equivalent methods, to obtain an estimator of $n(z)$ is especially popular \citep{cunha_estimating_2009, sheldon_photometric_2012}.
-However, photo-$z$ PDFs are probabilistic data products that must be handled in a fully consistent, probabilistic manner; hierarchical inference is the only mathematically valid way to do inference with photo-$z$ PDFs and other probabilistic quantities.
-
-It is desirable to create rigorous methods for using photo-$z$ PDFs in inference, beginning with the simplest one-point statistic of redshift, $n(z)$.
-In the spirit of \citet{hogg_inferring_2010, foreman-mackey_exoplanet_2014}, this paper derives a mathematically rigorous approach to inferring $n(z)$ in Sec.
-\ref{sec:method}.
-The presentation of a public code implementing this novel technique is given in Sec.
-\ref{sec:model_specifics}.
-The code is validated on mock data and tested on a subset of BOSS DR10 data in Sec.
-\ref{sec:experiments}.
-We conclude and discuss future directions in Sec.
-\ref{sec:discussion}.
+\begin{itemize}
+	\item How are photo-$z$ PDFs currently used in cosmology?
+	\item Why should we challenge existing methods?
+	\item How should photo-$z$ PDFs be used in inference?
+	\item How does the result of \chippr compare to established estimators 
+of $n(z)$?
+\end{itemize}
 
@@
 
-The redshift distribution $n(z)$ may be understood as the probability of finding a galaxy at a redshift $z$.
-We may express $n(z)$ in terms of some functional form $f_{\vec{\theta}}(z)$ with parameters comprising $\vec{\theta}$ according to Eq.
-\ref{eq:parametrization}.
-\begin{align}
-\label{eq:parametrization}
-n(z) &\equiv f_{\vec{\theta}}(z)\ \equiv\ p(z|\vec{\theta})
-\end{align}
-In this work, we wish to characterize the posterior distribution $p(\vec{\theta}|D)$ of the parameters contained in $\vec{\theta}$ defining $n(z)$ given all available data $D$.
-The likelihood $p(z|\vec{\theta})$ represents the probability of a random variable, in this case redshift $z$, given the parameters in $\vec{\theta}$ defining the distribution $n(z)$ from which it is drawn.
-Our model for inferring $p(\vec{\theta}|D)$ will be introduced in Sec.
-\ref{sec:derivation}, and alternative methods of estimating $\vec{\theta}$ will be presented in Sec.
-\ref{sec:alternatives}.
+\subsection{Model}
+\label{sec:model}
 
-\subsection{Model Specifics}
-\label{sec:derivation}
+This is where I will present the DAG and corresponding math, along with an 
+explicit discussion of the assumptions that must be satisfied in order for the 
+\chippr approach to be valid.
 
-Inference of $p(\vec{\theta}|D)$ is perfectly suited to a hierarchical Bayesian model.
-In this problem, the available data $D$ is a catalog of photometry $\{\vec{d}_{i}\}$ of individual galaxies $i$ where each $\vec{d}_{i}$ is a random variable that is a function of its redshift parameter $z_{i}$; the redshift parameters $\{z_{i}\}$ are independently drawn from a function defined by hyperparameters in $\vec{\theta}$.
-These physical relationships may be illustrated in Fig.
-\ref{fig:pgm} by a directed acyclic graph representing a probabilistic graphical model (PGM).
-\begin{figure}
-\begin{center}
-\includegraphics[width=0.5\textwidth]{pgm.png}
-\caption{This directed acyclic graph corresponds to a PGM for a hierarchical inference of $p(\vec{\theta}|\{\vec{d}_{i}\})$.
-In this graph, all random variables are shown in circles, with observed variables shown in shaded circles.
-Relationships between variables are indicated by arrows.
-The box indicates that there are a number of copies of the relationships between boxed parameters, each independent of all others.
-The hyperparameters $\vec{\theta}$ representing $n(z)$ are at the top.
-Independently drawn from a function of the hyperparameters $\vec{\theta}$ are galaxy redshifts $\{z_{i}\}$ below.
-The observed galaxy photometry $\{\vec{d}_{i}\}$, shown in shaded circles, is determined by the redshifts above.}
-\label{fig:pgm}
-\end{center}
-\end{figure}
-This PGM may be used as the basis for a derivation of the desired hyperposterior $p(\vec{\theta}|\{\vec{d}_{i}\})$, assuming our physical model is complete.
-From this point on, we will work solely with log probabilities.
+\subsection{Implementation}
+\label{sec:implementation}
 
-We begin with Eq.
-\ref{eq:bayes} by applying Bayes' Rule to express the log-hyperposterior $\ln\left[p(\vec{\theta}|\{\vec{d}_{i}\})\right]$ in terms of the log-hyperlikelihood $\ln\left[p(\{\vec{d}_{i}\}|\vec{\theta})\right]$.
-\begin{align}
-\label{eq:bayes}
-\ln\left[p(\vec{\theta}|\{\vec{d}_{i}\})\right] &\propto \ln\left[p(\vec{\theta})\right] + \ln\left[p(\{\vec{d}_{i}\}|\vec{\theta})\right]
-\end{align}
-To do this, we will choose a log-hyperprior distribution $\ln\left[p(\vec{\theta})\right]$ representing our beliefs about the distribution of the hyperparameters $\vec{\theta}$ in terms of fixed variables that we will assume are known.
+I will describe the \chippr code here.
 
-The log-hyperlikelihood $\ln\left[p(\{\vec{d}_{i}\}|\vec{\theta})\right]$ contains no explicit reference to redshifts, therefore, we employ the PGM to write the marginalization over the redshifts in Eq.
-\ref{eq:marginalization}.
-\begin{align}
-\label{eq:marginalization}
-\ln\left[p(\{\vec{d}_{i}\}|\vec{\theta})\right] &= \ln\left[\int\ p(\{\vec{d}_{i}\}|\{z_{i}\})\ p(\{z_{i}\}|\vec{\theta})\ d\{z_{i}\}\right]
-\end{align}
-We shall handle the two terms in the integral separately.
+\subsection{Metrics}
+\label{sec:metrics}
 
-The first term is the likelihood of all galaxy photometry given all galaxy redshifts.
-If we assume independence of galaxy photometry, such that $p(\vec{d}_{i}|\{\vec{d}_{i'\neq i}\}, \{z_{i}\})=p(\vec{d}_{i}|z_{i})$, then we may write Eq.
-\ref{eq:independence}.
-\begin{align}
-\label{eq:independence}
-\ln\left[p(\{\vec{d}_{i}\}|\{z_{i}\})\right] &= \sum_{i}\ln\left[p(\vec{d}_{i}|z_{i})\right]
-\end{align}
-The second term may also be expanded by assuming that all galaxy redshifts are drawn independently from $n(z)$.
-\begin{align}
-\label{eq:poisson}
-\ln\left[p(\{z_{i}\}|\vec{\theta})\right] &= \sum_{i}\ln\left[p(z_{i}|\vec{\theta})\right]
-\end{align}
+I will discuss the metrics that I use in \ref{sec:validation} and include in 
+the release version.
 
-Photo-$z$ PDFs, on the other hand, are posteriors, probabilities of a parameter, in this case redshift $z$, given the random variable derived from the parameter, in this case photometric data $\vec{d}$.
-Though they are commonly written as $p(z)$, a photo-$z$ PDF for galaxy $i$ in the catalog is truly a posterior $p(z_{i}|\vec{d}_{i})$ where $\vec{d}_{i}$ represents photometric data.
+\section{Validation}
+\label{sec:validation}
 
-This approach makes several assumptions that must be explicitly stated.
-\begin{enumerate}
-\item The PGM is an expression of our beliefs about the physics of the problem, and the inference will only be valid to the degree that the model is complete.
-\item We must choose a hyperprior $p(\vec{\theta})$; though we aim to choose a sufficiently general hyperprior that will not be a dominant source of information in the hyperposterior, it must still be assumed that whatever hyperprior is chosen will be accurate.
-\item We have assumed independence of galaxies in our catalog such that each galaxy's photometry $\vec{d}_{i}$ is independent from all other galaxies' photometry $\{\vec{d}_{i'\neq i}\}$ and all other galaxies' redshifts $\{z_{i'\neq i}\}$.
-\end{enumerate}
+This is where the setup and results of the simulated test cases will be 
+presented.
 
-\section{Alternative methods}
-\label{sec:alternatives}
+\subsection{Variations on Photo-$z$ Likelihoods}
+\label{sec:likelihoods}
 
-\section{Experiments}
-\label{sec:experiments}
+I think this is too many test cases!
 
-\section{Discussion}
-\label{sec:discussion}
+\begin{itemize}
+	\item A fiducial case of Gaussian likelihoods with narrow, 
+redshift-independent intrinsic scatter and no other systematics; also the 
+fiducial case with larger intrinsic scatter
+	\item The fiducial case with catastrophic outliers as seen in 
+template-based photo-$z$ estimators (attractor in the space of $z_{phot}$: some 
+galaxies at a range of $z_{spec}$ map to the same $z_{phot}$ because their 
+shared SED type does not have sufficiently strong features, leading galaxies of 
+that type at many $z_{spec}$ to have the same colors)
+	\item The fiducial case with catastrophic outliers as seen in 
+training-based photo-$z$ estimators (attractor in the space of $z_{spec}$: some 
+galaxies at a particular $z_{spec}$ map to a range of $z_{phot}$ because their 
+shared SED features fall between filters, leading to many galaxies at 
+particular $z_{spec}$ to have the same colors)
+\end{itemize}
 
-%\appendix{}
+\subsection{Variations on Interim Priors}
+\label{sec:priors}
+
+\subsection{Variations on True Hyperparameters}
+\label{sec:truth}
+
+\section{Application}
+\label{sec:application}
+
+I will apply \chippr to a realistic simulated dataset (probably 
+\texttt{Buzzard}) so the results can be compared to the truth.  The photo-$z$ 
+PDFs should be derived using the \texttt{SDSS}-approved $k$ nearest neighbors 
+algorithm because it has the most straightforward interim prior.
+
+\section{Conclusion}
+\label{sec:conclusion}
+
+I will review the questions from the introduction:
+
+\begin{itemize}
+	\item Existing methods have shortcomings that propagate to inaccuracies 
+in characterizing the cosmological parameters.\\
+	\item Photo-$z$ PDFs are probabilistic data products so must be handled 
+in a mathematically consistent manner; \chippr is one such method, conditioned 
+on some assumptions.
+	\item In addition to coming with its own error distribution, the $n(z)$ 
+estimator produced by \chippr is more accurate than established estimators; 
+furthermore, propagation of the \chippr result leads to a quantifiable 
+improvement in the constraints on cosmological parameters.
+\end{itemize}
+
+I may briefly discuss some obvious extensions of \chippr, such as a fully 
+hierarchical inference of the cosmological parameters that avoids tomographic 
+binning based on photo-$z$ point estimates.
 
 \begin{acknowledgements}
-AIM thanks Phil Marshall for advising on the production of usable code, Mohammadjavad Vakili for insightful input on statistics, Geoffrey Ryan for assistance in debugging code, and Boris Leistedt for helpful comments provided in the preparation of this paper.
+AIM thanks Phil Marshall for significant coaching on code development and paper 
+preparation, Mohammadjavad Vakili for insightful input on statistics, Geoffrey 
+Ryan for assistance in debugging, and Boris Leistedt for helpful comments 
+provided in the preparation of this paper.
 \end{acknowledgements}

### 2017-06-07 research/paper/cleaner.py

@@ -1,4 +1,5 @@
 import os, re
+#import subprocess
 #import argparse
 def clean(fileName):
 	remove_keys = ['abstract', 'annote', 'isbn', 'mendeley-groups', 'keywords', 'file', 'issn']
@@ -15,5 +16,16 @@ def clean(fileName):
 	cleanf.close()
 	os.remove(fileName)
 	os.rename(fileName + '.tmp', fileName)
+def readable(infile, outfile):
+	intfile = 'temp.tex'
+	os.system('sed \'/^%/ d\' < ' + infile + ' > ' + intfile)
+	os.system('fold -w80 -s ' + intfile + ' > ' + outfile)
 
 clean('references.bib')
+
+cwd = os.getcwd()
+infile = os.path.join(cwd, 'draft-editing.tex')
+outfile = os.path.join(cwd, 'draft.tex')
+print(infile, outfile)
+readable(infile, outfile)
+

### 2017-06-09 research/paper/draft.tex
wrapping up the outline

@@
 
+I will show that a really wacky interim prior, something very far from the true 
+$n(z)$, will imprint itself on the stacked estimator but will have no effect on 
+\chippr's result.
+
 \subsection{Variations on True Hyperparameters}
 \label{sec:truth}
+
+I will show that if $n(z)$ has significant features, stacking will not recover 
+them but \chippr will.
 
@@
 	\item Existing methods have shortcomings that propagate to inaccuracies 
-in characterizing the cosmological parameters.\\
+in characterizing the cosmological parameters.
 	\item Photo-$z$ PDFs are probabilistic data products so must be handled 

### 2017-06-10 research/paper/draft.tex
incorporating feedback

--- 2017-06-07 research/paper/draft.tex
@@
-\documentclass[preprint]{aastex}
+\documentclass[iop]{emulateapj}
 
@@
 estimates}
-\shortauthors{Malz and Hogg}
+\shortauthors{Malz, et al.}
 
@@
 
-\author{Alex Malz\altaffilmark{1} \& David W. Hogg\altaffilmark{1,2,3,4}}
+\author{Alex Malz\altaffilmark{1}, David W. Hogg\altaffilmark{1,2,3,4}, Phil 
+Marshall\altaffilmark{5}, \& Others}
 \email{aimalz@nyu.edu}
@@
 Heidelberg, Germany}
+\altaffiltext{5}{[SLAC]}
 
@@
 (PDFs).  However, analytic methods for utilizing these new data products in 
-cosmological inference are still evolving.
-
-This paper presents a novel approach to estimating the posterior distribution 
-over $n(z)$ from a survey of galaxy photo-$z$ PDFs based upon a probabilistic 
-graphical model of hierarchical inference.  We present the Cosmological 
-Hierarchical Inference with Probabilistic Photometric Redshifts (\chippr) code 
-implementing this technique, as well as its validation on mock data and testing 
-on the \texttt{Buzzard} simulations.  \chippr yields a more accurate 
-characterization of $n(z)$ containing information beyond the best-fit estimator 
-produced by traditional procedures.  The publicly available code is easily 
-extensible to other one-point statistics that depend on redshift.
+cosmological inference are still evolving.  This paper presents a novel 
+approach to estimating the posterior distribution over $n(z)$ from a survey of 
+galaxy photo-$z$ PDFs based upon a probabilistic graphical model of 
+hierarchical inference.  We present the Cosmological Hierarchical Inference 
+with Probabilistic Photometric Redshifts (\chippr) code implementing this 
+technique, as well as its validation on mock data and testing on the 
+\texttt{Buzzard} simulations.  \chippr yields a more accurate characterization 
+of $n(z)$ containing information beyond the best-fit estimator produced by 
+traditional procedures.  The publicly available code is easily extensible to 
+other one-point statistics that depend on redshift.
 
@@
 
-\chippr aims to answer the following questions:
+After a brief literature review addressing how photo-$z$ PDFs are currently 
+used in cosmology, this paper aims to answer the following questions:
 
 \begin{itemize}
-	\item How are photo-$z$ PDFs currently used in cosmology?
-	\item Why should we challenge existing methods?
-	\item How should photo-$z$ PDFs be used in inference?
+	\item Why should we question existing methods?
+	\item How can we improve the effectiveness of using photo-$z$ PDFs in 
+inference?
 	\item How does the result of \chippr compare to established estimators 
-of $n(z)$?
+in terms of the accuracy of $n(z)$?
+	\item How significant is the effect of the discrepancy between $n(z)$ 
+estimators on cosmological constraints?
 \end{itemize}
@@
 
+This paper presents a mathematically consistent method for obtaining the 
+posterior distribution over the redshift density function $n(z)$ using a 
+catalog of photo-$z$ PDFs.
+
 \subsection{Model}
@@
 
-This is where I will present the DAG and corresponding math, along with an 
-explicit discussion of the assumptions that must be satisfied in order for the 
-\chippr approach to be valid.
+The directed acyclic graph of Fig. \ref{fig:pgm} represents a probabilistic 
+graphical model for hierarchical inference of $n(z)$, the mathematical 
+interpretation of which will be presented below.
+
+\begin{figure}
+	\begin{center}
+		\includegraphics[width=0.5\textwidth]{pgm.png}
+		\caption{\textbf{Replace \texttt{daft} version with more 
+polished \texttt{tikz} version.}}
+	\label{fig:pgm}
+	\end{center}
+\end{figure}
+
+\textbf{Math from \texttt{prob-z} version will go here.}
+
+This framework entails a number of choices and assumptions that must be 
+addressed explicitly.
+
+\begin{enumerate}
+	\item While we advocate for the approach of hierarchical inference, the 
+probabilistic graphical model presented here is not the only one that could be 
+proposed.
+\end{enumerate}
+
+\subsection{Alternative Approaches}
+\label{sec:others}
+
+It is useful to translate some popular existing methods for deriving $n(z)$ 
+from photo-$z$ PDFs into the mathematical framework of Sec. \ref{sec:model}.
 
@@
 
-I will describe the \chippr code here.
+The publicly available \chippr code implements the probabilistic graphical 
+model presented in Sec. \ref{sec:model}.
 
-\subsection{Metrics}
-\label{sec:metrics}
+In addition to the choices and assumptions underlying the probabilistic 
+graphical model, the implementation of \chippr makes choices and assumptions of 
+its own.
 
-I will discuss the metrics that I use in \ref{sec:validation} and include in 
-the release version.
+\begin{enumerate}
+	\item \chippr currently only accepts photo-$z$ PDFs and produces $n(z)$ 
+samples of a single format, that of the piecewise constant parametrization, 
+also referred to as a binned histogram parametrization and a sum of top hat 
+functions.
+\end{enumerate}
 
-\section{Validation}
+\section{Validation on Simple Mock Data}
 \label{sec:validation}
 
-This is where the setup and results of the simulated test cases will be 
-presented.
+The superiority of \chippr over alternative approaches is demonstrated in a 
+number of compelling test cases on mock data.  Each experiment is characterized 
+by a single change to a fiducial case in order to isolate the influence of 
+systematic effects known to be relevant to photo-$z$ estimation and propagation 
+in analysis.
 
-\subsection{Variations on Photo-$z$ Likelihoods}
+\subsection{Mock Data \& Metrics}
+\label{sec:validintro}
+
+The mock data in these tests consists of photo-$z$ interim posteriors rather 
+than photometric data because the various existing methods for deriving 
+photo-$z$ interim posteriors do not in general yield results that are 
+consistent with one another, indicating that their systematics are not 
+well-understood.  Because the mock data is independent of any choice of 
+photo-$z$ PDF production method, we not only ensure that our photo-$z$ interim 
+priors are perfectly understood but also deter readers from assuming that 
+\chippr has any preference over the method by which photo-$z$ interim 
+posteriors are derived from photometric data.
+
+\subsubsection{Mock Data}
+\label{sec:mockdata}
+
+The mock data used here are produced by the following steps.  This method for 
+deriving mock data is referred to as the fiducial case, and variations on it 
+will refer directly to the steps that are altered.
+
+\subsubsection{$n(z)$ Accuracy Metric}
+\label{sec:accuracy}
+
+The Kullback-Leibler divergence will be used as the primary measure of the 
+accuracy of estimators of $n(z)$ in cases of mock data with known true 
+redshifts.
+
+\textbf{Review precision and bias from \texttt{kld.ipynb} and interpret in 
+terms of a \% difference.}
+
+\subsection{Emulated Data Quality Effects}
 \label{sec:likelihoods}
 
-I think this is too many test cases!
+In the following test cases, the properties of the mock photo-$z$ likelihoods 
+are varied in an effort to emulate known systematics in photo-$z$ estimation.
 
-\begin{itemize}
-	\item A fiducial case of Gaussian likelihoods with narrow, 
-redshift-independent intrinsic scatter and no other systematics; also the 
-fiducial case with larger intrinsic scatter
-	\item The fiducial case with catastrophic outliers as seen in 
-template-based photo-$z$ estimators (attractor in the space of $z_{phot}$: some 
-galaxies at a range of $z_{spec}$ map to the same $z_{phot}$ because their 
-shared SED type does not have sufficiently strong features, leading galaxies of 
-that type at many $z_{spec}$ to have the same colors)
-	\item The fiducial case with catastrophic outliers as seen in 
-training-based photo-$z$ estimators (attractor in the space of $z_{spec}$: some 
-galaxies at a particular $z_{spec}$ map to a range of $z_{phot}$ because their 
-shared SED features fall between filters, leading to many galaxies at 
-particular $z_{spec}$ to have the same colors)
-\end{itemize}
+\subsubsection{Intrinsic Scatter}
+\label{sec:intscat}
 
-\subsection{Variations on Interim Priors}
+One major concern about photo-$z$s is the intrinsic scatter of point 
+estimators, including those derived from photo-$z$ PDFs, that is observed to 
+varying extents with every existing photo-$z$ algorithm and illustrated in Fig. 
+\ref{fig:intscat}.
+
+\begin{figure}
+	\begin{center}
+		%\includegraphics[width=0.5\textwidth]{pgm.png}
+		\caption{[plot of $z_{phot}$ vs. $z_{spec}$ with varying 
+intrinsic scatter]}
+		\label{fig:intscat}
+	\end{center}
+\end{figure}
+
+To emulate intrinsic scatter, we modify the fiducial case to simply broaden the 
+single Gaussian component of the likelihood.  To enforce self-consistency, the 
+mean is drawn from a Gaussian distribution with the newly increased variance.
+
+\subsubsection{Template-like Catastrophic Outliers}
+\label{sec:tempcatout}
+
+In addition to intrinsic scatter, photo-$z$ methods employing template fitting 
+tend to produce catastrophic outliers that are distributed to be broad in 
+$z_{spec}$ and narrow in $z_{phot}$, as in Fig. \ref{fig:tempfail}.  The 
+systematic behind these catastrophic outliers may be described as an attractor 
+in the space of $z_{phot}$; some galaxies at a range of $z_{spec}$ map onto a 
+single $z_{phot}$ (with some scatter) if their true SED does not have 
+sufficiently strong features (as is the case for blue galaxies), leading 
+galaxies of that type at many $z_{spec}$ to have similar colors.
+
+\begin{figure}
+	\begin{center}
+		%\includegraphics[width=0.5\textwidth]{pgm.png}
+		\caption{[plot of $z_{phot}$ vs. $z_{spec}$ with template-like 
+catastrophic outliers]}
+		\label{fig:tempcatout}
+	\end{center}
+\end{figure}
+
+\subsubsection{Training-like Catastrophic Outliers}
+\label{sec:traincatout}
+
+Data driven photo-$z$ methods tend to suffer from a different form of 
+catastrophic outliers that are distributed to be narrow in $z_{spec}$ and broad 
+in $z_{phot}$, as in Fig. \ref{fig:trainfail}.  The systematic behind these 
+catastrophic outliers may be described as an attractor in the space of 
+$z_{spec}$; some galaxies near a particular $z_{spec}$ map to a range of 
+$z_{phot}$ if their true SED's features fall between photometric filters, 
+leading to many galaxies near that $z_{spec}$ to have similar colors.
+
+\begin{figure}
+	\begin{center}
+		%\includegraphics[width=0.5\textwidth]{pgm.png}
+		\caption{[plot of $z_{phot}$ vs. $z_{spec}$ with training-like 
+catastrophic outliers]}
+		\label{fig:traincatout}
+	\end{center}
+\end{figure}
+
+\subsection{Emulated Interim Prior Effects}
 \label{sec:priors}
 
-\subsection{Variations on True Hyperparameters}
+The interim prior encapsulates the information about the relationship between 
+observed photometry and redshift upon which a photo-$z$ estimate is based.  
+Interim priors are in general not identical to the true $n(z)$ we wish to 
+estimate; if they were, we would not need any data!  For template fitting 
+photo-$z$ methods, the interim prior is usually an input chosen by the 
+researcher.  However, for many machine learning methods, the interim prior is 
+some function of the training set data that may be influenced by random numbers 
+and is rarely output with the redshift estimates.  Interim priors for template 
+fitting methods tend to have incomplete coverage in the space of true 
+photometry, because they are limited by the choice of the library of SEDs.  
+Interim priors for machine learning methods tend to have incomplete coverage in 
+the space of redshifts, because there are fewer galaxies with spectroscopically 
+confirmed redshifts at high redshifts than low redshifts.
+
+Existing $n(z)$ estimation routines will always produce a biased estimator when 
+the interim prior is not equal to the true $n(z)$.  We demonstrate here that 
+regardless of the appropriateness of the interim prior as an approximation to 
+the true $n(z)$, \chippr is not affected by the choice of the interim prior so 
+long as it has nontrivial coverage.
+
+\subsubsection{Template-like Interim Prior}
+\label{sec:tempintpr}
+
+An interim prior based on a template library may be a sum of smooth functions 
+representing $n(z)$ for each SED type in the library.  Template libraries do 
+not include every possible galaxy SED, and the $n(z)$ used for each SED type 
+may not be accurate.  The interim prior shown in Fig. \ref{fig:tempintpr} is an 
+emulation of an interim prior corresponding to a template library of this type.
+
+\begin{figure}
+	\begin{center}
+		%\includegraphics[width=0.5\textwidth]{pgm.png}
+		\caption{[plot of template-like interim prior]}
+		\label{fig:tempintpr}
+	\end{center}
+\end{figure}
+
+\subsubsection{Training-like Interim Prior}
+\label{sec:trainintpr}
+
+An interim prior based on a training set may be biased toward low redshifts due 
+to the dearth of distant galaxies with spectroscopic redshifts.  The interim 
+prior shown in Fig. \ref{fig:trainintpr} is an emulation of an interim prior 
+corresponding to a training set biased in this way.
+
+\begin{figure}
+	\begin{center}
+		%\includegraphics[width=0.5\textwidth]{pgm.png}
+		\caption{[plot of training-like interim prior]}
+		\label{fig:trainintpr}
+	\end{center}
+\end{figure}
+
+\subsection{Underlying $n(z)$ Effects}
 \label{sec:truth}
 
-\section{Application}
+Existing $n(z)$ estimators are systematically smoother than the true $n(z)$.  
+Here we show that the traditional estimators of $n(z)$ cannot recover a highly 
+featured $n(z)$.  The implication of this failure is quite serious; the 
+consistently smooth, unimodal $n(z)$ estimates appearing in the literature 
+could result from much more featured true redshift distributions, and there 
+would be no way to catch this error without using a fully probabilistic method.
+
+\subsubsection{Featured $n(z)$}
+\label{sec:featured}
+
+In this test, we choose the true $n(z)$ of Fig. \ref{fig:featured}, which has 
+nontrivial structure.
+
+\begin{figure}
+	\begin{center}
+		%\includegraphics[width=0.5\textwidth]{pgm.png}
+		\caption{[plot of a highly featured true $n(z)$]}
+		\label{fig:featured}
+	\end{center}
+\end{figure}
+
+\section{Application to Realistic Mock Data}
 \label{sec:application}
 
-I will apply \chippr to a realistic simulated dataset (probably 
-\texttt{Buzzard}) so the results can be compared to the truth.  The photo-$z$ 
-PDFs should be derived using the \texttt{SDSS}-approved $k$ nearest neighbors 
-algorithm because it has the most straightforward interim prior.
+To show how the choice of $n(z)$ estimator propagates to cosmological 
+constraints, we apply \chippr to a data from a realistic cosmological 
+simulation (probably \texttt{Buzzard}).
+
+\subsection{Mock Data \& Metrics}
+\label{sec:appintro}
+
+As in Sec. \ref{sec:validintro}, the mock data takes the form of photo-$z$ 
+interim posteriors, but that is where the similarity ends.  These photo-$z$ 
+interim posteriors are derived from the photometry resulting from the 
+\texttt{Buzzard} simulation by way of [method used to make the interim 
+posteriors].  Because the simulation begins with setting true values of the 
+cosmological parameters, the different estimators of $n(z)$ are propagated 
+through a forecasting code to produce error ellipses on the cosmological 
+parameters.
+
+\subsubsection{Mock Data}
+\label{sec:buzzard}
+
+The details of the \texttt{Buzzard} simulation will be summarized here.
+
+\subsubsection{Cosmological Constraint Metric}
+\label{sec:cosmo}
+
+Because the mock data is associated with true values of the cosmological 
+parameters, we may compare the quality of cosmological constraints.
+
+\textbf{Check Dark Energy Task Force metrics on error ellipses, 
+multi-dimensional KLD, centroid offset, etc.}
+
+\subsection{Results}
+\label{sec:results}
+
+\begin{figure}
+	\begin{center}
+		%\includegraphics[width=0.5\textwidth]{pgm.png}
+		\caption{[moneyplot]}
+		\label{fig:money}
+	\end{center}
+\end{figure}
+
+\section{Discussion}
+\label{sec:discussion}
+
+Thee results of Fig \ref{fig:money} have significant implications for the 
+developing data analysis pipelines of next-generation telescope surveys.  
+However, the method presented here has its own limitations, which are 
+reiterated to discourage the community from applying our work inappropriately.
+
+There are a number of extensions of the work presented in this paper that will 
+be pursued in future work.
 
@@
 
-I will review the questions from the introduction:
+We now summarize answers to the questions posed in the introduction:
 
 \begin{itemize}
-	\item Existing methods have shortcomings that propagate to inaccuracies 
-in characterizing the cosmological parameters.\\
+	\item Existing $n(z)$ estimation methods produce biased estimators that 
+propagate to inaccuracies in characterizing the cosmological parameters.
 	\item Photo-$z$ PDFs are probabilistic data products so must be handled 
-in a mathematically consistent manner; \chippr is one such method, conditioned 
-on some assumptions.
+in a mathematically consistent manner such as the probabilistic graphical model 
+outlined in this paper.
 	\item In addition to coming with its own error distribution, the $n(z)$ 
-estimator produced by \chippr is more accurate than established estimators; 
-furthermore, propagation of the \chippr result leads to a quantifiable 
+estimator produced by \chippr is quantifiably more accurate than established 
+estimators.
+	\item Propagation of the \chippr result leads to a quantifiable 
 improvement in the constraints on cosmological parameters.
@@
 
-I may briefly discuss some obvious extensions of \chippr, such as a fully 
-hierarchical inference of the cosmological parameters that avoids tomographic 
-binning based on photo-$z$ point estimates.
+In conclusion, we discourage the community from continuing to use the stacked 
+estimator and reductions of photo-$z$ PDFs to redshift point estimates in 
+obtaining estimators of $n(z)$.  \chippr is freely available to the community 
+for incorporation into evolving data analysis pipelines.
 
 \begin{acknowledgements}
-AIM thanks Phil Marshall for significant coaching on code development and paper 
-preparation, Mohammadjavad Vakili for insightful input on statistics, Geoffrey 
+AIM thanks Mohammadjavad Vakili for insightful input on statistics, Geoffrey 
 Ryan for assistance in debugging, and Boris Leistedt for helpful comments 
@@
 
-\bibliographystyle{apj}
-\bibliography{references}
-
 \end{document}

### 2017-06-12 research/paper/draft.tex
fixed an error in reference names

@@
 
-The superiority of \chippr over alternative approaches is demonstrated in a 
+We demonstrate the superiority of \chippr over alternative approaches in a 
 number of compelling test cases on mock data.  Each experiment is characterized 
@@
 
-The mock data used here are produced by the following steps.  This method for 
-deriving mock data is referred to as the fiducial case, and variations on it 
-will refer directly to the steps that are altered.
+The mock data used here are produced by the following steps.
+
+This method for deriving mock data is referred to as the fiducial case, and 
+variations on it will refer directly to the steps that are altered.
 
@@
 
-The Kullback-Leibler divergence will be used as the primary measure of the 
-accuracy of estimators of $n(z)$ in cases of mock data with known true 
-redshifts.
+The Kullback-Leibler divergence is our primary measure of the accuracy of 
+estimators of $n(z)$ in cases of mock data with known true redshifts.
 
@@
 
-In the following test cases, the properties of the mock photo-$z$ likelihoods 
-are varied in an effort to emulate known systematics in photo-$z$ estimation.
+In the following test cases, we vary the properties of the mock photo-$z$ 
+likelihoods in an effort to emulate known systematics in photo-$z$ estimation.
 
@@
 single Gaussian component of the likelihood.  To enforce self-consistency, the 
-mean is drawn from a Gaussian distribution with the newly increased variance.
+mean is a random variable drawn from a Gaussian distribution with the newly 
+increased variance.
 
@@
 tend to produce catastrophic outliers that are distributed to be broad in 
-$z_{spec}$ and narrow in $z_{phot}$, as in Fig. \ref{fig:tempfail}.  The 
+$z_{spec}$ and narrow in $z_{phot}$, as in Fig. \ref{fig:tempcatout}.  The 
 systematic behind these catastrophic outliers may be described as an attractor 
@@
 sufficiently strong features (as is the case for blue galaxies), leading 
-galaxies of that type at many $z_{spec}$ to have similar colors.
+galaxies of that SED type at many $z_{spec}$ to have similar colors.
 
@@
 catastrophic outliers that are distributed to be narrow in $z_{spec}$ and broad 
-in $z_{phot}$, as in Fig. \ref{fig:trainfail}.  The systematic behind these 
+in $z_{phot}$, as in Fig. \ref{fig:traincatout}.  The systematic behind these 
 catastrophic outliers may be described as an attractor in the space of 
 $z_{spec}$; some galaxies near a particular $z_{spec}$ map to a range of 
-$z_{phot}$ if their true SED's features fall between photometric filters, 
-leading to many galaxies near that $z_{spec}$ to have similar colors.
+$z_{phot}$ if the training set galaxies at that $z_{spec}$ have inconsistent 
+$z_{phot}$, as might occur if their SED's features fall between photometric 
+filters.
 
@@
 
-The interim prior encapsulates the information about the relationship between 
-observed photometry and redshift upon which a photo-$z$ estimate is based.  
-Interim priors are in general not identical to the true $n(z)$ we wish to 
-estimate; if they were, we would not need any data!  For template fitting 
-photo-$z$ methods, the interim prior is usually an input chosen by the 
-researcher.  However, for many machine learning methods, the interim prior is 
-some function of the training set data that may be influenced by random numbers 
-and is rarely output with the redshift estimates.  Interim priors for template 
-fitting methods tend to have incomplete coverage in the space of true 
-photometry, because they are limited by the choice of the library of SEDs.  
-Interim priors for machine learning methods tend to have incomplete coverage in 
-the space of redshifts, because there are fewer galaxies with spectroscopically 
-confirmed redshifts at high redshifts than low redshifts.
+The interim prior encapsulates the the relationship between observed photometry 
+and redshift information upon which a photo-$z$ estimate is based.  Interim 
+priors are in general not identical to the true $n(z)$ we wish to estimate; if 
+they were, we would not need any data!  For template fitting photo-$z$ methods, 
+the interim prior is usually an input chosen by the researcher.  However, for 
+machine learning methods, the interim prior is some function of the training 
+set data that in many cases may be influenced by random numbers and is rarely 
+output with the redshift estimates.  Interim priors for template fitting 
+methods tend to have incomplete coverage in the space of true photometry, 
+because they are limited by the choice of the library of SEDs.  Interim priors 
+for machine learning methods tend to have incomplete coverage in the space of 
+redshifts, because there are fewer galaxies with spectroscopically confirmed 
+redshifts at high redshifts than low redshifts.
 
@@
 the true $n(z)$, \chippr is not affected by the choice of the interim prior so 
-long as it has nontrivial coverage.
+long as it has nontrivial coverage in the space of redshift.
 
@@
 		%\includegraphics[width=0.5\textwidth]{pgm.png}
-		\caption{[plot of template-like interim prior]}
+		\caption{[plot of template-like (multimodal) interim prior]}
 		\label{fig:tempintpr}
@@
 		%\includegraphics[width=0.5\textwidth]{pgm.png}
-		\caption{[plot of training-like interim prior]}
+		\caption{[plot of training-like (low-$z$ favoring) interim 
+prior]}
 		\label{fig:trainintpr}
@@
 		%\includegraphics[width=0.5\textwidth]{pgm.png}
-		\caption{[plot of a highly featured true $n(z)$]}
+		\caption{[plot of a highly featured (delta function or 
+multimodal?) true $n(z)$]}
 		\label{fig:featured}
@@
 constraints, we apply \chippr to a data from a realistic cosmological 
-simulation (probably \texttt{Buzzard}).
+simulation (probably \texttt{Buzzard}) with photo-$z$ PDFs produced by a 
+popular method (probably \texttt{BPZ}).
 
@@
 interim posteriors are derived from the photometry resulting from the 
-\texttt{Buzzard} simulation by way of [method used to make the interim 
-posteriors].  Because the simulation begins with setting true values of the 
-cosmological parameters, the different estimators of $n(z)$ are propagated 
-through a forecasting code to produce error ellipses on the cosmological 
-parameters.
+\texttt{Buzzard} simulation by way of \texttt{BPZ}.  Because the simulation 
+begins with setting true values of the cosmological parameters, we can 
+propagate the different estimators of $n(z)$ through a forecasting code (which 
+one?) to generate error ellipses on the cosmological parameters.
 
@@
 
-The details of the \texttt{Buzzard} simulation will be summarized here.
+We summarize the details of the \texttt{Buzzard} simulation here.
 
@@
 		%\includegraphics[width=0.5\textwidth]{pgm.png}
-		\caption{[moneyplot]}
+		\caption{[moneyplot of error ellipses resulting from different 
+$n(z)$ estimators]}
 		\label{fig:money}
@@
 However, the method presented here has its own limitations, which are 
-reiterated to discourage the community from applying our work inappropriately.
+reiterated to discourage the community from applying this work inappropriately.
 
-There are a number of extensions of the work presented in this paper that will 
-be pursued in future work.
+We intend to pursue a number of extensions of the work presented in this paper 
+in future work.
 

### 2017-06-12 research/paper/draft-editing.tex
periodic update

--- 2017-06-12 research/paper/draft.tex
@@
 \documentclass[iop]{emulateapj}
+%\documentclass[preprint]{aastex}
+%\documentclass[12pt, onecolumn]{emulateapj}
+%\documentstyle[aas2pp4,natbib209]{article}
 
@@
 
-\shorttitle{How to obtain the redshift distribution from probabilistic redshift 
-estimates}
+\shorttitle{How to obtain the redshift distribution from probabilistic redshift estimates}
 \shortauthors{Malz, et al.}
@@
 
-\title{How to obtain the redshift distribution from probabilistic redshift 
-estimates}
+\title{How to obtain the redshift distribution from probabilistic redshift estimates}
 
-\author{Alex Malz\altaffilmark{1}, David W. Hogg\altaffilmark{1,2,3,4}, Phil 
-Marshall\altaffilmark{5}, \& Others}
+\author{Alex Malz\altaffilmark{1}, David W. Hogg\altaffilmark{1,2,3,4}, Phil Marshall\altaffilmark{5}, \& Others}
 \email{aimalz@nyu.edu}
 
-\altaffiltext{1}{Center for Cosmology and Particle Physics, Department of 
-Physics,
+\altaffiltext{1}{Center for Cosmology and Particle Physics, Department of Physics,
   New York University, 726 Broadway, 9th floor, New York, NY 10003, USA}
-\altaffiltext{2}{Simons Center for Computational Astrophysics, 162 Fifth 
-Avenue, 7th floor, New York, NY 10010, USA}
-\altaffiltext{3}{Center for Data Science, New York University, 60 Fifth Avenue, 
-7th floor, New York, NY 10003, USA}
-\altaffiltext{4}{Max-Planck-Institut f\"ur Astronomie, K\"onigstuhl 17, D-69117 
-Heidelberg, Germany}
+\altaffiltext{2}{Simons Center for Computational Astrophysics, 162 Fifth Avenue, 7th floor, New York, NY 10010, USA}
+\altaffiltext{3}{Center for Data Science, New York University, 60 Fifth Avenue, 7th floor, New York, NY 10003, USA}
+\altaffiltext{4}{Max-Planck-Institut f\"ur Astronomie, K\"onigstuhl 17, D-69117 Heidelberg, Germany}
 \altaffiltext{5}{[SLAC]}
@@
 \begin{abstract}
-The redshift distribution $n(z)$ is a crucial ingredient for weak lensing 
-cosmology.  Spectroscopically confirmed redshifts for the dim and numerous 
-galaxies observed by weak lensing surveys are expected to be inaccessible, 
-making photometric redshifts (photo-$z$s) the next best alternative.  Because 
-of the nontrivial inference involved in their determination, photo-$z$ point 
-estimates are being superseded by photo-$z$ probability distribution functions 
-(PDFs).  However, analytic methods for utilizing these new data products in 
-cosmological inference are still evolving.  This paper presents a novel 
-approach to estimating the posterior distribution over $n(z)$ from a survey of 
-galaxy photo-$z$ PDFs based upon a probabilistic graphical model of 
-hierarchical inference.  We present the Cosmological Hierarchical Inference 
-with Probabilistic Photometric Redshifts (\chippr) code implementing this 
-technique, as well as its validation on mock data and testing on the 
-\texttt{Buzzard} simulations.  \chippr yields a more accurate characterization 
-of $n(z)$ containing information beyond the best-fit estimator produced by 
-traditional procedures.  The publicly available code is easily extensible to 
-other one-point statistics that depend on redshift.
+The redshift distribution $n(z)$ is a crucial ingredient for weak lensing cosmology.
+Spectroscopically confirmed redshifts for the dim and numerous galaxies observed by weak lensing surveys are expected to be inaccessible, making photometric redshifts (photo-$z$s) the next best alternative.
+Because of the nontrivial inference involved in their determination, photo-$z$ point estimates are being superseded by photo-$z$ probability distribution functions (PDFs).
+However, analytic methods for utilizing these new data products in cosmological inference are still evolving.
+This paper presents a novel approach to estimating the posterior distribution over $n(z)$ from a survey of galaxy photo-$z$ PDFs based upon a probabilistic graphical model of hierarchical inference.
+We present the Cosmological Hierarchical Inference with Probabilistic Photometric Redshifts (\chippr) code implementing this technique, as well as its validation on mock data and testing on the \texttt{Buzzard} simulations.
+\chippr yields a more accurate characterization of $n(z)$ containing information beyond the best-fit estimator produced by traditional procedures.
+The publicly available code is easily extensible to other one-point statistics that depend on redshift.
 
@@
 
-\keywords{catalogs --- cosmology: cosmological parameters --- galaxies: 
-statistics --- gravitational lensing: weak --- methods: analytical --- methods: 
-data analysis --- methods: statistical --- techniques: photometric}
+\keywords{catalogs --- cosmology: cosmological parameters --- galaxies: statistics --- gravitational lensing: weak --- methods: analytical --- methods: data analysis --- methods: statistical --- techniques: photometric}
 
@@
 
-After a brief literature review addressing how photo-$z$ PDFs are currently 
-used in cosmology, this paper aims to answer the following questions:
+After a brief literature review addressing how photo-$z$ PDFs are currently used in cosmology, this paper aims to answer the following questions:
 
@@
 	\item Why should we question existing methods?
-	\item How can we improve the effectiveness of using photo-$z$ PDFs in 
-inference?
-	\item How does the result of \chippr compare to established estimators 
-in terms of the accuracy of $n(z)$?
-	\item How significant is the effect of the discrepancy between $n(z)$ 
-estimators on cosmological constraints?
+	\item How can we improve the effectiveness of using photo-$z$ PDFs in inference?
+	\item How does the result of \chippr compare to established estimators in terms of the accuracy of $n(z)$?
+	\item How significant is the effect of the discrepancy between $n(z)$ estimators on cosmological constraints?
 \end{itemize}
@@
 
-This paper presents a mathematically consistent method for obtaining the 
-posterior distribution over the redshift density function $n(z)$ using a 
-catalog of photo-$z$ PDFs.
+This paper presents a mathematically consistent method for obtaining the posterior distribution over the redshift density function $n(z)$ using a catalog of photo-$z$ PDFs.
 
@@
 
-The directed acyclic graph of Fig. \ref{fig:pgm} represents a probabilistic 
-graphical model for hierarchical inference of $n(z)$, the mathematical 
-interpretation of which will be presented below.
+The directed acyclic graph of Fig. \ref{fig:pgm} represents a probabilistic graphical model for hierarchical inference of $n(z)$, the mathematical interpretation of which will be presented below.
 
@@
 		\includegraphics[width=0.5\textwidth]{pgm.png}
-		\caption{\textbf{Replace \texttt{daft} version with more 
-polished \texttt{tikz} version.}}
+		\caption{\textbf{Replace \texttt{daft} version with more polished \texttt{tikz} version.}}
+%		\caption{This directed acyclic graph corresponds to a PGM for a hierarchical inference of $p(\vec{\theta}|\{\vec{d}_{i}\})$.
+In this graph, all random variables are shown in circles, with observed random variables shown in shaded circles.
+Relationships between variables are indicated by arrows from parameters to the variables distributed according to functions of them.
+The box indicates that there are a number of copies of the relationships between boxed parameters, each independent of all others.
+The hyperparameters $\vec{\theta}$ representing $n(z)$ are at the top.
+Independently drawn from a function of the hyperparameters $\vec{\theta}$ are galaxy redshifts $\{z_{i}\}$ below.
+The observed galaxy photometry $\{\vec{d}_{i}\}$, shown in shaded circles, is determined by the redshifts above.}
 	\label{fig:pgm}
@@
 
-This framework entails a number of choices and assumptions that must be 
-addressed explicitly.
+This framework entails a number of choices and assumptions that must be addressed explicitly.
 
 \begin{enumerate}
-	\item While we advocate for the approach of hierarchical inference, the 
-probabilistic graphical model presented here is not the only one that could be 
-proposed.
+	\item While we advocate for the approach of hierarchical inference, the probabilistic graphical model presented here is not the only one that could be proposed.
 \end{enumerate}
@@
 
-It is useful to translate some popular existing methods for deriving $n(z)$ 
-from photo-$z$ PDFs into the mathematical framework of Sec. \ref{sec:model}.
+It is useful to translate some popular existing methods for deriving $n(z)$ from photo-$z$ PDFs into the mathematical framework of Sec. \ref{sec:model}.
 
@@
 
-The publicly available \chippr code implements the probabilistic graphical 
-model presented in Sec. \ref{sec:model}.
+The publicly available \chippr code implements the probabilistic graphical model presented in Sec. \ref{sec:model}.
 
-In addition to the choices and assumptions underlying the probabilistic 
-graphical model, the implementation of \chippr makes choices and assumptions of 
-its own.
+In addition to the choices and assumptions underlying the probabilistic graphical model, the implementation of \chippr makes choices and assumptions of its own.
 
 \begin{enumerate}
-	\item \chippr currently only accepts photo-$z$ PDFs and produces $n(z)$ 
-samples of a single format, that of the piecewise constant parametrization, 
-also referred to as a binned histogram parametrization and a sum of top hat 
-functions.
+	\item \chippr currently only accepts photo-$z$ PDFs and produces $n(z)$ samples of a single format, that of the piecewise constant parametrization, also referred to as a binned histogram parametrization and a sum of top hat functions.
 \end{enumerate}
@@
 
-We demonstrate the superiority of \chippr over alternative approaches in a 
-number of compelling test cases on mock data.  Each experiment is characterized 
-by a single change to a fiducial case in order to isolate the influence of 
-systematic effects known to be relevant to photo-$z$ estimation and propagation 
-in analysis.
+We demonstrate the superiority of \chippr over alternative approaches in a number of compelling test cases on mock data.
+Each experiment is characterized by a single change to a fiducial case in order to isolate the influence of systematic effects known to be relevant to photo-$z$ estimation and propagation in analysis.
 
@@
 
-The mock data in these tests consists of photo-$z$ interim posteriors rather 
-than photometric data because the various existing methods for deriving 
-photo-$z$ interim posteriors do not in general yield results that are 
-consistent with one another, indicating that their systematics are not 
-well-understood.  Because the mock data is independent of any choice of 
-photo-$z$ PDF production method, we not only ensure that our photo-$z$ interim 
-priors are perfectly understood but also deter readers from assuming that 
-\chippr has any preference over the method by which photo-$z$ interim 
-posteriors are derived from photometric data.
+The mock data in these tests consists of photo-$z$ interim posteriors rather than photometric data because the various existing methods for deriving photo-$z$ interim posteriors do not in general yield results that are consistent with one another, indicating that their systematics are not well-understood.
+Because the mock data is independent of any choice of photo-$z$ PDF production method, we not only ensure that our photo-$z$ interim priors are perfectly understood but also deter readers from assuming that \chippr has any preference over the method by which photo-$z$ interim posteriors are derived from photometric data.
 
@@
 
-This method for deriving mock data is referred to as the fiducial case, and 
-variations on it will refer directly to the steps that are altered.
+This method for deriving mock data is referred to as the fiducial case, and variations on it will refer directly to the steps that are altered.
 
@@
 
-The Kullback-Leibler divergence is our primary measure of the accuracy of 
-estimators of $n(z)$ in cases of mock data with known true redshifts.
+The Kullback-Leibler divergence is our primary measure of the accuracy of estimators of $n(z)$ in cases of mock data with known true redshifts.
 
-\textbf{Review precision and bias from \texttt{kld.ipynb} and interpret in 
-terms of a \% difference.}
+\textbf{Review precision and bias from \texttt{kld.ipynb} and interpret in terms of a \% difference.}
 
@@
 
-In the following test cases, we vary the properties of the mock photo-$z$ 
-likelihoods in an effort to emulate known systematics in photo-$z$ estimation.
+In the following test cases, we vary the properties of the mock photo-$z$ likelihoods in an effort to emulate known systematics in photo-$z$ estimation.
 
@@
 
-One major concern about photo-$z$s is the intrinsic scatter of point 
-estimators, including those derived from photo-$z$ PDFs, that is observed to 
-varying extents with every existing photo-$z$ algorithm and illustrated in Fig. 
+One major concern about photo-$z$s is the intrinsic scatter of point estimators, including those derived from photo-$z$ PDFs, that is observed to varying extents with every existing photo-$z$ algorithm and illustrated in Fig.
 \ref{fig:intscat}.
@@
 		%\includegraphics[width=0.5\textwidth]{pgm.png}
-		\caption{[plot of $z_{phot}$ vs. $z_{spec}$ with varying 
-intrinsic scatter]}
+		\caption{[plot of $z_{phot}$ vs. $z_{spec}$ with varying intrinsic scatter]}
 		\label{fig:intscat}
@@
 
-To emulate intrinsic scatter, we modify the fiducial case to simply broaden the 
-single Gaussian component of the likelihood.  To enforce self-consistency, the 
-mean is a random variable drawn from a Gaussian distribution with the newly 
-increased variance.
+To emulate intrinsic scatter, we modify the fiducial case to simply broaden the single Gaussian component of the likelihood.
+To enforce self-consistency, the mean is a random variable drawn from a Gaussian distribution with the newly increased variance.
 
@@
 
-In addition to intrinsic scatter, photo-$z$ methods employing template fitting 
-tend to produce catastrophic outliers that are distributed to be broad in 
-$z_{spec}$ and narrow in $z_{phot}$, as in Fig. \ref{fig:tempcatout}.  The 
-systematic behind these catastrophic outliers may be described as an attractor 
-in the space of $z_{phot}$; some galaxies at a range of $z_{spec}$ map onto a 
-single $z_{phot}$ (with some scatter) if their true SED does not have 
-sufficiently strong features (as is the case for blue galaxies), leading 
-galaxies of that SED type at many $z_{spec}$ to have similar colors.
+In addition to intrinsic scatter, photo-$z$ methods employing template fitting tend to produce catastrophic outliers that are distributed to be broad in $z_{spec}$ and narrow in $z_{phot}$, as in Fig.
+\ref{fig:tempcatout}.
+The systematic behind these catastrophic outliers may be described as an attractor in the space of $z_{phot}$; some galaxies at a range of $z_{spec}$ map onto a single $z_{phot}$ (with some scatter) if their true SED does not have sufficiently strong features (as is the case for blue galaxies), leading galaxies of that SED type at many $z_{spec}$ to have similar colors.
 
@@
 		%\includegraphics[width=0.5\textwidth]{pgm.png}
-		\caption{[plot of $z_{phot}$ vs. $z_{spec}$ with template-like 
-catastrophic outliers]}
+		\caption{[plot of $z_{phot}$ vs. $z_{spec}$ with template-like catastrophic outliers]}
 		\label{fig:tempcatout}
@@
 
-Data driven photo-$z$ methods tend to suffer from a different form of 
-catastrophic outliers that are distributed to be narrow in $z_{spec}$ and broad 
-in $z_{phot}$, as in Fig. \ref{fig:traincatout}.  The systematic behind these 
-catastrophic outliers may be described as an attractor in the space of 
-$z_{spec}$; some galaxies near a particular $z_{spec}$ map to a range of 
-$z_{phot}$ if the training set galaxies at that $z_{spec}$ have inconsistent 
-$z_{phot}$, as might occur if their SED's features fall between photometric 
-filters.
+Data driven photo-$z$ methods tend to suffer from a different form of catastrophic outliers that are distributed to be narrow in $z_{spec}$ and broad in $z_{phot}$, as in Fig.
+\ref{fig:traincatout}.
+The systematic behind these catastrophic outliers may be described as an attractor in the space of $z_{spec}$; some galaxies near a particular $z_{spec}$ map to a range of $z_{phot}$ if the training set galaxies at that $z_{spec}$ have inconsistent $z_{phot}$, as might occur if their SED's features fall between photometric filters.
 
@@
 		%\includegraphics[width=0.5\textwidth]{pgm.png}
-		\caption{[plot of $z_{phot}$ vs. $z_{spec}$ with training-like 
-catastrophic outliers]}
+		\caption{[plot of $z_{phot}$ vs. $z_{spec}$ with training-like catastrophic outliers]}
 		\label{fig:traincatout}
@@
 
-The interim prior encapsulates the the relationship between observed photometry 
-and redshift information upon which a photo-$z$ estimate is based.  Interim 
-priors are in general not identical to the true $n(z)$ we wish to estimate; if 
-they were, we would not need any data!  For template fitting photo-$z$ methods, 
-the interim prior is usually an input chosen by the researcher.  However, for 
-machine learning methods, the interim prior is some function of the training 
-set data that in many cases may be influenced by random numbers and is rarely 
-output with the redshift estimates.  Interim priors for template fitting 
-methods tend to have incomplete coverage in the space of true photometry, 
-because they are limited by the choice of the library of SEDs.  Interim priors 
-for machine learning methods tend to have incomplete coverage in the space of 
-redshifts, because there are fewer galaxies with spectroscopically confirmed 
-redshifts at high redshifts than low redshifts.
+The interim prior encapsulates the the relationship between observed photometry and redshift information upon which a photo-$z$ estimate is based.
+Interim priors are in general not identical to the true $n(z)$ we wish to estimate; if they were, we would not need any data!
+For template fitting photo-$z$ methods, the interim prior is usually an input chosen by the researcher.
+However, for machine learning methods, the interim prior is some function of the training set data that in many cases may be influenced by random numbers and is rarely output with the redshift estimates.
+Interim priors for template fitting methods tend to have incomplete coverage in the space of true photometry, because they are limited by the choice of the library of SEDs.
+Interim priors for machine learning methods tend to have incomplete coverage in the space of redshifts, because there are fewer galaxies with spectroscopically confirmed redshifts at high redshifts than low redshifts.
 
-Existing $n(z)$ estimation routines will always produce a biased estimator when 
-the interim prior is not equal to the true $n(z)$.  We demonstrate here that 
-regardless of the appropriateness of the interim prior as an approximation to 
-the true $n(z)$, \chippr is not affected by the choice of the interim prior so 
-long as it has nontrivial coverage in the space of redshift.
+Existing $n(z)$ estimation routines will always produce a biased estimator when the interim prior is not equal to the true $n(z)$.
+We demonstrate here that regardless of the appropriateness of the interim prior as an approximation to the true $n(z)$, \chippr is not affected by the choice of the interim prior so long as it has nontrivial coverage in the space of redshift.
 
@@
 
-An interim prior based on a template library may be a sum of smooth functions 
-representing $n(z)$ for each SED type in the library.  Template libraries do 
-not include every possible galaxy SED, and the $n(z)$ used for each SED type 
-may not be accurate.  The interim prior shown in Fig. \ref{fig:tempintpr} is an 
-emulation of an interim prior corresponding to a template library of this type.
+An interim prior based on a template library may be a sum of smooth functions representing $n(z)$ for each SED type in the library.
+Template libraries do not include every possible galaxy SED, and the $n(z)$ used for each SED type may not be accurate.
+The interim prior shown in Fig.
+\ref{fig:tempintpr} is an emulation of an interim prior corresponding to a template library of this type.
 
@@
 
-An interim prior based on a training set may be biased toward low redshifts due 
-to the dearth of distant galaxies with spectroscopic redshifts.  The interim 
-prior shown in Fig. \ref{fig:trainintpr} is an emulation of an interim prior 
-corresponding to a training set biased in this way.
+An interim prior based on a training set may be biased toward low redshifts due to the dearth of distant galaxies with spectroscopic redshifts.
+The interim prior shown in Fig.
+\ref{fig:trainintpr} is an emulation of an interim prior corresponding to a training set biased in this way.
 
@@
 		%\includegraphics[width=0.5\textwidth]{pgm.png}
-		\caption{[plot of training-like (low-$z$ favoring) interim 
-prior]}
+		\caption{[plot of training-like (low-$z$ favoring) interim prior]}
 		\label{fig:trainintpr}
@@
 Existing $n(z)$ estimators are systematically smoother than the true $n(z)$.
-Here we show that the traditional estimators of $n(z)$ cannot recover a highly 
-featured $n(z)$.  The implication of this failure is quite serious; the 
-consistently smooth, unimodal $n(z)$ estimates appearing in the literature 
-could result from much more featured true redshift distributions, and there 
-would be no way to catch this error without using a fully probabilistic method.
+Here we show that the traditional estimators of $n(z)$ cannot recover a highly featured $n(z)$.
+The implication of this failure is quite serious; the consistently smooth, unimodal $n(z)$ estimates appearing in the literature could result from much more featured true redshift distributions, and there would be no way to catch this error without using a fully probabilistic method.
 
@@
 
-In this test, we choose the true $n(z)$ of Fig. \ref{fig:featured}, which has 
-nontrivial structure.
+In this test, we choose the true $n(z)$ of Fig. \ref{fig:featured}, which has nontrivial structure.
 
@@
 		%\includegraphics[width=0.5\textwidth]{pgm.png}
-		\caption{[plot of a highly featured (delta function or 
-multimodal?) true $n(z)$]}
+		\caption{[plot of a highly featured (delta function or multimodal?) true $n(z)$]}
 		\label{fig:featured}
@@
 
-To show how the choice of $n(z)$ estimator propagates to cosmological 
-constraints, we apply \chippr to a data from a realistic cosmological 
-simulation (probably \texttt{Buzzard}) with photo-$z$ PDFs produced by a 
-popular method (probably \texttt{BPZ}).
+To show how the choice of $n(z)$ estimator propagates to cosmological constraints, we apply \chippr to a data from a realistic cosmological simulation (probably \texttt{Buzzard}) with photo-$z$ PDFs produced by a popular method (probably \texttt{BPZ}).
 
@@
 
-As in Sec. \ref{sec:validintro}, the mock data takes the form of photo-$z$ 
-interim posteriors, but that is where the similarity ends.  These photo-$z$ 
-interim posteriors are derived from the photometry resulting from the 
-\texttt{Buzzard} simulation by way of \texttt{BPZ}.  Because the simulation 
-begins with setting true values of the cosmological parameters, we can 
-propagate the different estimators of $n(z)$ through a forecasting code (which 
-one?) to generate error ellipses on the cosmological parameters.
+As in Sec.
+\ref{sec:validintro}, the mock data takes the form of photo-$z$ interim posteriors, but that is where the similarity ends.
+These photo-$z$ interim posteriors are derived from the photometry resulting from the \texttt{Buzzard} simulation by way of \texttt{BPZ}.
+Because the simulation begins with setting true values of the cosmological parameters, we can propagate the different estimators of $n(z)$ through a forecasting code (which one?) to generate error ellipses on the cosmological parameters.
 
@@
 
-Because the mock data is associated with true values of the cosmological 
-parameters, we may compare the quality of cosmological constraints.
+Because the mock data is associated with true values of the cosmological parameters, we may compare the quality of cosmological constraints.
 
-\textbf{Check Dark Energy Task Force metrics on error ellipses, 
-multi-dimensional KLD, centroid offset, etc.}
+\textbf{Check Dark Energy Task Force metrics on error ellipses, multi-dimensional KLD, centroid offset, etc.}
 
@@
 		%\includegraphics[width=0.5\textwidth]{pgm.png}
-		\caption{[moneyplot of error ellipses resulting from different 
-$n(z)$ estimators]}
+		\caption{[moneyplot of error ellipses resulting from different $n(z)$ estimators]}
 		\label{fig:money}
@@
 
-Thee results of Fig \ref{fig:money} have significant implications for the 
-developing data analysis pipelines of next-generation telescope surveys.  
-However, the method presented here has its own limitations, which are 
-reiterated to discourage the community from applying this work inappropriately.
+Thee results of Fig \ref{fig:money} have significant implications for the developing data analysis pipelines of next-generation telescope surveys.
+However, the method presented here has its own limitations, which are reiterated to discourage the community from applying this work inappropriately.
 
-We intend to pursue a number of extensions of the work presented in this paper 
-in future work.
+We intend to pursue a number of extensions of the work presented in this paper in future work.
 
@@
 \begin{itemize}
-	\item Existing $n(z)$ estimation methods produce biased estimators that 
-propagate to inaccuracies in characterizing the cosmological parameters.
-	\item Photo-$z$ PDFs are probabilistic data products so must be handled 
-in a mathematically consistent manner such as the probabilistic graphical model 
-outlined in this paper.
-	\item In addition to coming with its own error distribution, the $n(z)$ 
-estimator produced by \chippr is quantifiably more accurate than established 
-estimators.
-	\item Propagation of the \chippr result leads to a quantifiable 
-improvement in the constraints on cosmological parameters.
+	\item Existing $n(z)$ estimation methods produce biased estimators that propagate to inaccuracies in characterizing the cosmological parameters.
+	\item Photo-$z$ PDFs are probabilistic data products so must be handled in a mathematically consistent manner such as the probabilistic graphical model outlined in this paper.
+	\item In addition to coming with its own error distribution, the $n(z)$ estimator produced by \chippr is quantifiably more accurate than established estimators.
+	\item Propagation of the \chippr result leads to a quantifiable improvement in the constraints on cosmological parameters.
 \end{itemize}
 
-In conclusion, we discourage the community from continuing to use the stacked 
-estimator and reductions of photo-$z$ PDFs to redshift point estimates in 
-obtaining estimators of $n(z)$.  \chippr is freely available to the community 
-for incorporation into evolving data analysis pipelines.
+In conclusion, we discourage the community from continuing to use the stacked estimator and reductions of photo-$z$ PDFs to redshift point estimates in obtaining estimators of $n(z)$.
+\chippr is freely available to the community for incorporation into evolving data analysis pipelines.
+
+%\section{Introduction}
+%\label{sec:introduction}
+%
+%The redshift distribution $n(z)$ is necessary for calculating two-point statistics of galaxy properties used to determine the cosmological parameter values that inform our understanding of the evolution of large-scale structure in the universe \citep{masters_mapping_2015}.
+Inaccurate estimates of $n(z)$ can significantly impact the constraining power of a galaxy survey, biasing recovery of the cosmological parameters.
+For example, if the $n(z)$ used in analysis is offset from the true $n(z)$ by a small, negative constant, it results in underestimating $w_{0}$ and overestimating $\sigma_{8}$ \citep{samuroff_simultaneous_2017}.
+%
+%Though the redshift density function has traditionally been determined from spectroscopically observed redshifts, modern galaxy surveys including DES, LSST, Euclid, and WFIRST seek to obtain two-point statistics of redshift from samples of galaxies for which spectroscopic redshifts are unavailable, either due to their large numbers or their low brightnesses.
+For decades, photometrically estimated redshifts (photo-$z$s) have been the leading alternative to spectroscopically observed redshifts, though they suffer from issues of precision in the form of an intrinsic scatter and accuracy in the form of catastrophic outliers, as well as other systematics imparted by the properties of the survey, data reduction pipeline(s), and assumptions underlying the analysis \citep{baum_photoelectric_1962}.
+These weaknesses are illuminated by a probabilistic interpretation of photo-$zs$; if these nontrivial uncertainties were expressed as a probability distribution function (PDF) over redshift, photo-$z$s could be replaced by photo-$z$ PDFs containing more information than a simple point estimate \citep{koo_overview_1999}.
+Such data products have been commonly released by photometric galaxy surveys since SDSS DR7 \citep{abazajian_seventh_2009} using a great variety of methods.
+%
+%Methods for using photo-$z$ PDFs in cosmological inference remain underdeveloped, with many survey pipelines reducing them to familiar point estimators that are compatible with existing technology or engaging with them heuristically in a manner inconsistent with their probabilistic nature.
+Stacking photo-$z$ PDFs, or other mathematically equivalent methods, to obtain an estimator of $n(z)$ is especially popular \citep{cunha_estimating_2009, sheldon_photometric_2012}.
+However, photo-$z$ PDFs are probabilistic data products that must be handled in a fully consistent, probabilistic manner; hierarchical inference is the only mathematically valid way to do inference with photo-$z$ PDFs and other probabilistic quantities.
+%
+%It is desirable to create rigorous methods for using photo-$z$ PDFs in inference, beginning with the simplest one-point statistic of redshift, $n(z)$.
+In the spirit of \citet{hogg_inferring_2010, foreman-mackey_exoplanet_2014}, this paper derives a mathematically rigorous approach to inferring $n(z)$ in Sec.
+\ref{sec:method}.
+The presentation of a public code implementing this novel technique is given in Sec.
+\ref{sec:model_specifics}.
+The code is validated on mock data and tested on a subset of BOSS DR10 data in Sec.
+\ref{sec:experiments}.
+We conclude and discuss future directions in Sec.
+\ref{sec:discussion}.
+%
+%\section{Method}
+%\label{sec:method}
+%
+%The redshift distribution $n(z)$ may be understood as the probability of finding a galaxy at a redshift $z$.
+We may express $n(z)$ in terms of some functional form $f_{\vec{\theta}}(z)$ with parameters comprising $\vec{\theta}$ according to Eq.
+\ref{eq:parametrization}.
+%\begin{align}
+%\label{eq:parametrization}
+%n(z) &\equiv f_{\vec{\theta}}(z)\ \equiv\ p(z|\vec{\theta})
+%\end{align}
+%In this work, we wish to characterize the posterior distribution $p(\vec{\theta}|\textul{D})$ of the parameters contained in $\vec{\theta}$ defining $n(z)$ given all available data $\textul{D}$.
+The likelihood $p(z|\vec{\theta})$ represents the probability of a random variable, in this case redshift $z$, given the parameters in $\vec{\theta}$ defining the distribution $n(z)$ from which it is drawn.
+Our model for inferring $p(\vec{\theta}|\textul{D})$ will be introduced in Sec.
+\ref{sec:derivation}, and alternative methods of estimating $\vec{\theta}$ will be presented in Sec.
+\ref{sec:alternatives}.
+%
+%\subsection{Derivation}
+%\label{sec:derivation}
+%
+%Inference of $p(\vec{\theta}|\textul{D})$ is perfectly suited to a hierarchical Bayesian model.
+In this problem, the available data $\textul{D}$ is a catalog of photometry $\{\vec{d}_{i}\}$ of individual galaxies $i$ where each $\vec{d}_{i}$ is a random variable that is a function of its redshift parameter $z_{i}$; the redshift parameters $\{z_{i}\}$ are independently drawn from a function defined by hyperparameters in $\vec{\theta}$.
+These physical relationships may be illustrated in Fig.
+\ref{fig:pgm} by a directed acyclic graph representing a probabilistic graphical model (PGM).
+%\begin{figure}
+%\begin{center}
+%\includegraphics[width=0.5\textwidth]{pgm.png}
+%\caption{This directed acyclic graph corresponds to a PGM for a hierarchical inference of $p(\vec{\theta}|\{\vec{d}_{i}\})$.
+In this graph, all random variables are shown in circles, with observed variables shown in shaded circles.
+Relationships between variables are indicated by arrows.
+The box indicates that there are a number of copies of the relationships between boxed parameters, each independent of all others.
+The hyperparameters $\vec{\theta}$ representing $n(z)$ are at the top.
+Independently drawn from a function of the hyperparameters $\vec{\theta}$ are galaxy redshifts $\{z_{i}\}$ below.
+The observed galaxy photometry $\{\vec{d}_{i}\}$, shown in shaded circles, is determined by the redshifts above.}
+%\label{fig:pgm}
+%\end{center}
+%\end{figure}
+%This PGM may be used as the basis for a derivation of the desired hyperposterior $p(\vec{\theta}|\{\vec{d}_{i}\})$, assuming our physical model is complete.
+From this point on, we will work solely with log probabilities.
+%
+%We begin with Eq.
+\ref{eq:hyperbayes} by applying Bayes' Rule to express the log-hyperposterior $\ln\left[p(\vec{\theta}|\{\vec{d}_{i}\})\right]$ in terms of the log-hyperlikelihood $\ln\left[p(\{\vec{d}_{i}\}|\vec{\theta})\right]$.
+%\begin{align}
+%\label{eq:hyperbayes}
+%\ln\left[p(\vec{\theta}|\{\vec{d}_{i}\})\right] &\propto \ln\left[p(\vec{\theta})\right] + \ln\left[p(\{\vec{d}_{i}\}|\vec{\theta})\right]
+%\end{align}
+%To do this, we will choose a log-hyperprior distribution $\ln\left[p(\vec{\theta})\right]$ representing our beliefs about the distribution of the hyperparameters $\vec{\theta}$ in terms of fixed variables that we will assume are known.
+%
+%The log-hyperlikelihood $\ln\left[p(\{\vec{d}_{i}\}|\vec{\theta})\right]$ contains no explicit reference to redshifts, therefore, we employ the PGM to write the marginalization over the redshifts in Eq.
+\ref{eq:marginalization}.
+%\begin{align}
+%\label{eq:marginalization}
+%\ln\left[p(\{\vec{d}_{i}\}|\vec{\theta})\right] &= \ln\left[\int\ p(\{\vec{d}_{i}\}|\{z_{i}\})\ p(\{z_{i}\}|\vec{\theta})\ d\{z_{i}\}\right]
+%\end{align}
+%We shall handle the two terms in the integral separately.
+%
+%The first term is the likelihood of all galaxy photometry given all galaxy redshifts.
+If we assume independence of galaxy photometry, such that $p(\vec{d}_{i}|\{\vec{d}_{i'\neq i}\}, \{z_{i}\})=p(\vec{d}_{i}|z_{i})$, then we may write Eq.
+\ref{eq:independentdata}.
+%\begin{align}
+%\label{eq:independentdata}
+%\ln\left[p(\{\vec{d}_{i}\}|\{z_{i}\})\right] &= \sum_{i}\ln\left[p(\vec{d}_{i}|z_{i})\right]
+%\end{align}
+%The second term may also be expanded as in Eq. \ref{eq:independentredshifts} by assuming that all galaxy redshifts are drawn independently from $n(z)$.
+%\begin{align}
+%\label{eq:independentredshifts}
+%\ln\left[p(\{z_{i}\}|\vec{\theta})\right] &= \sum_{i}\ln\left[p(z_{i}|\vec{\theta})\right]
+%\end{align}
+%Since all galaxy redshifts are independent of one another, an integral over the aggregate $\int\dots d\{z_{i}\}$ is simply the product of the integrals over each one $\prod_{i}\int\dots dz_{i}$.
+Thus, the log-hyperlikelihood can be written according to Eq.
+\ref{eq:log-hyperlikelihood}.
+%\begin{align}
+%\label{eq:log-hyperlikelihood}
+%\ln\left[p(\{\vec{d}_{i}\}|\vec{\theta})\right] &= \sum_{i}\ln\left[\int\ p(\vec{d}_{i}|z_{i})\ p(z_{i}|\vec{\theta})\ dz_{i}\right]
+%\end{align}
+%The expression of Eq.
+\ref{eq:log-hyperlikelihood} contains two types of quantities.
+The $\{p(z_{i}|\vec{\theta})\}$ are known likelihoods that will be equal to $\{f_{\vec{\theta}}(z_{i})\}$.
+The $\{p(\vec{d}_{i}|z_{i})\}$, however, are likelihoods that are unknown and in general unknowable, and it is worth discussing these facts.
+%
+%Photo-$z$ PDFs are commonly written simply as $p(z)$, but this notation oversimplifies their substance.
+Whether determined by way of template-fitting or machine learning, photo-$z$ PDFs are dependent on the data $\vec{d}$ from which they are calculated.
+This data $\vec{d}$ must be considered to be a quantity upon which the redshift $z$ is conditioned, otherwise normalizing a photo-$z$ PDF would require integrating over all possible values of $\vec{d}$.
+This means that photo-$z$ PDFs are posteriors, probabilities of parameters $z$ conditioned on data $\vec{d}$.
+However, where there is a posterior, there is always a prior.
+Photo-$z$ PDFs are in general \textit{interim} posteriors, because in addition to being conditioned on observations, they are also conditioned on an interim prior in the form of a particular value $\vec{\theta}^{*}$ of $\vec{\theta}$ necessary for computing the photo-$z$ PDF.
+In the case of template-fitting methods, the interim prior is usually specified as an input to the photo-$z$ PDF production code, and it often takes the form of an initial guess for $\vec{\theta}$ based on the results of previous galaxy surveys or simulations thereof; in the case of machine learning methods, the interim prior may be explicitly derived from the training set or implicitly produced in the process of determining the photo-$z$ PDFs, and the interim prior is not always revealed to the user.
+Hierarchical inference may be performed only when the interim prior is known, which means it will only be possible for photo-$z$ PDFs made via certain methods.
+%
+%Since the data products from which we hope to probe the log-hyperposterior are themselves posteriors, we must express the likelihoods $\{p(\vec{d}_{i}|z_{i})\}$ of Eq.
+\ref{eq:log-hyperlikelihood} in terms of the photo-$z$ PDFs $\{p(z_{i}|\vec{d}_{i}, \vec{\theta}^{*})\}$.
+We start by multiplying the likelihood by an inspired factor of unity written in terms of the photo-$z$ interim posterior we have at hand, as in Eq.
+\ref{eq:unity}.
+%\begin{align}
+%\label{eq:unity}
+%p(\vec{d}_{i}|z_{i}) &= p(\vec{d}_{i}|z_{i})\ \frac{p(z_{i}|\vec{d}_{i}, \vec{\theta}^{*})}{p(z_{i}|\vec{d}_{i}, \vec{\theta}^{*})}
+%\end{align}
+%Next, we expand the denominator in terms of Bayes' Rule to obtain Eq. \ref{eq:bayes}.
+%\begin{align}
+%\label{eq:bayes}
+%p(\vec{d}_{i}|z_{i}) &= p(\vec{d}_{i}|z_{i})\ p(z_{i}|\vec{d}_{i}, \vec{\theta}^{*})\ \frac{p(\vec{d}_{i}|\vec{\theta}^{*})}{p(z_{i}|\vec{\theta}^{*})\ p(\vec{d}_{i}|z_{i}, \vec{\theta}^{*})}
+%\end{align}
+%Because the redshift $z_{i}$ is independent of the interim prior $\vec{\theta}^{*}$, the interim likelihood $p(\vec{d}_{i}|z_{i}, \vec{\theta}^{*})$ may be expanded further as in Eq.
+\ref{eq:expansion}.
+%\begin{align}
+%\label{eq:expansion}
+%p(\vec{d}_{i}|z_{i}) &= p(\vec{d}_{i}|z_{i})\ p(z_{i}|\vec{d}_{i}, \vec{\theta}^{*})\ \frac{p(\vec{d}_{i}|\vec{\theta}^{*})}{p(z_{i}|\vec{\theta}^{*})\ p(\vec{d}_{i}|z_{i})\ p(\vec{d}_{i}|\vec{\theta}^{*})}
+%\end{align}
+%We cancel the terms $p(\vec{d}_{i}|z_{i})$ and $p(\vec{d}_{i}|\vec{\theta}^{*})$ that appear in both the numerator and denominator of Eq. \ref{eq:expansion} to obtain Eq. \ref{eq:cancellation}.
+%\begin{align}
+%\label{eq:cancellation}
+%p(\vec{d}_{i}|z_{i}) &= \frac{p(z_{i}|\vec{d}_{i}, \vec{\theta}^{*})}{p(z_{i}|\vec{\theta}^{*})}
+%\end{align}
+%
+%Finally we may return to Eq.
+\ref{eq:hyperbayes} to express the log-hyperposterior $\ln\left[p(\vec{\theta}|\{\vec{d}_{i}\})\right]$ in terms of the photo-$z$ interim posteriors $\{p(z_{i}|\vec{d}_{i}, \vec{\theta}^{*})\}$, as in Eq.
+\ref{eq:final}.
+%\begin{align}
+%\label{eq:final}
+%\ln\left[p(\vec{\theta}|\{\vec{d}_{i}\})\right] &\propto \ln\left[p(\vec{\theta})\right] + \sum_{i}\ln\left[\int\ p(z_{i}|\vec{d}_{i}, \vec{\theta}^{*})\ \frac{p(z_{i}|\vec{\theta})}{p(z_{i}|\vec{\theta}^{*})}\ dz_{i}\right]
+%\end{align}
+%Now we have an expression for a quantity proportional to the log-hyperposterior we desire!
+Though the constant of proportionality, $p(\{\vec{d}_{i}\})$ is in general not possible to calculate, we may still characterize the log-hyperposterior by way of MCMC sampling and optimization, which will be discussed in Sec.
+\ref{sec:implementation}.
+%
+%Several assumptions were noted in the above derivation; for clarity, they will be enumerated below, where their limitations will be discussed.
+%\begin{enumerate}
+%\item The PGM is an expression of our beliefs about the physics of the problem, and the inference will only be valid to the degree that the model is correct.
+One can easily think of ways in which the PGM of Fig.
+\ref{fig:pgm} is incomplete; for example, intrinsic galaxy SEDs will evolve with redshift.
+However, we assume that physical processes not represented in the PGM of Fig.
+\ref{fig:pgm} are subdominant and may thus be neglected.
+%\item We must choose a hyperprior $p(\vec{\theta})$; we assume that our choice of a sufficiently general hyperprior that will not be a dominant source of information in the hyperposterior is successful.
+If we choose poorly, the hyperprior may dominate over the hyperlikelihood, downweighting the significance of the data in determining the hyperposterior or biasing the result based on a misunderstanding of the underlying physics.
+%\item We assume independence of galaxies in our catalog such that each galaxy's photometry $\vec{d}_{i}$ is independent from all other galaxies' photometry $\{\vec{d}_{i'\neq i}\}$ and all other galaxies' redshifts $\{z_{i'\neq i}\}$.
+However, the photometry $\{\vec{d}_{i}\}$ will inherently share instrumental and systematic effects due to being observed with the same telescope as part of a single survey project.
+Furthermore, there may be blended objects or other correlations between the redshifts of different galaxies.
+We must assume that such effects are negligible and will not consider them in this treatment.
+%\item We assume that the interim prior is known to us, either as input to or output of the method producing the photo-$z$ interim posteriors.
+Furthermore, it must be representable in the parametrization of $f_{\vec{\theta}}(z)$, i.e.
+our model $f_{\vec{\theta}}(z)$ for $n(z)$ must be flexible enough to encompass both the true distribution we aim to probe and the interim prior.
+%\item Finally, we assume that the photo-$z$ interim posteriors $\{p(z_{i}|\vec{d}_{i}, \vec{\theta}^{*})\}$ are accurate.
+There is some evidence that photo-$z$ PDFs derived by different methods disagree with one another or are strongly dependent on the interim prior.
+However, the choice of the technique by which the photo-$z$ interim posteriors are determined is outside the scope of this paper.
+%\end{enumerate}
+%
+%\subsection{Alternative methods}
+%\label{sec:alternatives}
+%
+%This method is presented in contrast to existing methods for inferring $n(z)$ from photo-$z$ PDFs, which fall into two broad categories: stacking and point estimation.
+Stacking assumes Eq.
+\ref{eq:stacking}, where the constant of proportionality is chosen to ensure $\int f_{\vec{\theta}}(z)\ dz$ is unity.
+%\begin{align}
+%\label{eq:stacking}
+%f_{\vec{\theta}}(z) &\propto \sum_{i}p(z_{i}|\vec{d}_{i}, \vec{\theta}^{*})
+%\end{align}
+%
+%\section{Implementation}
+%\label{sec:implementation}
+%
+%We introduce the Cosmological Hierarchical Inference with Probabilistic Photometric Redshifts (CHIPPR) code as a comprehensive package for estimating $n(z)$ from catalogs of photo-$z$ PDFs.
+In particular, CHIPPR implements both sampling and optimization of Eq.
+\ref{eq:final} and compares the results to those of the more traditional approaches presented in Sec.
+\ref{sec:alternatives}.
+%
+%\section{Experiments}
+%\label{sec:experiments}
+%
+%\section{Discussion}
+%\label{sec:discussion}
+
+%\appendix{}
 
 \begin{acknowledgements}
-AIM thanks Mohammadjavad Vakili for insightful input on statistics, Geoffrey 
-Ryan for assistance in debugging, and Boris Leistedt for helpful comments 
-provided in the preparation of this paper.
+AIM thanks Mohammadjavad Vakili for insightful input on statistics, Geoffrey Ryan for assistance in debugging, and Boris Leistedt for helpful comments provided in the preparation of this paper.
 \end{acknowledgements}
 
+%\bibliographystyle{apj}
+%\bibliography{references}
+
 \end{document}

### 2017-06-13 chippr/defaults.py
normalizing at the end

@@ -2,7 +2,8 @@ import sys
 
 seed = 42
 
-eps = sys.float_info.epsilon
+eps = sys.float_info.min
+log_eps = sys.float_info.min_exp
 
 min_x = 0.
 max_x = 1.

### 2017-06-13 chippr/log_z_dens.py

@@ -107,14 +107,20 @@ class log_z_dens(object):
         log_hyper_likelihood: float
             log likelihood probability associated with parameters in log_nz
         """
-        #log_hyper_lfs = np.logaddexp(self.log_pdfs - self.log_int_pr + self.log_bin_difs, log_nz)#u.safe_log(np.sum((self.log_pdfs + const_terms), axis=1))
-        #log_hyper_likelihood = np.sum(log_hyper_lfs)
-        nz = np.exp(log_nz)
-        n = np.dot(nz, self.bin_difs)
-        norm_nz = nz / n
-        #hyper_lfs = np.sum(np.exp(log_nz + self.log_pdfs - self.log_int_pr + self.log_bin_difs), axis=1)
-        hyper_lfs = np.dot(norm_nz[None,:] * self.pdfs / self.int_pr[None,:], self.bin_difs)#, axis=1)#log_nz + self.log_pdfs - self.int_pr#
-        log_hyper_likelihood = np.sum(u.safe_log(hyper_lfs))# - n
+        # #log_hyper_lfs = np.logaddexp(self.log_pdfs - self.log_int_pr + self.log_bin_difs, log_nz)#u.safe_log(np.sum((self.log_pdfs + const_terms), axis=1))
+        # #log_hyper_likelihood = np.sum(log_hyper_lfs)
+        # #nz = np.exp(log_nz)
+        # #norm_nz = nz / np.dot(nz, self.bin_difs)
+        # hyper_lfs = np.sum(np.exp(log_nz + self.log_pdfs - self.log_int_pr) * self.bin_difs, axis=1)
+        # #hyper_lfs = np.dot(norm_nz[None,:] * self.pdfs / self.int_pr[None,:], self.bin_difs)#, axis=1)#log_nz + self.log_pdfs - self.int_pr#
+        # log_hyper_likelihood = np.sum(u.safe_log(hyper_lfs))# - np.dot(np.exp(log_nz), self.bin_difs)
+        # terms = self.log_pdfs + log_nz[np.newaxis, :] - self.log_int_pr[np.newaxis, :]
+        # galterms = np.sum(np.exp(terms) * self.bin_difs[np.newaxis, :], axis = 1)
+        # log_hyper_likelihood = np.sum(u.safe_log(galterms)) - np.dot(np.exp(log_nz), self.bin_difs)
+        nz = np.exp(log_nz)# - np.max(log_nz))
[…]

### 2017-06-13 chippr/log_z_dens_plots.py

@@ -64,7 +64,7 @@ def plot_ivals(ivals, info, plot_dir):
     ival_integrals = np.dot(np.exp(ivals), bin_difs)
     log_ival_integrals = u.safe_log(ival_integrals)
     sps_sum.hist(log_ival_integrals, color='k', normed=1)
-    sps_sum.vlines(u.safe_log(np.dot(np.exp(info['log_interim_prior']), bin_difs)), 0., 1., linewidth=w_int, linestyle=s_int, alpha=a_int, color=c_int, dashes=d_int, label=l_int+nz)
+    sps_sum.vlines(np.log(np.dot(np.exp(info['log_interim_prior']), bin_difs)), 0., 1., linewidth=w_int, linestyle=s_int, alpha=a_int, color=c_int, dashes=d_int, label=l_int+nz)
     sps_sum.vlines(np.mean(log_ival_integrals), 0., 1., linewidth=w_bfe, linestyle=s_bfe, alpha=a_bfe, color=c_bfe, dashes=d_bfe, label=l_bfe+lnz)
 
     sps_sum.set_xlabel(r'$\ln\left[\int n(z)dz\right]$')

### 2017-06-13 chippr/utils.py

@@ -23,10 +23,12 @@ def safe_log(arr, threshold=d.eps):
     logged: ndarray
         logarithms, with approximation in place of zeros and negative numbers
     """
-    shape = np.shape(arr)
-    flat = arr.flatten()
-    logged = np.log(np.array([max(a,threshold) for a in flat])).reshape(shape)
-    return logged
+    # shape = np.shape(arr)
+    # flat = arr.flatten()
+    # logged = np.log(np.array([max(a,threshold) for a in flat])).reshape(shape)
+    arr[arr < threshold] = threshold
+    return np.log(arr)
+    # return logged
 
 def ingest(in_info):
     """

### 2017-06-13 chippr/defaults.py
final commit, hopefully!

@@ -13,7 +13,7 @@ n_bins = 10
 
 constant_sigma = 0.05
 
-gr_threshold = 1.02
+gr_threshold = 1.2
 
 n_accepted = 3
 n_burned = 2

### 2017-06-13 chippr/log_z_dens.py

@@ -20,16 +20,19 @@ class log_z_dens(object):
 
     def __init__(self, catalog, hyperprior, truth=None, loc='', vb=True):
         """
-        An object representing the redshift density function (normalized redshift distribution function)
+        An object representing the redshift density function (normalized
+        redshift distribution function)
 
         Parameters
         ----------
         catalog: chippr.catalog object
-            dict containing bin endpoints, interim prior bin values, and interim posterior PDF bin values
+            dict containing bin endpoints, interim prior bin values, and
+            interim posterior PDF bin values
         hyperprior: chippr.mvn object
             multivariate Gaussian distribution for hyperprior distribution
         truth: chippr.gmix object, optional
-            true redshift density function expressed as univariate Gaussian mixture
+            true redshift density function expressed as univariate Gaussian
+            mixture
         loc: string, optional
             directory into which to save results and plots made along the way
         vb: boolean, optional
@@ -55,7 +58,7 @@ class log_z_dens(object):
         self.info['log_interim_posteriors'] = self.log_pdfs
 
         if vb:
-            print(str(self.n_bins)+' bins, '+str(len(self.log_pdfs))+' interim posterior PDFs')
+            print(str(self.n_bins) + ' bins, ' + str(len(self.log_pdfs)) + ' interim posterior PDFs')
 
         self.hyper_prior = hyperprior
 
@@ -100,7 +103,8 @@ class log_z_dens(object):
         Parameters
         ----------
         log_nz: numpy.ndarray, float
-            vector of logged redshift density bin values at which to evaluate the hyperlikelihood
+            vector of logged redshift density bin values at which to evaluate
+            the hyperlikelihood
 
         Returns
         -------
@@ -130,7 +134,8 @@ class log_z_dens(object):
         Parameters
         ----------
         log_nz: numpy.ndarray, float
-            vector of logged redshift density bin values at which to evaluate the hyperprior
+            vector of logged redshift density bin values at which to evaluate
+            the hyperprior
 
         Returns
         -------
@@ -153,7 +158,8 @@ class log_z_dens(object):
         Parameters
         ----------
         log_nz: numpy.ndarray, float
-            vector of logged redshift density bin values at which to evaluate the full posterior
+            vector of logged redshift density bin values at which to evaluate
+            the full posterior
 
[…]

### 2017-06-13 chippr/catalog.py
major clean-up, hopefully will not break anything

@@ -16,14 +16,15 @@ from chippr import catalog_plots as plots
 
 class catalog(object):
 
-    def __init__(self, params={}, vb=True, loc=''):
+    def __init__(self, params={}, vb=True, loc='.'):
         """
         Object containing catalog of photo-z interim posteriors
 
         Parameters
         ----------
         params: dict or string, optional
-            dictionary containing parameter values for catalog creation or string containing location of parameter file
+            dictionary containing parameter values for catalog creation or
+            string containing location of parameter file
         vb: boolean, optional
             True to print progress messages to stdout, False to suppress
         loc: string, optional
@@ -61,19 +62,19 @@ class catalog(object):
             True to print progress messages to stdout, False to suppress
         """
         self.n_coarse = self.params['n_bins']
-        x_min = self.params['bin_min']
-        x_max = self.params['bin_max']
+        z_min = self.params['bin_min']
+        z_max = self.params['bin_max']
         self.n_fine = self.n_coarse
         self.n_tot = self.n_coarse * self.n_fine
-        x_range = x_max-x_min
+        z_range = z_max - z_min
 
-        self.dx_coarse = x_range / self.n_coarse
-        self.dx_fine = x_range / self.n_tot
+        self.dz_coarse = z_range / self.n_coarse
+        self.dz_fine = z_range / self.n_tot
 
-        self.x_coarse = np.arange(x_min + 0.5 * self.dx_coarse, x_max, self.dx_coarse)
-        self.x_fine = np.arange(x_min + 0.5 * self.dx_fine, x_max, self.dx_fine)
+        self.z_coarse = np.arange(z_min + 0.5 * self.dz_coarse, z_max, self.dz_coarse)
+        self.z_fine = np.arange(z_min + 0.5 * self.dz_fine, z_max, self.dz_fine)
 
-        self.bin_ends = np.arange(x_min, x_max + self.dx_coarse, self.dx_coarse)
+        self.bin_ends = np.arange(z_min, z_max + self.dz_coarse, self.dz_coarse)
 
         return
 
@@ -91,24 +92,24 @@ class catalog(object):
         coarse: numpy.ndarray, float
             vector of binned values of function
         """
-        coarse = fine / (np.sum(fine) * self.dx_fine)
-        coarse = np.array([np.sum(coarse[k * self.n_fine : (k+1) * self.n_fine]) * self.dx_fine for k in range(self.n_coarse)])
-        coarse /= self.dx_coarse
+        coarse = fine / (np.sum(fine) * self.dz_fine)
[…]

### 2017-06-13 chippr/defaults.py

@@ -22,7 +22,8 @@ plot_colors = 10
 
 def check_sim_params(params={}):
     """
-    Checks simulation parameter dictionary for various keywords and sets to default values if not present
+    Checks simulation parameter dictionary for various keywords and sets to
+    default values if not present
 
     Parameters
     ----------
@@ -98,7 +99,8 @@ def check_variable_sigmas(params):
 
 def check_catastrophic_outliers(params):
     """
-    Sets parameter values pertaining to presence of a catastrophic outlier population
+    Sets parameter values pertaining to presence of a catastrophic outlier
+    population
 
     Parameters
     ----------
@@ -123,11 +125,14 @@ def check_catastrophic_outliers(params):
 
 def check_inf_params(params={}):
     """
-    Checks inference parameter dictionary for various keywords and sets to default values if not present
+    Checks inference parameter dictionary for various keywords and sets to
+    default values if not present
+
     Parameters
     ----------
     params: dict, optional
         dictionary containing initial key/value pairs for inference
+        
     Returns
     -------
     params: dict

### 2017-06-13 chippr/distributor.py

@@ -1,99 +1,0 @@
-# """
-# distribute module sets up for multiprocessing with communication between workers and consumers
-# """
-
-# import multiprocessing as mp
-# import sys
-# import traceback
-
-# class distribute_key(object):
-#     """class to access data that's been serialized to disk"""
-#     def __init__(self):
-#         pass
-
-# class consumer(object):
-#     """consumer is the plotter, producer is MCMC calculation"""
-#     def __init__(self, *args):
-#         pass
-#     def loop(self, queue):
-#         while (True):
-#             key = queue.get()
-#             print('getting key '+str(key))
-#             if (key=='done'):
-#                 self.finish()
-#                 return
-#             self.handle(key)
-#     def finish(self):
-#         print "BAD: Generic finish called."
-
-# def run_offthread(func, *args):
-#    proc = mp.Process(target=func, args=args)
-#    proc.start()
-#    return proc
-# def run_offthread_sync(func, *args):
-#    run_offthread(func, *args).join()
-
-# # def run_offthread(pool, func, *args):
-# #     proc = pool.apply_async(func, args)
-# #     return proc
-# # def run_offthread_sync(pool, func, *args):
-# #     print 'func: ' + str(func)
-# #     return pool.apply(func, args)
-
-
-# def do_consume(ctor, q, args):
-#     """plot results of one run in loop"""
-#     #print(ctor,q,args)
-#     try:
-#         obj = ctor(**args)
-#         obj.loop(q)
-#     except:
-#         e = sys.exc_info()[0]
-#         print 'Nested Exception: '
-#         print traceback.format_exc()
-#         sys.stdout.flush()
-#         raise e
-# def run_consumer(ctor, q, args):
-#     return run_offthread(do_consume, ctor, q, args)
-
-# class distribute(object):
-#     """
-#     class distributes computation over multiple threads
-#     consumers is list of consumers called and added to queue whenever a new key is complete
-#     """
-#     def __init__(self, consumers, start = True, **args):
-#         self.queues = [mp.Queue() for _ in consumers]
-# #         for q in self.queues:
-# #             key = q.get()
-# #             print('original key '+str(key))
-
-#         self.consumer_lambdas = consumers
-#         self.consumers = [run_consumer(c,q, args) for (c,q) in zip(self.consumer_lambdas, self.queues)]
-#         self.started = False
-
-#     def complete_chunk(self, key):
-#         """called by producer when chunk of data has been produced"""
-#         for q in self.queues:
-#             q.put(key)
[…]

### 2017-06-13 chippr/io.py

@@ -1,153 +1,0 @@
-# """
-# key-mcmc module defines loading/saving/naming conventions for MCMC
-# """
-
-# # TO DO: redo state handling with dicts or generators rather than lists
-
-# from utilmcmc import path
-# import distributor
-# import cPickle
-# import hickle as hkl
-# import os
-
-# def safe_load_c(path, num_objs = None):
-#     """read cPickle"""
-#     print 'loading: ' + path
-#     if not os.path.exists(path):
-#         return None
-#     with open(path, "rb") as f:
-#         if num_objs is None:
-#             return cPickle.load(f)
-#         return [cPickle.load(f) for _ in xrange(num_objs)]
-#     f.close()
-# def safe_store_c(path, o):
-#     """write cPickle"""
-#     print 'storing: ' + path
-#     directory = path[:path.rfind('/')]
-#     if not os.path.exists(directory):
-#         os.makedirs(directory)
-#     with open(path, "wb") as f:
-#         cPickle.dump(o, f)
-#     f.close()
-
-# def safe_load_h(path, num_objs = None):
-#     """read hickle"""
-#     if not os.path.exists(path):
-#         return None
-#     with open(path, "r") as f:
-#         if num_objs is None:
-#             return hkl.load(f)
-#         return [hkl.load(f) for _ in xrange(num_objs)]
-#     f.close()
-# def safe_store_h(path, o):
-#     """write hickle"""
-# #     print 'storing hkl:' + path
-#     directory = path[:path.rfind('/')]
-#     if not os.path.exists(directory):
-#         os.makedirs(directory)
-#     with open(path, "w") as f:
-#         hkl.dump(o, f)
-#     f.close()
-
-# # define templates for path
-# state_builder = path("{topdir}/state-{r}.p")#hkl")
-# iterno_builder = path("{topdir}/iterno.p")
-# statistics_builder = path("{topdir}/stat_{stat_name}.p")
-
-# class key(distribute.distribute_key):
-#     """all the dict handling here"""
-#     def __init__(self, **kv):
-#         self.t = None
-#         self.r = None
-#         self.burnin = None
-#         self.update_dict(kv)
-
-#     def copy(self):
-#         ret = key()
-#         ret.t = self.t
-#         ret.r = self.r
-#         ret.burnin = self.burnin
-#         return ret
-
-#     def __hash__(self):
-#         return hash((self.t, self.r, self.burnin))
-
-#     def __eq__(self, other):
[…]

### 2017-06-13 chippr/locator.py

@@ -1,32 +1,0 @@
-# """
-# util-mcmc module defines handy tools for MCMC
-# """
-
-# import sys
-# import numpy as np
-# import statistics
-# import cPickle as cpkl
-# import scipy as sp
-
-# class path(object):
-#     """
-#     path object takes templates of path style and variables for it and makes os.path objects from them
-#     """
-#     def __init__(self, path_template, filled = None):
-#         self.path_template = path_template
-#         if filled is None:
-#             self.filled = {}
-#         else:
-#             self.filled = filled
-
-#     def construct(self, **args):
-#         """actually constructs the final path, as a string.  Optionally takes in any missing parameters"""
-#         nfilled = self.filled.copy()
-#         nfilled.update(args)
-#         return self.path_template.format(**nfilled)
-
-#     def fill(self, **args):
-#         """fills any number of missing parameters, returns new object"""
-#         dct = self.filled.copy()
-#         dct.update(args)
-#         return path(self.path_template, dct)

### 2017-06-13 chippr/log_z_dens.py

@@ -18,7 +18,7 @@ from chippr import log_z_dens_plots as plots
 
 class log_z_dens(object):
 
-    def __init__(self, catalog, hyperprior, truth=None, loc='', vb=True):
+    def __init__(self, catalog, hyperprior, truth=None, loc='.', vb=True):
         """
         An object representing the redshift density function (normalized
         redshift distribution function)
@@ -111,17 +111,7 @@ class log_z_dens(object):
         log_hyper_likelihood: float
             log likelihood probability associated with parameters in log_nz
         """
-        # #log_hyper_lfs = np.logaddexp(self.log_pdfs - self.log_int_pr + self.log_bin_difs, log_nz)#u.safe_log(np.sum((self.log_pdfs + const_terms), axis=1))
-        # #log_hyper_likelihood = np.sum(log_hyper_lfs)
-        # #nz = np.exp(log_nz)
-        # #norm_nz = nz / np.dot(nz, self.bin_difs)
-        # hyper_lfs = np.sum(np.exp(log_nz + self.log_pdfs - self.log_int_pr) * self.bin_difs, axis=1)
-        # #hyper_lfs = np.dot(norm_nz[None,:] * self.pdfs / self.int_pr[None,:], self.bin_difs)#, axis=1)#log_nz + self.log_pdfs - self.int_pr#
-        # log_hyper_likelihood = np.sum(u.safe_log(hyper_lfs))# - np.dot(np.exp(log_nz), self.bin_difs)
-        # terms = self.log_pdfs + log_nz[np.newaxis, :] - self.log_int_pr[np.newaxis, :]
-        # galterms = np.sum(np.exp(terms) * self.bin_difs[np.newaxis, :], axis = 1)
-        # log_hyper_likelihood = np.sum(u.safe_log(galterms)) - np.dot(np.exp(log_nz), self.bin_difs)
-        nz = np.exp(log_nz)# - np.max(log_nz))
+        nz = np.exp(log_nz)
         norm_nz = nz / np.dot(nz, self.bin_difs)
         hyper_lfs = np.sum(norm_nz[None,:] * self.pdfs / self.int_pr[None,:] * self.bin_difs, axis=1)
         log_hyper_likelihood = np.sum(u.safe_log(hyper_lfs)) - np.dot(norm_nz, self.bin_difs)
@@ -142,13 +132,7 @@ class log_z_dens(object):
         log_hyper_prior: float
[…]

### 2017-06-13 chippr/log_z_dens_plots.py

@@ -13,6 +13,8 @@ from chippr import utils as u
 from chippr import plot_utils as pu
 from chippr import stat_utils as s
 
+# defining some shared variables
+
 lnz, nz = r'$\ln[n(z)]$', r'$n(z)$'
 
 s_tru, w_tru, a_tru, c_tru, d_tru, l_tru = '--', 0.5, 1., 'k', [(0, (1, 1))], 'True '
@@ -82,6 +84,8 @@ def set_up_burn_in_plots(n_bins, n_walkers):
     ----------
     n_bins: int
         number of parameters defining n(z)
+    n_walkers: int
+        number of walkers for the sampler
 
     Returns
     -------
@@ -133,6 +137,7 @@ def set_up_burn_in_plots(n_bins, n_walkers):
     chain_evolution_plot = [f_chain_evolution, sps_chain_evolution, random_walkers]
 
     plot_information = (gelman_rubin_evolution_plot, autocorrelation_times_plot, acceptance_fractions_plot, posterior_probabilities_plot, chain_evolution_plot)
+
     return plot_information
 
 def plot_sampler_progress(plot_information, sampler_output, full_chain, burn_ins, plot_dir):
@@ -146,11 +151,14 @@ def plot_sampler_progress(plot_information, sampler_output, full_chain, burn_ins
         autocorrelation times, acceptance fractions, posterior probabilities,
         and chain evolution
     sampler_output: dict
-        dictionary containing array of sampled redshift density function bin values as well as posterior probabilities, acceptance fractions, and autocorrelation times
+        dictionary containing array of sampled redshift density function bin
+        values as well as posterior probabilities, acceptance fractions, and
+        autocorrelation times
     full_chain: ndarray, float
         array of all accepted samples so far
     burn_ins: int
-        number of between-convergence-check intervals that have already been performed
+        number of between-convergence-check intervals that have already been
+        performed
     plot_dir: string
         location in which to store the plots
 
@@ -245,6 +253,7 @@ def plot_sampler_progress(plot_information, sampler_output, full_chain, burn_ins
     f_chain_evolution.savefig(os.path.join(plot_dir, 'chain_evolution.png'), bbox_inches='tight', pad_inches = 0)
 
     plot_information = (gelman_rubin_evolution_plot, autocorrelation_times_plot, acceptance_fractions_plot, posterior_probabilities_plot, chain_evolution_plot)
+
     return plot_information
 
 def plot_estimators(info, plot_dir):
[…]

### 2017-06-13 chippr/mvn.py

@@ -29,41 +29,56 @@ class mvn(object):
     def norm_var(self):
         """
         Function to normalize covariance matrix
+
+        Returns
+        -------
+        det: float
+            determinant of variance
         """
-        return np.linalg.det(self.var)
+        det = np.linalg.det(self.var)
+        return det
 
     def invert_var(self):
         """
         Function to invert covariance matrix
-        """
-        return np.linalg.inv(self.var)
 
-    def evaluate_one(self, x):
+        Returns
+        -------
+        inv: numpy.ndarray, float
+            inverse variance
         """
-        Function to evaluate multivariate Gaussian probability distribution once
+        inv = np.linalg.inv(self.var)
+        return inv
+
+    def evaluate_one(self, z):
+        """
+        Function to evaluate multivariate Gaussian probability distribution
+        once
 
         Parameters
         ----------
-        x: numpy.ndarray, float
-            value at which to evaluate multivariate Gaussian probability distribution
+        z: numpy.ndarray, float
+            value at which to evaluate multivariate Gaussian probability
+            distribution
 
         Returns
         -------
         p: float
-            probability associated with x
+            probability associated with z
         """
-        norm_x = x - self.mean
+        norm_z = z - self.mean
         p = max(d.eps, 1. / (np.sqrt(2. * np.pi) * self.sigma) * \
-                np.exp(-0.5 * np.dot(np.dot(norm_x, self.invvar), norm_x)))
+                np.exp(-0.5 * np.dot(np.dot(norm_z, self.invvar), norm_z)))
         return p
 
-    def evaluate(self, xs):
+    def evaluate(self, zs):
         """
-        Function to evaluate multivariate Gaussian probability distribution at multiple points
+        Function to evaluate multivariate Gaussian probability distribution at
+        multiple points
 
         Parameters
         ----------
-        xs: ndarray, float
+        zs: ndarray, float
             input vectors at which to evaluate probability
 
         Returns
@@ -71,22 +86,23 @@ class mvn(object):
         ps: ndarray, float
             output probabilities
         """
-        ps = np.zeros(len(xs))
-        for n, x in enumerate(xs):
-            ps[n] += self.evaluate_one(x)
+        ps = np.zeros(len(zs))
+        for n, z in enumerate(zs):
+            ps[n] += self.evaluate_one(z)
         return ps
 
     def sample_one(self):
         """
-        Function to take one sample from multivariate Gaussian probability distribution
[…]

### 2017-06-13 chippr/plot_utils.py

@@ -20,7 +20,7 @@ def set_up_plot():
 
     return
 
-def plot_step(sub_plot, bin_ends, plot, s='--', c='k', a=1, w=1, d=[(0,(1,0.0001))], l=None, r=False):
+def plot_step(sub_plot, bin_ends, to_plot, s='--', c='k', a=1, w=1, d=[(0,(1,0.0001))], l=None, r=False):
     """
     Plots a step function
 
@@ -30,7 +30,7 @@ def plot_step(sub_plot, bin_ends, plot, s='--', c='k', a=1, w=1, d=[(0,(1,0.0001
         subplot into which step function is drawn
     bin_ends: list or ndarray
         list or array of endpoints of bins
-    plot: list or ndarray
+    to_plot: list or ndarray
         list or array of values within each bin
     s: string, optional
         matplotlib.pyplot linestyle
@@ -41,17 +41,20 @@ def plot_step(sub_plot, bin_ends, plot, s='--', c='k', a=1, w=1, d=[(0,(1,0.0001
     w: int or float, optional
         matplotlib.pyplot linewidth
     d: list of tuple, optional
-        matplotlib.pyplot dash style, of form [(start_point, (points_on, points_off, ...))]
+        matplotlib.pyplot dash style, of form
+        [(start_point, (points_on, points_off, ...))]
     l: string, optional
         label for function
     r: boolean, optional
         True for rasterized, False for vectorized
     """
 
-    plot_h(sub_plot, bin_ends, plot, s, c, a, w, d, l, r)
-    plot_v(sub_plot, bin_ends, plot, s, c, a, w, d, r)
+    plot_h(sub_plot, bin_ends, to_plot, s, c, a, w, d, l, r)
+    plot_v(sub_plot, bin_ends, to_plot, s, c, a, w, d, r)
 
-def plot_h(sub_plot, bin_ends, plot, s='--', c='k', a=1, w=1, d=[(0,(1,0.0001))], l=None, r=False):
+    return
+
+def plot_h(sub_plot, bin_ends, to_plot, s='--', c='k', a=1, w=1, d=[(0,(1,0.0001))], l=None, r=False):
     """
     Helper function to plot horizontal lines of a step function
 
@@ -61,7 +64,7 @@ def plot_h(sub_plot, bin_ends, plot, s='--', c='k', a=1, w=1, d=[(0,(1,0.0001))]
         subplot into which step function is drawn
     bin_ends: list or ndarray
         list or array of endpoints of bins
[…]

### 2017-06-13 chippr/stat_utils.py

@@ -58,9 +58,11 @@ def calculate_kld(pe, qe, vb=True):
     Parameters
     ----------
     pe: numpy.ndarray, float
-        probability distribution evaluated on a grid whose distance from `q` will be calculated.
+        probability distribution evaluated on a grid whose distance from `q`
+        will be calculated.
     qe: numpy.ndarray, float
-        probability distribution evaluated on a grid whose distance to `p` will be calculated.
+        probability distribution evaluated on a grid whose distance to `p` will
+        be calculated.
     vb: boolean
         report on progress to stdout?
 
@@ -87,9 +89,11 @@ def calculate_rms(pe, qe, vb=True):
     Parameters
     ----------
     pe: numpy.ndarray, float
-        probability distribution evaluated on a grid whose distance _from_ `q` will be calculated.
+        probability distribution evaluated on a grid whose distance _from_ `q`
+        will be calculated.
     qe: numpy.ndarray, float
-        probability distribution evaluated on a grid whose distance _to_ `p` will be calculated.
+        probability distribution evaluated on a grid whose distance _to_ `p`
+        will be calculated.
     vb: boolean
         report on progress to stdout?
 
@@ -106,7 +110,8 @@ def calculate_rms(pe, qe, vb=True):
 
 def single_parameter_gr_stat(chain):
     """
-    Calculates the Gelman-Rubin test statistic of convergence of an MCMC chain over one parameter
+    Calculates the Gelman-Rubin test statistic of convergence of an MCMC chain
+    over one parameter
 
     Parameters
     ----------
@@ -131,7 +136,8 @@ def single_parameter_gr_stat(chain):
 
 def multi_parameter_gr_stat(sample):
     """
-    Calculates the Gelman-Rubin test statistic of convergence of an MCMC chain over multiple parameters
+    Calculates the Gelman-Rubin test statistic of convergence of an MCMC chain
+    over multiple parameters
 
     Parameters
     ----------
@@ -172,9 +178,10 @@ def gr_test(sample, threshold=d.gr_threshold):
     """
     gr = multi_parameter_gr_stat(sample)
     print('Gelman-Rubin test statistic = '+str(gr))
-    return np.max(gr) > threshold
+    test_result = np.max(gr) > threshold
+    return test_result
 
-def cft(xtimes, lag):
+def cft(xtimes, lag):#xtimes has ntimes elements
     """
     Helper function to calculate autocorrelation time for chain of MCMC samples
 
[…]

### 2017-06-13 chippr/utils.py

@@ -5,6 +5,7 @@ import numpy as np
 import chippr
 from chippr import defaults as d
 
+# Choose random seed at import, hopefully carries through everywhere
 np.random.seed(d.seed)
 
 def safe_log(arr, threshold=d.eps):
@@ -13,36 +14,36 @@ def safe_log(arr, threshold=d.eps):
 
     Parameters
     ----------
-    arr: ndarray
-        values to be logged
+    arr: ndarray, float
+        array of values to be logged
     threshold: float, optional
         small, positive value to replace zeros and negative numbers
 
     Returns
     -------
     logged: ndarray
-        logarithms, with approximation in place of zeros and negative numbers
+        logged values, with small value replacing un-loggable values
     """
-    # shape = np.shape(arr)
-    # flat = arr.flatten()
-    # logged = np.log(np.array([max(a,threshold) for a in flat])).reshape(shape)
     arr[arr < threshold] = threshold
-    return np.log(arr)
-    # return logged
+    logged = np.log(arr)
+    return logged
 
 def ingest(in_info):
     """
-    Function reading in parameter file to define functions necessary for generation of posterior probability distributions
+    Function reading in parameter file to define functions necessary for
+    generation of posterior probability distributions
 
     Parameters
     ----------
     in_info: string or dict
-        string containing path to plaintext input file or dict containing likelihood input parameters
+        string containing path to plaintext input file or dict containing
+        likelihood input parameters
 
     Returns
     -------
     in_dict: dict
-        dict containing keys and values necessary for posterior probability distributions
+        dict containing keys and values necessary for posterior probability
+        distributions
     """
     if type(in_info) == str:
         with open(in_info) as infile:

### 2017-06-13 research/scripts/basic_inf_script.py

@@ -1,10 +1,12 @@
 def check_prob_params(params):
     """
     Sets parameter values pertaining to components of probability
+
     Parameters
     ----------
     params: dict
         dictionary containing key/value pairs for probability
+
     Returns
     -------
     params: dict
@@ -26,7 +28,8 @@ def check_prob_params(params):
 
 def make_true_nz(test_name):
     """
-    Function to create true redshift distribution to be shared among several test cases
+    Function to create true redshift distribution to be shared among several
+    test cases
 
     Parameters
     ----------
@@ -40,7 +43,8 @@ def make_true_nz(test_name):
 
     Notes
     -----
-    test_name is currently ignored but will soon be used to load parameters for making true_nz instead of hardcoded values.
+    test_name is currently ignored but will soon be used to load parameters for
+    making true_nz instead of hardcoded values.
     """
     true_amps = np.array([0.20, 0.35, 0.55])
     true_means = np.array([0.5, 0.2, 0.75])
@@ -57,7 +61,8 @@ def set_up_prior(data, params):
     Parameters
     ----------
     data: dict
-        catalog dictionary containing bin endpoints, log interim prior, and log interim posteriors
+        catalog dictionary containing bin endpoints, log interim prior, and log
+        interim posteriors
     params: dict
         dictionary of parameter values for creation of prior
 

### 2017-06-13 research/scripts/multi_inf_script.py

@@ -1,10 +1,12 @@
 def check_prob_params(params):
     """
     Sets parameter values pertaining to components of probability
+
     Parameters
     ----------
     params: dict
         dictionary containing key/value pairs for probability
+
     Returns
     -------
     params: dict
@@ -26,7 +28,8 @@ def check_prob_params(params):
 
 def make_true_nz(test_name):
     """
-    Function to create true redshift distribution to be shared among several test cases
+    Function to create true redshift distribution to be shared among several
+    test cases
 
     Parameters
     ----------
@@ -48,7 +51,7 @@ def make_true_nz(test_name):
 
     true_nz = chippr.gmix(true_amps, true_means, true_sigmas, limits=(0., 1.))
 
-    return(true_nz)
+    return true_nz
 
 def set_up_prior(data, params):
     """
@@ -57,7 +60,8 @@ def set_up_prior(data, params):
     Parameters
     ----------
     data: dict
-        catalog dictionary containing bin endpoints, log interim prior, and log interim posteriors
+        catalog dictionary containing bin endpoints, log interim prior, and log
+        interim posteriors
     params: dict
         dictionary of parameter values for creation of prior
 

### 2017-06-13 research/scripts/multi_sim_script.py

@@ -1,6 +1,7 @@
 def make_true_nz(test_name):
     """
-    Function to create true redshift distribution to be shared among several test cases
+    Function to create true redshift distribution to be shared among several
+    test cases
 
     Parameters
     ----------
@@ -14,7 +15,8 @@ def make_true_nz(test_name):
 
     Notes
     -----
-    test_name is currently ignored but will soon be used to load parameters for making true_nz instead of hardcoded values.
+    test_name is currently ignored but will soon be used to load parameters for
+    making true_nz instead of hardcoded values.
     """
     true_amps = np.array([0.20, 0.35, 0.55])
     true_means = np.array([0.5, 0.2, 0.75])

### 2017-06-13 research/paper/draft-editing.tex
saving this for now

@@
 
-The directed acyclic graph of Fig. \ref{fig:pgm} represents a probabilistic graphical model for hierarchical inference of $n(z)$, the mathematical interpretation of which will be presented below.
+We present the two attributes of a probabilistic graphical model (PGM) for the redshift distribution $n(z)$: the directed acyclic graph of Fig.
+\ref{fig:pgm} and its mathematical interpretation below.
+The assumption underlying the concept of photo-$z$ estimation is that each galaxy $i$ has some observed photometric data $\vec{d}_{i}$ (fluxes, magnitudes, or colors) that is drawn from a function of its redshift $z_{i}$, which is a parameter in this model.
+The redshifts $\{z_{i}\}$ for all galaxies in a survey are random draws from the redshift distribution, a function which itself may be parametrized by hyperparameters comprising $\vec{\phi}$ that are shared among all $N$ galaxies in the survey.
+From this perspective, $n(z)$ is itself a probability distribution for galaxy redshifts, i.e.
+$z_{i} \sim n(z)$.
+Given a functional form of $n(z)$, we can say $n(z) = p(z_{i}|\vec{\phi})$ for any $i = 1, \dots, N$, determining that $n(z)$ is itself a likelihood.
+This PGM enables hierarchical Bayesian inference; we will directly infer the posterior probability distribution $p(\vec{\phi} | \{\vec{d}_{i}\})$ of $\vec{\phi}$ given the observations.
 
@@
 		\includegraphics[width=0.5\textwidth]{pgm.png}
-		\caption{\textbf{Replace \texttt{daft} version with more polished \texttt{tikz} version.}}
-%		\caption{This directed acyclic graph corresponds to a PGM for a hierarchical inference of $p(\vec{\theta}|\{\vec{d}_{i}\})$.
+		\caption{This directed acyclic graph corresponds to a PGM for a hierarchical inference of $p(\vec{\phi}|\{\vec{d}_{i}\})$.
 In this graph, all random variables are shown in circles, with observed random variables shown in shaded circles.
@@
 The box indicates that there are a number of copies of the relationships between boxed parameters, each independent of all others.
-The hyperparameters $\vec{\theta}$ representing $n(z)$ are at the top.
-Independently drawn from a function of the hyperparameters $\vec{\theta}$ are galaxy redshifts $\{z_{i}\}$ below.
+The hyperparameters comprising $\vec{\phi}$, which define $n(z)$, are at the top.
+Independently drawn from a function of the hyperparameters $\vec{\phi}$ are galaxy redshifts $\{z_{i}\}$ below.
 The observed galaxy photometry $\{\vec{d}_{i}\}$, shown in shaded circles, is determined by the redshifts above.}
@@
 \end{figure}
-
-\textbf{Math from \texttt{prob-z} version will go here.}
 

### 2017-06-14 research/paper/draft.tex
periodic commit

@@
 
-The directed acyclic graph of Fig. \ref{fig:pgm} represents a probabilistic 
-graphical model for hierarchical inference of $n(z)$, the mathematical 
-interpretation of which will be presented below.
+We present the two attributes of a probabilistic graphical model (PGM) for the 
+redshift distribution $n(z)$: the directed acyclic graph of Fig. \ref{fig:pgm} 
+and its mathematical interpretation below.  The assumption underlying the 
+concept of photo-$z$ estimation is that each galaxy $i$ has some observed 
+photometric data $\vec{d}_{i}$ (fluxes, magnitudes, or colors) that is drawn 
+from a function of its redshift $z_{i}$, which is a parameter in this model.  
+The redshifts $\{z_{i}\}$ for all galaxies in a survey are random draws from 
+the redshift distribution, a function which itself may be parametrized by 
+hyperparameters comprising $\vec{\phi}$ that are shared among all $N$ galaxies 
+in the survey.  From this perspective, $n(z)$ is itself a probability 
+distribution for galaxy redshifts, i.e. $z_{i} \sim n(z)$.  Given a functional 
+form of $n(z)$, we can say $n(z) = p(z_{i}|\vec{\phi})$ for any $i = 1, \dots, 
+N$, determining that $n(z)$ is in particular a likelihood.
 
@@
 		\includegraphics[width=0.5\textwidth]{pgm.png}
-		\caption{\textbf{Replace \texttt{daft} version with more 
-polished \texttt{tikz} version.}}
+		\caption{This directed acyclic graph corresponds to a PGM for a 
+hierarchical inference of $p(\vec{\phi}|\{\vec{d}_{i}\})$.  In this graph, all 
+random variables are shown in circles, with observed random variables shown in 
+shaded circles.  Relationships between variables are indicated by arrows from 
+parameters to the variables distributed according to functions of them.  The 
+box indicates that there are a number of copies of the relationships between 
+boxed parameters, each independent of all others.  The hyperparameters 
+comprising $\vec{\phi}$, which define $n(z)$, are at the top.  Independently 
+drawn from a function of the hyperparameters $\vec{\phi}$ are galaxy redshifts 
+$\{z_{i}\}$ below.  The observed galaxy photometry $\{\vec{d}_{i}\}$, shown in 
+shaded circles, is determined by the redshifts above.}
 	\label{fig:pgm}
@@
 
-\textbf{Math from \texttt{prob-z} version will go here.}
+It would be easy to estimate $\vec{\phi}$ given $\{z_{i}\}_{N}$, however, what 
+we actually have is photometric data $\{\vec{d}_{i}\}_{N}$ which we used to 
+derive photo-$z$ PDFs.  Because a photo-$z$ PDF is a probability distribution 
+conditioned on the photometric data, it must be interpreted as a posterior 
+probability distribution.
+
+In addition to the data, the posterior probability distribution is conditioned 
+on other information about how it was made, which is in general specific to the 
+method that produced it.  This information only influences the posterior by its 
+implicit distribution over redshifts, which we call the interim prior over 
+redshift.  The interim prior, like the redshift distribution we wish to infer, 
+has some functional form with some parameters, but since the redshift 
+distribution and interim prior need not share the same parametrization, we will 
+use $\vec{\alpha}$ to represent the parameters for the interim prior.  Thus, 
+the posterior probability that is often oversimplified as "$p(z_{i})$" is 
+really the interim posterior probability $p(z_{i} | \vec{d}_{i}, 
+\vec{\alpha})$.  Note that we will assume that all galaxies whose photo-$z$ 
+interim posteriors are derived by a single method share the same interim prior 
+$p(z | \vec{\alpha})$; depending on the method, this may not always be true.
+
+Given a catalog of photo-$z$ interim posteriors $\{p(z_{i} | \vec{d}_{i}, 
+\vec{\alpha})\}$
+
+To make use of the photo-$z$ interim priors, it is absolutely essential that 
+$\vec{\alpha}$ be known.
 

### 2017-06-14 research/paper/draft-editing.tex
periodic backup

@@
 $z_{i} \sim n(z)$.
-Given a functional form of $n(z)$, we can say $n(z) = p(z_{i}|\vec{\phi})$ for any $i = 1, \dots, N$, determining that $n(z)$ is itself a likelihood.
-This PGM enables hierarchical Bayesian inference; we will directly infer the posterior probability distribution $p(\vec{\phi} | \{\vec{d}_{i}\})$ of $\vec{\phi}$ given the observations.
+Given a functional form of $n(z)$, we can say $n(z) = p(z_{i}|\vec{\phi})$ for any $i = 1, \dots, N$, determining that $n(z)$ is in particular a likelihood.
 
@@
 \end{figure}
+
+It would be easy to estimate $\vec{\phi}$ given $\{z_{i}\}_{N}$, however, what we actually have is photometric data $\{\vec{d}_{i}\}_{N}$ which we used to derive photo-$z$ PDFs.
+Because a photo-$z$ PDF is a probability distribution conditioned on the photometric data, it must be interpreted as a posterior probability distribution.
+
+In addition to the data, the posterior probability distribution is conditioned on other information about how it was made, which is in general specific to the method that produced it.
+This information only influences the posterior by its implicit distribution over redshifts, which we call the interim prior over redshift.
+The interim prior, like the redshift distribution we wish to infer, has some functional form with some parameters, but since the redshift distribution and interim prior need not share the same parametrization, we will use $\vec{\alpha}$ to represent the parameters for the interim prior.
+Thus, the posterior probability that is often oversimplified as "$p(z_{i})$" is really the interim posterior probability $p(z_{i} | \vec{d}_{i}, \vec{\alpha})$.
+Note that we will assume that all galaxies whose photo-$z$ interim posteriors are derived by a single method share the same interim prior $p(z | \vec{\alpha})$; depending on the method, this may not always be true.
+
+Given a catalog of photo-$z$ interim posteriors $\{p(z_{i} | \vec{d}_{i}, \vec{\alpha})\}$
+
+To make use of the photo-$z$ interim priors, it is absolutely essential that $\vec{\alpha}$ be known.
 

### 2017-06-15 research/paper/draft-editing.tex
progress in method and model sections

--- 2017-06-13 research/paper/draft-editing.tex
@@
 
-This paper presents a mathematically consistent method for obtaining the posterior distribution over the redshift density function $n(z)$ using a catalog of photo-$z$ PDFs.
+This paper presents a mathematically consistent method for obtaining the posterior probability distribution over the redshift distribution $n(z)$ using a catalog of photo-$z$ PDFs.
+We start by introducing some nomenclature, definitions, and symbols that will be used throughout Secs.
+\ref{sec:model}, \ref{sec:others}, and \ref{sec:implementation}.
+
+We shall say the parameters comprising $\vec{\phi}$ define the redshift distribution under some functional form about which this method is agnostic, so long as the functional form evaluated with some true values of the parameters accurately describes the true $n(z)$.
+Since the redshift distribution is itself a probability distribution, it may be written as $p(z | \vec{\phi})$.
+
+Before jumping into the details of the model, it is crucial to settle on an interpretation of what a photo-$z$ PDF actually is.
+When a photo-$z$ PDF is reported, it is an object containing information about galaxy $i$'s redshift $z_{i}$ based on its photometric data $\vec{d}_{i}$, making a photo-$z$ PDF a \textit{posterior} probability distribution.
+The data $\vec{d}_{i}$ may be fluxes, magnitudes, colors, or any combination thereof.
+We assume that the redshifts of the galaxies are not related to the redshifts or photometry of other galaxies in the survey, i.e.
+they are \textit{independent}.
+
+In addition to the data $\vec{d}_{i}$, the photo-$z$ PDF also contains information imparted by assumptions that went into the process by which the photo-$z$ was made.
+Those assumptions are only relevant insomuch as they impose a preference on the redshift distribution.
+Thus, they can be reduced to a sort of prior on the distribution of redshifts, which can be parametrized under the chosen functional form by parameters $\vec{\phi}^{*}$.
+The redshift distribution $p(z | \vec{\phi}^{*})$ associated with the parameters in $\vec{\phi}^{*}$ shall be called the interim prior because in many cases it was our best guess as to the true redshift distribution before observing any data.
+For example, in template-based photo-$z$ PDF methods, the interim prior is a linear combination of the redshift distributions of the classes of templates based on previous observations.
+(In training-based photo-$z$ PDF methods, the interim prior is related to the redshift distribution of the training set and thus may not be shared among all galaxies for which photo-$z$ PDFs are produced; however, in this work, we will assume that it is shared among all galaxies in a survey.)  Because the choice of parameters $\vec{\phi}^{*}$ is not causally connected to the data or the true redshift distribution but nonetheless contributes to the photo-$z$ PDF, we call it an \textit{interim prior}.
+
+A photo-$z$ PDF is then an \textit{interim posterior} probability distribution of the redshift of a galaxy given its observed photometric data and the interim prior: $p(z_{i} | \vec{d}_{i}, \vec{\phi}^{*})$.
+We would like to learn about $p(z | \vec{\phi})$ from a catalog of photo-$z$ interim posteriors.
 
@@
 
-We present the two attributes of a probabilistic graphical model (PGM) for the redshift distribution $n(z)$: the directed acyclic graph of Fig.
-\ref{fig:pgm} and its mathematical interpretation below.
-The assumption underlying the concept of photo-$z$ estimation is that each galaxy $i$ has some observed photometric data $\vec{d}_{i}$ (fluxes, magnitudes, or colors) that is drawn from a function of its redshift $z_{i}$, which is a parameter in this model.
-The redshifts $\{z_{i}\}$ for all galaxies in a survey are random draws from the redshift distribution, a function which itself may be parametrized by hyperparameters comprising $\vec{\phi}$ that are shared among all $N$ galaxies in the survey.
-From this perspective, $n(z)$ is itself a probability distribution for galaxy redshifts, i.e.
-$z_{i} \sim n(z)$.
-Given a functional form of $n(z)$, we can say $n(z) = p(z_{i}|\vec{\phi})$ for any $i = 1, \dots, N$, determining that $n(z)$ is itself a likelihood.
-This PGM enables hierarchical Bayesian inference; we will directly infer the posterior probability distribution $p(\vec{\phi} | \{\vec{d}_{i}\})$ of $\vec{\phi}$ given the observations.
+We present the two attributes of a probabilistic graphical model (PGM) for the redshift distribution $n(z)$: the directed acyclic graph (DAG) and its mathematical interpretation in terms of Bayesian hierarchical inference.
+
+The fundamental assumption underlying the concept of photo-$z$ estimation is that each galaxy $i$ has some observed photometric data $\vec{d}_{i}$ (fluxes, magnitudes, or colors) that is drawn from a function of its redshift $z_{i}$, which is a parameter in this model.
+The redshifts $\{z_{i}\}$ for all galaxies in a survey are random draws from the redshift distribution, whose parameters  $\vec{\phi}$ under a chosen functional form are called \textit{hyperparameters} because they are shared among all $N$ galaxies in the survey.
+Fig.
+\ref{fig:pgm} illustrates these relationships.
+PGMs of this structure are \textit{hierarchical} in that, though the data are only directly influenced by their redshift parameters, we can still use them to infer something about the global hyperparameters by way of a higher level in the graph.
 
@@
 		\includegraphics[width=0.5\textwidth]{pgm.png}
-		\caption{This directed acyclic graph corresponds to a PGM for a hierarchical inference of $p(\vec{\phi}|\{\vec{d}_{i}\})$.
+		\caption{This directed acyclic graph corresponds to a probabilistic graphical model for a Bayesian hierarchical inference of $p(\vec{\phi}|\{\vec{d}_{i}\})$.
 In this graph, all random variables are shown in circles, with observed random variables shown in shaded circles.
@@
 
+% It would be easy to estimate $\vec{\phi}$ given $\{z_{i}\}_{N}$, however, what we actually have is photometric data $\{\vec{d}_{i}\}_{N}$ which we used to derive photo-$z$ PDFs.
+Because a photo-$z$ PDF is a probability distribution conditioned on the photometric data, it must be interpreted as a posterior probability distribution.
+
+% In addition to the data, the posterior probability distribution is conditioned on other information about how it was made, which is in general specific to the method that produced it.
+This information only influences the posterior by its implicit distribution over redshifts, which we call the interim prior over redshift.
+The interim prior, like the redshift distribution we wish to infer, has some functional form with some parameters, but since the redshift distribution and interim prior need not share the same parametrization, we will use $\vec{\alpha}$ to represent the parameters for the interim prior.
+Thus, the posterior probability that is often oversimplified as "$p(z_{i})$" is really the interim posterior probability $p(z_{i} | \vec{d}_{i}, \vec{\alpha})$.
+Note that we will assume that all galaxies whose photo-$z$ interim posteriors are derived by a single method share the same interim prior $p(z | \vec{\alpha})$; depending on the method, this may not always be true.
+
+The DAG of Fig.
+\ref{fig:pgm} translates directly into a mathematical expression for the posterior distribution $p(\vec{\phi} | \{\vec{d}_{i}\})$ of the hyperparameters given the entire dataset given the set of interim posteriors $\{p(z_{i} | \vec{d}_{i}, \vec{\phi}^{*})\}$ by way of the following derivation.
+From this point forward, we will use log probabilities.
+
+According to Bayes' rule, the posterior distribution of interest is
+\begin{align}
+\log p(\vec{\phi} | \{\vec{d}_{i}\}) &\propto \log p(\{\vec{d}_{i}\} | \vec{\phi}) + \log p(\vec{\phi}),
+\end{align}
+where $p(\vec{\phi})$ is the textit{hyperprior} probability distribution over possible values of the hyperparameters.
+The hyperprior is a choice made by those performing the inference and is not to be confused with the interim prior parameters that influence the photo-$z$ PDF data product.
+
+Focusing on the likelihood term $p(\{\vec{d}_{i}\} | \vec{\phi})$, we employ the power of the hierarchical model through \textit{marginalization} of the unobserved redshift parameters by integrating over them to obtain
+\begin{align}
+% \log p(\{\vec{d}_{i}\} | \vec{\phi}) &= \log\[\int\ \exp\[\log\[p(\{\vec{d}_{i}\} | \{z_{i}\})\] + \log\[p(\{z_{i}\} | \vec{\phi})\]\]\ d\{z_{i}\}\].
+\end{align}
+
+Given a catalog of photo-$z$ interim posteriors $\{p(z_{i} | \vec{d}_{i}, \vec{\alpha})\}$
+
+To make use of the photo-$z$ interim priors, it is absolutely essential that $\vec{\alpha}$ be known.
+
 This framework entails a number of choices and assumptions that must be addressed explicitly.
@@
 \begin{enumerate}
+	\item The chosen functional form with parameters $\vec{\phi}$ must actually be able to describe the true redshift distribution.
+	\item The photometric data and redshift for each galaxy is independent of the photometric data and redshifts of all other galaxies.
+	\item There is one interim prior with parameters $\vec{\phi}^{*}$ shared among all galaxies in the survey.
+	\item We must choose a hyperprior probability distribution $p(\vec{\phi})$ over the hyperparameters.
 	\item While we advocate for the approach of hierarchical inference, the probabilistic graphical model presented here is not the only one that could be proposed.

### 2017-06-15 research/paper/draft.tex
progress in method and model sections

--- 2017-06-12 research/paper/draft.tex
@@
 This paper presents a mathematically consistent method for obtaining the 
-posterior distribution over the redshift density function $n(z)$ using a 
-catalog of photo-$z$ PDFs.
+posterior probability distribution over the redshift distribution $n(z)$ using 
+a catalog of photo-$z$ PDFs.  We start by introducing some nomenclature, 
+definitions, and symbols that will be used throughout Secs. \ref{sec:model}, 
+\ref{sec:others}, and \ref{sec:implementation}.
+
+We shall say the parameters comprising $\vec{\phi}$ define the redshift 
+distribution under some functional form about which this method is agnostic, so 
+long as the functional form evaluated with some true values of the parameters 
+accurately describes the true $n(z)$.  Since the redshift distribution is 
+itself a probability distribution, it may be written as $p(z | \vec{\phi})$.
+
+Before jumping into the details of the model, it is crucial to settle on an 
+interpretation of what a photo-$z$ PDF actually is.  When a photo-$z$ PDF is 
+reported, it is an object containing information about galaxy $i$'s redshift 
+$z_{i}$ based on its photometric data $\vec{d}_{i}$, making a photo-$z$ PDF a 
+\textit{posterior} probability distribution.  The data $\vec{d}_{i}$ may be 
+fluxes, magnitudes, colors, or any combination thereof.  We assume that the 
+redshifts of the galaxies are not related to the redshifts or photometry of 
+other galaxies in the survey, i.e. they are \textit{independent}.
+
+In addition to the data $\vec{d}_{i}$, the photo-$z$ PDF also contains 
+information imparted by assumptions that went into the process by which the 
+photo-$z$ was made.  Those assumptions are only relevant insomuch as they 
+impose a preference on the redshift distribution. Thus, they can be reduced to 
+a sort of prior on the distribution of redshifts, which can be parametrized 
+under the chosen functional form by parameters $\vec{\phi}^{*}$.  The redshift 
+distribution $p(z | \vec{\phi}^{*})$ associated with the parameters in 
+$\vec{\phi}^{*}$ shall be called the interim prior because in many cases it was 
+our best guess as to the true redshift distribution before observing any data.  
+For example, in template-based photo-$z$ PDF methods, the interim prior is a 
+linear combination of the redshift distributions of the classes of templates 
+based on previous observations.  (In training-based photo-$z$ PDF methods, the 
+interim prior is related to the redshift distribution of the training set and 
+thus may not be shared among all galaxies for which photo-$z$ PDFs are 
+produced; however, in this work, we will assume that it is shared among all 
+galaxies in a survey.)  Because the choice of parameters $\vec{\phi}^{*}$ is 
+not causally connected to the data or the true redshift distribution but 
+nonetheless contributes to the photo-$z$ PDF, we call it an \textit{interim 
+prior}.
+
+A photo-$z$ PDF is then an \textit{interim posterior} probability distribution 
+of the redshift of a galaxy given its observed photometric data and the interim 
+prior: $p(z_{i} | \vec{d}_{i}, \vec{\phi}^{*})$.  We would like to learn about 
+$p(z | \vec{\phi})$ from a catalog of photo-$z$ interim posteriors.
 
@@
 
-The directed acyclic graph of Fig. \ref{fig:pgm} represents a probabilistic 
-graphical model for hierarchical inference of $n(z)$, the mathematical 
-interpretation of which will be presented below.
+We present the two attributes of a probabilistic graphical model (PGM) for the 
+redshift distribution $n(z)$: the directed acyclic graph (DAG) and its 
+mathematical interpretation in terms of Bayesian hierarchical inference.
+
+The fundamental assumption underlying the concept of photo-$z$ estimation is 
+that each galaxy $i$ has some observed photometric data $\vec{d}_{i}$ (fluxes, 
+magnitudes, or colors) that is drawn from a function of its redshift $z_{i}$, 
+which is a parameter in this model.  The redshifts $\{z_{i}\}$ for all galaxies 
+in a survey are random draws from the redshift distribution, whose parameters  
+$\vec{\phi}$ under a chosen functional form are called \textit{hyperparameters} 
+because they are shared among all $N$ galaxies in the survey.  Fig. 
+\ref{fig:pgm} illustrates these relationships.  PGMs of this structure are 
+\textit{hierarchical} in that, though the data are only directly influenced by 
+their redshift parameters, we can still use them to infer something about the 
+global hyperparameters by way of a higher level in the graph.
 
@@
 		\includegraphics[width=0.5\textwidth]{pgm.png}
-		\caption{\textbf{Replace \texttt{daft} version with more 
-polished \texttt{tikz} version.}}
+		\caption{This directed acyclic graph corresponds to a 
+probabilistic graphical model for a Bayesian hierarchical inference of 
+$p(\vec{\phi}|\{\vec{d}_{i}\})$.  In this graph, all random variables are shown 
+in circles, with observed random variables shown in shaded circles.  
+Relationships between variables are indicated by arrows from parameters to the 
+variables distributed according to functions of them.  The box indicates that 
+there are a number of copies of the relationships between boxed parameters, 
+each independent of all others.  The hyperparameters comprising $\vec{\phi}$, 
+which define $n(z)$, are at the top.  Independently drawn from a function of 
+the hyperparameters $\vec{\phi}$ are galaxy redshifts $\{z_{i}\}$ below.  The 
+observed galaxy photometry $\{\vec{d}_{i}\}$, shown in shaded circles, is 
+determined by the redshifts above.}
 	\label{fig:pgm}
@@
 
-\textbf{Math from \texttt{prob-z} version will go here.}
+The DAG of Fig. \ref{fig:pgm} translates directly into a mathematical 
+expression for the posterior distribution $p(\vec{\phi} | \{\vec{d}_{i}\})$ of 
+the hyperparameters given the entire dataset given the set of interim 
+posteriors $\{p(z_{i} | \vec{d}_{i}, \vec{\phi}^{*})\}$ by way of the following 
+derivation.  From this point forward, we will use log probabilities.
+
+According to Bayes' rule, the posterior distribution of interest is
+\begin{align}
+\log p(\vec{\phi} | \{\vec{d}_{i}\}) &\propto \log p(\{\vec{d}_{i}\} | 
+\vec{\phi}) + \log p(\vec{\phi}),
+\end{align}
+where $p(\vec{\phi})$ is the textit{hyperprior} probability distribution over 
+possible values of the hyperparameters.  The hyperprior is a choice made by 
+those performing the inference and is not to be confused with the interim prior 
+parameters that influence the photo-$z$ PDF data product.
+
+Focusing on the likelihood term $p(\{\vec{d}_{i}\} | \vec{\phi})$, we employ 
+the power of the hierarchical model through \textit{marginalization} of the 
+unobserved redshift parameters by integrating over them to obtain 
+\begin{align}
+\end{align}
+
+Given a catalog of photo-$z$ interim posteriors $\{p(z_{i} | \vec{d}_{i}, 
+\vec{\alpha})\}$
+
+To make use of the photo-$z$ interim priors, it is absolutely essential that 
+$\vec{\alpha}$ be known.
 
@@
 \begin{enumerate}
+	\item The chosen functional form with parameters $\vec{\phi}$ must 
+actually be able to describe the true redshift distribution.
+	\item The photometric data and redshift for each galaxy is independent 
+of the photometric data and redshifts of all other galaxies.
+	\item There is one interim prior with parameters $\vec{\phi}^{*}$ 
+shared among all galaxies in the survey.
+	\item We must choose a hyperprior probability distribution 
+$p(\vec{\phi})$ over the hyperparameters.
 	\item While we advocate for the approach of hierarchical inference, the 

### 2017-06-15 research/paper/draft-editing.tex
updating paer with fiducial mock data process

--- 2017-06-13 research/paper/draft-editing.tex
@@
 
-We present the two attributes of a probabilistic graphical model (PGM) for the redshift distribution $n(z)$: the directed acyclic graph of Fig.
-\ref{fig:pgm} and its mathematical interpretation below.
-The assumption underlying the concept of photo-$z$ estimation is that each galaxy $i$ has some observed photometric data $\vec{d}_{i}$ (fluxes, magnitudes, or colors) that is drawn from a function of its redshift $z_{i}$, which is a parameter in this model.
-The redshifts $\{z_{i}\}$ for all galaxies in a survey are random draws from the redshift distribution, a function which itself may be parametrized by hyperparameters comprising $\vec{\phi}$ that are shared among all $N$ galaxies in the survey.
-From this perspective, $n(z)$ is itself a probability distribution for galaxy redshifts, i.e.
-$z_{i} \sim n(z)$.
-Given a functional form of $n(z)$, we can say $n(z) = p(z_{i}|\vec{\phi})$ for any $i = 1, \dots, N$, determining that $n(z)$ is itself a likelihood.
-This PGM enables hierarchical Bayesian inference; we will directly infer the posterior probability distribution $p(\vec{\phi} | \{\vec{d}_{i}\})$ of $\vec{\phi}$ given the observations.
+The directed acyclic graph of Fig. \ref{fig:pgm} represents a probabilistic graphical model for hierarchical inference of $n(z)$, the mathematical interpretation of which will be presented below.
 
@@
 	\begin{center}
-		\includegraphics[width=0.5\textwidth]{pgm.png}
-		\caption{This directed acyclic graph corresponds to a PGM for a hierarchical inference of $p(\vec{\phi}|\{\vec{d}_{i}\})$.
+% 		\includegraphics[width=0.5\textwidth]{pgm.png}
+		\caption{\textbf{Replace \texttt{daft} version with more polished \texttt{tikz} version.}}
+%		\caption{This directed acyclic graph corresponds to a PGM for a hierarchical inference of $p(\vec{\theta}|\{\vec{d}_{i}\})$.
 In this graph, all random variables are shown in circles, with observed random variables shown in shaded circles.
@@
 The box indicates that there are a number of copies of the relationships between boxed parameters, each independent of all others.
-The hyperparameters comprising $\vec{\phi}$, which define $n(z)$, are at the top.
-Independently drawn from a function of the hyperparameters $\vec{\phi}$ are galaxy redshifts $\{z_{i}\}$ below.
+The hyperparameters $\vec{\theta}$ representing $n(z)$ are at the top.
+Independently drawn from a function of the hyperparameters $\vec{\theta}$ are galaxy redshifts $\{z_{i}\}$ below.
 The observed galaxy photometry $\{\vec{d}_{i}\}$, shown in shaded circles, is determined by the redshifts above.}
@@
 \end{figure}
+
+\textbf{Math from \texttt{prob-z} version will go here.}
 
@@
 
+\begin{enumerate}
+	\item Choose a true $n(z)$ that is a continuous function with known parameters $\vec{\theta}$.
+	\item Sample the true $n(z)$ to create a catalog of $N$ true redshifts $z_{i}$.
+	\item Choose a true intrinsic scatter $\sigma$, and sample Gaussians $p(z'_{i} | z_{i}, \sigma) = \mathcal{N}(z_{i}, \sigma)$ to get "observed" redshifts $z'_{i}$ defining Gaussian likelihoods $p(\vec{d}_{i} | z'_{i}, \sigma) = \mathcal{N}(z'_{i}, \sigma)$.
+	\item Choose a parametrization and the parameters $\vec{\phi}^{*}$ of the interim prior.
+	\item Create interim posteriors $p(z_{i} | \vec{d}_{i}, \vec{\phi}^{*}) = p(\vec{d}_{i} | z_{i})\ p(z_{i} | \vec{\phi}^{*})$
+\end{enumerate}
+
 This method for deriving mock data is referred to as the fiducial case, and variations on it will refer directly to the steps that are altered.
@@
 \textbf{Review precision and bias from \texttt{kld.ipynb} and interpret in terms of a \% difference.}
+
+\subsection{Underlying $n(z)$ Effects}
+\label{sec:truth}
+
+Existing $n(z)$ estimators are systematically smoother than the true $n(z)$.
+Here we show that the traditional estimators of $n(z)$ cannot recover a highly featured $n(z)$.
+The implication of this failure is quite serious; the consistently smooth, unimodal $n(z)$ estimates appearing in the literature could result from much more featured true redshift distributions, and there would be no way to catch this error without using a fully probabilistic method.
+
+\subsubsection{Featureless $n(z)$}
+\label{sec:smooth}
+
+In this test, we choose a weakly featured true $n(z)$ of Fig. \ref{fig:smooth} with a smooth, unimodal shape.
+
+\begin{figure}
+	\begin{center}
+		%\includegraphics[width=0.5\textwidth]{pgm.png}
+		\caption{[plot of a weakly featured true $n(z)$, probably SDSS interim prior]}
+		\label{fig:featured}
+	\end{center}
+\end{figure}
+
+\subsubsection{Featured $n(z)$}
+\label{sec:featured}
+
+In this test, we choose the true $n(z)$ of Fig. \ref{fig:featured} with nontrivial structure.
+
+\begin{figure}
+	\begin{center}
+		%\includegraphics[width=0.5\textwidth]{pgm.png}
+		\caption{[plot of a highly featured, multimodal true $n(z)$]}
+		\label{fig:featured}
+	\end{center}
+\end{figure}
+
+This featured true $n(z)$ will henceforth be referred to as the fiducial $n(z)$
 
@@
 		\label{fig:trainintpr}
-	\end{center}
-\end{figure}
-
-\subsection{Underlying $n(z)$ Effects}
-\label{sec:truth}
-
-Existing $n(z)$ estimators are systematically smoother than the true $n(z)$.
-Here we show that the traditional estimators of $n(z)$ cannot recover a highly featured $n(z)$.
-The implication of this failure is quite serious; the consistently smooth, unimodal $n(z)$ estimates appearing in the literature could result from much more featured true redshift distributions, and there would be no way to catch this error without using a fully probabilistic method.
-
-\subsubsection{Featured $n(z)$}
-\label{sec:featured}
-
-In this test, we choose the true $n(z)$ of Fig. \ref{fig:featured}, which has nontrivial structure.
-
-\begin{figure}
-	\begin{center}
-		%\includegraphics[width=0.5\textwidth]{pgm.png}
-		\caption{[plot of a highly featured (delta function or multimodal?) true $n(z)$]}
-		\label{fig:featured}
 	\end{center}

### 2017-06-15 chippr/log_z_dens.py
fixing merge errors

@@ -94,13 +94,6 @@ class log_z_dens(object):
         if not os.path.exists(self.res_dir):
             os.makedirs(self.res_dir)
 
-        if type(params) == str:
-            self.params = u.ingest(params)
-        else:
-            self.params = params
-        self.params = d.check_inf_params(self.params)
-        if vb:
-            print self.dir + ': ' + str(self.params)
         return
 
     def evaluate_log_hyper_likelihood(self, log_nz):
@@ -356,7 +349,7 @@ class log_z_dens(object):
         mcmc_outputs['acors'] = acors
         return mcmc_outputs
 
-    def calculate_samples(self, ivals, n_accepted=10**d.n_accepted, n_burned=10**d.n_burned, vb=True, n_procs=1, no_data=0, no_prior=0):
+    def calculate_samples(self, ivals, n_accepted=d.n_accepted, n_burned=d.n_burned, vb=True, n_procs=1, no_data=0, no_prior=0):
         """
         Calculates samples estimating the redshift density function
 
@@ -408,7 +401,7 @@ class log_z_dens(object):
             while self.burning_in:
                 if vb:
                     print('beginning sampling '+str(self.burn_ins))
-                burn_in_mcmc_outputs = self.sample(vals, n_burned)
+                burn_in_mcmc_outputs = self.sample(vals, 10**n_burned)
                 chain = burn_in_mcmc_outputs['chains']
                 burn_in_mcmc_outputs['chains'] -= u.safe_log(np.sum(np.exp(chain) * self.bin_difs[np.newaxis, np.newaxis, :], axis=2))[:, :, np.newaxis]
                 with open(os.path.join(self.res_dir, 'mcmc'+str(self.burn_ins)+'.p'), 'wb') as file_location:
@@ -420,7 +413,7 @@ class log_z_dens(object):
                 vals = np.array([item[-1] for item in burn_in_mcmc_outputs['chains']])
                 self.burn_ins += 1
 
-            mcmc_outputs = self.sample(vals, n_accepted)
+            mcmc_outputs = self.sample(vals, 10**n_accepted)
             chain = mcmc_outputs['chains']
             mcmc_outputs['chains'] -= u.safe_log(np.sum(np.exp(chain) * self.bin_difs[np.newaxis, np.newaxis, :], axis=2))[:, :, np.newaxis]
             full_chain = np.concatenate((full_chain, mcmc_outputs['chains']), axis=1)

### 2017-06-15 chippr/stat_utils.py

@@ -160,7 +160,7 @@ def multi_parameter_gr_stat(sample):
         Rs[i] = single_parameter_gr_stat(chains)
     return Rs
 
-def gr_test(sample, threshold):
+def gr_test(sample, threshold=d.eps):
     """
     Performs the Gelman-Rubin test of convergence of an MCMC chain
 
@@ -168,7 +168,7 @@ def gr_test(sample, threshold):
     ----------
     sample: numpy.ndarray, float
         chain output
-    threshold: float
+    threshold: float, optional
         Gelman-Rubin test statistic criterion (usually around 1)
 
     Returns

### 2017-06-15 research/scripts/basic_sim_script.py
forgot to comment slow step for travis

@@ -83,19 +83,14 @@ if __name__ == "__main__":
     from chippr import *
 
     result_dir = os.path.join('..', 'results')
-    name_file = 'which_sim_tests.txt'
-
-    with open(name_file) as tests_to_run:
-        all_tests = {}
-        for test_name in tests_to_run:
-            true_nz = make_true_nz(test_name)
-            true_zs = make_true_zs(true_nz)
-            test_info = {}
-            test_info['name'] = test_name
-            test_info['true_nz'] = true_nz
-            test_info['true_zs'] = true_zs
-            all_tests[test_name] = test_info
-
-    nps = mp.cpu_count()-1
-    pool = mp.Pool(nps)
-    pool.map(make_catalog, all_tests.keys())
+    test_name = 'null_test\n'
+    all_tests = {}
+    true_nz = make_true_nz(test_name)
+    true_zs = make_true_zs(true_nz)
+    test_info = {}
+    test_info['name'] = test_name
+    test_info['true_nz'] = true_nz
+    test_info['true_zs'] = true_zs
+    all_tests[test_name] = test_info
+    
+    make_catalog(test_name)

### 2017-06-15 chippr/catalog.py
simulation panel of paper plot format

@@ -125,26 +125,33 @@ class catalog(object):
         self.samp_range = range(self.n_items)
 
         self.obs_samps = self.sample_obs()
-        if vb:
-            plots.plot_obs_scatter(self.true_samps, self.obs_samps, plot_loc=self.plot_dir)
 
         self.int_pr = int_pr
         self.proc_bins()
 
         self.obs_lfs = self.evaluate_lfs()
 
-        int_pr_fine = int_pr.evaluate(self.z_fine)
+        int_pr_fine = self.int_pr.evaluate(self.z_fine)
         int_pr_coarse = self.coarsify(int_pr_fine)
 
-        pfs = np.zeros((self.n_items, self.n_coarse))
+        # rewrite to take advantage of numpy array manipulation
+        pfs_fine = np.zeros((self.n_items, self.n_tot))
         for n in self.samp_range:
-            pf = int_pr_fine * self.obs_lfs[n]
-            pf = self.coarsify(pf)
-            pfs[n] += pf
+            pfs_fine[n] += int_pr_fine * self.obs_lfs[n]
+            pfs_fine[n] /= np.sum(pfs_fine[n]) * self.dz_fine
+
+        if vb:
+            plots.plot_obs_scatter(self.true_samps, pfs_fine, self.z_fine, plot_loc=self.plot_dir)
+
+        # rewrite to take advantage of numpy array manipulation
+        pfs_coarse = np.zeros((self.n_items, self.n_coarse))
+        for n in self.samp_range:
+            pfs_coarse[n] += self.coarsify(pfs_fine[n])
+            pfs_coarse[n] /= np.sum(pfs_coarse[n]) * self.dz_coarse
 
         self.cat['bin_ends'] = self.bin_ends
         self.cat['log_interim_prior'] = u.safe_log(int_pr_coarse)
-        self.cat['log_interim_posteriors'] = u.safe_log(pfs)
+        self.cat['log_interim_posteriors'] = u.safe_log(pfs_coarse)
 
         return self.cat
 

### 2017-06-15 chippr/catalog_plots.py

@@ -6,6 +6,7 @@ mpl.use('PS')
 import matplotlib.pyplot as plt
 
 import chippr
+from chippr import defaults as d
 from chippr import utils as u
 from chippr import plot_utils as pu
 
@@ -30,32 +31,51 @@ def plot_true_histogram(true_samps, n_bins=50, plot_loc='', plot_name='true_hist
     sps.hist(true_samps, bins=n_bins, normed=1, color='k')
     sps.set_xlabel(r'$z_{true}$')
     sps.set_ylabel(r'$n(z_{true})$')
-    f.savefig(os.path.join(plot_loc, plot_name), bbox_inches='tight', pad_inches = 0, dpi=100)
+    f.savefig(os.path.join(plot_loc, plot_name), bbox_inches='tight', pad_inches = 0, dpi=d.dpi)
 
     return
 
-def plot_obs_scatter(true_samps, obs_samps, plot_loc='', plot_name='obs_scatter.png'):
+def plot_obs_scatter(true_zs, pfs, z_grid, plot_loc='', plot_name='scatter.png'):
     """
     Plots a scatterplot of true and observed redshift values
 
     Parameters
     ----------
-    true_samps: numpy.ndarray, float
+    true_zs: numpy.ndarray, float
         vector of true values of scalar input
-    obs_samps: numpy.ndarray, float
-        vector of observed values of scalar input
+    pfs: numpy.ndarray, float
+        matrix of interim posteriors evaluated on a fine grid
+    z_grid: numpy.ndarray, float
+        vector of redshifts at which
     plot_loc: string, optional
         location in which to store plot
     plot_name: string, optional
         filename for plot
     """
+    obs_zs = np.array([z_grid[np.argmax(pf)] for pf in pfs])
+    max_pfs = np.max(pfs)
+    n = len(obs_zs)
+    dz = (max(z_grid) - min(z_grid)) / len(z_grid)
+    jitters = np.random.uniform(-1. * dz / 2., dz / 2., n)
+    obs_zs = obs_zs + jitters
     pu.set_up_plot()
     f = plt.figure(figsize=(5, 5))
     sps = f.add_subplot(1, 1, 1)
-    sps.scatter(true_samps, obs_samps, c='k', linewidth=0., alpha=0.1)
[…]

### 2017-06-15 chippr/defaults.py

@@ -18,7 +18,8 @@ gr_threshold = 1.2
 n_accepted = 3
 n_burned = 2
 
-plot_colors = 10
+plot_colors = 5
+dpi = 250
 
 def check_sim_params(params={}):
     """

### 2017-06-15 chippr/log_z_dens_plots.py

@@ -5,7 +5,6 @@ import scipy as sp
 import matplotlib as mpl
 mpl.use('PS')
 import matplotlib.pyplot as plt
-import matplotlib.cm as cm
 
 import chippr
 from chippr import defaults as d
@@ -25,9 +24,6 @@ s_exp, w_exp, a_exp, c_exp, d_exp, l_exp = '--', 0.5, 1., 'r', [(0, (2.5, 2.5))]
 s_mle, w_mle, a_mle, c_mle, d_mle, l_mle = '-', 2.5, 1., 'b', [(0, (1, 1))], 'MMLE '
 s_smp, w_smp, a_smp, c_smp, d_smp, l_smp = '-', 1., 1., 'k', [(0, (1, 1))], 'Sampled '
 s_bfe, w_bfe, a_bfe, c_bfe, d_bfe, l_bfe = '-', 1.5, 1., 'b', [(0, (1, 1))], 'Mean of\n Samples '
-
-cmap = np.linspace(0., 1., d.plot_colors)
-colors = [cm.viridis(i) for i in cmap]
 
 def plot_ivals(ivals, info, plot_dir):
     """
@@ -54,7 +50,7 @@ def plot_ivals(ivals, info, plot_dir):
     f = plt.figure(figsize=(10, 5))
     sps_samp = f.add_subplot(1, 2, 1)
     for i in range(d.plot_colors):
-        pu.plot_step(sps_samp, info['bin_ends'], ivals[walkers[i]], c=colors[i])
+        pu.plot_step(sps_samp, info['bin_ends'], ivals[walkers[i]], c=pu.colors[i])
     pu.plot_step(sps_samp, info['bin_ends'], info['log_interim_prior'], w=w_int, s=s_int, a=a_int, c=c_int, d=d_int, l=l_int+nz)
     if info['truth'] is not None:
         sps_samp.plot(info['truth']['z_grid'], np.log(info['truth']['nz_grid']), linewidth=w_tru, alpha=a_tru, color=c_tru, label=l_tru+nz)
@@ -204,7 +200,7 @@ def plot_sampler_progress(plot_information, sampler_output, full_chain, burn_ins
                            s=2)
     sps_autocorrelation_times.set_xlim(0, (burn_ins + 2) * n_burn_test)
     autocorrelation_times_plot = [f_autocorrelation_times, sps_autocorrelation_times]
[…]

### 2017-06-15 chippr/defaults.py
paper plot for inference ready to go

@@ -9,7 +9,6 @@ min_x = 0.
 max_x = 1.
 
 n_bins = 10
-# n_items = 4
 
 constant_sigma = 0.03
 

### 2017-06-15 chippr/log_z_dens.py

@@ -188,7 +188,7 @@ class log_z_dens(object):
                 return -2. * self.evaluate_log_hyper_posterior(log_nz)
 
         if vb:
-            print(self.dir + 'starting at', start, _objective(start))
+            print(self.dir + ' starting at ', start, _objective(start))
 
         res = op.minimize(_objective, start, method="Nelder-Mead", options={"maxfev": 1e5, "maxiter":1e5})
 

### 2017-06-15 chippr/log_z_dens_plots.py

@@ -17,13 +17,14 @@ from chippr import stat_utils as s
 lnz, nz = r'$\ln[n(z)]$', r'$n(z)$'
 
 s_tru, w_tru, a_tru, c_tru, d_tru, l_tru = '-', 0.5, 1., 'k', [(0, (1, 1))], 'True '
+s_tbp, w_tbp, a_tbp, c_tbp, d_tbp, l_tbp = '-', 0.5, 0.75, 'k', [(0, (1, 1))], 'Binned True '
 s_int, w_int, a_int, c_int, d_int, l_int = '-', 0.5, 0.5, 'k', [(0, (1, 1))], 'Interim '
-s_stk, w_stk, a_stk, c_stk, d_stk, l_stk = '--', 1.5, 1., 'g', [(0, (7.5, 7.5))], 'Stacked '
-s_map, w_map, a_map, c_map, d_map, l_map = '--', 1., 1., 'r', [(0, (5., 5.))], 'MMAP '
-s_exp, w_exp, a_exp, c_exp, d_exp, l_exp = '--', 0.5, 1., 'r', [(0, (2.5, 2.5))], 'MExp '
-s_mle, w_mle, a_mle, c_mle, d_mle, l_mle = '-', 2.5, 1., 'b', [(0, (1, 1))], 'MMLE '
+s_stk, w_stk, a_stk, c_stk, d_stk, l_stk = '--', 1.5, 0.75, pu.colors[4], [(0, (7.5, 7.5))], 'Stacked '#'g'
+s_map, w_map, a_map, c_map, d_map, l_map = '--', 1., 0.75, pu.colors[3], [(0, (5., 5.))], 'MMAP '#'r'
+s_exp, w_exp, a_exp, c_exp, d_exp, l_exp = '--', 0.5, 0.75, pu.colors[2], [(0, (2.5, 2.5))], 'MExp '#'r'
+s_mle, w_mle, a_mle, c_mle, d_mle, l_mle = '-', 1.5, 0.75, pu.colors[1], [(0, (1, 1))], 'MMLE '#'b'
 s_smp, w_smp, a_smp, c_smp, d_smp, l_smp = '-', 1., 1., 'k', [(0, (1, 1))], 'Sampled '
-s_bfe, w_bfe, a_bfe, c_bfe, d_bfe, l_bfe = '-', 1.5, 1., 'b', [(0, (1, 1))], 'Mean of\n Samples '
[…]

### 2017-06-15 chippr/stat_utils.py

@@ -160,7 +160,7 @@ def multi_parameter_gr_stat(sample):
         Rs[i] = single_parameter_gr_stat(chains)
     return Rs
 
-def gr_test(sample, threshold=d.eps):
+def gr_test(sample, threshold=d.gr_threshold):
     """
     Performs the Gelman-Rubin test of convergence of an MCMC chain
 

### 2017-06-15 research/scripts/basic_inf_script.py

@@ -128,7 +128,7 @@ def do_inference(given_key):
 
     (prior, cov) = set_up_prior(data, params)
 
-    nz = log_z_dens(data, prior, truth=true_nz, loc=test_dir, params=param_file_name, vb=True)
+    nz = log_z_dens(data, prior, truth=true_nz, loc=test_dir, vb=True)
 
     nz_stacked = nz.calculate_stacked()
     print('stacked: '+str(np.dot(np.exp(nz_stacked), z_difs)))

### 2017-06-15 research/scripts/basic_replot_script.py

@@ -80,7 +80,7 @@ def just_plot(given_key):
     n_ivals = 2 * n_bins
     initial_values = prior.sample(n_ivals)
 
-    nz = log_z_dens(data, prior, truth=true_nz, loc=test_dir, params=param_file_name, vb=True)
+    nz = log_z_dens(data, prior, truth=true_nz, loc=test_dir, vb=True)
 
     nz.info = nz.read('nz.p')
     nz.plot_estimators()
@@ -97,17 +97,13 @@ if __name__ == "__main__":
     from chippr import *
 
     result_dir = os.path.join('..', 'results')
-    name_file = 'which_inf_tests.txt'
+    test_name = 'null_test\n'
 
-    with open(name_file) as tests_to_run:
-        all_tests = {}
-        for test_name in tests_to_run:
-            true_nz = make_true_nz(test_name)
-            test_info = {}
-            test_info['name'] = test_name
-            test_info['truth'] = true_nz
-            all_tests[test_name] = test_info
+    all_tests = {}
+    true_nz = make_true_nz(test_name)
+    test_info = {}
+    test_info['name'] = test_name
+    test_info['truth'] = true_nz
+    all_tests[test_name] = test_info
 
-    nps = mp.cpu_count()-1
-    pool = mp.Pool(nps)
-    pool.map(just_plot, all_tests.keys())
+    just_plot(test_name)

### 2017-06-15 research/paper/draft.tex
putting shiny new figures in paper

--- 2017-06-12 research/paper/draft.tex
@@
 	\begin{center}
-		\includegraphics[width=0.5\textwidth]{pgm.png}
-		\caption{\textbf{Replace \texttt{daft} version with more 
-polished \texttt{tikz} version.}}
+ 		\includegraphics[width=0.5\textwidth]{fig/pgm.png}
+		\caption{This directed acyclic graph corresponds to a PGM for a 
+hierarchical inference of $p(\vec{\theta}|\{\vec{d}_{i}\})$.  In this graph, 
+all random variables are shown in circles, with observed random variables shown 
+in shaded circles.  Relationships between variables are indicated by arrows 
+from parameters to the variables distributed according to functions of them.  
+The box indicates that there are a number of copies of the relationships 
+between boxed parameters, each independent of all others.  The hyperparameters 
+$\vec{\theta}$ representing $n(z)$ are at the top.  Independently drawn from a 
+function of the hyperparameters $\vec{\theta}$ are galaxy redshifts $\{z_{i}\}$ 
+below.  The observed galaxy photometry $\{\vec{d}_{i}\}$, shown in shaded 
+circles, is determined by the redshifts above.}
 	\label{fig:pgm}
@@
 
+\begin{enumerate}
+	\item Choose a true $n(z)$ that is a continuous function with known 
+parameters $\vec{\theta}$.
+	\item Sample the true $n(z)$ to create a catalog of $N$ true redshifts 
+$z_{i}$.
+	\item Choose a true intrinsic scatter $\sigma$, and sample Gaussians 
+$p(z'_{i} | z_{i}, \sigma) = \mathcal{N}(z_{i}, \sigma)$ to get "observed" 
+redshifts $z'_{i}$ defining Gaussian likelihoods $p(\vec{d}_{i} | z'_{i}, 
+\sigma) = \mathcal{N}(z'_{i}, \sigma)$.
+	\item Choose a parametrization and the parameters $\vec{\phi}^{*}$ of 
+the interim prior.
+	\item Create interim posteriors $p(z_{i} | \vec{d}_{i}, \vec{\phi}^{*}) 
+= p(\vec{d}_{i} | z_{i})\ p(z_{i} | \vec{\phi}^{*})$ in this parametrization.
+\end{enumerate}
+
 This method for deriving mock data is referred to as the fiducial case, and 
@@
 terms of a \% difference.}
+
+\subsection{Underlying $n(z)$ Effects}
+\label{sec:truth}
+
+Existing $n(z)$ estimators are systematically smoother than the true $n(z)$.  
+Here we show that the traditional estimators of $n(z)$ cannot recover a highly 
+featured $n(z)$.  The implication of this failure is quite serious; the 
+consistently smooth, unimodal $n(z)$ estimates appearing in the literature 
+could result from much more featured true redshift distributions, and there 
+would be no way to catch this error without using a fully probabilistic method.
+
+\subsubsection{Featureless $n(z)$}
+\label{sec:smooth}
+
+In this test, we choose a weakly featured true $n(z)$ of Fig. \ref{fig:smooth} 
+with a smooth, unimodal shape.
+
+\begin{figure}
+	\begin{center}
+		%\includegraphics[width=0.5\textwidth]{pgm.png}
+		\caption{[plot of a weakly featured true $n(z)$, probably SDSS 
+interim prior]}
+		\label{fig:featured}
+	\end{center}
+\end{figure}
+
+\subsubsection{Featured $n(z)$}
+\label{sec:featured}
+
+In this test, we choose the true $n(z)$ of Fig. \ref{fig:featured} with 
+nontrivial structure.
+
+\begin{figure}
+	\begin{center}
+
+\includegraphics[width=0.5\textwidth]{fig/fiducial/scatter.png}\\
+
+\includegraphics[width=0.5\textwidth]{fig/fiducial/estimators.png}		
+		\caption{[plot of a highly featured, multimodal true $n(z)$]}
+		\label{fig:featured}
+	\end{center}
+\end{figure}
+
+This featured true $n(z)$ will henceforth be referred to as the fiducial $n(z)$
 
@@
 
-\subsection{Underlying $n(z)$ Effects}
-\label{sec:truth}
-
-Existing $n(z)$ estimators are systematically smoother than the true $n(z)$.  
-Here we show that the traditional estimators of $n(z)$ cannot recover a highly 
-featured $n(z)$.  The implication of this failure is quite serious; the 
-consistently smooth, unimodal $n(z)$ estimates appearing in the literature 
-could result from much more featured true redshift distributions, and there 
-would be no way to catch this error without using a fully probabilistic method.
-
-\subsubsection{Featured $n(z)$}
-\label{sec:featured}
-
-In this test, we choose the true $n(z)$ of Fig. \ref{fig:featured}, which has 
-nontrivial structure.
-
-\begin{figure}
-	\begin{center}
-		%\includegraphics[width=0.5\textwidth]{pgm.png}
-		\caption{[plot of a highly featured (delta function or 
-multimodal?) true $n(z)$]}
-		\label{fig:featured}
-	\end{center}
-\end{figure}
-
 \section{Application to Realistic Mock Data}

### 2017-06-15 research/paper/draft.tex
adding figures to paper

@@
 \begin{enumerate}
-	\item Choose a true $n(z)$ that is a continuous function with known 
-parameters $\vec{\theta}$.
-	\item Sample the true $n(z)$ to create a catalog of $N$ true redshifts 
-$z_{i}$.
-	\item Choose a true intrinsic scatter $\sigma$, and sample Gaussians 
-$p(z'_{i} | z_{i}, \sigma) = \mathcal{N}(z_{i}, \sigma)$ to get "observed" 
-redshifts $z'_{i}$ defining Gaussian likelihoods $p(\vec{d}_{i} | z'_{i}, 
-\sigma) = \mathcal{N}(z'_{i}, \sigma)$.
-	\item Choose a parametrization and the parameters $\vec{\phi}^{*}$ of 
-the interim prior.
-	\item Create interim posteriors $p(z_{i} | \vec{d}_{i}, \vec{\phi}^{*}) 
-= p(\vec{d}_{i} | z_{i})\ p(z_{i} | \vec{\phi}^{*})$ in this parametrization.
+	\item \label{itm:true} Choose a true $n(z)$ that is a continuous 
+function with known parameters $\vec{\theta}$.
+	\item \label{itm:sample} Sample the true $n(z)$ to create a catalog of 
+$N$ true redshifts $z_{i}$.
+	\item \label{itm:scatter} Choose a true intrinsic scatter $\sigma$, and 
+sample Gaussians $p(z'_{i} | z_{i}, \sigma) = \mathcal{N}(z_{i}, \sigma)$ to 
+get "observed" redshifts $z'_{i}$ defining Gaussian likelihoods $p(\vec{d}_{i} 
+| z'_{i}, \sigma) = \mathcal{N}(z'_{i}, \sigma)$.
+	\item \label{itm:interim} Choose a parametrization and the parameters 
+$\vec{\phi}^{*}$ of the interim prior.
+	\item \label{itm:posterior} Create interim posteriors $p(z_{i} | 
+\vec{d}_{i}, \vec{\phi}^{*}) = p(\vec{d}_{i} | z_{i})\ p(z_{i} | 
+\vec{\phi}^{*})$ in this parametrization.
 \end{enumerate}
 
-This method for deriving mock data is referred to as the fiducial case, and 
-variations on it will refer directly to the steps that are altered.
+In all of the following validation tests, we use $10$ bins and $N = 10,000$ 
+galaxies.  In Secs. \ref{sec:truth} and \ref{sec:likelihoods}, we use a flat 
+interim prior.  This method for deriving mock data is referred to as the 
+fiducial case, and variations on it will refer directly to the steps that are 
+altered.
 
@@
 Existing $n(z)$ estimators are systematically smoother than the true $n(z)$.  
-Here we show that the traditional estimators of $n(z)$ cannot recover a highly 
-featured $n(z)$.  The implication of this failure is quite serious; the 
-consistently smooth, unimodal $n(z)$ estimates appearing in the literature 
-could result from much more featured true redshift distributions, and there 
-would be no way to catch this error without using a fully probabilistic method.
+Here we show that the traditional estimators perform better when the true 
+$n(z)$ is weakly featured than when the true $n(z)$ is strongly featured by 
+experimenting with Step \ref{itm:true} of Sec. \ref{sec:mockdata}.  The 
+implication of this issue is quite serious; the consistently smooth, unimodal 
+$n(z)$ estimates appearing in the literature could result from much more 
+featured true redshift distributions, and there would be no way to catch this 
+error without using a fully probabilistic method.
 
@@
 \includegraphics[width=0.5\textwidth]{fig/fiducial/estimators.png}		
-		\caption{[plot of a highly featured, multimodal true $n(z)$]}
+		\caption{[This plot doesn't really show what I want because the 
+intrinsic scatter is too low!  I used $\sigma=0.03$ because that's what has 
+been quoted as what LSST, etc. needs, but that's not what comes out of 
+photo-$z$ estimation methods before they impose aggressive cuts.]  Top panel: 
+The traditional MAP reduction of a photo-$z$ PDF against the true redshifts 
+with a few rescaled photo-$z$ interim posteriors are overplotted in solid 
+lines, with a dotted line indicating zero probability.  Bottom panel: Various 
+estimators of $\ln[n(z)]$, the interim prior, and the true $\ln[n(z)]$ as a 
+continuous function and under a binned parametrization.}
 		\label{fig:featured}
@@
 likelihoods in an effort to emulate known systematics in photo-$z$ estimation.  
+These tests vary Step \ref{itm:scatter} of Sec. \ref{sec:mockdata}.
 
@@
 	\begin{center}
-		%\includegraphics[width=0.5\textwidth]{pgm.png}
-		\caption{[plot of $z_{phot}$ vs. $z_{spec}$ with varying 
-intrinsic scatter]}
+
+\includegraphics[width=0.5\textwidth]{fig/high_scatter/scatter.png}\\
+
+\includegraphics[width=0.5\textwidth]{fig/high_scatter/estimators.png}		
+		\caption{[This would be a lot more compelling with more 
+galaxies.  Also, the weird edge effects in the top panel are real, because the 
+point estimator is the MAP, not the center of the Gaussian likelihood, and 
+there's no requirement that the mean of the likelihood be within the true 
+redshift range.]  Top panel: The traditional MAP reduction of a photo-$z$ PDF 
+against the true redshifts with a few rescaled photo-$z$ interim posteriors are 
+overplotted in solid lines, with a dotted line indicating zero probability.  
+Bottom panel: Various estimators of $\ln[n(z)]$, the interim prior, and the 
+true $\ln[n(z)]$ as a continuous function and under a binned parametrization.}
 		\label{fig:intscat}
@@
 	\begin{center}
-		%\includegraphics[width=0.5\textwidth]{pgm.png}
-		\caption{[plot of $z_{phot}$ vs. $z_{spec}$ with template-like 
-catastrophic outliers]}
+
+\includegraphics[width=0.5\textwidth]{fig/template_outliers/scatter.png}\\
+
+\includegraphics[width=0.5\textwidth]{fig/template_outliers/estimators.png}
+
+		\caption{[]  Top panel: The traditional MAP reduction of a 
+photo-$z$ PDF against the true redshifts with a few rescaled photo-$z$ interim 
+posteriors are overplotted in solid lines, with a dotted line indicating zero 
+probability.  Bottom panel: Various estimators of $\ln[n(z)]$, the interim 
+prior, and the true $\ln[n(z)]$ as a continuous function and under a binned 
+parametrization.}
 		\label{fig:tempcatout}
@@
 the true $n(z)$, \chippr is not affected by the choice of the interim prior so 
-long as it has nontrivial coverage in the space of redshift.
+long as it has nontrivial coverage in the space of redshift.  These tests 
+modify Step \ref{itm:interim} of Sec. \ref{sec:mockdata}.
 

### 2017-06-15 chippr/catalog.py

@@ -168,12 +168,15 @@ class catalog(object):
             true_lfs = [gauss(self.true_samps[n], self.params['constant_sigma']**2) for n in self.samp_range]
         if self.params['outlier_fraction'] > 0.:
             outlier_lf = gauss(self.params['outlier_mean'], self.params['outlier_sigma']**2)
-        obs_samps = np.zeros(self.n_items)
+
+        obs_samps = np.zeros(self.n_items) - 1.
         for n in self.samp_range:
-            if np.random.uniform() < self.params['outlier_fraction']:
-                obs_samps[n] = outlier_lf.sample_one()
-            else:
-                obs_samps[n] = true_lfs[n].sample_one()
+            while obs_samps[n] < self.params['bin_min'] or obs_samps[n] > self.params['bin_max']:
+                if np.random.uniform() < self.params['outlier_fraction']:
+                    obs_samps[n] = outlier_lf.sample_one()
+                else:
+                    obs_samps[n] = true_lfs[n].sample_one()
+
         return obs_samps
 
     def evaluate_lfs(self):

### 2017-06-15 chippr/log_z_dens_plots.py

@@ -304,7 +304,7 @@ def plot_estimators(info, plot_dir):
         pu.plot_step(sps_log, info['bin_ends'], info['estimators']['log_mexp_nz'], w=w_exp, s=s_exp, a=a_exp, c=c_exp, d=d_exp, l=l_exp+lnz)
 
     sps_log.legend(fontsize='x-small', loc='upper left')
-    f.savefig(os.path.join(plot_dir, 'all_estimators.png'), bbox_inches='tight', pad_inches = 0, dpi=d.dpi)
+    f.savefig(os.path.join(plot_dir, 'estimators.png'), bbox_inches='tight', pad_inches = 0, dpi=d.dpi)
 
     return
 

### 2017-06-15 research/scripts/multi_sim_script.py

@@ -48,6 +48,7 @@ def make_catalog(given_key):
     """
     test_info = all_tests[given_key]
     test_name = test_info['name']
+
     true_nz = test_info['true_nz']
     true_zs = test_info['true_zs']
 
@@ -98,9 +99,6 @@ if __name__ == "__main__":
             test_info['true_zs'] = true_zs
             all_tests[test_name] = test_info
 
-#     nps = mp.cpu_count()-1
-#     pool = mp.Pool(nps)
-#     pool.map(make_catalog, all_tests.keys())
-
-    for key in all_tests.keys():
-        make_catalog(key)
+    nps = mp.cpu_count()-1
+    pool = mp.Pool(nps)
+    pool.map(make_catalog, all_tests.keys())

### 2017-06-15 research/scripts/multi_inf_script.py
minor changes in scripts

@@ -140,6 +140,8 @@ def do_inference(given_key):
     print('MExp: '+str(np.dot(np.exp(nz_mexp), z_difs)))
     nz_mmle = nz.calculate_mmle(nz_stacked, no_data=params['no_data'], no_prior=params['no_prior'])
     print('MMLE: '+str(np.dot(np.exp(nz_mmle), z_difs)))
+
+    nz_stats = nz.compare()
     nz.plot_estimators()
     nz.write('nz.p')
 

### 2017-06-16 research/scripts/multi_sim_script.py
preparing for more test cases

@@ -1,12 +1,42 @@
-def make_true_nz(test_name):
+def check_extra_params(params):
+    """
+    Sets parameter values pertaining to true n(z)
+
+    Parameters
+    ----------
+    params: dict
+        dictionary containing key/value pairs for simulation
+
+    Returns
+    -------
+    params: dict
+        dictionary containing key/value pairs for simulation
+    """
+    if 'smooth_truth' not in params:
+        params['smooth_truth'] = 0
+    else:
+        params['smooth_truth'] = int(params['smooth_truth'][0])
+
+    if 'interim_prior' not in params:
+        params['interim_prior'] = 'flat'
+    else:
+        params['interim_prior'] = str(params['interim_prior'][0])
+    if 'n_galaxies' not in params:
+        params['n_galaxies'] = 10**4
+    else:
+        params['n_galaxies'] = 10**int(params['n_galaxies'][0])
+
+    return(params)
+
+def make_true(given_key):
     """
     Function to create true redshift distribution to be shared among several
     test cases
 
     Parameters
     ----------
-    test_name: string
-        name used to look up parameters for making true_nz
+    given_key: string
+        name of test case for which true n(z) is to be made
 
     Returns
     -------
@@ -18,24 +48,81 @@ def make_true_nz(test_name):
     test_name is currently ignored but will soon be used to load parameters for
     making true_nz instead of hardcoded values.
     """
-    true_amps = np.array([0.20, 0.35, 0.55])
-    true_means = np.array([0.5, 0.2, 0.75])
-    true_sigmas = np.array([0.4, 0.2, 0.1])
-    true_nz = chippr.gmix(true_amps, true_means, true_sigmas, limits=(0., 1.))
-    return(true_nz)
+    test_info = all_tests[given_key]
 
-def make_true_zs(true_nz):
+    if test_info['params']['smooth_truth'] == 1:
+        bin_mids = (test_info['bin_ends'][1:] + test_info['bin_ends'][:-1]) / 2.
+        bin_difs = test_info['bin_ends'][1:] - test_info['bin_ends'][:-1]
+        true_amps = np.array([0.150,0.822,1.837,2.815,3.909,
+                              5.116,6.065,6.477,6.834,7.304,
+                              7.068,6.771,6.587,6.089,5.165,
+                              4.729,4.228,3.664,3.078,2.604,
+                              2.130,1.683,1.348,0.977,0.703,
[…]

### 2017-06-16 chippr/catalog.py
periodic backup

@@ -171,11 +171,11 @@ class catalog(object):
 
         obs_samps = np.zeros(self.n_items) - 1.
         for n in self.samp_range:
-            while obs_samps[n] < self.params['bin_min'] or obs_samps[n] > self.params['bin_max']:
-                if np.random.uniform() < self.params['outlier_fraction']:
-                    obs_samps[n] = outlier_lf.sample_one()
-                else:
-                    obs_samps[n] = true_lfs[n].sample_one()
+            # while obs_samps[n] < self.params['bin_min'] or obs_samps[n] > self.params['bin_max']:
+            if np.random.uniform() < self.params['outlier_fraction']:
+                obs_samps[n] = outlier_lf.sample_one()
+            else:
+                obs_samps[n] = true_lfs[n].sample_one()
 
         return obs_samps
 

### 2017-06-16 chippr/log_z_dens.py

@@ -77,6 +77,7 @@ class log_z_dens(object):
                 self.fine_nz.extend(fine_n)
                 coarse_nz = np.sum(fine_n) * fine_dz
                 self.tru_nz[b] += coarse_nz
+            self.tru_nz /= np.dot(self.tru_nz, self.bin_difs)
             self.log_tru_nz = u.safe_log(self.tru_nz)
             self.info['log_tru_nz'] = self.log_tru_nz
             self.info['truth']['z_grid'] = np.array(self.fine_zs)

### 2017-06-16 chippr/log_z_dens.py
changes to plotting

@@ -499,6 +499,8 @@ class log_z_dens(object):
             print('The following quantities were read from '+read_loc+' in the '+style+' format:')
             for key in self.info:
                 print(key)
+            if 'estimators' in self.info:
+                print(self.info['estimators'].keys())
         return self.info
 
     def write(self, write_loc, style='pickle', vb=True):

### 2017-06-16 chippr/log_z_dens_plots.py

@@ -5,6 +5,7 @@ import scipy as sp
 import matplotlib as mpl
 mpl.use('PS')
 import matplotlib.pyplot as plt
+from matplotlib import gridspec
 
 import chippr
 from chippr import defaults as d
@@ -16,15 +17,15 @@ from chippr import stat_utils as s
 
 lnz, nz = r'$\ln[n(z)]$', r'$n(z)$'
 
-s_tru, w_tru, a_tru, c_tru, d_tru, l_tru = '-', 0.5, 1., 'k', [(0, (1, 1))], 'True '
-s_tbp, w_tbp, a_tbp, c_tbp, d_tbp, l_tbp = '-', 0.5, 0.75, 'k', [(0, (1, 1))], 'Binned True '
-s_int, w_int, a_int, c_int, d_int, l_int = '-', 0.5, 0.5, 'k', [(0, (1, 1))], 'Interim '
-s_stk, w_stk, a_stk, c_stk, d_stk, l_stk = '--', 1.5, 0.75, pu.colors[4], [(0, (7.5, 7.5))], 'Stacked '#'g'
-s_map, w_map, a_map, c_map, d_map, l_map = '--', 1., 0.75, pu.colors[3], [(0, (5., 5.))], 'MMAP '#'r'
-s_exp, w_exp, a_exp, c_exp, d_exp, l_exp = '--', 0.5, 0.75, pu.colors[2], [(0, (2.5, 2.5))], 'MExp '#'r'
-s_mle, w_mle, a_mle, c_mle, d_mle, l_mle = '-', 1.5, 0.75, pu.colors[1], [(0, (1, 1))], 'MMLE '#'b'
-s_smp, w_smp, a_smp, c_smp, d_smp, l_smp = '-', 1., 1., 'k', [(0, (1, 1))], 'Sampled '
-s_bfe, w_bfe, a_bfe, c_bfe, d_bfe, l_bfe = '-', 2.5, 0.75, pu.colors[0], [(0, (1, 1))], 'Mean of\n Samples '#'b'
+s_tru, w_tru, a_tru, c_tru, d_tru, l_tru = '-', 2., 1., 'k', [(0, (1, 1))], 'True '
+s_tbp, w_tbp, a_tbp, c_tbp, d_tbp, l_tbp = ':', 2., 0.75, 'k', [(0, (1, 1))], 'Binned True '
[…]

### 2017-06-16 research/scripts/multi_replot_script.py

@@ -1,0 +1,113 @@
+def make_true_nz(test_name):
+    """
+    Function to create true redshift distribution to be shared among several test cases
+
+    Parameters
+    ----------
+    test_name: string
+        name used to look up parameters for making true_nz
+
+    Returns
+    -------
+    true_nz: chippr.gmix object
+        gaussian mixture probability distribution
+
+    Notes
+    -----
+    test_name is currently ignored but will soon be used to load parameters for making true_nz instead of hardcoded values.
+    """
+    true_amps = np.array([0.20, 0.35, 0.55])
+    true_means = np.array([0.5, 0.2, 0.75])
+    true_sigmas = np.array([0.4, 0.2, 0.1])
+
+    true_nz = chippr.gmix(true_amps, true_means, true_sigmas, limits=(0., 1.))
+
+    return(true_nz)
+
+def set_up_prior(data):
+    """
+    Function to create prior distribution from data
+
+    Parameters
+    ----------
+    data: dict
+        catalog dictionary containing bin endpoints, log interim prior, and log interim posteriors
+
+    Returns
+    -------
+    prior: chippr.mvn object
+        prior distribution as multivariate normal
+    """
+    zs = data['bin_ends']
+    log_nz_intp = np.exp(data['log_interim_prior'])
+    log_z_posts = np.exp(data['log_interim_posteriors'])
+
+    z_difs = zs[1:]-zs[:-1]
+    z_mids = (zs[1:]+zs[:-1])/2.
+    n_bins = len(z_mids)
+
+    prior_var = np.eye(n_bins)
+    for k in range(n_bins):
+        prior_var[k] = 1. * np.exp(-0.5 * (z_mids[k] - z_mids) ** 2 / 0.05 ** 2)
+    prior_mean = log_nz_intp
+    prior = mvn(prior_mean, prior_var)
+    return prior
+
+def just_plot(given_key):
+    """
+    Function to make plot from a previously generate log_z_dens object
+
+    Parameters
+    ----------
+    given_key: string
+        name of test case to be run
+    """
+    test_info = all_tests[given_key]
+    test_name = test_info['name']
+    true_nz = test_info['truth']
+
+    test_name = test_name[:-1]
+    param_file_name = test_name + '.txt'
+
+    test_dir = os.path.join(result_dir, test_name)
+    simulated_posteriors = catalog(params=param_file_name, loc=test_dir)
+    saved_location = 'data'
[…]

### 2017-06-16 chippr/log_z_dens_plots.py
fiddled with legend to no avail

@@ -20,12 +20,12 @@ lnz, nz = r'$\ln[n(z)]$', r'$n(z)$'
 s_tru, w_tru, a_tru, c_tru, d_tru, l_tru = '-', 2., 1., 'k', [(0, (1, 1))], 'True '
 s_tbp, w_tbp, a_tbp, c_tbp, d_tbp, l_tbp = ':', 2., 0.75, 'k', [(0, (1, 1))], 'Binned True '
 s_int, w_int, a_int, c_int, d_int, l_int = '-', 2., 0.5, 'k', [(0, (1, 1))], 'Interim '
-s_stk, w_stk, a_stk, c_stk, d_stk, l_stk = '--', 2.5, 0.75, 'g', [(0, (7.5, 7.5))], 'Stacked '
-s_map, w_map, a_map, c_map, d_map, l_map = '--', 2., 0.75, 'r', [(0, (5., 5.))], 'MMAP '
-s_exp, w_exp, a_exp, c_exp, d_exp, l_exp = '--', 2., 0.75, 'r', [(0, (2.5, 2.5))], 'MExp '
-s_mle, w_mle, a_mle, c_mle, d_mle, l_mle = '-', 3., 0.75, 'b', [(0, (1, 1))], 'MMLE '
+s_stk, w_stk, a_stk, c_stk, d_stk, l_stk = '--', 2.5, 0.75, 'g', [(0, (3, 3))], 'Stacked '
+s_map, w_map, a_map, c_map, d_map, l_map = '--', 2., 0.5, 'r', [(0, (4, 4))], 'MMAP '
+s_exp, w_exp, a_exp, c_exp, d_exp, l_exp = '--', 2., 0.5, 'm', [(0, (5, 5))], 'MExp '
+s_mle, w_mle, a_mle, c_mle, d_mle, l_mle = '-', 3., 1., 'c', [(0, (2, 2))], 'MMLE '
+s_bfe, w_bfe, a_bfe, c_bfe, d_bfe, l_bfe = '-', 3.5, 1., 'b', [(0, (1, 1))], 'Mean of\n Samples '
 s_smp, w_smp, a_smp, c_smp, d_smp, l_smp = '-', 2., 1., 'k', [(0, (1, 1))], 'Sampled '
-s_bfe, w_bfe, a_bfe, c_bfe, d_bfe, l_bfe = '-', 3.5, 0.75, 'b', [(0, (1, 1))], 'Mean of\n Samples '
 
[…]

### 2017-06-22 research/paper/draft-editing.tex
expanding cosmological constraint metric discussion

--- 2017-06-15 research/paper/draft-editing.tex
@@
 The fundamental assumption underlying the concept of photo-$z$ estimation is that each galaxy $i$ has some observed photometric data $\vec{d}_{i}$ (fluxes, magnitudes, or colors) that is drawn from a function of its redshift $z_{i}$, which is a parameter in this model.
+This function constitutes a forward model for the observations.
 The redshifts $\{z_{i}\}$ for all galaxies in a survey are random draws from the redshift distribution, whose parameters  $\vec{\phi}$ under a chosen functional form are called \textit{hyperparameters} because they are shared among all $N$ galaxies in the survey.
@@
 The DAG of Fig.
-\ref{fig:pgm} translates directly into a mathematical expression for the posterior distribution $p(\vec{\phi} | \{\vec{d}_{i}\})$ of the hyperparameters given the entire dataset given the set of interim posteriors $\{p(z_{i} | \vec{d}_{i}, \vec{\phi}^{*})\}$ by way of the following derivation.
+\ref{fig:pgm} translates directly into a mathematical expression for the posterior distribution $p(\vec{\phi} | \{\vec{d}_{i}\})$ of the hyperparameters given the entire set of interim posteriors $\{p(z_{i} | \vec{d}_{i}, \vec{\phi}^{*})\}$ by way of the following derivation.
 From this point forward, we will use log probabilities.
@@
 \end{align}
-where $p(\vec{\phi})$ is the textit{hyperprior} probability distribution over possible values of the hyperparameters.
+where $p(\vec{\phi})$ is the \textit{hyperprior} probability distribution over possible values of the hyperparameters.
 The hyperprior is a choice made by those performing the inference and is not to be confused with the interim prior parameters that influence the photo-$z$ PDF data product.
@@
 \begin{align}
-% \log p(\{\vec{d}_{i}\} | \vec{\phi}) &= \log\[\int\ \exp\[\log\[p(\{\vec{d}_{i}\} | \{z_{i}\})\] + \log\[p(\{z_{i}\} | \vec{\phi})\]\]\ d\{z_{i}\}\].
+\log[p(\{\vec{d}_{i}\} | \vec{\phi})] &= \log[\int\exp[\log[p(\{\vec{d}_{i}\} | \{z_{i}\})]\\
+& + \log[p(\{z_{i}\} | \vec{\phi})]]\ d\{z_{i}\}].
 \end{align}
@@
 
-Because the mock data is associated with true values of the cosmological parameters, we may compare the quality of cosmological constraints.
+Because the realistic mock data of Sec.
+\ref{sec:buzzard} is associated with true values of the cosmological parameters, we may compare the quality of cosmological constraints under different estimators of $n(z)$, creating a figure like Fig.
+\ref{eq:scgsr}.
+We perform this analysis using a forecasting code that takes $n(z)$ and produces projected error ellipses in the space of cosmological parameters.
 
-\textbf{Check Dark Energy Task Force metrics on error ellipses, multi-dimensional KLD, centroid offset, etc.}
+\begin{figure}
+	\begin{center}
+		\includegraphics[width=0.5\textwidth]{cosmo_constraints.png}
+		\caption{[I'd like to be able to make a plot like this with the different $n(z)$ estimators produced by \texttt{chippr} from the same set of photo-$z$ PDFs.
+Then I can calculate some metrics of the accuracy and precision of the error distributions relative to the true values of the cosmological parameters that produced the mock data.]}
+		\label{fig:scgsr}
+	\end{center}
+\end{figure}
+
+% \textbf{Check Dark Energy Task Force metrics on error ellipses, multi-dimensional KLD, centroid offset, etc.}
 

### 2017-06-22 research/paper/draft.tex
expanding cosmological constraint metric discussion

--- 2017-06-15 research/paper/draft.tex
@@
 magnitudes, or colors) that is drawn from a function of its redshift $z_{i}$, 
-which is a parameter in this model.  The redshifts $\{z_{i}\}$ for all galaxies 
-in a survey are random draws from the redshift distribution, whose parameters  
-$\vec{\phi}$ under a chosen functional form are called \textit{hyperparameters} 
-because they are shared among all $N$ galaxies in the survey.  Fig. 
-\ref{fig:pgm} illustrates these relationships.  PGMs of this structure are 
+which is a parameter in this model.  This function constitutes a forward model 
+for the observations.  The redshifts $\{z_{i}\}$ for all galaxies in a survey 
+are random draws from the redshift distribution, whose parameters  $\vec{\phi}$ 
+under a chosen functional form are called \textit{hyperparameters} because they 
+are shared among all $N$ galaxies in the survey.  Fig. \ref{fig:pgm} 
+illustrates these relationships.  PGMs of this structure are 
 \textit{hierarchical} in that, though the data are only directly influenced by 
@@
 expression for the posterior distribution $p(\vec{\phi} | \{\vec{d}_{i}\})$ of 
-the hyperparameters given the entire dataset given the set of interim 
-posteriors $\{p(z_{i} | \vec{d}_{i}, \vec{\phi}^{*})\}$ by way of the following 
-derivation.  From this point forward, we will use log probabilities.
+the hyperparameters given the entire set of interim posteriors $\{p(z_{i} | 
+\vec{d}_{i}, \vec{\phi}^{*})\}$ by way of the following derivation.  From this 
+point forward, we will use log probabilities.
 
@@
 \end{align}
-where $p(\vec{\phi})$ is the textit{hyperprior} probability distribution over 
+where $p(\vec{\phi})$ is the \textit{hyperprior} probability distribution over 
 possible values of the hyperparameters.  The hyperprior is a choice made by 
@@
 \begin{align}
+\log[p(\{\vec{d}_{i}\} | \vec{\phi})] &= \log[\int\exp[\log[p(\{\vec{d}_{i}\} | 
+\{z_{i}\})]\\
+& + \log[p(\{z_{i}\} | \vec{\phi})]]\ d\{z_{i}\}].
 \end{align}
@@
 
-Because the mock data is associated with true values of the cosmological 
-parameters, we may compare the quality of cosmological constraints.
+Because the realistic mock data of Sec. \ref{sec:buzzard} is associated with 
+true values of the cosmological parameters, we may compare the quality of 
+cosmological constraints under different estimators of $n(z)$, creating a 
+figure like Fig. \ref{eq:scgsr}.  We perform this analysis using a forecasting 
+code that takes $n(z)$ and produces projected error ellipses in the space of 
+cosmological parameters.
 
-\textbf{Check Dark Energy Task Force metrics on error ellipses, 
-multi-dimensional KLD, centroid offset, etc.}
+\begin{figure}
+	\begin{center}
+		\includegraphics[width=0.5\textwidth]{cosmo_constraints.png}
+		\caption{[I'd like to be able to make a plot like this with the 
+different $n(z)$ estimators produced by \texttt{chippr} from the same set of 
+photo-$z$ PDFs.  Then I can calculate some metrics of the accuracy and 
+precision of the error distributions relative to the true values of the 
+cosmological parameters that produced the mock data.]}
+		\label{fig:scgsr}
+	\end{center}
+\end{figure}
 

### 2017-06-23 research/paper/draft-editing.tex

--- 2017-06-15 research/paper/draft-editing.tex
@@
 \documentclass[iop]{emulateapj}
-%\documentclass[preprint]{aastex}
-%\documentclass[12pt, onecolumn]{emulateapj}
-%\documentstyle[aas2pp4,natbib209]{article}
+% \documentclass[preprint]{aastex}
+% \documentclass[12pt, onecolumn]{emulateapj}
+% \documentstyle[aas2pp4,natbib209]{article}
 
@@
 	\begin{center}
-% 		\includegraphics[width=0.5\textwidth]{pgm.png}
-		\caption{\textbf{Replace \texttt{daft} version with more polished \texttt{tikz} version.}}
-%		\caption{This directed acyclic graph corresponds to a PGM for a hierarchical inference of $p(\vec{\theta}|\{\vec{d}_{i}\})$.
+ 		\includegraphics[width=0.5\textwidth]{fig/pgm.png}
+		\caption{This directed acyclic graph corresponds to a PGM for a hierarchical inference of $p(\vec{\theta}|\{\vec{d}_{i}\})$.
 In this graph, all random variables are shown in circles, with observed random variables shown in shaded circles.
@@
 \begin{enumerate}
-	\item Choose a true $n(z)$ that is a continuous function with known parameters $\vec{\theta}$.
-	\item Sample the true $n(z)$ to create a catalog of $N$ true redshifts $z_{i}$.
-	\item Choose a true intrinsic scatter $\sigma$, and sample Gaussians $p(z'_{i} | z_{i}, \sigma) = \mathcal{N}(z_{i}, \sigma)$ to get "observed" redshifts $z'_{i}$ defining Gaussian likelihoods $p(\vec{d}_{i} | z'_{i}, \sigma) = \mathcal{N}(z'_{i}, \sigma)$.
-	\item Choose a parametrization and the parameters $\vec{\phi}^{*}$ of the interim prior.
-	\item Create interim posteriors $p(z_{i} | \vec{d}_{i}, \vec{\phi}^{*}) = p(\vec{d}_{i} | z_{i})\ p(z_{i} | \vec{\phi}^{*})$
+	\item \label{itm:true} Choose a true $n(z)$ that is a continuous function with known parameters $\vec{\theta}$.
+	\item \label{itm:sample} Sample the true $n(z)$ to create a catalog of $N$ true redshifts $z_{i}$.
+	\item \label{itm:scatter} Choose a true intrinsic scatter $\sigma$, and sample Gaussians $p(z'_{i} | z_{i}, \sigma) = \mathcal{N}(z_{i}, \sigma)$ to get "observed" redshifts $z'_{i}$ defining Gaussian likelihoods $p(\vec{d}_{i} | z'_{i}, \sigma) = \mathcal{N}(z'_{i}, \sigma)$.
+	\item \label{itm:interim} Choose a parametrization and the parameters $\vec{\phi}^{*}$ of the interim prior.
+	\item \label{itm:posterior} Create interim posteriors $p(z_{i} | \vec{d}_{i}, \vec{\phi}^{*}) = p(\vec{d}_{i} | z_{i})\ p(z_{i} | \vec{\phi}^{*})$ in this parametrization.
 \end{enumerate}
 
+In all of the following validation tests, we use $10$ bins and $N = 10,000$ galaxies.
+In Secs.
+\ref{sec:truth} and \ref{sec:likelihoods}, we use a flat interim prior.
 This method for deriving mock data is referred to as the fiducial case, and variations on it will refer directly to the steps that are altered.
@@
 Existing $n(z)$ estimators are systematically smoother than the true $n(z)$.
-Here we show that the traditional estimators of $n(z)$ cannot recover a highly featured $n(z)$.
-The implication of this failure is quite serious; the consistently smooth, unimodal $n(z)$ estimates appearing in the literature could result from much more featured true redshift distributions, and there would be no way to catch this error without using a fully probabilistic method.
+Here we show that the traditional estimators perform better when the true $n(z)$ is weakly featured than when the true $n(z)$ is strongly featured by experimenting with Step \ref{itm:true} of Sec.
+\ref{sec:mockdata}.
+The implication of this issue is quite serious; the consistently smooth, unimodal $n(z)$ estimates appearing in the literature could result from much more featured true redshift distributions, and there would be no way to catch this error without using a fully probabilistic method.
 
@@
 
-In this test, we choose a weakly featured true $n(z)$ of Fig. \ref{fig:smooth} with a smooth, unimodal shape.
+In this test, we choose a weakly featured true $n(z)$ of Fig.
+\ref{fig:smooth} with a smooth, unimodal shape, based on the interim prior used for the SDSS DR8 photo-$z$ PDFs.
+[check and include citation!]
 
@@
 	\begin{center}
-		%\includegraphics[width=0.5\textwidth]{pgm.png}
-		\caption{[plot of a weakly featured true $n(z)$, probably SDSS interim prior]}
+		\includegraphics[width=0.5\textwidth]{fig/smooth_truth/scatter.png}\\
+		\includegraphics[width=0.5\textwidth]{fig/smooth_truth/estimators.png}	
+		\caption{All estimators perform well when the true $n(z)$ is well behaved, exhibiting significant deviation only when $n(z)$ is very small, as the sample size of true redshifts in that range will be small.
+Top panel: The traditional MAP reduction of a photo-$z$ PDF against the true redshifts with a few rescaled photo-$z$ interim posteriors are overplotted in solid lines, with a dotted line indicating zero probability.
+Bottom panel: Various estimators of $\ln[n(z)]$, the interim prior, and the true $\ln[n(z)]$ as a continuous function and under a binned parametrization.}
 		\label{fig:featured}
@@
 	\begin{center}
-		%\includegraphics[width=0.5\textwidth]{pgm.png}
-		\caption{[plot of a highly featured, multimodal true $n(z)$]}
+		\includegraphics[width=0.5\textwidth]{fig/fiducial/scatter.png}\\
+		\includegraphics[width=0.5\textwidth]{fig/fiducial/estimators.png}		
+		\caption{[This plot doesn't really show what I want because the intrinsic scatter is too low!
+I used $\sigma=0.03$ because that's what has been quoted as what LSST, etc.
+needs, but that's not what comes out of photo-$z$ estimation methods before they impose aggressive cuts\dots]  Top panel: The traditional MAP reduction of a photo-$z$ PDF against the true redshifts with a few rescaled photo-$z$ interim posteriors are overplotted in solid lines, with a dotted line indicating zero probability.
+Bottom panel: Various estimators of $\ln[n(z)]$, the interim prior, and the true $\ln[n(z)]$ as a continuous function and under a binned parametrization.}
 		\label{fig:featured}
@@
 In the following test cases, we vary the properties of the mock photo-$z$ likelihoods in an effort to emulate known systematics in photo-$z$ estimation.
+These tests vary Step \ref{itm:scatter} of Sec.
+\ref{sec:mockdata}.
 
@@
 \ref{fig:intscat}.
+To emulate intrinsic scatter, we modify the fiducial case to simply broaden the single Gaussian component of the likelihood.
+To enforce self-consistency, the mean is a random variable drawn from a Gaussian distribution with the newly increased variance.
 
@@
 	\begin{center}
-		%\includegraphics[width=0.5\textwidth]{pgm.png}
-		\caption{[plot of $z_{phot}$ vs. $z_{spec}$ with varying intrinsic scatter]}
+		\includegraphics[width=0.5\textwidth]{fig/high_scatter/scatter.png}\\
+		\includegraphics[width=0.5\textwidth]{fig/high_scatter/estimators.png}		
+		\caption{As the intrinsic scatter increases, the discrepancy between estimators increases.
+In particular, the stacked estimator and marginalized point estimators predict $\ln[n(z)]$ to be smoother than the truth, while the  [This would be a lot more compelling with more galaxies.
+Also, the weird edge effects in the top panel are real, because the point estimator is the MAP, not the center of the Gaussian likelihood, and there's no requirement that the mean of the likelihood be within the true redshift range.]  Top panel: The traditional MAP reduction of a photo-$z$ PDF against the true redshifts with a few rescaled photo-$z$ interim posteriors are overplotted in solid lines, with a dotted line indicating zero probability.
+Bottom panel: Various estimators of $\ln[n(z)]$, the interim prior, and the true $\ln[n(z)]$ as a continuous function and under a binned parametrization.}
 		\label{fig:intscat}
@@
 
-To emulate intrinsic scatter, we modify the fiducial case to simply broaden the single Gaussian component of the likelihood.
-To enforce self-consistency, the mean is a random variable drawn from a Gaussian distribution with the newly increased variance.
+As the intrinsic scatter increases, the discrepancy between estimators increases.
 
@@
 	\begin{center}
-		%\includegraphics[width=0.5\textwidth]{pgm.png}
-		\caption{[plot of $z_{phot}$ vs. $z_{spec}$ with template-like catastrophic outliers]}
+		\includegraphics[width=0.5\textwidth]{fig/template_outliers/scatter.png}\\
+		\includegraphics[width=0.5\textwidth]{fig/template_outliers/estimators.png}		
+		\caption{Top panel: The traditional MAP reduction of a photo-$z$ PDF against the true redshifts with a few rescaled photo-$z$ interim posteriors are overplotted in solid lines, with a dotted line indicating zero probability.
+Bottom panel: Various estimators of $\ln[n(z)]$, the interim prior, and the true $\ln[n(z)]$ as a continuous function and under a binned parametrization.}
 		\label{fig:tempcatout}
@@
 	\begin{center}
-		%\includegraphics[width=0.5\textwidth]{pgm.png}
-		\caption{[plot of $z_{phot}$ vs. $z_{spec}$ with training-like catastrophic outliers]}
+		\includegraphics[width=0.5\textwidth]{fig/training_outliers/scatter.png}\\
+		\includegraphics[width=0.5\textwidth]{fig/training_outliers/estimators.png}	
+		\caption{[This has proven to be the most challenging test case to implement, and there's clearly still a bug in the function that makes the likelihoods.]}
 		\label{fig:traincatout}
@@
 We demonstrate here that regardless of the appropriateness of the interim prior as an approximation to the true $n(z)$, \chippr is not affected by the choice of the interim prior so long as it has nontrivial coverage in the space of redshift.
+These tests modify Step \ref{itm:interim} of Sec.
+\ref{sec:mockdata}.
 
@@
 	\begin{center}
-		%\includegraphics[width=0.5\textwidth]{pgm.png}
-		\caption{[plot of template-like (multimodal) interim prior]}
+		\includegraphics[width=0.5\textwidth]{fig/template_prior/scatter.png}\\
+		\includegraphics[width=0.5\textwidth]{fig/template_prior/estimators.png}	
+		\caption{[The case of a multimodal interim prior was a very compelling test in the previous version but somehow isn't anymore.]}
 		\label{fig:tempintpr}
@@
 \ref{fig:trainintpr} is an emulation of an interim prior corresponding to a training set biased in this way.
+We chose an interpolation of the interim prior used for the SDSS DR8 photo-$z$ PDFs.
 
@@
 	\begin{center}
-		%\includegraphics[width=0.5\textwidth]{pgm.png}
-		\caption{[plot of training-like (low-$z$ favoring) interim prior]}
+		\includegraphics[width=0.5\textwidth]{fig/training_prior/scatter.png}\\
+		\includegraphics[width=0.5\textwidth]{fig/training_prior/estimators.png}	
+		\caption{[The case of a low-$z$ favoring interim prior was a very compelling test in the previous version but somehow isn't anymore.]}
 		\label{fig:trainintpr}
@@
 	\begin{center}
-		%\includegraphics[width=0.5\textwidth]{pgm.png}
+		%\includegraphics[width=0.5\textwidth]{money.png}
 		\caption{[moneyplot of error ellipses resulting from different $n(z)$ estimators]}

### 2017-06-23 research/paper/draft.tex

--- 2017-06-15 research/paper/draft.tex
@@
 In this test, we choose a weakly featured true $n(z)$ of Fig. \ref{fig:smooth} 
-with a smooth, unimodal shape.
+with a smooth, unimodal shape, based on the interim prior used for the SDSS DR8 
+photo-$z$ PDFs.  [check and include citation!]
 
@@
 	\begin{center}
-		%\includegraphics[width=0.5\textwidth]{pgm.png}
-		\caption{[plot of a weakly featured true $n(z)$, probably SDSS 
-interim prior]}
+
+\includegraphics[width=0.5\textwidth]{fig/smooth_truth/scatter.png}\\
+
+\includegraphics[width=0.5\textwidth]{fig/smooth_truth/estimators.png}	
+		\caption{All estimators perform well when the true $n(z)$ is 
+well behaved, exhibiting significant deviation only when $n(z)$ is very small, 
+as the sample size of true redshifts in that range will be small.  Top panel: 
+The traditional MAP reduction of a photo-$z$ PDF against the true redshifts 
+with a few rescaled photo-$z$ interim posteriors are overplotted in solid 
+lines, with a dotted line indicating zero probability.  Bottom panel: Various 
+estimators of $\ln[n(z)]$, the interim prior, and the true $\ln[n(z)]$ as a 
+continuous function and under a binned parametrization.}
 		\label{fig:featured}
@@
 been quoted as what LSST, etc. needs, but that's not what comes out of 
-photo-$z$ estimation methods before they impose aggressive cuts.]  Top panel: 
-The traditional MAP reduction of a photo-$z$ PDF against the true redshifts 
-with a few rescaled photo-$z$ interim posteriors are overplotted in solid 
-lines, with a dotted line indicating zero probability.  Bottom panel: Various 
-estimators of $\ln[n(z)]$, the interim prior, and the true $\ln[n(z)]$ as a 
-continuous function and under a binned parametrization.}
+photo-$z$ estimation methods before they impose aggressive cuts\dots]  Top 
+panel: The traditional MAP reduction of a photo-$z$ PDF against the true 
+redshifts with a few rescaled photo-$z$ interim posteriors are overplotted in 
+solid lines, with a dotted line indicating zero probability.  Bottom panel: 
+Various estimators of $\ln[n(z)]$, the interim prior, and the true $\ln[n(z)]$ 
+as a continuous function and under a binned parametrization.}
 		\label{fig:featured}
@@
 varying extents with every existing photo-$z$ algorithm and illustrated in Fig. 
-\ref{fig:intscat}.
+\ref{fig:intscat}.  To emulate intrinsic scatter, we modify the fiducial case 
+to simply broaden the single Gaussian component of the likelihood.  To enforce 
+self-consistency, the mean is a random variable drawn from a Gaussian 
+distribution with the newly increased variance.
 
@@
 \includegraphics[width=0.5\textwidth]{fig/high_scatter/estimators.png}		
-		\caption{[This would be a lot more compelling with more 
-galaxies.  Also, the weird edge effects in the top panel are real, because the 
-point estimator is the MAP, not the center of the Gaussian likelihood, and 
-there's no requirement that the mean of the likelihood be within the true 
-redshift range.]  Top panel: The traditional MAP reduction of a photo-$z$ PDF 
-against the true redshifts with a few rescaled photo-$z$ interim posteriors are 
-overplotted in solid lines, with a dotted line indicating zero probability.  
-Bottom panel: Various estimators of $\ln[n(z)]$, the interim prior, and the 
-true $\ln[n(z)]$ as a continuous function and under a binned parametrization.}
+		\caption{As the intrinsic scatter increases, the discrepancy 
+between estimators increases.  In particular, the stacked estimator and 
+marginalized point estimators predict $\ln[n(z)]$ to be smoother than the 
+truth, while the  [This would be a lot more compelling with more galaxies.  
+Also, the weird edge effects in the top panel are real, because the point 
+estimator is the MAP, not the center of the Gaussian likelihood, and there's no 
+requirement that the mean of the likelihood be within the true redshift range.] 
+ Top panel: The traditional MAP reduction of a photo-$z$ PDF against the true 
+redshifts with a few rescaled photo-$z$ interim posteriors are overplotted in 
+solid lines, with a dotted line indicating zero probability.  Bottom panel: 
+Various estimators of $\ln[n(z)]$, the interim prior, and the true $\ln[n(z)]$ 
+as a continuous function and under a binned parametrization.}
 		\label{fig:intscat}
@@
 
-To emulate intrinsic scatter, we modify the fiducial case to simply broaden the 
-single Gaussian component of the likelihood.  To enforce self-consistency, the 
-mean is a random variable drawn from a Gaussian distribution with the newly 
-increased variance.
+As the intrinsic scatter increases, the discrepancy between estimators 
+increases.
 
@@
 
-		\caption{[]  Top panel: The traditional MAP reduction of a 
+		\caption{Top panel: The traditional MAP reduction of a 
 photo-$z$ PDF against the true redshifts with a few rescaled photo-$z$ interim 
@@
 	\begin{center}
-		%\includegraphics[width=0.5\textwidth]{pgm.png}
-		\caption{[plot of $z_{phot}$ vs. $z_{spec}$ with training-like 
-catastrophic outliers]}
+
+\includegraphics[width=0.5\textwidth]{fig/training_outliers/scatter.png}\\
+
+\includegraphics[width=0.5\textwidth]{fig/training_outliers/estimators.png}	
+		\caption{[This has proven to be the most challenging test case 
+to implement, and there's clearly still a bug in the function that makes the 
+likelihoods.]}
 		\label{fig:traincatout}
@@
 	\begin{center}
-		%\includegraphics[width=0.5\textwidth]{pgm.png}
-		\caption{[plot of template-like (multimodal) interim prior]}
+
+\includegraphics[width=0.5\textwidth]{fig/template_prior/scatter.png}\\
+
+\includegraphics[width=0.5\textwidth]{fig/template_prior/estimators.png}	
+		\caption{[The case of a multimodal interim prior was a very 
+compelling test in the previous version but somehow isn't anymore.]}
 		\label{fig:tempintpr}
@@
 prior shown in Fig. \ref{fig:trainintpr} is an emulation of an interim prior 
-corresponding to a training set biased in this way.
+corresponding to a training set biased in this way.  We chose an interpolation 
+of the interim prior used for the SDSS DR8 photo-$z$ PDFs.
 
@@
 	\begin{center}
-		%\includegraphics[width=0.5\textwidth]{pgm.png}
-		\caption{[plot of training-like (low-$z$ favoring) interim 
-prior]}
+
+\includegraphics[width=0.5\textwidth]{fig/training_prior/scatter.png}\\
+
+\includegraphics[width=0.5\textwidth]{fig/training_prior/estimators.png}	
+		\caption{[The case of a low-$z$ favoring interim prior was a 
+very compelling test in the previous version but somehow isn't anymore.]}
 		\label{fig:trainintpr}
@@
 	\begin{center}
-		%\includegraphics[width=0.5\textwidth]{pgm.png}
+		%\includegraphics[width=0.5\textwidth]{money.png}
 		\caption{[moneyplot of error ellipses resulting from different 

### 2017-06-23 research/paper/draft-editing.tex

--- 2017-06-22 research/paper/draft-editing.tex
@@
 
+\begin{enumerate}
+	\item \label{itm:true} Choose a true $n(z)$ that is a continuous function with known parameters $\vec{\theta}$.
+	\item \label{itm:sample} Sample the true $n(z)$ to create a catalog of $N$ true redshifts $z_{i}$.
+	\item \label{itm:scatter} Choose a true intrinsic scatter $\sigma$, and sample Gaussians $p(z'_{i} | z_{i}, \sigma) = \mathcal{N}(z_{i}, \sigma)$ to get "observed" redshifts $z'_{i}$ defining Gaussian likelihoods $p(\vec{d}_{i} | z'_{i}, \sigma) = \mathcal{N}(z'_{i}, \sigma)$.
+	\item \label{itm:interim} Choose a parametrization and the parameters $\vec{\phi}^{*}$ of the interim prior.
+	\item \label{itm:posterior} Create interim posteriors $p(z_{i} | \vec{d}_{i}, \vec{\phi}^{*}) = p(\vec{d}_{i} | z_{i})\ p(z_{i} | \vec{\phi}^{*})$ in this parametrization.
+\end{enumerate}
+
+In all of the following validation tests, we use $10$ bins and $N = 10,000$ galaxies.
+In Secs.
+\ref{sec:truth} and \ref{sec:likelihoods}, we use a flat interim prior.
 This method for deriving mock data is referred to as the fiducial case, and variations on it will refer directly to the steps that are altered.
@@
 
+\subsection{Underlying $n(z)$ Effects}
+\label{sec:truth}
+
+Existing $n(z)$ estimators are systematically smoother than the true $n(z)$.
+Here we show that the traditional estimators perform better when the true $n(z)$ is weakly featured than when the true $n(z)$ is strongly featured by experimenting with Step \ref{itm:true} of Sec.
+\ref{sec:mockdata}.
+The implication of this issue is quite serious; the consistently smooth, unimodal $n(z)$ estimates appearing in the literature could result from much more featured true redshift distributions, and there would be no way to catch this error without using a fully probabilistic method.
+
+\subsubsection{Featureless $n(z)$}
+\label{sec:smooth}
+
+In this test, we choose a weakly featured true $n(z)$ of Fig.
+\ref{fig:smooth} with a smooth, unimodal shape, based on the interim prior used for the SDSS DR8 photo-$z$ PDFs.
+[check and include citation!]
+
+\begin{figure}
+	\begin{center}
+		\includegraphics[width=0.5\textwidth]{fig/smooth_truth/scatter.png}\\
+		\includegraphics[width=0.5\textwidth]{fig/smooth_truth/estimators.png}	
+		\caption{All estimators perform well when the true $n(z)$ is well behaved, exhibiting significant deviation only when $n(z)$ is very small, as the sample size of true redshifts in that range will be small.
+Top panel: The traditional MAP reduction of a photo-$z$ PDF against the true redshifts with a few rescaled photo-$z$ interim posteriors are overplotted in solid lines, with a dotted line indicating zero probability.
+Bottom panel: Various estimators of $\ln[n(z)]$, the interim prior, and the true $\ln[n(z)]$ as a continuous function and under a binned parametrization.}
+		\label{fig:featured}
+	\end{center}
+\end{figure}
+
+\subsubsection{Featured $n(z)$}
+\label{sec:featured}
+
+In this test, we choose the true $n(z)$ of Fig. \ref{fig:featured} with nontrivial structure.
+
+\begin{figure}
+	\begin{center}
+		\includegraphics[width=0.5\textwidth]{fig/fiducial/scatter.png}\\
+		\includegraphics[width=0.5\textwidth]{fig/fiducial/estimators.png}		
+		\caption{[This plot doesn't really show what I want because the intrinsic scatter is too low!
+I used $\sigma=0.03$ because that's what has been quoted as what LSST, etc.
+needs, but that's not what comes out of photo-$z$ estimation methods before they impose aggressive cuts\dots]  Top panel: The traditional MAP reduction of a photo-$z$ PDF against the true redshifts with a few rescaled photo-$z$ interim posteriors are overplotted in solid lines, with a dotted line indicating zero probability.
+Bottom panel: Various estimators of $\ln[n(z)]$, the interim prior, and the true $\ln[n(z)]$ as a continuous function and under a binned parametrization.}
+		\label{fig:featured}
+	\end{center}
+\end{figure}
+
+This featured true $n(z)$ will henceforth be referred to as the fiducial $n(z)$
+
 \subsection{Emulated Data Quality Effects}
@@
 	\begin{center}
-		%\includegraphics[width=0.5\textwidth]{pgm.png}
-		\caption{[plot of training-like (low-$z$ favoring) interim prior]}
+		\includegraphics[width=0.5\textwidth]{fig/training_prior/scatter.png}\\
+		\includegraphics[width=0.5\textwidth]{fig/training_prior/estimators.png}	
+		\caption{[The case of a low-$z$ favoring interim prior was a very compelling test in the previous version but somehow isn't anymore.]}
 		\label{fig:trainintpr}
-	\end{center}
-\end{figure}
-
-\subsection{Underlying $n(z)$ Effects}
-\label{sec:truth}
-
-Existing $n(z)$ estimators are systematically smoother than the true $n(z)$.
-Here we show that the traditional estimators of $n(z)$ cannot recover a highly featured $n(z)$.
-The implication of this failure is quite serious; the consistently smooth, unimodal $n(z)$ estimates appearing in the literature could result from much more featured true redshift distributions, and there would be no way to catch this error without using a fully probabilistic method.
-
-\subsubsection{Featured $n(z)$}
-\label{sec:featured}
-
-In this test, we choose the true $n(z)$ of Fig. \ref{fig:featured}, which has nontrivial structure.
-
-\begin{figure}
-	\begin{center}
-		%\includegraphics[width=0.5\textwidth]{pgm.png}
-		\caption{[plot of a highly featured (delta function or multimodal?) true $n(z)$]}
-		\label{fig:featured}
 	\end{center}
@@
 	\begin{center}
-		%\includegraphics[width=0.5\textwidth]{pgm.png}
+% 		\includegraphics[width=0.5\textwidth]{money.png}
 		\caption{[moneyplot of error ellipses resulting from different $n(z)$ estimators]}

### 2017-06-23 research/paper/draft.tex

--- 2017-06-22 research/paper/draft.tex
@@
 
-This method for deriving mock data is referred to as the fiducial case, and 
-variations on it will refer directly to the steps that are altered.
+\begin{enumerate}
+	\item \label{itm:true} Choose a true $n(z)$ that is a continuous 
+function with known parameters $\vec{\theta}$.
+	\item \label{itm:sample} Sample the true $n(z)$ to create a catalog of 
+$N$ true redshifts $z_{i}$.
+	\item \label{itm:scatter} Choose a true intrinsic scatter $\sigma$, and 
+sample Gaussians $p(z'_{i} | z_{i}, \sigma) = \mathcal{N}(z_{i}, \sigma)$ to 
+get "observed" redshifts $z'_{i}$ defining Gaussian likelihoods $p(\vec{d}_{i} 
+| z'_{i}, \sigma) = \mathcal{N}(z'_{i}, \sigma)$.
+	\item \label{itm:interim} Choose a parametrization and the parameters 
+$\vec{\phi}^{*}$ of the interim prior.
+	\item \label{itm:posterior} Create interim posteriors $p(z_{i} | 
+\vec{d}_{i}, \vec{\phi}^{*}) = p(\vec{d}_{i} | z_{i})\ p(z_{i} | 
+\vec{\phi}^{*})$ in this parametrization.
+\end{enumerate}
+
+In all of the following validation tests, we use $10$ bins and $N = 10,000$ 
+galaxies.  In Secs. \ref{sec:truth} and \ref{sec:likelihoods}, we use a flat 
+interim prior.  This method for deriving mock data is referred to as the 
+fiducial case, and variations on it will refer directly to the steps that are 
+altered.
 
@@
 
+\subsection{Underlying $n(z)$ Effects}
+\label{sec:truth}
+
+Existing $n(z)$ estimators are systematically smoother than the true $n(z)$.  
+Here we show that the traditional estimators perform better when the true 
+$n(z)$ is weakly featured than when the true $n(z)$ is strongly featured by 
+experimenting with Step \ref{itm:true} of Sec. \ref{sec:mockdata}.  The 
+implication of this issue is quite serious; the consistently smooth, unimodal 
+$n(z)$ estimates appearing in the literature could result from much more 
+featured true redshift distributions, and there would be no way to catch this 
+error without using a fully probabilistic method.
+
+\subsubsection{Featureless $n(z)$}
+\label{sec:smooth}
+
+In this test, we choose a weakly featured true $n(z)$ of Fig. \ref{fig:smooth} 
+with a smooth, unimodal shape, based on the interim prior used for the SDSS DR8 
+photo-$z$ PDFs.  [check and include citation!]
+
+\begin{figure}
+	\begin{center}
+
+\includegraphics[width=0.5\textwidth]{fig/smooth_truth/scatter.png}\\
+
+\includegraphics[width=0.5\textwidth]{fig/smooth_truth/estimators.png}	
+		\caption{All estimators perform well when the true $n(z)$ is 
+well behaved, exhibiting significant deviation only when $n(z)$ is very small, 
+as the sample size of true redshifts in that range will be small.  Top panel: 
+The traditional MAP reduction of a photo-$z$ PDF against the true redshifts 
+with a few rescaled photo-$z$ interim posteriors are overplotted in solid 
+lines, with a dotted line indicating zero probability.  Bottom panel: Various 
+estimators of $\ln[n(z)]$, the interim prior, and the true $\ln[n(z)]$ as a 
+continuous function and under a binned parametrization.}
+		\label{fig:featured}
+	\end{center}
+\end{figure}
+
+\subsubsection{Featured $n(z)$}
+\label{sec:featured}
+
+In this test, we choose the true $n(z)$ of Fig. \ref{fig:featured} with 
+nontrivial structure.
+
+\begin{figure}
+	\begin{center}
+
+\includegraphics[width=0.5\textwidth]{fig/fiducial/scatter.png}\\
+
+\includegraphics[width=0.5\textwidth]{fig/fiducial/estimators.png}		
+		\caption{[This plot doesn't really show what I want because the 
+intrinsic scatter is too low!  I used $\sigma=0.03$ because that's what has 
+been quoted as what LSST, etc. needs, but that's not what comes out of 
+photo-$z$ estimation methods before they impose aggressive cuts\dots]  Top 
+panel: The traditional MAP reduction of a photo-$z$ PDF against the true 
+redshifts with a few rescaled photo-$z$ interim posteriors are overplotted in 
+solid lines, with a dotted line indicating zero probability.  Bottom panel: 
+Various estimators of $\ln[n(z)]$, the interim prior, and the true $\ln[n(z)]$ 
+as a continuous function and under a binned parametrization.}
+		\label{fig:featured}
+	\end{center}
+\end{figure}
+
+This featured true $n(z)$ will henceforth be referred to as the fiducial $n(z)$
+
 \subsection{Emulated Data Quality Effects}
@@
 	\begin{center}
-		%\includegraphics[width=0.5\textwidth]{pgm.png}
-		\caption{[plot of $z_{phot}$ vs. $z_{spec}$ with varying 
-intrinsic scatter]}
+
+\includegraphics[width=0.5\textwidth]{fig/high_scatter/scatter.png}\\
+
+\includegraphics[width=0.5\textwidth]{fig/high_scatter/estimators.png}		
+		\caption{As the intrinsic scatter increases, the discrepancy 
+between estimators increases.  In particular, the stacked estimator and 
+marginalized point estimators predict $\ln[n(z)]$ to be smoother than the 
+truth, while the  [This would be a lot more compelling with more galaxies.  
+Also, the weird edge effects in the top panel are real, because the point 
+estimator is the MAP, not the center of the Gaussian likelihood, and there's no 
+requirement that the mean of the likelihood be within the true redshift range.] 
+ Top panel: The traditional MAP reduction of a photo-$z$ PDF against the true 
+redshifts with a few rescaled photo-$z$ interim posteriors are overplotted in 
+solid lines, with a dotted line indicating zero probability.  Bottom panel: 
+Various estimators of $\ln[n(z)]$, the interim prior, and the true $\ln[n(z)]$ 
+as a continuous function and under a binned parametrization.}
 		\label{fig:intscat}
@@
 	\begin{center}
-		%\includegraphics[width=0.5\textwidth]{pgm.png}
-		\caption{[plot of $z_{phot}$ vs. $z_{spec}$ with template-like 
-catastrophic outliers]}
+
+\includegraphics[width=0.5\textwidth]{fig/template_outliers/scatter.png}\\
+
+\includegraphics[width=0.5\textwidth]{fig/template_outliers/estimators.png}
+
+		\caption{Top panel: The traditional MAP reduction of a 
+photo-$z$ PDF against the true redshifts with a few rescaled photo-$z$ interim 
+posteriors are overplotted in solid lines, with a dotted line indicating zero 
+probability.  Bottom panel: Various estimators of $\ln[n(z)]$, the interim 
+prior, and the true $\ln[n(z)]$ as a continuous function and under a binned 
+parametrization.}
 		\label{fig:tempcatout}
@@
 	\begin{center}
-		%\includegraphics[width=0.5\textwidth]{pgm.png}
-		\caption{[plot of training-like (low-$z$ favoring) interim 
-prior]}
+
+\includegraphics[width=0.5\textwidth]{fig/training_prior/scatter.png}\\
+
+\includegraphics[width=0.5\textwidth]{fig/training_prior/estimators.png}	
+		\caption{[The case of a low-$z$ favoring interim prior was a 
+very compelling test in the previous version but somehow isn't anymore.]}
 		\label{fig:trainintpr}
-	\end{center}
-\end{figure}
-
-\subsection{Underlying $n(z)$ Effects}
-\label{sec:truth}
-
-Existing $n(z)$ estimators are systematically smoother than the true $n(z)$.  
-Here we show that the traditional estimators of $n(z)$ cannot recover a highly 
-featured $n(z)$.  The implication of this failure is quite serious; the 
-consistently smooth, unimodal $n(z)$ estimates appearing in the literature 
-could result from much more featured true redshift distributions, and there 
-would be no way to catch this error without using a fully probabilistic method.
-
-\subsubsection{Featured $n(z)$}
-\label{sec:featured}
-
-In this test, we choose the true $n(z)$ of Fig. \ref{fig:featured}, which has 
-nontrivial structure.
-
-\begin{figure}
-	\begin{center}
-		%\includegraphics[width=0.5\textwidth]{pgm.png}
-		\caption{[plot of a highly featured (delta function or 
-multimodal?) true $n(z)$]}
-		\label{fig:featured}
 	\end{center}
@@
 	\begin{center}
-		%\includegraphics[width=0.5\textwidth]{pgm.png}
 		\caption{[moneyplot of error ellipses resulting from different 

### 2017-06-26 research/paper/draft.tex
merged test cases
math added

@@
 interpretation of what a photo-$z$ PDF actually is.  When a photo-$z$ PDF is 
-reported, it is an object containing information about galaxy $i$'s redshift 
-$z_{i}$ based on its photometric data $\vec{d}_{i}$, making a photo-$z$ PDF a 
-\textit{posterior} probability distribution.  The data $\vec{d}_{i}$ may be 
+reported, it is an object containing information about the redshift $z_{i}$ of 
+galaxy $i$ based on its photometric data $\vec{d}_{i}$, making a photo-$z$ PDF 
+a \textit{posterior} probability distribution.  The data $\vec{d}_{i}$ may be 
 fluxes, magnitudes, colors, or any combination thereof.  We assume that the 
@@
 \begin{align}
-\log p(\vec{\phi} | \{\vec{d}_{i}\}) &\propto \log p(\{\vec{d}_{i}\} | 
-\vec{\phi}) + \log p(\vec{\phi}),
+\label{eq:hyper_bayes}
+\log[p(\vec{\phi} | \{\vec{d}_{i}\})] &\propto \log[p(\{\vec{d}_{i}\} | 
+\vec{\phi})] + \log[p(\vec{\phi})] ,
 \end{align}
@@
 \begin{align}
-\log[p(\{\vec{d}_{i}\} | \vec{\phi})] &= \log[\int\exp[\log[p(\{\vec{d}_{i}\} | 
-\{z_{i}\})]\\
-& + \log[p(\{z_{i}\} | \vec{\phi})]]\ d\{z_{i}\}].
+\label{eq:marginalize}
+\begin{split}
+\log[p(\{\vec{d}_{i}\} | \vec{\phi})] = & 
+\log\left[\int\exp\left[\log[p(\{\vec{d}_{i}\} | \{z_{i}\})]\right.\right. \\
+& \left.\left.+ \log[p(\{z_{i}\} | \vec{\phi})]\right]\ d\{z_{i}\}\right] .
+\end{split}
+\end{align}
+The redshifts $\{z_{i}\}$ are statistically independent of each other, and each 
+galaxy's photometry depends solely on its own redshift, so we may invoke 
+statistical independence to get
+\begin{align}
+\label{eq:independence_hyper}
+\log[p(\{z_{i}\} | \vec{\phi})] &= \sum_{i}^{N} \log[p(z_{i} | \vec{\phi})]
+\end{align}
+and
+\begin{align}
+\label{eq:independence}
+\log[p(\{\vec{d}_{i}\} | \{z_{i}\})] &= \sum_{i}^{N} \log[p(\vec{d}_{i} | 
+z_{i})] .
+\end{align}
+Having chosen the functional form mapping $\vec{\phi}$ to $n(z)$, the terms on 
+the righthand side of Eq. \ref{eq:independence_hyper} are known.  However, the 
+terms on the righthand side of Eq. \ref{eq:independence} are likelihoods, and 
+the photo-$z$ PDFs we have are interim posteriors.
+
+To transform the interim posteriors we have into the likelihoods we need, we 
+start with the vacuously true statement
+\begin{align}
+\begin{split}
+\label{eq:unity}
+\log[p(\vec{d}_{i} | z_{i})] =& \log[p(\vec{d}_{i} | z_{i})] + \log[p(z_{i} | 
+\vec{d}_{i}, \vec{\phi}^{*})]\\
+& - \log[p(z_{i} | \vec{d}_{i}, \vec{\phi}^{*})].
+\end{split}
+\end{align}
+We expand the last term using Bayes' Rule to get
+\begin{align}
+\begin{split}
+\label{eq:bayes}
+\log[p(\vec{d}_{i} | z_{i})] =& \log[p(\vec{d}_{i} | z_{i})] + \log[p(z_{i} | 
+\vec{d}_{i}, \vec{\phi}^{*})]\\
+& - \log[p(z_{i} | \vec{\phi}^{*})] - \log[p(\vec{d}_{i} | z_{i}, 
+\vec{\phi}^{*})] + \log[p(\vec{d}_{i} | \vec{\phi}^{*})].
+\end{split}
+\end{align}
+Fig. \ref{fig:pgm} has no direct connection between $\vec{d}_{i}$ and 
+$\vec{\phi}$, so we can break up $\log[p(\vec{d}_{i} | z_{i}, \vec{\phi}^{*})]$ 
+as follows:
+\begin{align}
+\begin{split}
+\label{eq:split}
+\log[p(\vec{d}_{i} | z_{i})] =& \log[p(\vec{d}_{i} | z_{i})] + \log[p(z_{i} | 
+\vec{d}_{i}, \vec{\phi}^{*})]\\
+& - \log[p(z_{i} | \vec{\phi}^{*})] - \log[p(\vec{d}_{i} | z_{i})] - 
+\log[p(\vec{d}_{i} | \vec{\phi}^{*})]\\
+& + \log[p(\vec{d}_{i} | \vec{\phi}^{*})].
+\end{split}
+\end{align}
+Canceling terms, we find
+\begin{align}
+\label{eq:cancel}
+\log[p(\vec{d}_{i} | z_{i})] =& \log[p(z_{i} | \vec{d}_{i}, \vec{\phi}^{*})] - 
+\log[p(z_{i} | \vec{\phi}^{*})].
 \end{align}
 
-Given a catalog of photo-$z$ interim posteriors $\{p(z_{i} | \vec{d}_{i}, 
-\vec{\alpha})\}$
-
-To make use of the photo-$z$ interim priors, it is absolutely essential that 
-$\vec{\alpha}$ be known.
+Now we combine Eqs. \ref{eq:hyper_bayes}, \ref{eq:marginalize}, 
+\ref{eq:independence_hyper}, \ref{eq:independence}, and \ref{eq:cancel} to 
+arrive at 
+\begin{align}
+\begin{split}
+\label{eq:final}
+\log[p(\vec{\phi} | \{\vec{d}_{i}\})] \propto& \log[p(\vec{\phi})] + 
+\log\left[\int \exp\left[\sum_{i}^{N} \left(\right.\right.\right.\\
+& \left.\left.\left. + \log[p(z_{i} | \vec{\phi})] - \log[p(z_{i} | 
+\vec{\phi}^{*})]\right.\right.\right.\\
+& \left.\left.\left. + \log[p(z_{i} | \vec{d}_{i}, \vec{\phi}^{*})]
+\right)\right] d\{z_{i}\}\right].
+\end{split}
+\end{align}
+Thus, if we have a catalog of photo-$z$ interim posteriors $\{p(z_{i} | 
+\vec{d}_{i}, \vec{\phi}^{*})\}$, the interim prior parameters $\vec{\phi}^{*}$, 
+and a prior distribution $p(\vec{\phi})$, we may find the posterior 
+$\log[p(\vec{\phi} | \{\vec{d}_{i}\})]$ on the redshift distribution $n(z)$.
 
 This framework entails a number of choices and assumptions that must be 
-addressed explicitly.
-
+addressed explicitly:
 \begin{enumerate}
-	\item The chosen functional form with parameters $\vec{\phi}$ must 
-actually be able to describe the true redshift distribution.
+	\item The chosen functional form with parameters $\vec{\phi}$ must be 
+capable of describing the true redshift distribution.
 	\item The photometric data and redshift for each galaxy is independent 
 of the photometric data and redshifts of all other galaxies.
-	\item There is one interim prior with parameters $\vec{\phi}^{*}$ 
-shared among all galaxies in the survey.
+	\item There is one interim prior shared among all galaxies in the 
+survey, and its parameters with known parameters $\vec{\phi}^{*}$ are known.
 	\item We must choose a hyperprior probability distribution 
@@
 cosmological constraints under different estimators of $n(z)$, creating a 
-figure like Fig. \ref{eq:scgsr}.  We perform this analysis using a forecasting 
+figure like Fig. \ref{fig:scgsr}.  We perform this analysis using a forecasting 
 code that takes $n(z)$ and produces projected error ellipses in the space of 
@@
 	\begin{center}
+		Not my figure! (N. MacCrann via J. DeRose)
 		\includegraphics[width=0.5\textwidth]{cosmo_constraints.png}
@@
 
-Thee results of Fig \ref{fig:money} have significant implications for the 
+The results of Fig \ref{fig:money} have significant implications for the 
 developing data analysis pipelines of next-generation telescope surveys.  

### 2017-06-26 research/paper/draft.tex
added alternatives, completed assumptions

@@
 \vec{d}_{i}, \vec{\phi}^{*})]\\
-& - \log[p(z_{i} | \vec{\phi}^{*})] - \log[p(\vec{d}_{i} | z_{i}, 
-\vec{\phi}^{*})] + \log[p(\vec{d}_{i} | \vec{\phi}^{*})].
+& - \log[p(\vec{d}_{i} | z_{i}, \vec{\phi}^{*})]\\
+& + \log[p(\vec{d}_{i} | \vec{\phi}^{*})] - \log[p(z_{i} | \vec{\phi}^{*})].
 \end{split}
@@
 \vec{d}_{i}, \vec{\phi}^{*})]\\
-& - \log[p(z_{i} | \vec{\phi}^{*})] - \log[p(\vec{d}_{i} | z_{i})] - 
-\log[p(\vec{d}_{i} | \vec{\phi}^{*})]\\
-& + \log[p(\vec{d}_{i} | \vec{\phi}^{*})].
+& - \log[p(\vec{d}_{i} | z_{i})] - \log[p(\vec{d}_{i} | \vec{\phi}^{*})]\\
+& + \log[p(\vec{d}_{i} | \vec{\phi}^{*})] - \log[p(z_{i} | \vec{\phi}^{*})].
 \end{split}
@@
 \begin{enumerate}
-	\item The chosen functional form with parameters $\vec{\phi}$ must be 
-capable of describing the true redshift distribution.
-	\item The photometric data and redshift for each galaxy is independent 
-of the photometric data and redshifts of all other galaxies.
+	\item The chosen functional form of $n(z)$ with parameters $\vec{\phi}$ 
+must be capable of describing the true redshift distribution.
+	\item The photometric data and redshift for each galaxy must be 
+independent of the photometric data and redshifts of all other galaxies.
 	\item There is one interim prior shared among all galaxies in the 
@@
 from photo-$z$ PDFs into the mathematical framework of Sec. \ref{sec:model}.  
+We do not advocate for these methods for reasons addressed in the Appendix.
+
+\subsubsection{Stacking}
+\label{sec:stacking}
+
+The most common way to combine photo-$z$ PDFs into an estimator 
+$\hat{n}_{stack}(z)$ for $n(z)$ is to stack them, which corresponds to
+\begin{align}
+\label{eq:stack}
+\hat{n}^{stack}(z) &= \frac{1}{N}\ \sum_{i}^{N} p(z_{i} | \vec{d}_{i}, 
+\vec{\phi}^{*}).
+\end{align}
+
+\subsubsection{Point Estimation}
+\label{sec:point_estimation}
+
+In some cases, estimators $\hat{n}(z)$ are obtained by assuming the photo-$z$ 
+PDFs are effectively delta functions by reducing them to point estimates 
+according to
+\begin{align}
+\label{eq:delta}
+p(z_{i} | \vec{d}_{i}, \vec{\phi}^{*}) &\approx \delta(z, \hat{z}_{i}^{point})
+\end{align}
+before applying Eq. \ref{eq:stack}.  The most common redshift point estimators 
+are the mean $\hat{z}^{mean}_{i}$, median $\hat{z}^{median}_{i}$, and mode 
+$\hat{z}^{mode}_{i}$ of the original photo-$z$ interim posterior.
 
@@
 \begin{enumerate}
+	\item \chippr is only applicable if the redshift interim posteriors and 
+interim prior are accurate.
 	\item \chippr currently only accepts photo-$z$ PDFs and produces $n(z)$ 
@@
 	\item \label{itm:true} Choose a true $n(z)$ that is a continuous 
-function with known parameters $\vec{\theta}$.
+function with known parameters $\vec{\phi}'$.
 	\item \label{itm:sample} Sample the true $n(z)$ to create a catalog of 
-$N$ true redshifts $z_{i}$.
-	\item \label{itm:scatter} Choose a true intrinsic scatter $\sigma$, and 
-sample Gaussians $p(z'_{i} | z_{i}, \sigma) = \mathcal{N}(z_{i}, \sigma)$ to 
-get "observed" redshifts $z'_{i}$ defining Gaussian likelihoods $p(\vec{d}_{i} 
-| z'_{i}, \sigma) = \mathcal{N}(z'_{i}, \sigma)$.
+$N$ true redshifts $z'_{i}$.
+	\item \label{itm:scatter} Choose a true intrinsic scatter parameter 
+$\sigma$, and sample Gaussians $p(z''_{i} | z'_{i}, \sigma) = 
+\mathcal{N}(z'_{i}, \sigma)$ to get "observed" redshifts $z'_{i}$ defining 
+Gaussian likelihoods $p(\vec{d}_{i} | z''_{i}, \sigma) = \mathcal{N}(z''_{i}, 
+\sigma)$.
 	\item \label{itm:interim} Choose a parametrization and the parameters 
-$\vec{\phi}^{*}$ of the interim prior.
-	\item \label{itm:posterior} Create interim posteriors $p(z_{i} | 
-\vec{d}_{i}, \vec{\phi}^{*}) = p(\vec{d}_{i} | z_{i})\ p(z_{i} | 
-\vec{\phi}^{*})$ in this parametrization.
+$\vec{\phi}^{*}$ of the interim prior $p(z_{i} | \vec{\phi}^{*})$.
+	\item \label{itm:posterior} Create interim posteriors $\log[p(z_{i} | 
+\vec{d}_{i}, \vec{\phi}^{*})] = \log[p(\vec{d}_{i} | z'_{i})]\ \log[p(z_{i} | 
+\vec{\phi}^{*})]$ in this parametrization.
 \end{enumerate}
 
-In all of the following validation tests, we use $10$ bins and $N = 10,000$ 
-galaxies.  In Secs. \ref{sec:truth} and \ref{sec:likelihoods}, we use a flat 
-interim prior.  This method for deriving mock data is referred to as the 
-fiducial case, and variations on it will refer directly to the steps that are 
-altered.
+In all of the following validation tests, we use a piecewise constant 
+parametrization in log-space with $10$ bins.  $N = 10,000$ galaxies.  In Secs. 
+\ref{sec:truth} and \ref{sec:likelihoods}, we use a flat interim prior.  This 
+method for deriving mock data is referred to as the fiducial case, and 
+variations on it will refer directly to the steps that are altered.
 
@@
 
+\appendix{}
+
 \begin{acknowledgements}

### 2017-06-27 chippr/catalog.py
enabling replotting of simulation plots

@@ -255,6 +255,10 @@ class catalog(object):
                 out.writerow(self.cat['log_interim_prior'])
                 for line in self.cat['log_interim_posteriors']:
                     out.writerow(line)
+            with open(os.path.join(self.data_dir, 'true_zs' + style), 'wb') as csvfile:
+                out = csv.writer(csvfile, delimiter=' ')
+                for item in self.true_samps:
+                    out.writerow(item)
         return
 
     def read(self, loc='data', style='.txt'):

### 2017-06-27 research/scripts/basic_replot_script.py

@@ -47,6 +47,13 @@ def just_plot(given_key):
     saved_location = 'data'
     saved_type = '.txt'
     data = simulated_posteriors.read(loc=saved_location, style=saved_type)
+
+    with open(os.path.join(test_dir, 'true_zs.txt'), 'rb') as csvfile:
+        tuples = (line.split(None) for line in csvfile)
+        true_zs = np.array([float(item) for item in tuples])
+    bin_mids = (data['bin_ends'][1:] + data['bin_ends'][:-1]) / 2.
+    catalog_plots.plot_obs_scatter(true_zs, data['log_interim_posteriors'], bin_mids)
+
     with open(os.path.join(os.path.join(test_dir, saved_location), 'true_params.p'), 'r') as true_file:
         true_nz_params = pickle.load(true_file)
     true_nz = chippr.gmix(true_nz_params['amps'], true_nz_params['means'], true_nz_params['sigmas'],

### 2017-06-27 chippr/catalog.py
fixing bug in writeout

@@ -255,10 +255,10 @@ class catalog(object):
                 out.writerow(self.cat['log_interim_prior'])
                 for line in self.cat['log_interim_posteriors']:
                     out.writerow(line)
+
             with open(os.path.join(self.data_dir, 'true_zs' + style), 'wb') as csvfile:
                 out = csv.writer(csvfile, delimiter=' ')
-                for item in self.true_samps:
-                    out.writerow(item)
+                out.writerow(self.true_samps)
         return
 
     def read(self, loc='data', style='.txt'):

### 2017-06-28 research/paper/draft.tex
periodic update

@@
 from photo-$z$ PDFs into the mathematical framework of Sec. \ref{sec:model}.  
-We do not advocate for these methods for reasons addressed in the Appendix.
+We briefly discuss the conditions under which they are equivalent to Eq. 
+\ref{eq:final}, which are elaborated upon in the Appendix.
 
@@
 
-The most common way to combine photo-$z$ PDFs into an estimator 
-$\hat{n}_{stack}(z)$ for $n(z)$ is to stack them, which corresponds to
+The most common way to combine photo-$z$ interim posteriors into an estimator 
+$\hat{n}_{stack}(z)$ for $n(z)$ is to "stack" them, which corresponds to
 \begin{align}
 \label{eq:stack}
-\hat{n}^{stack}(z) &= \frac{1}{N}\ \sum_{i}^{N} p(z_{i} | \vec{d}_{i}, 
-\vec{\phi}^{*}).
+\hat{n}^{stack}(z) &= \frac{1}{N}\ \sum_{i}^{N} \exp\left[\log[p(z_{i} | 
+\vec{d}_{i}, \vec{\phi}^{*})]\right].
 \end{align}
@@
 
-In some cases, estimators $\hat{n}(z)$ are obtained by assuming the photo-$z$ 
-PDFs are effectively delta functions by reducing them to point estimates 
-according to
+In some cases, estimators $\hat{n}^{point}(z)$ are obtained by assuming the 
+photo-$z$ PDFs are effectively delta functions by reducing them to point 
+estimates according to
 \begin{align}
@@
 \end{align}
-before applying Eq. \ref{eq:stack}.  The most common redshift point estimators 
+before applying Eq. \ref{eq:stack}.  The most popular redshift point estimators 
 are the mean $\hat{z}^{mean}_{i}$, median $\hat{z}^{median}_{i}$, and mode 

### 2017-06-28 research/paper/draft-editing.tex
merged test cases
fixed typos

@@
 Before jumping into the details of the model, it is crucial to settle on an interpretation of what a photo-$z$ PDF actually is.
-When a photo-$z$ PDF is reported, it is an object containing information about galaxy $i$'s redshift $z_{i}$ based on its photometric data $\vec{d}_{i}$, making a photo-$z$ PDF a \textit{posterior} probability distribution.
+When a photo-$z$ PDF is reported, it is an object containing information about the redshift $z_{i}$ of galaxy $i$ based on its photometric data $\vec{d}_{i}$, making a photo-$z$ PDF a \textit{posterior} probability distribution.
 The data $\vec{d}_{i}$ may be fluxes, magnitudes, colors, or any combination thereof.
@@
 
-% It would be easy to estimate $\vec{\phi}$ given $\{z_{i}\}_{N}$, however, what we actually have is photometric data $\{\vec{d}_{i}\}_{N}$ which we used to derive photo-$z$ PDFs.
-Because a photo-$z$ PDF is a probability distribution conditioned on the photometric data, it must be interpreted as a posterior probability distribution.
-
-% In addition to the data, the posterior probability distribution is conditioned on other information about how it was made, which is in general specific to the method that produced it.
-This information only influences the posterior by its implicit distribution over redshifts, which we call the interim prior over redshift.
-The interim prior, like the redshift distribution we wish to infer, has some functional form with some parameters, but since the redshift distribution and interim prior need not share the same parametrization, we will use $\vec{\alpha}$ to represent the parameters for the interim prior.
-Thus, the posterior probability that is often oversimplified as "$p(z_{i})$" is really the interim posterior probability $p(z_{i} | \vec{d}_{i}, \vec{\alpha})$.
-Note that we will assume that all galaxies whose photo-$z$ interim posteriors are derived by a single method share the same interim prior $p(z | \vec{\alpha})$; depending on the method, this may not always be true.
-
 The DAG of Fig.
@@
 \begin{align}
-\log p(\vec{\phi} | \{\vec{d}_{i}\}) &\propto \log p(\{\vec{d}_{i}\} | \vec{\phi}) + \log p(\vec{\phi}),
+\label{eq:hyper_bayes}
+\log[p(\vec{\phi} | \{\vec{d}_{i}\})] &\propto \log[p(\{\vec{d}_{i}\} | \vec{\phi})] + \log[p(\vec{\phi})] ,
 \end{align}
@@
 \begin{align}
-\log[p(\{\vec{d}_{i}\} | \vec{\phi})] &= \log[\int\exp[\log[p(\{\vec{d}_{i}\} | \{z_{i}\})]\\
-& + \log[p(\{z_{i}\} | \vec{\phi})]]\ d\{z_{i}\}].
+\label{eq:marginalize}
+\begin{split}
+\log[p(\{\vec{d}_{i}\} | \vec{\phi})] = & \log\left[\int\exp\left[\log[p(\{\vec{d}_{i}\} | \{z_{i}\})]\right.\right. \\
+& \left.\left.+ \log[p(\{z_{i}\} | \vec{\phi})]\right]\ d\{z_{i}\}\right] .
+\end{split}
+\end{align}
+The redshifts $\{z_{i}\}$ are statistically independent of each other, and each galaxy's photometry depends solely on its own redshift, so we may invoke statistical independence to get
+\begin{align}
+\label{eq:independence_hyper}
+\log[p(\{z_{i}\} | \vec{\phi})] &= \sum_{i}^{N} \log[p(z_{i} | \vec{\phi})]
+\end{align}
+and
+\begin{align}
+\label{eq:independence}
+\log[p(\{\vec{d}_{i}\} | \{z_{i}\})] &= \sum_{i}^{N} \log[p(\vec{d}_{i} | z_{i})] .
+\end{align}
+Having chosen the functional form mapping $\vec{\phi}$ to $n(z)$, the terms on the righthand side of Eq.
+\ref{eq:independence_hyper} are known.
+However, the terms on the righthand side of Eq.
+\ref{eq:independence} are likelihoods, and the photo-$z$ PDFs we have are interim posteriors.
+
+To transform the interim posteriors we have into the likelihoods we need, we start with the vacuously true statement
+\begin{align}
+\begin{split}
+\label{eq:unity}
+\log[p(\vec{d}_{i} | z_{i})] =& \log[p(\vec{d}_{i} | z_{i})] + \log[p(z_{i} | \vec{d}_{i}, \vec{\phi}^{*})]\\
+& - \log[p(z_{i} | \vec{d}_{i}, \vec{\phi}^{*})].
+\end{split}
+\end{align}
+We expand the last term using Bayes' Rule to get
+\begin{align}
+\begin{split}
+\label{eq:bayes}
+\log[p(\vec{d}_{i} | z_{i})] =& \log[p(\vec{d}_{i} | z_{i})] + \log[p(z_{i} | \vec{d}_{i}, \vec{\phi}^{*})]\\
+& - \log[p(\vec{d}_{i} | z_{i}, \vec{\phi}^{*})]\\
+& + \log[p(\vec{d}_{i} | \vec{\phi}^{*})] - \log[p(z_{i} | \vec{\phi}^{*})].
+\end{split}
+\end{align}
+Fig. \ref{fig:pgm} has no direct connection between $\vec{d}_{i}$ and $\vec{\phi}$, so we can break up $\log[p(\vec{d}_{i} | z_{i}, \vec{\phi}^{*})]$ as follows:
+\begin{align}
+\begin{split}
+\label{eq:split}
+\log[p(\vec{d}_{i} | z_{i})] =& \log[p(\vec{d}_{i} | z_{i})] + \log[p(z_{i} | \vec{d}_{i}, \vec{\phi}^{*})]\\
+& - \log[p(\vec{d}_{i} | z_{i})] - \log[p(\vec{d}_{i} | \vec{\phi}^{*})]\\
+& + \log[p(\vec{d}_{i} | \vec{\phi}^{*})] - \log[p(z_{i} | \vec{\phi}^{*})].
+\end{split}
+\end{align}
+Canceling terms, we find
+\begin{align}
+\label{eq:cancel}
+\log[p(\vec{d}_{i} | z_{i})] =& \log[p(z_{i} | \vec{d}_{i}, \vec{\phi}^{*})] - \log[p(z_{i} | \vec{\phi}^{*})].
 \end{align}
 
-Given a catalog of photo-$z$ interim posteriors $\{p(z_{i} | \vec{d}_{i}, \vec{\alpha})\}$
+Now we combine Eqs. \ref{eq:hyper_bayes}, \ref{eq:marginalize}, \ref{eq:independence_hyper}, \ref{eq:independence}, and \ref{eq:cancel} to arrive at 
+\begin{align}
+\begin{split}
+\label{eq:final}
+\log[p(\vec{\phi} | \{\vec{d}_{i}\})] \propto& \log[p(\vec{\phi})] + \log\left[\int \exp\left[\sum_{i}^{N} \left(\right.\right.\right.\\
+& \left.\left.\left. + \log[p(z_{i} | \vec{\phi})] - \log[p(z_{i} | \vec{\phi}^{*})]\right.\right.\right.\\
+& \left.\left.\left. + \log[p(z_{i} | \vec{d}_{i}, \vec{\phi}^{*})]
+\right)\right] d\{z_{i}\}\right].
+\end{split}
+\end{align}
+Thus, if we have a catalog of photo-$z$ interim posteriors $\{p(z_{i} | \vec{d}_{i}, \vec{\phi}^{*})\}$, the interim prior parameters $\vec{\phi}^{*}$, and a prior distribution $p(\vec{\phi})$, we may find the posterior $\log[p(\vec{\phi} | \{\vec{d}_{i}\})]$ on the redshift distribution $n(z)$.
 
-To make use of the photo-$z$ interim priors, it is absolutely essential that $\vec{\alpha}$ be known.
-
-This framework entails a number of choices and assumptions that must be addressed explicitly.
-
+This framework entails a number of choices and assumptions that must be addressed explicitly:
 \begin{enumerate}
-	\item The chosen functional form with parameters $\vec{\phi}$ must actually be able to describe the true redshift distribution.
-	\item The photometric data and redshift for each galaxy is independent of the photometric data and redshifts of all other galaxies.
-	\item There is one interim prior with parameters $\vec{\phi}^{*}$ shared among all galaxies in the survey.
+	\item The chosen functional form of $n(z)$ with parameters $\vec{\phi}$ must be capable of describing the true redshift distribution.
+	\item The photometric data and redshift for each galaxy must be independent of the photometric data and redshifts of all other galaxies.
+	\item There is one interim prior shared among all galaxies in the survey, and its parameters with known parameters $\vec{\phi}^{*}$ are known.
 	\item We must choose a hyperprior probability distribution $p(\vec{\phi})$ over the hyperparameters.
@@
 
-It is useful to translate some popular existing methods for deriving $n(z)$ from photo-$z$ PDFs into the mathematical framework of Sec. \ref{sec:model}.
+It is useful to translate some popular existing methods for deriving $n(z)$ from photo-$z$ PDFs into the mathematical framework of Sec.
+\ref{sec:model}.
+We briefly discuss the conditions under which they are equivalent to Eq.
+\ref{eq:final}, which are elaborated upon in the Appendix.
+
+\subsubsection{Stacking}
+\label{sec:stacking}
+
+The most common way to combine photo-$z$ interim posteriors into an estimator $\hat{n}_{stack}(z)$ for $n(z)$ is to "stack" them, which corresponds to
+\begin{align}
+\label{eq:stack}
+\hat{n}^{stack}(z) &= \frac{1}{N}\ \sum_{i}^{N} \exp\left[\log[p(z_{i} | \vec{d}_{i}, \vec{\phi}^{*})]\right].
+\end{align}
+
+\subsubsection{Point Estimation}
+\label{sec:point_estimation}
+
+In some cases, estimators $\hat{n}^{point}(z)$ are obtained by assuming the photo-$z$ PDFs are effectively delta functions by reducing them to point estimates according to
+\begin{align}
+\label{eq:delta}
+p(z_{i} | \vec{d}_{i}, \vec{\phi}^{*}) &\approx \delta(z, \hat{z}_{i}^{point})
+\end{align}
+before applying Eq.
+\ref{eq:stack}.
+The most popular redshift point estimators are the mean $\hat{z}^{mean}_{i}$, median $\hat{z}^{median}_{i}$, and mode $\hat{z}^{mode}_{i}$ of the original photo-$z$ interim posterior.
 
@@
 \begin{enumerate}
+	\item \chippr is only applicable if the redshift interim posteriors and interim prior are accurate.
 	\item \chippr currently only accepts photo-$z$ PDFs and produces $n(z)$ samples of a single format, that of the piecewise constant parametrization, also referred to as a binned histogram parametrization and a sum of top hat functions.
@@
 \begin{enumerate}
-	\item \label{itm:true} Choose a true $n(z)$ that is a continuous function with known parameters $\vec{\theta}$.
-	\item \label{itm:sample} Sample the true $n(z)$ to create a catalog of $N$ true redshifts $z_{i}$.
-	\item \label{itm:scatter} Choose a true intrinsic scatter $\sigma$, and sample Gaussians $p(z'_{i} | z_{i}, \sigma) = \mathcal{N}(z_{i}, \sigma)$ to get "observed" redshifts $z'_{i}$ defining Gaussian likelihoods $p(\vec{d}_{i} | z'_{i}, \sigma) = \mathcal{N}(z'_{i}, \sigma)$.
-	\item \label{itm:interim} Choose a parametrization and the parameters $\vec{\phi}^{*}$ of the interim prior.
-	\item \label{itm:posterior} Create interim posteriors $p(z_{i} | \vec{d}_{i}, \vec{\phi}^{*}) = p(\vec{d}_{i} | z_{i})\ p(z_{i} | \vec{\phi}^{*})$ in this parametrization.
+	\item \label{itm:true} Choose a true $n(z)$ that is a continuous function with known parameters $\vec{\phi}'$.
+	\item \label{itm:sample} Sample the true $n(z)$ to create a catalog of $N$ true redshifts $z'_{i}$.
+	\item \label{itm:scatter} Choose a true intrinsic scatter parameter $\sigma$, and sample Gaussians $p(z''_{i} | z'_{i}, \sigma) = \mathcal{N}(z'_{i}, \sigma)$ to get "observed" redshifts $z'_{i}$ defining Gaussian likelihoods $p(\vec{d}_{i} | z''_{i}, \sigma) = \mathcal{N}(z''_{i}, \sigma)$.
+	\item \label{itm:interim} Choose a parametrization and the parameters $\vec{\phi}^{*}$ of the interim prior $p(z_{i} | \vec{\phi}^{*})$.
+	\item \label{itm:posterior} Create interim posteriors $\log[p(z_{i} | \vec{d}_{i}, \vec{\phi}^{*})] = \log[p(\vec{d}_{i} | z'_{i})]\ \log[p(z_{i} | \vec{\phi}^{*})]$ in this parametrization.
 \end{enumerate}
 
-In all of the following validation tests, we use $10$ bins and $N = 10,000$ galaxies.
+In all of the following validation tests, we use a piecewise constant parametrization in log-space with $10$ bins.
+$N = 10,000$ galaxies.
 In Secs.
@@
 \ref{sec:buzzard} is associated with true values of the cosmological parameters, we may compare the quality of cosmological constraints under different estimators of $n(z)$, creating a figure like Fig.
-\ref{eq:scgsr}.
+\ref{fig:scgsr}.
 We perform this analysis using a forecasting code that takes $n(z)$ and produces projected error ellipses in the space of cosmological parameters.
@@
 
-Thee results of Fig \ref{fig:money} have significant implications for the developing data analysis pipelines of next-generation telescope surveys.
+The results of Fig \ref{fig:money} have significant implications for the developing data analysis pipelines of next-generation telescope surveys.
 However, the method presented here has its own limitations, which are reiterated to discourage the community from applying this work inappropriately.
@@
 
-%\appendix{}
+\appendix{}
+%showing stacking is always too broad
 

### 2017-06-28 research/paper/draft.tex
now describing new mock data approach in appendix

@@
 
-\appendix{}
+\appendix{Catalog Production}
+\label{app:catalog}
+
+Sec. \ref{sec:mockdata} outlines one way to make mock data, but it's not 
+necessarily a logically consistent way to do it, a problem identified in trying 
+to use it for the test cases in Sec. \ref{sec:likelihoods}.  If the goal is to 
+emulate the traditional $z_{spec}$ vs. $z_{phot}$ scatterplots, we need to 
+start from probability distributions in that space.  This means there are two 
+parameters $z_{spec}$ and $z_{phot}$ drawn from a two-dimensional distribution 
+$p(z_{spec}, z_{phot} | \textul{\sigma}, \vec{\phi})$, where $\textul{\sigma}$ 
+contains parameters concerning the relationship between $z_{spec}$ and 
+$z_{phot}$, and $\vec{\phi}$ contains hyperparameters concerning $n(z)$.  We 
+know from basic probability that
+\begin{align}
+\label{eq:trivial}
+p(z_{spec}, z_{phot} | \textul{\sigma}, \vec{\phi}) =& p(z_{phot} | z_{spec}, 
+\textul{\sigma}, \vec{\phi})\ p(z_{sped} | \textul{\sigma}, \vec{\phi}).
+\end{align}
+Because $z_{spec}$ depends only on $\vec{\phi}$ and $z_{phot}$ is independent 
+of $\vec{\phi}$, this becomes
+\begin{align}
+\label{eq:simpler}
+p(z_{spec}, z_{phot} | \textul{\sigma}, \vec{\phi}) =& p(z_{phot} | z_{spec}, 
+\textul{\sigma})\ p(z_{spec} | \vec{\phi}).
+\end{align}
+We know how to make reasonable choices for both terms on the righthand side of 
+Eq. \ref{eq:simpler}.
+
+In the $z_{spec}$ vs. $z_{phot}$ scatterplot, $p(z_{phot} | z_{spec}, 
+\textul{\sigma}, \vec{\phi})$ would be represented by vertical slices, but we 
+are interested in the horizontal slices that represent $p(z_{spec} | z_{phot}, 
+\textul{\sigma}, \vec{\phi})$.  Now we must distinguish the true 
+hyperparameters $\vec{\phi}_{true}$ from the interim hyperparameters 
+$\vec{\phi}_{int}$.  If the horizontal slices carry information about 
+$\vec{\phi}$, how can we get a catalog of $p(z_{spec} | z_{phot}, 
+\textul{\sigma}, \vec{\phi}_{int})$ that's consistent with $p(z_{spec} | 
+z_{phot}, \textul{\sigma}, \vec{\phi}_{true})$?  Somehow that information needs 
+to be propagated through in order to recover it.
+
+We achieve this goal by dividing out the true $n(z)$ and multiplying by the 
+interim $n(z)$ as in
+\begin{align}
+\label{eq:transformation}
+p(z_{spec} | z_{phot}, \textul{\sigma}, \vec{\phi}_{int}) =& p(z_{spec} | 
+z_{phot}, \textul{\sigma}, \vec{\phi}_{true})\ \frac{p(z_{spec} | 
+\vec{\phi}_{int})}{p(z_{spec} | \vec{\phi}_{true}},
+\end{align}
+since $p(z_{spec} | z_{phot}, \textul{\sigma}, \vec{\phi})$ is separable and 
+the term $p(z_{spec} | z_{phot}, \textul{\sigma})$ changes independently of the 
+$p(z_{spec} | \vec{\phi})$ term.  The information about $\vec{\phi}_{true}$ is 
+retained by the elements of the catalog: only $(z_{spec}, z_{phot})$ pairs 
+drawn from the true distribution will be represented in the catalog!
+
+It appears as if this is the only fully self-consistent way to ensure that a 
+draw from the posterior distribution $p(z_{spec} | z_{phot}, \textul{\sigma}, 
+\vec{\phi})$ does in fact follow that distribution.  In the cases with fiducial 
+posteriors $p(z_{spec} | z_{phot}, \sigma, \vec{\phi})$, the method of Sec. 
+\ref{sec:mockdata} was equivalent to this, but Sec. \ref{sec:likelihoods} 
+requires a more sophisticated approach.
 

### 2017-06-28 chippr/catalog.py
changed gaussian mixture object

@@ -255,10 +255,6 @@ class catalog(object):
                 out.writerow(self.cat['log_interim_prior'])
                 for line in self.cat['log_interim_posteriors']:
                     out.writerow(line)
-
-            with open(os.path.join(self.data_dir, 'true_zs' + style), 'wb') as csvfile:
-                out = csv.writer(csvfile, delimiter=' ')
-                out.writerow(self.true_samps)
         return
 
     def read(self, loc='data', style='.txt'):

### 2017-06-28 chippr/gmix.py

@@ -7,29 +7,25 @@ from chippr import utils as u
 
 class gmix(object):
 
-    def __init__(self, amps, means, sigmas, limits=(d.min_x, d.max_x)):
+    def __init__(self, amps, funcs, limits=(d.min_x, d.max_x)):
         """
-        Object to define a Gaussian mixture probability distribution
+        Object to define a mixture probability distribution
 
         Parameters
         ----------
         amps: ndarray, float
             array with one relative amplitude per component
-        means: ndarray, float
-            array with one mean per component
-        sigmas: ndarray, float
-            array with one standard deviation per component
+        funcs: list, chippr.gauss or chippr.discrete objects
+            list of components
         limits: tuple or list or numpy.ndarray, float, optional
             minimum and maximum sample values to return
         """
 
         self.amps = amps/np.sum(amps)
         self.cumamps = np.cumsum(self.amps)
-        self.means = means
-        self.sigmas = sigmas
         self.n_comps = len(self.amps)
 
-        self.funcs = [chippr.gauss(self.means[c], self.sigmas[c]**2) for c in range(self.n_comps)]
+        self.funcs = funcs#[chippr.gauss(self.means[c], self.sigmas[c]**2) for c in range(self.n_comps)]
 
         self.min_x = limits[0]
         self.max_x = limits[1]

### 2017-06-28 chippr/utils.py

@@ -24,7 +24,10 @@ def safe_log(arr, threshold=d.eps):
     logged: ndarray
         logged values, with small value replacing un-loggable values
     """
+    # if type(arr) == np.ndarray:
     arr[arr < threshold] = threshold
+    # else:
+    #     arr = max(threshold, arr)
     logged = np.log(arr)
     return logged
 

### 2017-06-28 research/scripts/basic_inf_script.py

@@ -98,7 +98,14 @@ def do_inference(given_key):
     z_difs = zs[1:]-zs[:-1]
     with open(os.path.join(os.path.join(test_dir, saved_location), 'true_params.p'), 'r') as true_file:
         true_nz_params = pickle.load(true_file)
-    true_nz = chippr.gmix(true_nz_params['amps'], true_nz_params['means'], true_nz_params['sigmas'],
+    true_amps = true_nz_params['amps']
+    true_means = true_nz_params['means']
+    true_sigmas =  true_nz_params['sigmas']
+    n_mix_comps = len(true_amps)
+    true_funcs = []
+    for c in range(n_mix_comps):
+        true_funcs.append(chippr.gauss(true_means[c], true_sigmas[c]**2))
+    true_nz = chippr.gmix(true_amps, true_funcs,
             limits=(min(zs), max(zs)))
 
     (prior, cov) = set_up_prior(data, params)
@@ -126,7 +133,7 @@ def do_inference(given_key):
         n_ivals = 10 * n_bins
     initial_values = start.sample(n_ivals)
 
-    nz_samps = nz.calculate_samples(initial_values, no_data=params['no_data'], no_prior=params['no_prior'])
+    # nz_samps = nz.calculate_samples(initial_values, no_data=params['no_data'], no_prior=params['no_prior'])
 
     nz_stats = nz.compare()
 
@@ -143,7 +150,7 @@ if __name__ == "__main__":
     from chippr import *
 
     result_dir = os.path.join('..', 'results')
-    test_name = 'template_prior'
+    test_name = 'fiducial'
     all_tests = {}
     test_info = {}
     test_info['name'] = test_name

### 2017-06-28 research/scripts/basic_replot_script.py

@@ -48,16 +48,21 @@ def just_plot(given_key):
     saved_type = '.txt'
     data = simulated_posteriors.read(loc=saved_location, style=saved_type)
 
-    with open(os.path.join(test_dir, 'true_zs.txt'), 'rb') as csvfile:
-        tuples = (line.split(None) for line in csvfile)
-        true_zs = np.array([float(item) for item in tuples])
-    bin_mids = (data['bin_ends'][1:] + data['bin_ends'][:-1]) / 2.
-    catalog_plots.plot_obs_scatter(true_zs, data['log_interim_posteriors'], bin_mids)
-
     with open(os.path.join(os.path.join(test_dir, saved_location), 'true_params.p'), 'r') as true_file:
         true_nz_params = pickle.load(true_file)
-    true_nz = chippr.gmix(true_nz_params['amps'], true_nz_params['means'], true_nz_params['sigmas'],
-        limits=(simulated_posteriors.params['bin_min'], simulated_posteriors.params['bin_min']))
+    true_amps = true_nz_params['amps']
+    true_means = true_nz_params['means']
+    true_sigmas =  true_nz_params['sigmas']
+    n_mix_comps = len(true_amps)
+    true_funcs = []
+    for c in range(n_mix_comps):
+        true_funcs.append(chippr.gauss(true_means[c], true_sigmas[c]**2))
+    true_nz = chippr.gmix(true_amps, true_funcs,
+            limits=(simulated_posteriors.params['bin_min'], simulated_posteriors.params['bin_max']))
+
+    true_zs = true_nz_params['zs']
+    bin_mids = (data['bin_ends'][1:] + data['bin_ends'][:-1]) / 2.
+    catalog_plots.plot_obs_scatter(true_zs, data['log_interim_posteriors'], bin_mids)
 
     prior = set_up_prior(data)
     n_bins = len(data['log_interim_prior'])

### 2017-06-28 research/scripts/basic_sim_script.py

@@ -72,7 +72,11 @@ def make_true(given_key):
         true_means = np.array([0.5, 0.2, 0.75]) * bin_range + min(test_info['bin_ends'])
         true_sigmas = np.array([0.4, 0.2, 0.1]) * bin_range
 
-    true_nz = chippr.gmix(true_amps, true_means, true_sigmas,
+    n_mix_comps = len(true_amps)
+    true_funcs = []
+    for c in range(n_mix_comps):
+        true_funcs.append(chippr.gauss(true_means[c], true_sigmas[c]**2))
+    true_nz = chippr.gmix(true_amps, true_funcs,
             limits=(min(test_info['bin_ends']), max(test_info['bin_ends'])))
 
     true_dict = {'amps': true_amps, 'means': true_means, 'sigmas': true_sigmas}
@@ -105,7 +109,11 @@ def make_interim_prior(given_key):
         int_amps = np.array([0.35, 0.5, 0.15])
         int_means = np.array([0.1, 0.5, 0.9]) * bin_range + min(test_info['bin_ends'])
         int_sigmas = np.array([0.1, 0.1, 0.1]) * bin_range
-        interim_prior = chippr.gmix(int_amps, int_means, int_sigmas,
+        n_mix_comps = len(int_amps)
+        int_funcs = []
+        for c in range(n_mix_comps):
+            int_funcs.append(chippr.gauss(int_means[c], int_sigmas[c]**2))
+        interim_prior = chippr.gmix(int_amps, int_funcs,
             limits=(min(test_info['bin_ends']), max(test_info['bin_ends'])))
     elif test_info['params']['interim_prior'] == 'training':
         int_amps = np.array([0.150,0.822,1.837,2.815,3.909,
@@ -123,7 +131,11 @@ def make_interim_prior(given_key):
         int_amps = f(bin_mids)
         int_means = bin_mids
         int_sigmas = bin_difs
-        interim_prior = chippr.gmix(int_amps, int_means, int_sigmas,
+        n_mix_comps = len(int_amps)
+        int_funcs = []
+        for c in range(n_mix_comps):
+            int_funcs.append(chippr.gauss(int_means[c], int_sigmas[c]**2))
+        interim_prior = chippr.gmix(int_amps, int_funcs,
                 limits=(min(test_info['bin_ends']), max(test_info['bin_ends'])))
     else:
         bin_ends = np.array([test_info['params']['bin_min'], test_info['params']['bin_max']])
[…]

### 2017-06-28 research/scripts/multi_inf_script.py

@@ -128,7 +128,14 @@ def do_inference(given_key):
     z_difs = zs[1:]-zs[:-1]
     with open(os.path.join(os.path.join(test_dir, saved_location), 'true_params.p'), 'r') as true_file:
         true_nz_params = pickle.load(true_file)
-    true_nz = chippr.gmix(true_nz_params['amps'], true_nz_params['means'], true_nz_params['sigmas'],
+    true_amps = true_nz_params['amps']
+    true_means = true_nz_params['means']
+    true_sigmas =  true_nz_params['sigmas']
+    n_mix_comps = len(true_amps)
+    true_funcs = []
+    for c in range(n_mix_comps):
+        true_funcs.append(chippr.gauss(true_means[c], true_sigmas[c]**2))
+    true_nz = chippr.gmix(true_amps, true_funcs,
             limits=(min(zs), max(zs)))
 
     (prior, cov) = set_up_prior(data, params)
@@ -155,7 +162,7 @@ def do_inference(given_key):
         n_ivals = 10 * n_bins
     initial_values = prior.sample(n_ivals)
     log_z_dens_plots.plot_ivals(initial_values, nz.info, nz.plot_dir)
-    nz_samps = nz.calculate_samples(initial_values, no_data=params['no_data'], no_prior=params['no_prior'])
+    # nz_samps = nz.calculate_samples(initial_values, no_data=params['no_data'], no_prior=params['no_prior'])
 
     nz_stats = nz.compare()
 

### 2017-06-28 research/scripts/multi_replot_script.py

@@ -47,16 +47,22 @@ def just_plot(given_key):
     saved_location = 'data'
     saved_type = '.txt'
     data = simulated_posteriors.read(loc=saved_location, style=saved_type)
-    with open(os.path.join(test_dir, 'true_zs.txt'), 'rb') as csvfile:
-        tuples = (line.split(None) for line in csvfile)
-        true_zs = np.array([float(item) for item in tuples])
-    bin_mids = (data['bin_ends'][1:] + data['bin_ends'][:-1]) / 2.
-    catalog_plots.plot_obs_scatter(true_zs, data['log_interim_posteriors'], bin_mids)
 
     with open(os.path.join(os.path.join(test_dir, saved_location), 'true_params.p'), 'r') as true_file:
         true_nz_params = pickle.load(true_file)
-    true_nz = chippr.gmix(true_nz_params['amps'], true_nz_params['means'], true_nz_params['sigmas'],
-        limits=(simulated_posteriors.params['bin_min'], simulated_posteriors.params['bin_min']))
+    true_amps = true_nz_params['amps']
+    true_means = true_nz_params['means']
+    true_sigmas =  true_nz_params['sigmas']
+    n_mix_comps = len(true_amps)
+    true_funcs = []
+    for c in range(n_mix_comps):
+        true_funcs.append(chippr.gauss(true_means[c], true_sigmas[c]**2))
+    true_nz = chippr.gmix(true_amps, true_funcs,
+            limits=(simulated_posteriors.params['bin_min'], simulated_posteriors.params['bin_max']))
+
+    true_zs = true_nz_params['zs']
+    bin_mids = (data['bin_ends'][1:] + data['bin_ends'][:-1]) / 2.
+    catalog_plots.plot_obs_scatter(true_zs, data['log_interim_posteriors'], bin_mids)
 
     prior = set_up_prior(data)
     n_bins = len(data['log_interim_prior'])

### 2017-06-28 research/scripts/multi_sim_script.py

@@ -72,7 +72,11 @@ def make_true(given_key):
         true_means = np.array([0.5, 0.2, 0.75]) * bin_range + min(test_info['bin_ends'])
         true_sigmas = np.array([0.4, 0.2, 0.1]) * bin_range
 
-    true_nz = chippr.gmix(true_amps, true_means, true_sigmas,
+    n_mix_comps = len(true_amps)
+    true_funcs = []
+    for c in range(n_mix_comps):
+        true_funcs.append(chippr.gauss(true_means[c], true_sigmas[c]**2))
+    true_nz = chippr.gmix(true_amps, true_funcs,
             limits=(min(test_info['bin_ends']), max(test_info['bin_ends'])))
 
     true_dict = {'amps': true_amps, 'means': true_means, 'sigmas': true_sigmas}
@@ -105,7 +109,11 @@ def make_interim_prior(given_key):
         int_amps = np.array([0.35, 0.5, 0.15])
         int_means = np.array([0.1, 0.5, 0.9]) * bin_range + min(test_info['bin_ends'])
         int_sigmas = np.array([0.1, 0.1, 0.1]) * bin_range
-        interim_prior = chippr.gmix(int_amps, int_means, int_sigmas,
+        n_mix_comps = len(int_amps)
+        int_funcs = []
+        for c in range(n_mix_comps):
+            int_funcs.append(chippr.gauss(int_means[c], int_sigmas[c]**2))
+        interim_prior = chippr.gmix(int_amps, int_funcs,
             limits=(min(test_info['bin_ends']), max(test_info['bin_ends'])))
     elif test_info['params']['interim_prior'] == 'training':
         int_amps = np.array([0.150,0.822,1.837,2.815,3.909,
@@ -123,7 +131,11 @@ def make_interim_prior(given_key):
         int_amps = f(bin_mids)
         int_means =bin_mids
         int_sigmas = bin_difs
-        interim_prior = chippr.gmix(int_amps, int_means, int_sigmas,
+        n_mix_comps = len(int_amps)
+        int_funcs = []
+        for c in range(n_mix_comps):
+            int_funcs.append(chippr.gauss(int_means[c], int_sigmas[c]**2))
+        interim_prior = chippr.gmix(int_amps, int_funcs,
                 limits=(min(test_info['bin_ends']), max(test_info['bin_ends'])))
     else:
         bin_ends = np.array([test_info['params']['bin_min'], test_info['params']['bin_max']])

### 2017-06-28 research/scripts/multi_replot_script.py
fixed tiny bug of where plots are saved

@@ -62,7 +62,7 @@ def just_plot(given_key):
 
     true_zs = true_nz_params['zs']
     bin_mids = (data['bin_ends'][1:] + data['bin_ends'][:-1]) / 2.
-    catalog_plots.plot_obs_scatter(true_zs, data['log_interim_posteriors'], bin_mids)
+    catalog_plots.plot_obs_scatter(true_zs, data['log_interim_posteriors'], bin_mids, plot_loc=os.path.join(test_dir, saved_location))
 
     prior = set_up_prior(data)
     n_bins = len(data['log_interim_prior'])

### 2017-06-28 chippr/catalog.py
now catalog takes parametric truth and draws the redshifts

@@ -98,15 +98,16 @@ class catalog(object):
 
         return coarse
 
-    def create(self, truth, int_pr, vb=True):
+    def create(self, truth, int_pr, N=d.n_gals, vb=True):
         """
         Function creating a catalog of interim posterior probability
         distributions, will split this up into helper functions
 
         Parameters
         ----------
-        truth: numpy.ndarray, float
-            vector of true redshifts
+        truth: chippr.gmix object or chippr.gauss object or chippr.discrete
+        object
+            true redshift distribution object
         int_pr: chippr.gmix object or chippr.gauss object or chippr.discrete
         object
             interim prior distribution object
@@ -118,7 +119,8 @@ class catalog(object):
         self.cat: dict
             dictionary comprising catalog information
         """
-        self.true_samps = truth
+        self.truth = truth
+        self.true_samps = self.truth.sample(10**N)
         if vb:
             plots.plot_true_histogram(self.true_samps, plot_loc=self.plot_dir)
         self.n_items = len(self.true_samps)

### 2017-06-28 chippr/defaults.py

@@ -10,6 +10,8 @@ max_x = 1.
 
 n_bins = 10
 
+n_gals = 4
+
 constant_sigma = 0.03
 
 gr_threshold = 1.2
@@ -54,10 +56,10 @@ def check_basic_setup(params):
     params: dict
         dictionary containing key/value pairs for simulation
     """
-    # if 'n_items' not in params:
-    #     params['n_items'] = n_items
-    # else:
-    #     params['n_items'] = 10 ** int(params['n_items'][0])
+    if 'n_gals' not in params:
+        params['n_gals'] = n_gals
+    else:
+        params['n_gals'] = int(params['n_gals'][0])
     if 'n_bins' not in params:
         params['n_bins'] = n_bins
     else:

### 2017-06-28 research/scripts/basic_sim_script.py

@@ -174,11 +174,20 @@ def make_catalog(given_key):
 
     test_info = make_true(given_key)
     true_zs = test_info['truth']['zs']
+    true_amps = test_info['truth']['amps']
+    true_means = test_info['truth']['means']
+    true_sigmas =  test_info['truth']['sigmas']
+    n_mix_comps = len(true_amps)
+    true_funcs = []
+    for c in range(n_mix_comps):
+        true_funcs.append(chippr.gauss(true_means[c], true_sigmas[c]**2))
+    true_nz = chippr.gmix(true_amps, true_funcs,
+            limits=(test_info['params']['bin_min'], test_info['params']['bin_max']))
 
     interim_prior = make_interim_prior(given_key)
 
     posteriors = chippr.catalog(param_file_name, loc=test_dir)
-    output = posteriors.create(true_zs, interim_prior)
+    output = posteriors.create(true_nz, interim_prior, N=test_info['params']['n_gals'])
     # data = np.exp(output['log_interim_posteriors'])
     posteriors.write()
     data_dir = posteriors.data_dir

### 2017-06-29 chippr/catalog.py
not working but pushing progress

@@ -12,6 +12,7 @@ from chippr import defaults as d
 from chippr import utils as u
 from chippr import sim_utils as su
 from chippr import gauss
+from chippr import discrete
 from chippr import catalog_plots as plots
 
 class catalog(object):
@@ -120,21 +121,32 @@ class catalog(object):
             dictionary comprising catalog information
         """
         self.truth = truth
-        self.true_samps = self.truth.sample(10**N)
-        if vb:
-            plots.plot_true_histogram(self.true_samps, plot_loc=self.plot_dir)
-        self.n_items = len(self.true_samps)
-        self.samp_range = range(self.n_items)
+        # self.true_samps = self.truth.sample(10**N)
+        # if vb:
+        #     plots.plot_true_histogram(self.true_samps, plot_loc=self.plot_dir)
+        # self.n_items = len(self.true_samps)
+        # self.samp_range = range(self.n_items)
 
         self.proc_bins()
-        self.obs_samps = self.sample_obs()
+        # self.obs_samps = self.sample_obs()
 
-        self.int_pr = int_pr
+        self.prob_space = self.make_probs()
+        if vb:
+            plots.plot_prob_space(self.z_fine, self.prob_space, plot_loc=self.plot_dir)
+
+        ## this far!
+        ## next, sample discrete to get z_true, z_obs
+        ## then literally take slices!
 
         self.obs_lfs = self.evaluate_lfs()
         for n in self.samp_range:
             self.obs_lfs[n] /= np.sum(self.obs_lfs[n]) * self.dz_fine
 
+        # if vb:
+        #     plots.plot_scatter(self.true_samps, self.obs_samps, self.obs_lfs, self.z_fine, plot_loc=self.plot_dir)
+        #
+
+        self.int_pr = int_pr
         int_pr_fine = self.int_pr.evaluate(self.z_fine)
         int_pr_coarse = self.coarsify(int_pr_fine)
 
@@ -159,51 +171,109 @@ class catalog(object):
 
         return self.cat
 
-    def sample_obs(self):
+    def make_probs(self, vb=True):
+        """
+        Makes the continuous 2D probability distribution over z_spec, z_phot
+
+        Parameters
+        ----------
+        vb: boolean
+            print progress to stdout?
+
+        Returns
+        -------
+
+        Notes
+        -----
+        Does not currently support variable sigmas, only one outlier population at a time
+        """
+        p_space = np.zeros((self.n_tot, self.n_tot))
+
[…]

### 2017-06-29 chippr/catalog_plots.py

@@ -35,7 +35,75 @@ def plot_true_histogram(true_samps, n_bins=50, plot_loc='', plot_name='true_hist
 
     return
 
-def plot_obs_scatter(true_zs, pfs, z_grid, plot_loc='', plot_name='scatter.png'):
+def plot_prob_space(z_grid, p_space, plot_loc='', plot_name='prob_space.png'):
+    """
+    Plots the 2D probability space of z_spec, z_phot
+
+    Parameters
+    ----------
+    p_space: numpy.ndarray, float
+        probabilities on the grid
+    z_grid: numpy.ndarray, float
+        fine grid of redshifts
+    plot_loc: string, optional
+        location in which to store plot
+    plot_name: string, optional
+        filename for plot
+    """
+    pu.set_up_plot()
+    f = plt.figure(figsize=(5, 5))
+    plt.subplot(1, 1, 1)
+    plt.pcolormesh(z_grid, z_grid, p_space.T, cmap='viridis')
+    plt.colorbar()
+    plt.xlabel(r'$z_{spec}$')
+    plt.ylabel(r'$z_{phot}$')
+    plt.axis([z_grid[0], z_grid[-1], z_grid[0], z_grid[-1]])
+    f.savefig(os.path.join(plot_loc, plot_name), bbox_inches='tight', pad_inches = 0, dpi=d.dpi)
+    return
+
+def plot_scatter(true_zs, obs_zs, pfs, z_grid, plot_loc='', plot_name='scatter.png'):
+    """
+    Plots a scatterplot of true and observed redshift values
+
+    Parameters
+    ----------
+    true_zs: numpy.ndarray, float
+        vector of true values of scalar input
+    obs_zs: numpy.ndarray, float
+        vector of observed values of scalar input
+    z_grid: numpy.ndarray, float
+        fine grid of redshifts
+    pfs: numpy.ndarray, float
+        matrix of posteriors evaluated on a fine grid
+    plot_loc: string, optional
+        location in which to store plot
+    plot_name: string, optional
+        filename for plot
+    """
+    n = len(obs_zs)
+    pu.set_up_plot()
+    f = plt.figure(figsize=(5, 5))
+    sps = f.add_subplot(1, 1, 1)
+    sps.scatter(true_zs, obs_zs, c='k', marker='.', s = 1., alpha=0.1)
[…]

### 2017-06-29 chippr/catalog.py
working code making posteriors for all cases, but slow and failing inference

@@ -13,6 +13,8 @@ from chippr import utils as u
 from chippr import sim_utils as su
 from chippr import gauss
 from chippr import discrete
+from chippr import multi_dist
+from chippr import gmix
 from chippr import catalog_plots as plots
 
 class catalog(object):
@@ -120,15 +122,15 @@ class catalog(object):
         self.cat: dict
             dictionary comprising catalog information
         """
+        self.N = 10**N
+        self.N_range = range(self.N)
+
         self.truth = truth
         # self.true_samps = self.truth.sample(10**N)
-        # if vb:
-        #     plots.plot_true_histogram(self.true_samps, plot_loc=self.plot_dir)
         # self.n_items = len(self.true_samps)
         # self.samp_range = range(self.n_items)
 
         self.proc_bins()
-        # self.obs_samps = self.sample_obs()
 
         self.prob_space = self.make_probs()
         if vb:
@@ -136,32 +138,34 @@ class catalog(object):
 
         ## this far!
         ## next, sample discrete to get z_true, z_obs
-        ## then literally take slices!
+        self.samps = self.sample(self.N)
+        if vb:
+            plots.plot_true_histogram(self.samps.T[0], plot_loc=self.plot_dir)
 
+        ## then literally take slices (evaluate at constant z_phot)
         self.obs_lfs = self.evaluate_lfs()
-        for n in self.samp_range:
+        for n in range(self.N):
             self.obs_lfs[n] /= np.sum(self.obs_lfs[n]) * self.dz_fine
 
-        # if vb:
-        #     plots.plot_scatter(self.true_samps, self.obs_samps, self.obs_lfs, self.z_fine, plot_loc=self.plot_dir)
-        #
+        if vb:
+            plots.plot_scatter(self.samps, self.obs_lfs, self.z_fine, plot_loc=self.plot_dir)
 
         self.int_pr = int_pr
         int_pr_fine = self.int_pr.evaluate(self.z_fine)
         int_pr_coarse = self.coarsify(int_pr_fine)
 
         # rewrite to take advantage of numpy array manipulation
-        pfs_fine = np.zeros((self.n_items, self.n_tot))
-        for n in self.samp_range:
+        pfs_fine = np.zeros((self.N, self.n_tot))
+        for n in self.N_range:
             pfs_fine[n] += int_pr_fine * self.obs_lfs[n]
             pfs_fine[n] /= np.sum(pfs_fine[n]) * self.dz_fine
-
-        if vb:
[…]

### 2017-06-29 chippr/catalog_plots.py

@@ -53,7 +53,9 @@ def plot_prob_space(z_grid, p_space, plot_loc='', plot_name='prob_space.png'):
     pu.set_up_plot()
     f = plt.figure(figsize=(5, 5))
     plt.subplot(1, 1, 1)
-    plt.pcolormesh(z_grid, z_grid, p_space.T, cmap='viridis')
+    # all_points = [[(z_grid[kk], z_grid[jj]) for kk in range(len(z_grid))] for jj in range(len(z_grid))]
+    all_vals = np.array([[p_space.evaluate_one(np.array([z_grid[jj], z_grid[kk]])) for jj in range(len(z_grid))] for kk in range(len(z_grid))])
+    plt.pcolormesh(z_grid, z_grid, all_vals, cmap='viridis')
     plt.colorbar()
     plt.xlabel(r'$z_{spec}$')
     plt.ylabel(r'$z_{phot}$')
@@ -61,16 +63,14 @@ def plot_prob_space(z_grid, p_space, plot_loc='', plot_name='prob_space.png'):
     f.savefig(os.path.join(plot_loc, plot_name), bbox_inches='tight', pad_inches = 0, dpi=d.dpi)
     return
 
-def plot_scatter(true_zs, obs_zs, pfs, z_grid, plot_loc='', plot_name='scatter.png'):
+def plot_scatter(zs, pfs, z_grid, plot_loc='', plot_name='scatter.png'):
     """
     Plots a scatterplot of true and observed redshift values
 
     Parameters
     ----------
-    true_zs: numpy.ndarray, float
-        vector of true values of scalar input
-    obs_zs: numpy.ndarray, float
-        vector of observed values of scalar input
+    zs: numpy.ndarray, float
+        matrix of spec, phot values
     z_grid: numpy.ndarray, float
         fine grid of redshifts
     pfs: numpy.ndarray, float
@@ -80,7 +80,10 @@ def plot_scatter(true_zs, obs_zs, pfs, z_grid, plot_loc='', plot_name='scatter.p
     plot_name: string, optional
         filename for plot
     """
-    n = len(obs_zs)
+    n = len(zs)
+    zs = zs.T
+    true_zs = zs[0]
+    obs_zs = zs[1]
     pu.set_up_plot()
     f = plt.figure(figsize=(5, 5))
     sps = f.add_subplot(1, 1, 1)
[…]

### 2017-06-29 chippr/gmix.py

@@ -26,9 +26,32 @@ class gmix(object):
         self.n_comps = len(self.amps)
 
         self.funcs = funcs#[chippr.gauss(self.means[c], self.sigmas[c]**2) for c in range(self.n_comps)]
+        # for c in range(self.n_comps):
+        #     print('gmix '+str((c, type(self.funcs[c]))))
 
+        self.dims = np.shape(np.array(limits).T)[0]
         self.min_x = limits[0]
         self.max_x = limits[1]
+
+    def evaluate_one(self, x):
+        """
+        Function to evaluate Gaussian mixture
+        once
+
+        Parameters
+        ----------
+        x: float
+            value at which to evaluate Gaussian mixture
+
+        Returns
+        -------
+        p: float
+            probability associated with x
+        """
+        p = 0.
+        for c in range(self.n_comps):
+            p += self.amps[c] * self.funcs[c].evaluate_one(x)
+        return p
 
     def evaluate(self, xs):
         """
@@ -44,7 +67,7 @@ class gmix(object):
         ps: ndarray, float
             values of Gaussian mixture probability distribution at xs
         """
-        ps = np.zeros_like(xs)
+        ps = np.zeros(len(xs))
         for c in range(self.n_comps):
             ps += self.amps[c] * self.funcs[c].evaluate(xs)
         return ps
@@ -59,8 +82,12 @@ class gmix(object):
             a single point sampled from the Gaussian mixture probability distribution
         """
 
-        x = -1.
-        while x < self.min_x or x > self.max_x:
+        x = -1. * np.ones(self.dims)
+        #don't do this every time!
+        min_x = self.min_x * np.ones(self.dims)
+        max_x = self.max_x * np.ones(self.dims)
+
+        while np.any(np.less(x, min_x)) or np.any(np.greater(x, max_x)):
             r = np.random.uniform(0., self.cumamps[-1])
             c = 0
             for k in range(1, self.n_comps):

### 2017-06-30 chippr/multi_dist.py
forgot the new distribution file

@@ -1,0 +1,97 @@
+import numpy as np
+import sys
+
+import chippr
+from chippr import defaults as d
+from chippr import utils as u
+
+class multi_dist(object):
+
+    def __init__(self, funcs):
+        """
+        A multidimensional freeform probability distribution object for
+        independent dimensions
+
+        Parameters
+        ----------
+        funcs: list, chippr.gauss or chippr.discrete or chippr.gmix objects
+            1D functions, one per dimension
+        """
+        self.funcs = funcs
+        self.dims = len(funcs)
+        # for d in range(self.dims):
+        #     print('multi dist '+str((d, type(self.funcs[d]))))
+
+    def evaluate_one(self, point):
+        """
+        Function to evaluate the probability at a point in multidimensional
+        space
+
+        Parameters
+        ----------
+        point: numpy.ndarray, float
+            coordinate at which to evaluate the probability
+
+        Returns
+        -------
+        prob: float
+            probability associated with point
+        """
+        prob = 1.
+        for d in range(self.dims):
+            prob *= self.funcs[d].evaluate_one(point[d])
+        return prob
+
+    def evaluate(self, points):
+        """
+        Function to evaluate the probability at a point in multidimensional
+        space
+
+        Parameters
+        ----------
+        points: numpy.ndarray, float
+            coordinates at which to evaluate the probability
+
+        Returns
+        -------
+        probs: float
+            probabilities associated with points
+        """
+        if len(np.shape(points.T)) == 1:
+            return self.evaluate_one(points)
+        probs = np.ones(len(points))
+        points = points.T
+        for d in range(self.dims):
+            probs *= self.funcs[d].evaluate(points[d])
+        return probs
+
+    def sample_one(self):
+        """
+        Function to take one sample from independent distributions
+
+        Returns
+        -------
+        point: numpy.ndarray, float
+            single sample from independent distributions
+        """
+        point = np.zeros(self.dims)
+        for d in range(self.dims):
+            point[d] = self.funcs[d].sample_one()
+        return point
+
+    def sample(self, n_samps):
+        """
+        Function to sample from multivariate Gaussian probability distribution
+
+        Parameters
+        ----------
+        n_samps: positive int
+            number of samples to take
+
+        Returns
+        -------
+        points: ndarray, float
+            array of n_samps samples from independent distributions
+        """
+        points = np.array([self.sample_one() for n in range(n_samps)])
+        return points

### 2017-07-03 chippr/catalog.py
preparing for test on broiler

@@ -273,8 +273,6 @@ class catalog(object):
             points = zip(self.z_fine, [self.samps[n][1]] * self.n_tot)
             lfs.append(self.prob_space.evaluate(np.array(points)))
         lfs = np.array(lfs)
-        if vb:
-            print('lf shape '+str(np.shape(lfs))+' should be n_gals * n_zs_fine')
         lfs /= np.sum(lfs, axis=-1)[:, np.newaxis] * self.dz_fine
         return lfs
 

### 2017-07-03 research/scripts/basic_inf_script.py

@@ -54,7 +54,7 @@ def set_up_prior(data, params):
     n_pdfs = len(log_z_posts)
 
     a = 1.# / n_bins
-    b = 30.#1. / z_difs ** 2
+    b = 20.#1. / z_difs ** 2
     c = a / n_pdfs
     prior_var = np.eye(n_bins)
     for k in range(n_bins):

### 2017-07-03 research/scripts/multi_inf_script.py

@@ -81,7 +81,7 @@ def set_up_prior(data, params):
     n_pdfs = len(log_z_posts)
 
     a = 1.# / n_bins
-    b = 30.#1. / z_difs ** 2
+    b = 20.#1. / z_difs ** 2
     c = a / n_pdfs
     prior_var = np.eye(n_bins)
     for k in range(n_bins):

### 2017-08-01 research/paper/draft-editing.tex

@@
 A photo-$z$ PDF is then an \textit{interim posterior} probability distribution of the redshift of a galaxy given its observed photometric data and the interim prior: $p(z_{i} | \vec{d}_{i}, \vec{\phi}^{*})$.
-We would like to learn about $p(z | \vec{\phi})$ from a catalog of photo-$z$ interim posteriors.
-
-\subsection{Model}
-\label{sec:model}
-
 We present the two attributes of a probabilistic graphical model (PGM) for the redshift distribution $n(z)$: the directed acyclic graph (DAG) and its mathematical interpretation in terms of Bayesian hierarchical inference.
@@
 \begin{figure}
-	\begin{center}
-		\includegraphics[width=0.5\textwidth]{pgm.png}
+	\begin{center}We would like to learn about $p(z | \vec{\phi})$ from a catalog of photo-$z$ interim posteriors.
+
+		\subsection{Model}
+		\label{sec:model}
+
+		\includegraphics[width=0.5\textwidth]{fig/pgm.png}
 		\caption{This directed acyclic graph corresponds to a probabilistic graphical model for a Bayesian hierarchical inference of $p(\vec{\phi}|\{\vec{d}_{i}\})$.
@@
 
-\appendix{}
+\appendix{Catalog Production}
+\label{app:catalog}
 %showing stacking is always too broad
+
+Sec.
+\ref{sec:mockdata} outlines one way to make mock data, but it's not necessarily a logically consistent way to do it, a problem identified in trying to use it for the test cases in Sec.
+\ref{sec:likelihoods}.
+If the goal is to emulate the traditional $z_{spec}$ vs.
+$z_{phot}$ scatterplots, we need to start from probability distributions in that space.
+This means there are two parameters $z_{spec}$ and $z_{phot}$ drawn from a two-dimensional distribution $p(z_{spec}, z_{phot} | \textul{\sigma}, \vec{\phi})$, where $\textul{\sigma}$ contains parameters concerning the relationship between $z_{spec}$ and $z_{phot}$, and $\vec{\phi}$ contains hyperparameters concerning $n(z)$.
+We know from basic probability that
+\begin{align}
+\label{eq:trivial}
+p(z_{spec}, z_{phot} | \textul{\sigma}, \vec{\phi}) =& p(z_{phot} | z_{spec}, \textul{\sigma}, \vec{\phi})\ p(z_{sped} | \textul{\sigma}, \vec{\phi}).
+\end{align}
+Because $z_{spec}$ depends only on $\vec{\phi}$ and $z_{phot}$ is independent of $\vec{\phi}$, this becomes
+\begin{align}
+\label{eq:simpler}
+p(z_{spec}, z_{phot} | \textul{\sigma}, \vec{\phi}) =& p(z_{phot} | z_{spec}, \textul{\sigma})\ p(z_{spec} | \vec{\phi}).
+\end{align}
+We know how to make reasonable choices for both terms on the righthand side of Eq. \ref{eq:simpler}.
+
+In the $z_{spec}$ vs.
+$z_{phot}$ scatterplot, $p(z_{phot} | z_{spec}, \textul{\sigma}, \vec{\phi})$ would be represented by vertical slices, but we are interested in the horizontal slices that represent $p(z_{spec} | z_{phot}, \textul{\sigma}, \vec{\phi})$.
+Now we must distinguish the true hyperparameters $\vec{\phi}_{true}$ from the interim hyperparameters $\vec{\phi}_{int}$.
+If the horizontal slices carry information about $\vec{\phi}$, how can we get a catalog of $p(z_{spec} | z_{phot}, \textul{\sigma}, \vec{\phi}_{int})$ that's consistent with $p(z_{spec} | z_{phot}, \textul{\sigma}, \vec{\phi}_{true})$?
+Somehow that information needs to be propagated through in order to recover it.
+
+We achieve this goal by dividing out the true $n(z)$ and multiplying by the interim $n(z)$ as in
+\begin{align}
+\label{eq:transformation}
+p(z_{spec} | z_{phot}, \textul{\sigma}, \vec{\phi}_{int}) =& p(z_{spec} | z_{phot}, \textul{\sigma}, \vec{\phi}_{true})\ \frac{p(z_{spec} | \vec{\phi}_{int})}{p(z_{spec} | \vec{\phi}_{true})},
+\end{align}
+since $p(z_{spec} | z_{phot}, \textul{\sigma}, \vec{\phi})$ is separable and the term $p(z_{spec} | z_{phot}, \textul{\sigma})$ changes independently of the $p(z_{spec} | \vec{\phi})$ term.
+The information about $\vec{\phi}_{true}$ is retained by the elements of the catalog: only $(z_{spec}, z_{phot})$ pairs drawn from the true distribution will be represented in the catalog!
+
+It appears as if this is the only fully self-consistent way to ensure that a draw from the posterior distribution $p(z_{spec} | z_{phot}, \textul{\sigma}, \vec{\phi})$ does in fact follow that distribution.
+In the cases with fiducial posteriors $p(z_{spec} | z_{phot}, \sigma, \vec{\phi})$, the method of Sec.
+\ref{sec:mockdata} was equivalent to this, but Sec.
+\ref{sec:likelihoods} requires a more sophisticated approach.
 

### 2017-08-01 research/scripts/scaling.py
starting the scaling script

@@ -1,0 +1,365 @@
+def set_params(n_gals):
+    """
+    The parameters of all these cases will be the same (and vanilla), except for the number of galaxies being different
+    """
+    params = {}
+
+    params['smooth_truth'] = 0
+    params['interim_prior'] = 'flat'
+    params['n_galaxies'] = n_gals
+    params['n_bins'] = 25
+    params['bin_min'] = 0.
+    params['bin_max'] = 1.
+    params['variable_sigmas'] = 0
+    params['constant_sigma'] = 0.03
+    params['catastrophic_outliers'] = '0'
+    params['gr_threshold'] = 1.1
+    params['n_accepted'] = 10000
+    params['n_burned'] = 1000
+    params['n_walkers'] = 100
+
+    return params
+
+def make_true(given_key):
+    """
+    Function to create true redshift distribution to be shared among several
+    test cases
+
+    Parameters
+    ----------
+    given_key: string
+        name of test case for which true n(z) is to be made
+
+    Returns
+    -------
+    true_nz: chippr.gmix object
+        gaussian mixture probability distribution
+
+    Notes
+    -----
+    test_name is currently ignored but will soon be used to load parameters for
+    making true_nz instead of hardcoded values.
+    """
+    test_info = all_tests[given_key]
+
+    if test_info['params']['smooth_truth'] == 1:
+        true_amps = np.array([0.150,0.822,1.837,2.815,3.909,
+                              5.116,6.065,6.477,6.834,7.304,
+                              7.068,6.771,6.587,6.089,5.165,
+                              4.729,4.228,3.664,3.078,2.604,
+                              2.130,1.683,1.348,0.977,0.703,
+                              0.521,0.339,0.283,0.187,0.141,
+                              0.104,0.081,0.055,0.043,0.034])
+        true_grid = np.linspace(test_info['bin_ends'][0], test_info['bin_ends'][-1], len(true_amps) + 1)
+        true_grid_mids = (true_grid[1:] + true_grid[:-1]) / 2.
+        f = spi.interp1d(true_grid_mids, true_amps)
+        bin_mids = (test_info['bin_ends'][1:] + test_info['bin_ends'][:-1]) / 2.
+        bin_difs = test_info['bin_ends'][1:] - test_info['bin_ends'][:-1]
+        true_amps = f(bin_mids)
+        true_means = bin_mids
[…]

### 2017-08-01 research/scripts/scaling.py
finished first go at scaling script for nersc

@@ -1,26 +1,4 @@
-def set_params(n_gals):
-    """
-    The parameters of all these cases will be the same (and vanilla), except for the number of galaxies being different
-    """
-    params = {}
-
-    params['smooth_truth'] = 0
-    params['interim_prior'] = 'flat'
-    params['n_galaxies'] = n_gals
-    params['n_bins'] = 25
-    params['bin_min'] = 0.
-    params['bin_max'] = 1.
-    params['variable_sigmas'] = 0
-    params['constant_sigma'] = 0.03
-    params['catastrophic_outliers'] = '0'
-    params['gr_threshold'] = 1.1
-    params['n_accepted'] = 10000
-    params['n_burned'] = 1000
-    params['n_walkers'] = 100
-
-    return params
-
-def make_true(given_key):
+def set_shared_params():
     """
     Function to create true redshift distribution to be shared among several
     test cases
@@ -40,103 +18,60 @@ def make_true(given_key):
     test_name is currently ignored but will soon be used to load parameters for
     making true_nz instead of hardcoded values.
     """
-    test_info = all_tests[given_key]
+    params = {}
 
-    if test_info['params']['smooth_truth'] == 1:
-        true_amps = np.array([0.150,0.822,1.837,2.815,3.909,
-                              5.116,6.065,6.477,6.834,7.304,
-                              7.068,6.771,6.587,6.089,5.165,
-                              4.729,4.228,3.664,3.078,2.604,
-                              2.130,1.683,1.348,0.977,0.703,
-                              0.521,0.339,0.283,0.187,0.141,
-                              0.104,0.081,0.055,0.043,0.034])
-        true_grid = np.linspace(test_info['bin_ends'][0], test_info['bin_ends'][-1], len(true_amps) + 1)
-        true_grid_mids = (true_grid[1:] + true_grid[:-1]) / 2.
-        f = spi.interp1d(true_grid_mids, true_amps)
-        bin_mids = (test_info['bin_ends'][1:] + test_info['bin_ends'][:-1]) / 2.
-        bin_difs = test_info['bin_ends'][1:] - test_info['bin_ends'][:-1]
-        true_amps = f(bin_mids)
-        true_means = bin_mids
-        true_sigmas = bin_difs
-    else:
-        bin_range = max(test_info['bin_ends']) - min(test_info['bin_ends'])
-        true_amps = np.array([0.20, 0.35, 0.55])
[…]

### 2017-08-01 research/scripts/scaling.py
finished first go at scaling script for nersc

@@ -217,7 +217,7 @@ def do_inference(params):
     initial_values = prior.sample(n_ivals)
     log_z_dens_plots.plot_ivals(initial_values, nz.info, nz.plot_dir)
 
-    nz_samps = nz.calculate_samples(initial_values, no_data=params['no_data'], no_prior=params['no_prior'])
+    # nz_samps = nz.calculate_samples(initial_values, no_data=params['no_data'], no_prior=params['no_prior'])
 
     nz_stats = nz.compare()
     nz.plot_estimators()
@@ -231,6 +231,7 @@ def one_scale(n_gals):
     """
     params = set_unique_params(n_gals)
     make_catalog(params)
+    print('STARTED ' + str(n_gals))
     do_inference(params)
     print('FINISHED ' + str(n_gals))
     return

### 2017-08-01 chippr/defaults.py
partial fix for bug in reading parameters from dict as opposed to file

@@ -37,9 +37,13 @@ def check_sim_params(params={}):
     params: dict
         dictionary containing final key/value pairs for simulation of catalog
     """
+    print('checking sim params')
     params = check_basic_setup(params)
+    print('basic setup params ok')
     params = check_variable_sigmas(params)
+    print('variable sigma params ok')
     params = check_catastrophic_outliers(params)
+    print('catastrophic outlier params ok')
     return params
 
 def check_basic_setup(params):
@@ -55,22 +59,26 @@ def check_basic_setup(params):
     -------
     params: dict
         dictionary containing key/value pairs for simulation
+
+    Notes
+    -----
+    Currently, the number of galaxies is really log10 the number of galaxies.  This will one day be changed to permit arbitrary numbers of galaxies!
     """
     if 'n_gals' not in params:
         params['n_gals'] = n_gals
-    else:
+    elif params['raw'] == 1:
         params['n_gals'] = int(params['n_gals'][0])
     if 'n_bins' not in params:
         params['n_bins'] = n_bins
-    else:
+    elif params['raw'] == 1:
         params['n_bins'] = int(params['n_bins'][0])
     if 'bin_min' not in params:
         params['bin_min'] = min_x
-    else:
+    elif params['raw'] == 1:
         params['bin_min'] = float(params['bin_min'][0])
     if 'bin_max' not in params:
         params['bin_max'] = max_x
-    else:
+    elif params['raw'] == 1:
         params['bin_max'] = float(params['bin_max'][0])
     return params
 
@@ -90,12 +98,12 @@ def check_variable_sigmas(params):
     """
     if 'variable_sigmas' not in params:
         params['variable_sigmas'] = 0
-    else:
+    elif params['raw'] == 1:
         params['variable_sigmas'] = int(params['variable_sigmas'][0])
     if not params['variable_sigmas']:
         if 'constant_sigma' not in params:
             params['constant_sigma'] = constant_sigma
-        else:
+        elif params['raw'] == 1:
             params['constant_sigma'] = float(params['constant_sigma'][0])
     return params
 
@@ -116,13 +124,13 @@ def check_catastrophic_outliers(params):
     """
     if 'catastrophic_outliers' not in params:
         params['catastrophic_outliers'] = '0'
-    else:
+    elif params['raw'] == 1:
         params['catastrophic_outliers'] = str(params['catastrophic_outliers'][0])
[…]

### 2017-08-01 chippr/utils.py

@@ -52,6 +52,8 @@ def ingest(in_info):
         with open(in_info) as infile:
             lines = (line.split(None) for line in infile)
             in_dict = {defn[0] : defn[1:] for defn in lines}
+        in_dict['raw'] = 1
     else:
         in_dict = in_info
+        in_dict['raw'] = 0
     return in_dict

### 2017-08-01 research/scripts/scaling.py
added profiling

@@ -18,21 +18,27 @@ def set_shared_params():
     test_name is currently ignored but will soon be used to load parameters for
     making true_nz instead of hardcoded values.
     """
-    params = {}
-
-    params['smooth_truth'] = 0
-    params['interim_prior'] = 'flat'
-    params['variable_sigmas'] = 0
-    params['constant_sigma'] = 0.03
-    params['catastrophic_outliers'] = '0'
-    params['gr_threshold'] = 1.1
-    params['n_accepted'] = 10000
-    params['n_burned'] = 1000
-    params['n_walkers'] = 100
-
-    params['n_bins'] = 25
-    params['bin_min'] = 0.
-    params['bin_max'] = 1.
+    # params = {}
+    #
+    # params['n_bins'] = 25
+    # params['bin_min'] = 0.
+    # params['bin_max'] = 1.
+    #
+    # params['smooth_truth'] = 0
+    # params['interim_prior'] = 'flat'
+    # params['variable_sigmas'] = 0
+    # params['constant_sigma'] = 0.03
+    # params['catastrophic_outliers'] = '0'
+    # params['outlier_fraction'] = 0.
+    #
+    # params['gr_threshold'] = 1.1
+    # params['n_accepted'] = 10000
+    # params['n_burned'] = 1000
+    # params['n_walkers'] = 100
+    paramfile = 'scaling.txt'
+    params = chippr.utils.ingest(paramfile)
+    params = chippr.defaults.check_sim_params(params)
+    params = chippr.defaults.check_inf_params(params)
 
     bin_range = params['bin_max'] - params['bin_min']
     true_amps = np.array([0.20, 0.35, 0.55])
@@ -66,8 +72,14 @@ def set_unique_params(n_gals):
     The parameters of all these cases will be the same (and vanilla), except for the number of galaxies being different
     """
     params = start_params
-    params['n_galaxies'] = n_gals
-    params['name'] = str(n_gals)
+    params['n_gals'] = n_gals
+    params['name'] = str(10**n_gals)
+
+    test_dir = os.path.join(result_dir, params['name'])
+    params['dir'] = test_dir
+    if os.path.exists(test_dir):
+        shutil.rmtree(test_dir)
+    os.makedirs(test_dir)
 
     return params
 
@@ -83,12 +95,7 @@ def make_catalog(params):
     test_info = {}
     test_info['params'] = params
     test_info['name'] = params['name']
-
[…]

### 2017-08-01 research/scripts/scaling.py
uncommented things to run on nersc

@@ -225,7 +225,7 @@ def do_inference(params):
     initial_values = prior.sample(n_ivals)
     log_z_dens_plots.plot_ivals(initial_values, nz.info, nz.plot_dir)
 
-    # nz_samps = nz.calculate_samples(initial_values, no_data=params['no_data'], no_prior=params['no_prior'])
+    nz_samps = nz.calculate_samples(initial_values, no_data=params['no_data'], no_prior=params['no_prior'])
 
     nz_stats = nz.compare()
     nz.plot_estimators()
@@ -283,7 +283,7 @@ if __name__ == "__main__":
     from chippr import *
 
     result_dir = os.path.join('..', 'results/scaling')
-    catalog_sizes = [2]#, 3, 4, 5]
+    catalog_sizes = [2, 3, 4, 5]
 
     start_params = set_shared_params()
     start_params['raw'] = 0

### 2017-08-01 chippr/defaults.py
removing print statements

@@ -37,13 +37,13 @@ def check_sim_params(params={}):
     params: dict
         dictionary containing final key/value pairs for simulation of catalog
     """
-    print('checking sim params')
+    # print('checking sim params')
     params = check_basic_setup(params)
-    print('basic setup params ok')
+    # print('basic setup params ok')
     params = check_variable_sigmas(params)
-    print('variable sigma params ok')
+    # print('variable sigma params ok')
     params = check_catastrophic_outliers(params)
-    print('catastrophic outlier params ok')
+    # print('catastrophic outlier params ok')
     return params
 
 def check_basic_setup(params):

### 2017-08-01 research/scripts/scaling.py

@@ -225,7 +225,7 @@ def do_inference(params):
     initial_values = prior.sample(n_ivals)
     log_z_dens_plots.plot_ivals(initial_values, nz.info, nz.plot_dir)
 
-    nz_samps = nz.calculate_samples(initial_values, no_data=params['no_data'], no_prior=params['no_prior'])
+    # nz_samps = nz.calculate_samples(initial_values, no_data=params['no_data'], no_prior=params['no_prior'])
 
     nz_stats = nz.compare()
     nz.plot_estimators()
@@ -242,6 +242,7 @@ def one_scale(n_gals):
     print('SET PARAMS ' + str(params['name']))
 
     sim_profile = os.path.join(params['dir'], 'sim_profile.txt')
+    sys.stdout = open(sim_profile, 'w')
     pr = cProfile.Profile()
     pr.enable()
     make_catalog(params)
@@ -249,11 +250,12 @@ def one_scale(n_gals):
     s = StringIO.StringIO()
     sortby = 'cumtime'
     ps = pstats.Stats(pr, stream=s).sort_stats(sortby)
-    with open(sim_profile, 'w') as profiler:
-        profiler.write(ps.print_stats())
+    ps.print_stats()
+    sys.stdout = sys.__stdout__
     print('MADE CATALOG ' + str(params['name']))
 
     inf_profile = os.path.join(params['dir'], 'inf_profile.txt')
+    sys.stdout = open(inf_profile, 'w')
     pr = cProfile.Profile()
     pr.enable()
     do_inference(params)
@@ -261,8 +263,10 @@ def one_scale(n_gals):
     s = StringIO.StringIO()
     sortby = 'cumtime'
     ps = pstats.Stats(pr, stream=s).sort_stats(sortby)
-    with open(inf_profile, 'w') as profiler:
-        profiler.write(ps.print_stats())
+    ps.print_stats()
+    # with open(inf_profile, 'w') as profiler:
+    #     profiler.write(str(ps.print_stats()))
+    sys.stdout = sys.__stdout__
     print('FINISHED INFERENCE ' + str(params['name']))
 
     return
@@ -272,6 +276,7 @@ if __name__ == "__main__":
     import numpy as np
     import pickle
     import os
+    import sys
     import shutil
     import cProfile
     import StringIO
@@ -283,7 +288,7 @@ if __name__ == "__main__":
     from chippr import *
 
[…]

### 2017-08-01 research/scripts/scaling.py
profiling actually works now

@@ -225,7 +225,7 @@ def do_inference(params):
     initial_values = prior.sample(n_ivals)
     log_z_dens_plots.plot_ivals(initial_values, nz.info, nz.plot_dir)
 
-    # nz_samps = nz.calculate_samples(initial_values, no_data=params['no_data'], no_prior=params['no_prior'])
+    nz_samps = nz.calculate_samples(initial_values, no_data=params['no_data'], no_prior=params['no_prior'])
 
     nz_stats = nz.compare()
     nz.plot_estimators()
@@ -242,7 +242,6 @@ def one_scale(n_gals):
     print('SET PARAMS ' + str(params['name']))
 
     sim_profile = os.path.join(params['dir'], 'sim_profile.txt')
-    sys.stdout = open(sim_profile, 'w')
     pr = cProfile.Profile()
     pr.enable()
     make_catalog(params)
@@ -251,11 +250,11 @@ def one_scale(n_gals):
     sortby = 'cumtime'
     ps = pstats.Stats(pr, stream=s).sort_stats(sortby)
     ps.print_stats()
-    sys.stdout = sys.__stdout__
+    with open(sim_profile, 'w') as profiler:
+        profiler.write(s.getvalue())
     print('MADE CATALOG ' + str(params['name']))
 
     inf_profile = os.path.join(params['dir'], 'inf_profile.txt')
-    sys.stdout = open(inf_profile, 'w')
     pr = cProfile.Profile()
     pr.enable()
     do_inference(params)
@@ -264,9 +263,8 @@ def one_scale(n_gals):
     sortby = 'cumtime'
     ps = pstats.Stats(pr, stream=s).sort_stats(sortby)
     ps.print_stats()
-    # with open(inf_profile, 'w') as profiler:
-    #     profiler.write(str(ps.print_stats()))
-    sys.stdout = sys.__stdout__
+    with open(inf_profile, 'w') as profiler:
+        profiler.write(s.getvalue())
     print('FINISHED INFERENCE ' + str(params['name']))
 
     return
@@ -276,7 +274,7 @@ if __name__ == "__main__":
     import numpy as np
     import pickle
     import os
-    import sys
+    # import sys
     import shutil
     import cProfile
     import StringIO
@@ -288,7 +286,7 @@ if __name__ == "__main__":
     from chippr import *
 
     result_dir = os.path.join('..', 'results/scaling')
-    catalog_sizes = [2]#, 3, 4, 5]
[…]

### 2017-08-01 chippr/catalog.py
trying to fix bug on nersc

@@ -37,8 +37,11 @@ class catalog(object):
             self.params = u.ingest(params)
         else:
             self.params = params
+            self.params['raw'] = 0
+        print('ingested params')
 
         self.params = d.check_sim_params(self.params)
+        print('params okay')
 
         if vb:
             print self.params

### 2017-08-01 chippr/catalog.py
removing comments

@@ -38,10 +38,8 @@ class catalog(object):
         else:
             self.params = params
             self.params['raw'] = 0
-        print('ingested params')
 
         self.params = d.check_sim_params(self.params)
-        print('params okay')
 
         if vb:
             print self.params

### 2017-08-01 research/paper/draft.tex
some figures updated
added mention of scaling investigation

@@
 of the redshift of a galaxy given its observed photometric data and the interim 
-prior: $p(z_{i} | \vec{d}_{i}, \vec{\phi}^{*})$.  We would like to learn about 
-$p(z | \vec{\phi})$ from a catalog of photo-$z$ interim posteriors.
-
-\subsection{Model}
-\label{sec:model}
-
+prior: $p(z_{i} | \vec{d}_{i}, \vec{\phi}^{*})$.  
 We present the two attributes of a probabilistic graphical model (PGM) for the 
@@
 \begin{figure}
-	\begin{center}
-		\includegraphics[width=0.5\textwidth]{pgm.png}
+	\begin{center}We would like to learn about $p(z | \vec{\phi})$ from a 
+catalog of photo-$z$ interim posteriors.
+
+		\subsection{Model}
+		\label{sec:model}
+
+		\includegraphics[width=0.5\textwidth]{fig/pgm.png}
 		\caption{This directed acyclic graph corresponds to a 
@@
 
+The tests presented here are conducted with samples of $N = 10^{4}$ with
+
 \begin{figure}
@@
 in future work.
+
+In comparison with use of the stacked estimator of the redshift distribution 
+function, the hierarchical approach is computationally expensive.  In a future 
+revision, the \chippr code will be optimized for speed and resource efficiency. 
+ At this stage, we establish scaling behavior to project the expense in terms 
+of computational resources.
 
@@
 z_{phot}, \textul{\sigma}, \vec{\phi}_{true})\ \frac{p(z_{spec} | 
-\vec{\phi}_{int})}{p(z_{spec} | \vec{\phi}_{true}},
+\vec{\phi}_{int})}{p(z_{spec} | \vec{\phi}_{true})},
 \end{align}

### 2017-09-07 research/paper/draft.tex
updating text for new tests

--- 2017-06-28 research/paper/draft.tex
@@
-\documentclass[iop]{emulateapj}
+\documentclass[12pt, onecolumn]{emulateapj}
 
@@
 \newcommand{\textul}{\underline}
-\newcommand{\chippr}{\texttt{CHIPPR} }
+\newcommand{\chippr}{\texttt{chippr}}
 
@@
 Heidelberg, Germany}
-\altaffiltext{5}{[SLAC]}
+\altaffiltext{5}{SLAC National Accelerator Laboratory, Menlo Park, CA 94025, 
+USA}
 
@@
 with Probabilistic Photometric Redshifts (\chippr) code implementing this 
-technique, as well as its validation on mock data and testing on the 
-\texttt{Buzzard} simulations.  \chippr yields a more accurate characterization 
-of $n(z)$ containing information beyond the best-fit estimator produced by 
-traditional procedures.  The publicly available code is easily extensible to 
-other one-point statistics that depend on redshift.
+technique, as well as its validation on mock data and testing on realistic mock 
+data.  \chippr\ yields a more accurate characterization of $n(z)$ containing 
+information beyond the best-fit estimator produced by traditional procedures.  
+The publicly available code is easily extensible to other one-point statistics 
+that depend on redshift.
 
@@
 
-After a brief literature review addressing how photo-$z$ PDFs are currently 
-used in cosmology, this paper aims to answer the following questions:
+\textbf{Obviously I need to write the lit. review!}
+
+This paper aims to answer the following questions:
 
@@
 inference?
-	\item How does the result of \chippr compare to established estimators 
+	\item How does the result of \chippr\ compare to established estimators 
 in terms of the accuracy of $n(z)$?
@@
 
-The publicly available \chippr code implements the probabilistic graphical 
+The publicly available \chippr\ code implements the probabilistic graphical 
 model presented in Sec. \ref{sec:model}.
@@
 In addition to the choices and assumptions underlying the probabilistic 
-graphical model, the implementation of \chippr makes choices and assumptions of 
-its own.
+graphical model, the implementation of \chippr\ makes choices and assumptions 
+of its own.
 
 \begin{enumerate}
-	\item \chippr is only applicable if the redshift interim posteriors and 
-interim prior are accurate.
-	\item \chippr currently only accepts photo-$z$ PDFs and produces $n(z)$ 
-samples of a single format, that of the piecewise constant parametrization, 
-also referred to as a binned histogram parametrization and a sum of top hat 
-functions.
+	\item \chippr\ is only applicable if the redshift interim posteriors 
+and interim prior are accurate.
+	\item \chippr\ currently only accepts photo-$z$ PDFs and produces 
+$n(z)$ samples of a single format, that of the piecewise constant 
+parametrization, also referred to as a binned histogram parametrization and a 
+sum of top hat functions.
 \end{enumerate}
@@
 
-We demonstrate the superiority of \chippr over alternative approaches in a 
+We demonstrate the superiority of \chippr\ over alternative approaches in a 
 number of compelling test cases on mock data.  Each experiment is characterized 
@@
 priors are perfectly understood but also deter readers from assuming that 
-\chippr has any preference over the method by which photo-$z$ interim 
+\chippr\ has any preference over the method by which photo-$z$ interim 
 posteriors are derived from photometric data.
@@
 \label{sec:mockdata}
+
+\textbf{This needs reorganization and clarification.}
 
@@
 function with known parameters $\vec{\phi}'$.
-	\item \label{itm:sample} Sample the true $n(z)$ to create a catalog of 
-$N$ true redshifts $z'_{i}$.
-	\item \label{itm:scatter} Choose a true intrinsic scatter parameter 
-$\sigma$, and sample Gaussians $p(z''_{i} | z'_{i}, \sigma) = 
-\mathcal{N}(z'_{i}, \sigma)$ to get "observed" redshifts $z'_{i}$ defining 
-Gaussian likelihoods $p(\vec{d}_{i} | z''_{i}, \sigma) = \mathcal{N}(z''_{i}, 
-\sigma)$.
-	\item \label{itm:interim} Choose a parametrization and the parameters 
-$\vec{\phi}^{*}$ of the interim prior $p(z_{i} | \vec{\phi}^{*})$.
-	\item \label{itm:posterior} Create interim posteriors $\log[p(z_{i} | 
-\vec{d}_{i}, \vec{\phi}^{*})] = \log[p(\vec{d}_{i} | z'_{i})]\ \log[p(z_{i} | 
-\vec{\phi}^{*})]$ in this parametrization.
+	\item \label{itm:consistent}  Choose an empirical distribution 
+$p(z_{spec}, z_{phot} | \textul{\sigma}, \vec{\phi}')$ emulating the error 
+model with parameters $\textul{\sigma}$ of photo-$z$s $z_{phot}$ and 
+spectroscopic redshifts $z_{spec}$.
+	\item \label{itm:sample} Sample $N$ pairs $(z_{phot}, z_{spec})$ from 
+$p(z_{spec}, z_{phot} | \textul{\sigma}, \vec{\phi}')$.
+	\item \label{itm:interim} Choose an interim $n(z)$ with known 
+parameters $\vec{\phi}^{*}$ and construct its empirical distribution 
+$p(z_{spec}, z_{phot} | \textul{\sigma}, \vec{\phi}^{*})$.
+	\item \label{itm:propagate} Follow Eq. \ref{eq:transformation} to 
+create the interim empirical distribution.
+	\item \label{itm:posterior} Create interim posteriors $p(z | z_{phot}, 
+\textul{\sigma}, \vec{\phi}^{*})$ by evaluating the interim empirical 
+distribution at values of $z_{phot}$ drawn in It. \ref{itm:sample}.
 \end{enumerate}
+
+With a goal of emulating the systematics of traditional $z_{spec}$ vs. 
+$z_{phot}$ scatterplots, we start from probability distributions in that space. 
+ This means there are two parameters $z_{spec}$ and $z_{phot}$ drawn from a 
+two-dimensional distribution $p(z_{spec}, z_{phot} | \textul{\sigma}, 
+\vec{\phi})$, where $\textul{\sigma}$ contains parameters concerning the 
+relationship between $z_{spec}$ and $z_{phot}$, and $\vec{\phi}$ contains 
+hyperparameters concerning $n(z)$.  We know from basic probability that
+\begin{align}
+\label{eq:trivial}
+p(z_{spec}, z_{phot} | \textul{\sigma}, \vec{\phi}) =& p(z_{phot} | z_{spec}, 
+\textul{\sigma}, \vec{\phi})\ p(z_{spec} | \textul{\sigma}, \vec{\phi}).
+\end{align}
+Because $z_{spec}$ depends only on $\vec{\phi}$ and $z_{phot}$ is independent 
+of $\vec{\phi}$, this becomes
+\begin{align}
+\label{eq:simpler}
+p(z_{spec}, z_{phot} | \textul{\sigma}, \vec{\phi}) =& p(z_{phot} | z_{spec}, 
+\textul{\sigma})\ p(z_{spec} | \vec{\phi}).
+\end{align}
+We know how to make reasonable choices for both terms on the righthand side of 
+Eq. \ref{eq:simpler} and note that this is equivalent to using $z_{phot}$ as a 
+proxy for photometric data and thinking of $z_{spec}$ as the $z$ we seek to 
+characterize.
+
+In the $z_{spec}$ vs. $z_{phot}$ scatterplot, $p(z_{phot} | z_{spec}, 
+\textul{\sigma}, \vec{\phi})$ would be represented by vertical slices, but we 
+are interested in the horizontal slices that represent $p(z_{spec} | z_{phot}, 
+\textul{\sigma}, \vec{\phi})$.  Now we must distinguish the true 
+hyperparameters $\vec{\phi}'$ we wish to estimate from the interim 
+hyperparameters $\vec{\phi}^{*}$.  If the horizontal slices carry information 
+about $\vec{\phi}$, we would like to get a catalog of $p(z_{spec} | z_{phot}, 
+\textul{\sigma}, \vec{\phi}^{*})$ that's consistent with $p(z_{spec} | 
+z_{phot}, \textul{\sigma}, \vec{\phi}')$.
+
+We achieve the goal of propagating that information by dividing out the true 
+$n(z)$ and multiplying by the interim $n(z)$ as in
+\begin{align}
+\label{eq:transformation}
+p(z_{spec} | z_{phot}, \textul{\sigma}, \vec{\phi}^{*}) =& p(z_{spec} | 
+z_{phot}, \textul{\sigma}, \vec{\phi}')\ \frac{p(z_{spec} | 
+\vec{\phi}^{*})}{p(z_{spec} | \vec{\phi}')},
+\end{align}
+since $p(z_{spec} | z_{phot}, \textul{\sigma}, \vec{\phi})$ is separable and 
+the term $p(z_{spec} | z_{phot}, \textul{\sigma})$ changes independently of the 
+$p(z_{spec} | \vec{\phi})$ term.  The information about $\vec{\phi}'$ is 
+retained by the elements of the catalog: only $(z_{spec}, z_{phot})$ pairs 
+drawn from the true distribution will be represented in the catalog.
+
+This is the only fully self-consistent way to ensure that a draw from the 
+posterior distribution $p(z_{spec} | z_{phot}, \textul{\sigma}, \vec{\phi})$ 
+does in fact follow that distribution.
 
@@
 
-The Kullback-Leibler divergence is our primary measure of the accuracy of 
-estimators of $n(z)$ in cases of mock data with known true redshifts.
+\textbf{This needs explication.}
 
-\textbf{Review precision and bias from \texttt{kld.ipynb} and interpret in 
-terms of a \% difference.}
+The Kullback-Leibler divergence (KLD) is our primary measure of the accuracy of 
+estimators of $n(z)$ in cases of mock data with known true redshifts.  The KLD
+\begin{align}
+\label{eq:kld}
+\mathrm{KLD} &= \int_{-\infty}^{\infty}\ n(z)\ 
+\log\left[\frac{n(z)}{\hat{n}(z)}\right]\ dz
+\end{align}
+measures the loss of information when using an estimator of a probability 
+distribution instead of the true probability distribution, where the 
+probability distribution in question is, in our case, the redshift distribution 
+function.
+
+\subsubsection{Cosmological Parameter Covariance}
+\label{sec:cosmoparams}
+
+We also perform a test with mock data derived from a self-consistent cosmology 
+likelihood analysis using 
+\texttt{CosmoLike}\footnote{https://github.com/CosmoLike} 
+\citep{krause_cosmolike_2017}.
 
@@
 regardless of the appropriateness of the interim prior as an approximation to 
-the true $n(z)$, \chippr is not affected by the choice of the interim prior so 
+the true $n(z)$, \chippr\ is not affected by the choice of the interim prior so 
 long as it has nontrivial coverage in the space of redshift.  These tests 
@@
 To show how the choice of $n(z)$ estimator propagates to cosmological 
-constraints, we apply \chippr to a data from a realistic cosmological 
-simulation (probably \texttt{Buzzard}) with photo-$z$ PDFs produced by a 
-popular method (probably \texttt{BPZ}).
+constraints, we apply \chippr\ to a data from a realistic cosmological 
+simulation.
 
@@
 
-As in Sec. \ref{sec:validintro}, the mock data takes the form of photo-$z$ 
-interim posteriors, but that is where the similarity ends.  These photo-$z$ 
-interim posteriors are derived from the photometry resulting from the 
-\texttt{Buzzard} simulation by way of \texttt{BPZ}.  Because the simulation 
-begins with setting true values of the cosmological parameters, we can 
-propagate the different estimators of $n(z)$ through a forecasting code (which 
-one?) to generate error ellipses on the cosmological parameters.
+\textbf{I need to write this section!}
+
+A cosmological analysis is conducted with \texttt{CosmoLike}.
 
 \subsubsection{Mock Data}
-\label{sec:buzzard}
+\label{sec:realistic}
 
-We summarize the details of the \texttt{Buzzard} simulation here.
+\textbf{This test will combine several systematics discussed above.}
+
+We summarize the details of the realistic data here.
 
@@
 
-Because the realistic mock data of Sec. \ref{sec:buzzard} is associated with 
+Because the realistic mock data of Sec. \ref{sec:realistic} is associated with 
 true values of the cosmological parameters, we may compare the quality of 
@@
 	\item In addition to coming with its own error distribution, the $n(z)$ 
-estimator produced by \chippr is quantifiably more accurate than established 
+estimator produced by \chippr\ is quantifiably more accurate than established 
 estimators.
-	\item Propagation of the \chippr result leads to a quantifiable 
+	\item Propagation of the \chippr\ result leads to a quantifiable 
 improvement in the constraints on cosmological parameters.
@@
 estimator and reductions of photo-$z$ PDFs to redshift point estimates in 
-obtaining estimators of $n(z)$.  \chippr is freely available to the community 
+obtaining estimators of $n(z)$.  \chippr\ is freely available to the community 
 for incorporation into evolving data analysis pipelines.
-
-\appendix{}
 

### 2017-09-11 research/paper/draft.tex
small updates

@@
-\documentclass[12pt, onecolumn]{emulateapj}
+\documentclass[iop]{emulateapj}
 
@@
 
-\textbf{Obviously I need to write the lit. review!}
-
-This paper aims to answer the following questions:
+After a brief review of the literature, this paper aims to answer the following 
+questions:
 
@@
 
-\textbf{This needs reorganization and clarification.}
-
 The mock data used here are produced by the following steps.
@@
 \label{sec:accuracy}
-
-\textbf{This needs explication.}
 
@@
 
-\textbf{I need to write this section!}
-
 A cosmological analysis is conducted with \texttt{CosmoLike}.
@@
 
-\textbf{This test will combine several systematics discussed above.}
-
-We summarize the details of the realistic data here.
+The realistic data combines several of the systematic effects of Sec. 
+\ref{sec:mockdata}.
 
@@
 
+\bibliographystyle{apj}
+\bibliography{references}
+
 \end{document}

### 2017-09-11 chippr/log_z_dens_plots.py

@@ -17,15 +17,15 @@ from chippr import stat_utils as s
 
 lnz, nz = '', ''#r'$\ln[n(z)]$', r'$n(z)$'
 
-s_tru, w_tru, a_tru, c_tru, d_tru, l_tru = '-', 2., 1., 'k', [(0, (1, 1))], 'Truth '
-s_tbp, w_tbp, a_tbp, c_tbp, d_tbp, l_tbp = ':', 2., 0.75, 'k', [(0, (1, 1))], 'Binned True '
-s_int, w_int, a_int, c_int, d_int, l_int = '-', 2., 0.5, 'k', [(0, (1, 1))], 'Interim '
-s_stk, w_stk, a_stk, c_stk, d_stk, l_stk = '--', 2.5, 1., 'g', [(0, (2, 2))], 'Stacked '
-s_map, w_map, a_map, c_map, d_map, l_map = '--', 2., 1., 'r', [(0, (2, 1))], 'Modes '
-s_exp, w_exp, a_exp, c_exp, d_exp, l_exp = '--', 2., 1., 'r', [(0, (1, 2))], 'Means '
-s_mle, w_mle, a_mle, c_mle, d_mle, l_mle = '-', 3., 1., 'c', [(0, (2, 2))], 'HB Optimization '
-s_bfe, w_bfe, a_bfe, c_bfe, d_bfe, l_bfe = '-', 3.5, 1., 'b', [(0, (1, 1))], 'HB Samples '
-s_smp, w_smp, a_smp, c_smp, d_smp, l_smp = '-', 2., 1., 'k', [(0, (1, 1))], 'Sampled '
+s_tru, w_tru, a_tru, c_tru, d_tru, l_tru = '-', 2., 1., 'k', '-', 'True '#[(0, (1, 1))], 'Truth '
+s_tbp, w_tbp, a_tbp, c_tbp, d_tbp, l_tbp = ':', 2., 0.75, 'k', '--', 'Binned True '#[(0, (1, 1))], 'Binned True '
+s_int, w_int, a_int, c_int, d_int, l_int = '-', 2., 0.5, 'k', ':', 'Interim '#[(0, (1, 1))], 'Interim '
+s_stk, w_stk, a_stk, c_stk, d_stk, l_stk = '--', 2.5, 1., 'r', '--', 'Stacked '#[(0, (2, 2))], 'Stacked '
[…]

### 2017-09-11 chippr/plot_utils.py

@@ -15,17 +15,23 @@ def set_up_plot():
     """
     Sets up plots to look decent
     """
-    title = 10
-    label = 10
+    title = 16
+    label = 14
     mpl.rcParams['text.usetex'] = True
     mpl.rcParams['axes.titlesize'] = title
     mpl.rcParams['axes.labelsize'] = label
-    mpl.rcParams['figure.subplot.left'] = 0.2
-    mpl.rcParams['figure.subplot.right'] = 0.9
-    mpl.rcParams['figure.subplot.bottom'] = 0.2
-    mpl.rcParams['figure.subplot.top'] = 0.9
-    mpl.rcParams['figure.subplot.wspace'] = 0.5
-    mpl.rcParams['figure.subplot.hspace'] = 0.5
+    # mpl.rcParams['figure.subplot.left'] = 0.2
+    # mpl.rcParams['figure.subplot.right'] = 0.9
+    # mpl.rcParams['figure.subplot.bottom'] = 0.2
+    # mpl.rcParams['figure.subplot.top'] = 0.9
+    # mpl.rcParams['figure.subplot.wspace'] = 0.5
+    # mpl.rcParams['figure.subplot.hspace'] = 0.5
+    mpl.rcParams['mathtext.rm'] = 'serif'
+    mpl.rcParams['font.family'] = 'serif'
+    mpl.rcParams['font.serif'] = 'Times New Roman'
+    mpl.rcParams['savefig.dpi'] = 250
+    mpl.rcParams['savefig.format'] = 'png'
+    mpl.rcParams['savefig.bbox'] = 'tight'
 
     return
 

### 2018-01-18 research/paper/draft.tex
updated abstract

--- 2017-08-01 research/paper/draft.tex
@@
 \author{Alex Malz\altaffilmark{1}, David W. Hogg\altaffilmark{1,2,3,4}, Phil 
-Marshall\altaffilmark{5}, \& Others}
+Marshall\altaffilmark{5}}
 \email{aimalz@nyu.edu}
@@
 \begin{abstract}
-The redshift distribution $n(z)$ is a crucial ingredient for weak lensing 
-cosmology.  Spectroscopically confirmed redshifts for the dim and numerous 
-galaxies observed by weak lensing surveys are expected to be inaccessible, 
-making photometric redshifts (photo-$z$s) the next best alternative.  Because 
-of the nontrivial inference involved in their determination, photo-$z$ point 
-estimates are being superseded by photo-$z$ probability distribution functions 
-(PDFs).  However, analytic methods for utilizing these new data products in 
-cosmological inference are still evolving.  This paper presents a novel 
-approach to estimating the posterior distribution over $n(z)$ from a survey of 
-galaxy photo-$z$ PDFs based upon a probabilistic graphical model of 
-hierarchical inference.  We present the Cosmological Hierarchical Inference 
-with Probabilistic Photometric Redshifts (\chippr) code implementing this 
-technique, as well as its validation on mock data and testing on the 
-\texttt{Buzzard} simulations.  \chippr yields a more accurate characterization 
-of $n(z)$ containing information beyond the best-fit estimator produced by 
-traditional procedures.  The publicly available code is easily extensible to 
-other one-point statistics that depend on redshift.
+A trustworthy estimate of the redshift distribution $n(z)$ is crucial for weak 
+lensing cosmology.  Spectroscopically confirmed redshifts for the dim and 
+numerous galaxies of weak lensing surveys are expected to be inaccessible, 
+making photometric redshifts (photo-$z$s) the next best alternative.  The 
+nontrivial systematics affecting photo-$z$ estimation have motivated the weak 
+lensing community to favor photo-$z$ probability distribution functions (PDFs) 
+as a more comprehensive alternative to photo-$z$ point estimates.  However, 
+analytic methods for utilizing these new data products in cosmological 
+inference are still evolving.  This paper presents the Cosmological 
+Hierarchical Inference with Probabilistic Photometric Redshifts (CHIPPR) model, 
+a novel approach to estimating the $n(z)$ from a catalog of galaxy photo-$z$ 
+PDFs based upon a probabilistic graphical model of hierarchical inference.  We 
+present the publicly available \chippr code implementing this technique, as 
+well as its validation on mock data.  The CHIPPR model yields a more accurate 
+characterization of $n(z)$ containing information beyond the best-fit estimator 
+produced by traditional procedures.
 
@@
 
-\keywords{catalogs --- cosmology: cosmological parameters --- galaxies: 
-statistics --- gravitational lensing: weak --- methods: analytical --- methods: 
-data analysis --- methods: statistical --- techniques: photometric}
+\keywords{cosmology: cosmological parameters --- galaxies: statistics --- 
+gravitational lensing: weak --- methods: data analysis --- methods: statistical}
 
@@
 
-The tests presented here are conducted with samples of $N = 10^{4}$ with
-
 \begin{figure}
@@
 in future work.
-
-In comparison with use of the stacked estimator of the redshift distribution 
-function, the hierarchical approach is computationally expensive.  In a future 
-revision, the \chippr code will be optimized for speed and resource efficiency. 
- At this stage, we establish scaling behavior to project the expense in terms 
-of computational resources.
 
@@
 \begin{acknowledgements}
-AIM thanks Mohammadjavad Vakili for insightful input on statistics, Geoffrey 
-Ryan for assistance in debugging, and Boris Leistedt for helpful comments 
-provided in the preparation of this paper.
+AIM thanks Elisabeth Krause for assistance with the \texttt{CosmoLike} code, 
+Mohammadjavad Vakili for insightful input on statistics, Geoffrey Ryan for 
+advice on debugging, and Boris Leistedt for helpful comments provided in the 
+preparation of this paper.
 \end{acknowledgements}

### 2018-06-27 chippr/catalog.py
periodic backup

@@ -233,17 +233,13 @@ class catalog(object):
                     # out_amps = uniform_lf.pdf(grid_means)
                 elif self.params['catastrophic_outliers'] == 'training':
                     full_pdf = np.exp(u.safe_log(outlier_lf.pdf(self.z_fine)))
-                    # print(full_pdf)
                     intermediate = np.dot(full_pdf, np.ones(self.n_tot) * self.dz_fine)
-                    # print(intermediate)
                     full_pdf = full_pdf / intermediate[np.newaxis]
-                    # print(full_pdf)
                     # flat_pdf = np.exp(u.safe_log(uniform_lf.pdf(self.z_fine)))
                     # flat_pdf = flat_pdf / np.dot(flat_pdf, self.dz_fine)
                     # items = np.array([vert_funcs[kk].pdf(self.z_fine[kk]) for kk in range(self.n_tot)])
                     fracs = np.array([full_pdf, np.ones(self.n_tot)]).T
                     fracs = fracs * np.array([frac, 1.-frac])[np.newaxis, :]
-                    # print(fracs)
                     grid_funcs = [gmix(fracs[kk], [uniform_lf, vert_funcs[kk]], limits=(self.bin_ends[0], self.bin_ends[-1])) for kk in range(self.n_tot)]
                     # out_funcs = [multi_dist([uniform_lfs[kk], uniform_lf]) for kk in range(self.n_tot)]
                     # out_amps = self.outlier_lf.pdf(grid_means)

### 2018-06-27 chippr/discrete.py

@@ -33,8 +33,14 @@ class discrete(object):
         self.bin_range = range(self.n_bins)
 
         self.weights = weights
+        # print('dbins: '+str(self.dbins))
+        # print('weights: '+str((self.weights)))
+        # print('sumweights: '+str((np.sum(self.weights))))
+        # print('dotweights: '+str((np.dot(self.weights, self.dbins))))
         self.normweights = np.cumsum(self.weights) / np.sum(self.weights)
+        # print('normweights: '+str(self.normweights))
         self.distweights = np.cumsum(self.weights) / np.dot(self.weights, self.dbins)
+        # print('distweights: '+str(self.distweights))
 
         self.funcs = [UD(self.bin_ends[i], self.bin_ends[i+1]) for i in self.bin_range]
         if self.n_bins > 1:

### 2018-06-27 research/scripts/basic_inf_script.py

@@ -176,7 +176,7 @@ if __name__ == "__main__":
     from chippr import *
 
     result_dir = os.path.join('..', 'results')
-    test_name = 'single_routhi'
+    test_name = 'single'
     all_tests = {}
     test_info = {}
     test_info['name'] = test_name

### 2018-06-27 research/scripts/basic_sim_script.py

@@ -258,7 +258,7 @@ if __name__ == "__main__":
     from chippr import *
 
     result_dir = os.path.join('..', 'results')
-    test_name = 'single_routhi'
+    test_name = 'single'
     all_tests = {}
     test_info = {}
     test_info['name'] = test_name

### 2018-06-28 research/paper/draft.tex
significant progress

@@
 
+\usepackage{soul}
+\usepackage[title]{appendix}
+
 \newcommand{\myemail}{aimalz@nyu.edu}
-\newcommand{\textul}{\underline}
-\newcommand{\chippr}{\texttt{CHIPPR} }
-
-\shorttitle{How to obtain the redshift distribution from probabilistic redshift 
-estimates}
-\shortauthors{Malz, et al.}
+\newcommand{\mathul}{\underline}
+\newcommand{\chippr}{CHIPPR}
+\newcommand{\nz}{$n(z)$}
+\newcommand{\pz}{photo-$z$}
+\newcommand{\pzpdf}{photo-$z$ PDF}
 
@@
 \title{How to obtain the redshift distribution from probabilistic redshift 
-estimates}
+estimates}%:\\
 
-\author{Alex Malz\altaffilmark{1}, David W. Hogg\altaffilmark{1,2,3,4}, Phil 
-Marshall\altaffilmark{5}}
+\author{Alex Malz\altaffilmark{1}}
+\author{David W. Hogg\altaffilmark{1,2,3,4}}
 \email{aimalz@nyu.edu}
@@
 Heidelberg, Germany}
-\altaffiltext{5}{[SLAC]}
 
 \begin{abstract}
-A trustworthy estimate of the redshift distribution $n(z)$ is crucial for weak 
-lensing cosmology.  Spectroscopically confirmed redshifts for the dim and 
-numerous galaxies of weak lensing surveys are expected to be inaccessible, 
-making photometric redshifts (photo-$z$s) the next best alternative.  The 
-nontrivial systematics affecting photo-$z$ estimation have motivated the weak 
-lensing community to favor photo-$z$ probability distribution functions (PDFs) 
-as a more comprehensive alternative to photo-$z$ point estimates.  However, 
-analytic methods for utilizing these new data products in cosmological 
-inference are still evolving.  This paper presents the Cosmological 
-Hierarchical Inference with Probabilistic Photometric Redshifts (CHIPPR) model, 
-a novel approach to estimating the $n(z)$ from a catalog of galaxy photo-$z$ 
-PDFs based upon a probabilistic graphical model of hierarchical inference.  We 
-present the publicly available \chippr code implementing this technique, as 
-well as its validation on mock data.  The CHIPPR model yields a more accurate 
-characterization of $n(z)$ containing information beyond the best-fit estimator 
-produced by traditional procedures.
-
+A trustworthy estimate of the redshift distribution $n(z)$ is crucial for 
+weak-lensing cosmology as we know it.
+Spectroscopic redshifts for the dim and numerous galaxies of weak-lensing 
+surveys are expected to be inaccessible, making photometric redshifts 
+(photo-$z$s) the next-best alternative.
+The nontrivial systematics affecting photo-$z$ estimation have motivated the 
+weak-lensing community to favor photo-$z$ probability density functions (PDFs) 
+as a more comprehensive alternative to photo-$z$ point estimates.
+However, analytic methods for utilizing these new data products in cosmological 
+inference are still evolving.
+The ubiquitous methodology known as stacking produces a systematically biased 
+estimator of $n(z)$ that worsens with decreasing signal-to-noise, the very 
+regime where photo-$z$ PDFs are most necessary.
+We introduce a mathematically rigorous probabilistic graphical model (PGM) of 
+hierarchical inference of $n(z)$, which is provably the only self-consistent 
+way to combine photo-$z$ PDFs to produce an estimator of $n(z)$.
+The novel Cosmological Hierarchical Inference with Probabilistic Photometric 
+Redshifts (CHIPPR) model yields a more accurate characterization of $n(z)$ by 
+correctly propagating the redshift uncertainty information beyond the best-fit 
+estimator produced by traditional procedures.
+We conclude by propagating these effects to constraints in the space of 
+cosmological parameters.
 \end{abstract}
@@
 
+\maketitle
+
 \section{Introduction}
@@
 
-After a brief literature review addressing how photo-$z$ PDFs are currently 
-used in cosmology, this paper aims to answer the following questions:
+\textcolor{blue}{Brief literature review addressing how \pzpdf s are currently 
+used in cosmology}
 
-\begin{itemize}
-	\item Why should we question existing methods?
-	\item How can we improve the effectiveness of using photo-$z$ PDFs in 
-inference?
-	\item How does the result of \chippr compare to established estimators 
-in terms of the accuracy of $n(z)$?
-	\item How significant is the effect of the discrepancy between $n(z)$ 
-estimators on cosmological constraints?
-\end{itemize}
+\textcolor{blue}{Q: Why should we question existing methods?\\
+A: Stacking is bad and only looks like it works because of assumptions that 
+don't hold when the data is as bad as we anticipate.  Cite pedantic doc?}
 
-\section{Method}
-\label{sec:method}
+overview of next questions: How can we improve the effectiveness of using 
+\pzpdf s in inference?  How does the result of \chippr\ compare to established 
+estimators in terms of the accuracy of \nz ?  \textcolor{blue}{Reach goal: How 
+significant is the effect of the discrepancy between \nz\ estimators on 
+cosmological constraints?}
 
-This paper presents a mathematically consistent method for obtaining the 
-posterior probability distribution over the redshift distribution $n(z)$ using 
-a catalog of photo-$z$ PDFs.  We start by introducing some nomenclature, 
-definitions, and symbols that will be used throughout Secs. \ref{sec:model}, 
-\ref{sec:others}, and \ref{sec:implementation}.
+\section{Methods}
+\label{sec:methods}
 
-We shall say the parameters comprising $\vec{\phi}$ define the redshift 
-distribution under some functional form about which this method is agnostic, so 
-long as the functional form evaluated with some true values of the parameters 
-accurately describes the true $n(z)$.  Since the redshift distribution is 
-itself a probability distribution, it may be written as $p(z | \vec{\phi})$.
+\textcolor{blue}{Q: How can we improve the effectiveness of using \pzpdf s in 
+inference?\\
+A: Hierarchical inference is the only self-consistent way.}
 
-Before jumping into the details of the model, it is crucial to settle on an 
-interpretation of what a photo-$z$ PDF actually is.  When a photo-$z$ PDF is 
-reported, it is an object containing information about the redshift $z_{i}$ of 
-galaxy $i$ based on its photometric data $\vec{d}_{i}$, making a photo-$z$ PDF 
-a \textit{posterior} probability distribution.  The data $\vec{d}_{i}$ may be 
-fluxes, magnitudes, colors, or any combination thereof.  We assume that the 
-redshifts of the galaxies are not related to the redshifts or photometry of 
-other galaxies in the survey, i.e. they are \textit{independent}.
+This paper presents the Cosmological Hierarchical Inference with Probabilistic 
+Photometric Redshifts (\chippr ) model as a mathematically consistent method 
+for constraining \nz s using a catalog of \pzpdf s.
+In Section~\ref{sec:model}, we introduce the \chippr\ formalism; in 
+Section~\ref{sec:application}, we outline how to use \chippr\ to constrain \nz\ 
+from a catalog of \pzpdf s; in Section~\ref{sec:limitations} we document the 
+assumptions that must hold for \chippr\ to be valid.
 
-In addition to the data $\vec{d}_{i}$, the photo-$z$ PDF also contains 
-information imparted by assumptions that went into the process by which the 
-photo-$z$ was made.  Those assumptions are only relevant insomuch as they 
-impose a preference on the redshift distribution. Thus, they can be reduced to 
-a sort of prior on the distribution of redshifts, which can be parametrized 
-under the chosen functional form by parameters $\vec{\phi}^{*}$.  The redshift 
-distribution $p(z | \vec{\phi}^{*})$ associated with the parameters in 
-$\vec{\phi}^{*}$ shall be called the interim prior because in many cases it was 
-our best guess as to the true redshift distribution before observing any data.  
-For example, in template-based photo-$z$ PDF methods, the interim prior is a 
-linear combination of the redshift distributions of the classes of templates 
-based on previous observations.  (In training-based photo-$z$ PDF methods, the 
-interim prior is related to the redshift distribution of the training set and 
-thus may not be shared among all galaxies for which photo-$z$ PDFs are 
-produced; however, in this work, we will assume that it is shared among all 
-galaxies in a survey.)  Because the choice of parameters $\vec{\phi}^{*}$ is 
-not causally connected to the data or the true redshift distribution but 
-nonetheless contributes to the photo-$z$ PDF, we call it an \textit{interim 
-prior}.
-
-A photo-$z$ PDF is then an \textit{interim posterior} probability distribution 
-of the redshift of a galaxy given its observed photometric data and the interim 
-prior: $p(z_{i} | \vec{d}_{i}, \vec{\phi}^{*})$.  
-We present the two attributes of a probabilistic graphical model (PGM) for the 
-redshift distribution $n(z)$: the directed acyclic graph (DAG) and its 
-mathematical interpretation in terms of Bayesian hierarchical inference.
-
-The fundamental assumption underlying the concept of photo-$z$ estimation is 
-that each galaxy $i$ has some observed photometric data $\vec{d}_{i}$ (fluxes, 
-magnitudes, or colors) that is drawn from a function of its redshift $z_{i}$, 
-which is a parameter in this model.  This function constitutes a forward model 
-for the observations.  The redshifts $\{z_{i}\}$ for all galaxies in a survey 
-are random draws from the redshift distribution, whose parameters  $\vec{\phi}$ 
-under a chosen functional form are called \textit{hyperparameters} because they 
-are shared among all $N$ galaxies in the survey.  Fig. \ref{fig:pgm} 
-illustrates these relationships.  PGMs of this structure are 
-\textit{hierarchical} in that, though the data are only directly influenced by 
-their redshift parameters, we can still use them to infer something about the 
-global hyperparameters by way of a higher level in the graph.
-
-\begin{figure}
-	\begin{center}We would like to learn about $p(z | \vec{\phi})$ from a 
-catalog of photo-$z$ interim posteriors.
-
-		\subsection{Model}
+\subsection{Model generalities}
 \label{sec:model}
 
-		\includegraphics[width=0.5\textwidth]{fig/pgm.png}
-		\caption{This directed acyclic graph corresponds to a 
-probabilistic graphical model for a Bayesian hierarchical inference of 
-$p(\vec{\phi}|\{\vec{d}_{i}\})$.  In this graph, all random variables are shown 
-in circles, with observed random variables shown in shaded circles.  
-Relationships between variables are indicated by arrows from parameters to the 
-variables distributed according to functions of them.  The box indicates that 
-there are a number of copies of the relationships between boxed parameters, 
-each independent of all others.  The hyperparameters comprising $\vec{\phi}$, 
-which define $n(z)$, are at the top.  Independently drawn from a function of 
-the hyperparameters $\vec{\phi}$ are galaxy redshifts $\{z_{i}\}$ below.  The 
-observed galaxy photometry $\{\vec{d}_{i}\}$, shown in shaded circles, is 
-determined by the redshifts above.}
+We begin by reframing the redshift distribution \nz\ from a probabilistic 
+perspective.
+Here we define a redshift distribution \nz\ as the normalized probability 
+density
+\begin{equation}
+\label{eqn:nz}
+\int_{-\infty}^{\infty}\ n(z)\ dz\ \equiv\ \frac{1}{J}\ 
+\int_{-\infty}^{\infty}\ \sum_{j=1}^{J}\ \delta(z_{j},\ z)\ dz = 1
+\end{equation}
+of finding a galaxy $j$ in a catalog of $J$ galaxies having a redshift $z$.
+We may without loss of generality impose a parametrization
+\begin{equation}
+\label{eqn:fz}
+f_{\phi}(z)\ \equiv\ n(z)
+\end{equation}
+in terms of some parameter vector $\phi$.
+We believe that galaxy redshifts are indeed drawn from \nz, making it a 
+probability density over redshift; this fact can also be confirmed from 
+dimensional analysis of Equation \ref{eqn:nz}.
+Therefore, it can be rewritten as
+\begin{equation}
+\label{eqn:pz}
+z_{j}\ \sim\ \Pr(z \mid \phi)\ \equiv\ f_{\phi}(z),
+\end{equation}
+a probability density over redshift conditioned on the parameters defining \nz.
+Note that $z_{j}$ does not depend on $z_{j'}$, a statement of the causal 
+independence of galaxy redshifts from one another.
+
+In addition to believing \nz\ is a PDF from which redshifts are drawn, we also 
+believe that there is some PDF from which photometric data $d$, which may be 
+any combination of fluxes, magnitudes, colors, and their observational errors, 
+are drawn.
+Such a PDF over data is a likelihood
+\begin{equation}
+\label{eqn:pzpdf}
+d\ \sim\ \Pr(d \mid z)
+\end{equation}
+conditioned on redshift.
+This assumption that the data are drawn from some function of the redshift 
+forms the foundation upon which \pz\ estimation is based.
+Note that galaxies may have different observational data $d$ despite sharing 
+the same redshift and that the data $d_{j}$ of one galaxy is causally 
+independent of the redshifts $z_{j'}$ and data $d_{j'}$ of other galaxies.
+
+This description of the physical system corresponds to a forward model by which 
+we actually believe photometry is generated:
+\begin{enumerate}
+	\item There exists a redshift distribution \nz\ with parameters $\phi$.
+	\item Galaxy redshifts $\{z_{j}\}$ are independent draws from $\Pr(z 
+\mid \phi)$.
+	\item Galaxy photometry $d_{j}$ is drawn from the likelihoods $\Pr(d | 
+z_{j})$.
+\end{enumerate}
+A forward model such as this corresponds to a probabilistic graphical model 
+(PGM), represented by a directed acyclic graph (DAG) as in Figure~\ref{fig:pgm}.
+\begin{figure}
+	\begin{center}
+		\includegraphics[width=0.25\textwidth]{fig/pgm.png}
+		\caption{The PGM}
 		\label{fig:pgm}
@@
 \end{figure}
+A DAG conveys the causal relationships between physical parameters and, like a 
+Feynman diagram in the context of particle physics, is a shorthand for 
+mathematical relationships between variables.
 
-The DAG of Fig. \ref{fig:pgm} translates directly into a mathematical 
-expression for the posterior distribution $p(\vec{\phi} | \{\vec{d}_{i}\})$ of 
-the hyperparameters given the entire set of interim posteriors $\{p(z_{i} | 
-\vec{d}_{i}, \vec{\phi}^{*})\}$ by way of the following derivation.  From this 
-point forward, we will use log probabilities.
+\subsection{Application to cosmology}
+\label{sec:application}
 
-According to Bayes' rule, the posterior distribution of interest is
+The problem facing cosmologists is to determine the true parameters $\phi_{0}$ 
+of \nz\ from observing the photometry $\{d_{j}\}$ of a large sample of galaxies 
+$j$.
+To self-consistently propagate the uncertainty in the redshifts, however, it is 
+more appropriate to estimate the posterior $\Pr(\phi \mid \{d_{j}\})$ over all 
+possible parameters $\phi$ (and thus potential redshift distributions \nz ) 
+conditioned on all the observed data $\{d_{j}\}$ available in a generic catalog.
+
+In order to use the DAG of Figure~\ref{fig:pgm} to derive an expression for 
+$\Pr(\phi \mid \{d_{j}\})$ in terms of \pzpdf s, we must introduce two more 
+concepts, confusingly named the implicit prior and the prior probability 
+density.
+
+When we constrain the redshift of a galaxy using its observed photometric data 
+$d_{j}$, we are effectively estimating a posterior $\Pr(z \mid d_{j})$.
+However, to do this, we must have a model for the general relationship between 
+redshifts and photometry, whether empirical, as is the case for machine 
+learning \pzpdf\ methods, or analytic, as is the case for template-based 
+\pzpdf\ methods.
+Such a relationship is defined in the space of probability density over 
+redshift, so it must be able to be parameterized by the same functional form 
+$f_{\phi}(z)$ as \nz .
+It is thus natural to write it as $\Pr(z \mid \phi^{*})$, where $\phi^{*}$ is 
+the parameters for this relationship under some generic \pzpdf\ method.
+We call $\Pr(z \mid \phi^{*})$ the \textit{implicit prior}, as it is rarely 
+explicitly known nor chosen by the researcher; for template-based methods where 
+it can be chosen to be ``realistic,'' it may be appropriate to call it an 
+\textit{interim prior}.
+Because the implicit prior is unavoidable, the \pzpdf s reported by any method 
+are really \textit{implicit-weighted posteriors} $\Pr(z \mid d, \phi^{*})$.
+
+The prior probability density $\Pr(\phi)$ is a more familiar concept in 
+astronomy; to progress, we will have to choose a prior probability density over 
+all possible parameters $\phi$.
+This prior need not be excessively proscriptive; for example, it may be chosen 
+to enforce smoothness at physically motivated scales in redshift without 
+imposing any particular region as over- or under-dense.
+
+With these definitions, we obtain the desired expression for $\Pr(\phi \mid 
+\{d_{j}\})$,
 \begin{align}
-\label{eq:hyper_bayes}
-\log[p(\vec{\phi} | \{\vec{d}_{i}\})] &\propto \log[p(\{\vec{d}_{i}\} | 
-\vec{\phi})] + \log[p(\vec{\phi})] ,
-\end{align}
-where $p(\vec{\phi})$ is the \textit{hyperprior} probability distribution over 
-possible values of the hyperparameters.  The hyperprior is a choice made by 
-those performing the inference and is not to be confused with the interim prior 
-parameters that influence the photo-$z$ PDF data product.
-
-Focusing on the likelihood term $p(\{\vec{d}_{i}\} | \vec{\phi})$, we employ 
-the power of the hierarchical model through \textit{marginalization} of the 
-unobserved redshift parameters by integrating over them to obtain 
-\begin{align}
-\label{eq:marginalize}
 \begin{split}
-\log[p(\{\vec{d}_{i}\} | \vec{\phi})] = & 
-\log\left[\int\exp\left[\log[p(\{\vec{d}_{i}\} | \{z_{i}\})]\right.\right. \\
-& \left.\left.+ \log[p(\{z_{i}\} | \vec{\phi})]\right]\ d\{z_{i}\}\right] .
+  \label{eqn:fullpost}
+  \ln[\Pr(\phi \mid \{d_{j}\})] & \propto \ln[\Pr(\phi)] + \ln \left[\int dz 
+\right.\\
+  & \left. \exp \left[\sum_{j=1}^{J} \left(\ln[\Pr(z \mid d_{j}, \phi^{*})] 
+\right. \right. \right.\\
+  & \left. \left. \left. + \ln[\Pr(z \mid \phi)] - \ln[\Pr(z \mid \phi^{*})]\ 
+\right)\ \right]\ \right] ,
 \end{split}
 \end{align}
-The redshifts $\{z_{i}\}$ are statistically independent of each other, and each 
-galaxy's photometry depends solely on its own redshift, so we may invoke 
-statistical independence to get
-\begin{align}
-\label{eq:independence_hyper}
-\log[p(\{z_{i}\} | \vec{\phi})] &= \sum_{i}^{N} \log[p(z_{i} | \vec{\phi})]
-\end{align}
-and
-\begin{align}
-\label{eq:independence}
-\log[p(\{\vec{d}_{i}\} | \{z_{i}\})] &= \sum_{i}^{N} \log[p(\vec{d}_{i} | 
-z_{i})] .
-\end{align}
-Having chosen the functional form mapping $\vec{\phi}$ to $n(z)$, the terms on 
-the righthand side of Eq. \ref{eq:independence_hyper} are known.  However, the 
-terms on the righthand side of Eq. \ref{eq:independence} are likelihoods, and 
-the photo-$z$ PDFs we have are interim posteriors.
-
-To transform the interim posteriors we have into the likelihoods we need, we 
-start with the vacuously true statement
-\begin{align}
-\begin{split}
-\label{eq:unity}
-\log[p(\vec{d}_{i} | z_{i})] =& \log[p(\vec{d}_{i} | z_{i})] + \log[p(z_{i} | 
-\vec{d}_{i}, \vec{\phi}^{*})]\\
-& - \log[p(z_{i} | \vec{d}_{i}, \vec{\phi}^{*})].
-\end{split}
-\end{align}
-We expand the last term using Bayes' Rule to get
-\begin{align}
-\begin{split}
-\label{eq:bayes}
-\log[p(\vec{d}_{i} | z_{i})] =& \log[p(\vec{d}_{i} | z_{i})] + \log[p(z_{i} | 
-\vec{d}_{i}, \vec{\phi}^{*})]\\
-& - \log[p(\vec{d}_{i} | z_{i}, \vec{\phi}^{*})]\\
-& + \log[p(\vec{d}_{i} | \vec{\phi}^{*})] - \log[p(z_{i} | \vec{\phi}^{*})].
-\end{split}
-\end{align}
-Fig. \ref{fig:pgm} has no direct connection between $\vec{d}_{i}$ and 
-$\vec{\phi}$, so we can break up $\log[p(\vec{d}_{i} | z_{i}, \vec{\phi}^{*})]$ 
-as follows:
-\begin{align}
-\begin{split}
-\label{eq:split}
-\log[p(\vec{d}_{i} | z_{i})] =& \log[p(\vec{d}_{i} | z_{i})] + \log[p(z_{i} | 
-\vec{d}_{i}, \vec{\phi}^{*})]\\
-& - \log[p(\vec{d}_{i} | z_{i})] - \log[p(\vec{d}_{i} | \vec{\phi}^{*})]\\
-& + \log[p(\vec{d}_{i} | \vec{\phi}^{*})] - \log[p(z_{i} | \vec{\phi}^{*})].
-\end{split}
-\end{align}
-Canceling terms, we find
-\begin{align}
-\label{eq:cancel}
-\log[p(\vec{d}_{i} | z_{i})] =& \log[p(z_{i} | \vec{d}_{i}, \vec{\phi}^{*})] - 
-\log[p(z_{i} | \vec{\phi}^{*})].
-\end{align}
-
-Now we combine Eqs. \ref{eq:hyper_bayes}, \ref{eq:marginalize}, 
-\ref{eq:independence_hyper}, \ref{eq:independence}, and \ref{eq:cancel} to 
-arrive at 
-\begin{align}
-\begin{split}
-\label{eq:final}
-\log[p(\vec{\phi} | \{\vec{d}_{i}\})] \propto& \log[p(\vec{\phi})] + 
-\log\left[\int \exp\left[\sum_{i}^{N} \left(\right.\right.\right.\\
-& \left.\left.\left. + \log[p(z_{i} | \vec{\phi})] - \log[p(z_{i} | 
-\vec{\phi}^{*})]\right.\right.\right.\\
-& \left.\left.\left. + \log[p(z_{i} | \vec{d}_{i}, \vec{\phi}^{*})]
-\right)\right] d\{z_{i}\}\right].
-\end{split}
-\end{align}
-Thus, if we have a catalog of photo-$z$ interim posteriors $\{p(z_{i} | 
-\vec{d}_{i}, \vec{\phi}^{*})\}$, the interim prior parameters $\vec{\phi}^{*}$, 
-and a prior distribution $p(\vec{\phi})$, we may find the posterior 
-$\log[p(\vec{\phi} | \{\vec{d}_{i}\})]$ on the redshift distribution $n(z)$.
-
-This framework entails a number of choices and assumptions that must be 
-addressed explicitly:
-\begin{enumerate}
-	\item The chosen functional form of $n(z)$ with parameters $\vec{\phi}$ 
-must be capable of describing the true redshift distribution.
-	\item The photometric data and redshift for each galaxy must be 
-independent of the photometric data and redshifts of all other galaxies.
-	\item There is one interim prior shared among all galaxies in the 
-survey, and its parameters with known parameters $\vec{\phi}^{*}$ are known.
-	\item We must choose a hyperprior probability distribution 
-$p(\vec{\phi})$ over the hyperparameters.
-	\item While we advocate for the approach of hierarchical inference, the 
-probabilistic graphical model presented here is not the only one that could be 
-proposed.
-\end{enumerate}
-
-\subsection{Alternative Approaches}
-\label{sec:others}
-
-It is useful to translate some popular existing methods for deriving $n(z)$ 
-from photo-$z$ PDFs into the mathematical framework of Sec. \ref{sec:model}.  
-We briefly discuss the conditions under which they are equivalent to Eq. 
-\ref{eq:final}, which are elaborated upon in the Appendix.
-
-\subsubsection{Stacking}
-\label{sec:stacking}
-
-The most common way to combine photo-$z$ interim posteriors into an estimator 
-$\hat{n}_{stack}(z)$ for $n(z)$ is to "stack" them, which corresponds to
-\begin{align}
-\label{eq:stack}
-\hat{n}^{stack}(z) &= \frac{1}{N}\ \sum_{i}^{N} \exp\left[\log[p(z_{i} | 
-\vec{d}_{i}, \vec{\phi}^{*})]\right].
-\end{align}
-
-\subsubsection{Point Estimation}
-\label{sec:point_estimation}
-
-In some cases, estimators $\hat{n}^{point}(z)$ are obtained by assuming the 
-photo-$z$ PDFs are effectively delta functions by reducing them to point 
-estimates according to
-\begin{align}
-\label{eq:delta}
-p(z_{i} | \vec{d}_{i}, \vec{\phi}^{*}) &\approx \delta(z, \hat{z}_{i}^{point})
-\end{align}
-before applying Eq. \ref{eq:stack}.  The most popular redshift point estimators 
-are the mean $\hat{z}^{mean}_{i}$, median $\hat{z}^{median}_{i}$, and mode 
-$\hat{z}^{mode}_{i}$ of the original photo-$z$ interim posterior.
+which is the heart of \chippr.
+The entire derivation of Equation~\ref{eqn:fullpost} is provided in 
+Appendix~\ref{app:math}.
 
@@
 
-The publicly available \chippr code implements the probabilistic graphical 
-model presented in Sec. \ref{sec:model}.
+In this study, we compare the results of Equation~\ref{eqn:fullpost} to those 
+of the two most common approaches to estimating \nz\ from a catalog of \pzpdf 
+s: the distribution $n(z_{\mathrm{max}})$ of the redshifts at maximum posterior 
+probability (i.e. the modes of the \pzpdf s), and the stacked estimator
+\begin{align}
+  \label{eqn:stacked}
+  \hat{n}(z) &\equiv \frac{1}{J}\ \sum_{j=1}^{J}\ \Pr(z \mid d_{j}, \phi^{*}) .
+\end{align}
 
-In addition to the choices and assumptions underlying the probabilistic 
-graphical model, the implementation of \chippr makes choices and assumptions of 
-its own.
-
-\begin{enumerate}
-	\item \chippr is only applicable if the redshift interim posteriors and 
-interim prior are accurate.
-	\item \chippr currently only accepts photo-$z$ PDFs and produces $n(z)$ 
-samples of a single format, that of the piecewise constant parametrization, 
-also referred to as a binned histogram parametrization and a sum of top hat 
-functions.
-\end{enumerate}
-
-\section{Validation on Simple Mock Data}
-\label{sec:validation}
-
-We demonstrate the superiority of \chippr over alternative approaches in a 
-number of compelling test cases on mock data.  Each experiment is characterized 
-by a single change to a fiducial case in order to isolate the influence of 
-systematic effects known to be relevant to photo-$z$ estimation and propagation 
-in analysis.
-
-\subsection{Mock Data \& Metrics}
-\label{sec:validintro}
-
-The mock data in these tests consists of photo-$z$ interim posteriors rather 
-than photometric data because the various existing methods for deriving 
-photo-$z$ interim posteriors do not in general yield results that are 
-consistent with one another, indicating that their systematics are not 
-well-understood.  Because the mock data is independent of any choice of 
-photo-$z$ PDF production method, we not only ensure that our photo-$z$ interim 
-priors are perfectly understood but also deter readers from assuming that 
-\chippr has any preference over the method by which photo-$z$ interim 
-posteriors are derived from photometric data.
-
-\subsubsection{Mock Data}
-\label{sec:mockdata}
-
-The mock data used here are produced by the following steps.
-
-\begin{enumerate}
-	\item \label{itm:true} Choose a true $n(z)$ that is a continuous 
-function with known parameters $\vec{\phi}'$.
-	\item \label{itm:sample} Sample the true $n(z)$ to create a catalog of 
-$N$ true redshifts $z'_{i}$.
-	\item \label{itm:scatter} Choose a true intrinsic scatter parameter 
-$\sigma$, and sample Gaussians $p(z''_{i} | z'_{i}, \sigma) = 
-\mathcal{N}(z'_{i}, \sigma)$ to get "observed" redshifts $z'_{i}$ defining 
-Gaussian likelihoods $p(\vec{d}_{i} | z''_{i}, \sigma) = \mathcal{N}(z''_{i}, 
-\sigma)$.
-	\item \label{itm:interim} Choose a parametrization and the parameters 
-$\vec{\phi}^{*}$ of the interim prior $p(z_{i} | \vec{\phi}^{*})$.
-	\item \label{itm:posterior} Create interim posteriors $\log[p(z_{i} | 
-\vec{d}_{i}, \vec{\phi}^{*})] = \log[p(\vec{d}_{i} | z'_{i})]\ \log[p(z_{i} | 
-\vec{\phi}^{*})]$ in this parametrization.
-\end{enumerate}
-
-In all of the following validation tests, we use a piecewise constant 
-parametrization in log-space with $10$ bins.  $N = 10,000$ galaxies.  In Secs. 
-\ref{sec:truth} and \ref{sec:likelihoods}, we use a flat interim prior.  This 
-method for deriving mock data is referred to as the fiducial case, and 
-variations on it will refer directly to the steps that are altered.
-
-\subsubsection{$n(z)$ Accuracy Metric}
-\label{sec:accuracy}
-
-The Kullback-Leibler divergence is our primary measure of the accuracy of 
-estimators of $n(z)$ in cases of mock data with known true redshifts.
-
-\textbf{Review precision and bias from \texttt{kld.ipynb} and interpret in 
-terms of a \% difference.}
-
-\subsection{Underlying $n(z)$ Effects}
-\label{sec:truth}
-
-Existing $n(z)$ estimators are systematically smoother than the true $n(z)$.  
-Here we show that the traditional estimators perform better when the true 
-$n(z)$ is weakly featured than when the true $n(z)$ is strongly featured by 
-experimenting with Step \ref{itm:true} of Sec. \ref{sec:mockdata}.  The 
-implication of this issue is quite serious; the consistently smooth, unimodal 
-$n(z)$ estimates appearing in the literature could result from much more 
-featured true redshift distributions, and there would be no way to catch this 
-error without using a fully probabilistic method.
-
-\subsubsection{Featureless $n(z)$}
-\label{sec:smooth}
-
-In this test, we choose a weakly featured true $n(z)$ of Fig. \ref{fig:smooth} 
-with a smooth, unimodal shape, based on the interim prior used for the SDSS DR8 
-photo-$z$ PDFs.  [check and include citation!]
-
-\begin{figure}
+We perform this comparison on mock data in the form of catalogs of emulated 
+\pzpdf s generated via the forward model discussed above and presented in 
+detail in Appendix~\ref{app:data}.
+The mock data emulates the three sources of error of highest concern to the 
+\pz\ community: intrinsic scatter, catastrophic outliers, and systematic bias.
+Figure~\ref{fig:mega_scatter} illustrates these three effects individually at 
+ten times the tolerance of the upcoming Large Synoptic Survey Telescope (LSST).
+\begin{figure*}
 	\begin{center}
 
-\includegraphics[width=0.5\textwidth]{fig/smooth_truth/scatter.png}\\
+\includegraphics[width=0.3\textwidth]{fig/forward_model_example.png}
+    \includegraphics[width=0.3\textwidth]{fig/forward_model_example.png}
+    \includegraphics[width=0.3\textwidth]{fig/forward_model_example.png}
+		\caption{Show this for each individually.}
+		\label{fig:mega_scatter}
+	\end{center}
+\end{figure*}
+Tests including all three effects at the tolerance levels of LSST and at twice 
+the tolerance levels of LSST are presented in Section~\ref{sec:results}.
 
-\includegraphics[width=0.5\textwidth]{fig/smooth_truth/estimators.png}	
-		\caption{All estimators perform well when the true $n(z)$ is 
-well behaved, exhibiting significant deviation only when $n(z)$ is very small, 
-as the sample size of true redshifts in that range will be small.  Top panel: 
-The traditional MAP reduction of a photo-$z$ PDF against the true redshifts 
-with a few rescaled photo-$z$ interim posteriors are overplotted in solid 
-lines, with a dotted line indicating zero probability.  Bottom panel: Various 
-estimators of $\ln[n(z)]$, the interim prior, and the true $\ln[n(z)]$ as a 
-continuous function and under a binned parametrization.}
-		\label{fig:featured}
-	\end{center}
-\end{figure}
+\subsection{Limitations}
+\label{sec:limitations}
 
-\subsubsection{Featured $n(z)$}
-\label{sec:featured}
+Finally, we explicitly review the assumptions made by this approach.
+\begin{itemize}
+  \item prior $\Pr(\phi)$
+  \item implicit prior $\phi^{*}$ known
+	\item \pzpdf s are accurate posteriors %draws from photo-$z$ PDFs do 
+not match marginal distributions in $z_true$ vs. $z_phot$ space, i.e. toy model 
+with $z_phot$ drawn from an uglier space but pdfs evaluated in lsst-limits space
+	\end{itemize}
 
-In this test, we choose the true $n(z)$ of Fig. \ref{fig:featured} with 
-nontrivial structure.
+\section{Results}
+\label{sec:results}
 
-\begin{figure}
+\textcolor{blue}{Q: How does the result of \chippr compare to established 
+estimators in terms of the accuracy of $n(z)$?\\
+A: \chippr yields the best possible $n(z)$, conditional on the accuracy of the 
+photo-$z$ PDFs used.}
+
+\subsection{Idealized example}
+\label{sec:idealized}
+
+\textcolor{blue}{LSST specs: unbiased, uniform outliers, medium z-dependent 
+scatter}
+
+\begin{figure*}
 	\begin{center}
 
-\includegraphics[width=0.5\textwidth]{fig/fiducial/scatter.png}\\
+\includegraphics[width=0.45\textwidth]{fig/forward_model_example.png}
+		\includegraphics[width=0.45\textwidth]{fig/results_example.png}
+		\caption{Show this for idealized case.}
+		\label{fig:idealized}
+	\end{center}
+\end{figure*}
 
-\includegraphics[width=0.5\textwidth]{fig/fiducial/estimators.png}		
-		\caption{[This plot doesn't really show what I want because the 
-intrinsic scatter is too low!  I used $\sigma=0.03$ because that's what has 
-been quoted as what LSST, etc. needs, but that's not what comes out of 
-photo-$z$ estimation methods before they impose aggressive cuts\dots]  Top 
-panel: The traditional MAP reduction of a photo-$z$ PDF against the true 
-redshifts with a few rescaled photo-$z$ interim posteriors are overplotted in 
-solid lines, with a dotted line indicating zero probability.  Bottom panel: 
-Various estimators of $\ln[n(z)]$, the interim prior, and the true $\ln[n(z)]$ 
-as a continuous function and under a binned parametrization.}
-		\label{fig:featured}
-	\end{center}
-\end{figure}
+\subsection{Realistic example}
+\label{sec:realistic}
 
-This featured true $n(z)$ will henceforth be referred to as the fiducial $n(z)$
+\textcolor{blue}{DES(?) specs: biased to enhance true $n(z)$, systematic 
+outliers, high z-dependent scatter}
 
-\subsection{Emulated Data Quality Effects}
-\label{sec:likelihoods}
-
-In the following test cases, we vary the properties of the mock photo-$z$ 
-likelihoods in an effort to emulate known systematics in photo-$z$ estimation.  
-These tests vary Step \ref{itm:scatter} of Sec. \ref{sec:mockdata}.
-
-\subsubsection{Intrinsic Scatter}
-\label{sec:intscat}
-
-One major concern about photo-$z$s is the intrinsic scatter of point 
-estimators, including those derived from photo-$z$ PDFs, that is observed to 
-varying extents with every existing photo-$z$ algorithm and illustrated in Fig. 
-\ref{fig:intscat}.  To emulate intrinsic scatter, we modify the fiducial case 
-to simply broaden the single Gaussian component of the likelihood.  To enforce 
-self-consistency, the mean is a random variable drawn from a Gaussian 
-distribution with the newly increased variance.
-
-\begin{figure}
+\begin{figure*}
 	\begin{center}
 
-\includegraphics[width=0.5\textwidth]{fig/high_scatter/scatter.png}\\
+\includegraphics[width=0.45\textwidth]{fig/forward_model_example.png}
+		\includegraphics[width=0.45\textwidth]{fig/results_example.png}
+		\caption{Show this for realistic case.}
+		\label{fig:realistic}
+	\end{center}
+\end{figure*}
 
-\includegraphics[width=0.5\textwidth]{fig/high_scatter/estimators.png}		
-		\caption{As the intrinsic scatter increases, the discrepancy 
-between estimators increases.  In particular, the stacked estimator and 
-marginalized point estimators predict $\ln[n(z)]$ to be smoother than the 
-truth, while the  [This would be a lot more compelling with more galaxies.  
-Also, the weird edge effects in the top panel are real, because the point 
-estimator is the MAP, not the center of the Gaussian likelihood, and there's no 
-requirement that the mean of the likelihood be within the true redshift range.] 
- Top panel: The traditional MAP reduction of a photo-$z$ PDF against the true 
-redshifts with a few rescaled photo-$z$ interim posteriors are overplotted in 
-solid lines, with a dotted line indicating zero probability.  Bottom panel: 
-Various estimators of $\ln[n(z)]$, the interim prior, and the true $\ln[n(z)]$ 
-as a continuous function and under a binned parametrization.}
-		\label{fig:intscat}
-	\end{center}
-\end{figure}
+\subsection{Violations of the model}
+\label{sec:violations}
 
-To emulate intrinsic scatter, we modify the fiducial case to simply broaden the 
-single Gaussian component of the likelihood.  To enforce self-consistency, the 
-mean is a random variable drawn from a Gaussian distribution with the newly 
-increased variance.
+\textcolor{blue}{mischaracterized interim prior, with idealized or realistic 
+data?}
 
-\subsubsection{Template-like Catastrophic Outliers}
-\label{sec:tempcatout}
-
-In addition to intrinsic scatter, photo-$z$ methods employing template fitting 
-tend to produce catastrophic outliers that are distributed to be broad in 
-$z_{spec}$ and narrow in $z_{phot}$, as in Fig. \ref{fig:tempcatout}.  The 
-systematic behind these catastrophic outliers may be described as an attractor 
-in the space of $z_{phot}$; some galaxies at a range of $z_{spec}$ map onto a 
-single $z_{phot}$ (with some scatter) if their true SED does not have 
-sufficiently strong features (as is the case for blue galaxies), leading 
-galaxies of that SED type at many $z_{spec}$ to have similar colors.
-
-\begin{figure}
+\begin{figure*}
 	\begin{center}
 
-\includegraphics[width=0.5\textwidth]{fig/template_outliers/scatter.png}\\
-
-\includegraphics[width=0.5\textwidth]{fig/template_outliers/estimators.png}
-
-		\caption{Top panel: The traditional MAP reduction of a 
-photo-$z$ PDF against the true redshifts with a few rescaled photo-$z$ interim 
-posteriors are overplotted in solid lines, with a dotted line indicating zero 
-probability.  Bottom panel: Various estimators of $\ln[n(z)]$, the interim 
-prior, and the true $\ln[n(z)]$ as a continuous function and under a binned 
-parametrization.}
-		\label{fig:tempcatout}
+\includegraphics[width=0.45\textwidth]{fig/forward_model_example.png}
+		\includegraphics[width=0.45\textwidth]{fig/results_example.png}
+		\caption{Show this for mischaracterized case.}
+		\label{fig:mischaracterized}
 	\end{center}
-\end{figure}
-
-\subsubsection{Training-like Catastrophic Outliers}
-\label{sec:traincatout}
-
-Data driven photo-$z$ methods tend to suffer from a different form of 
-catastrophic outliers that are distributed to be narrow in $z_{spec}$ and broad 
-in $z_{phot}$, as in Fig. \ref{fig:traincatout}.  The systematic behind these 
-catastrophic outliers may be described as an attractor in the space of 
-$z_{spec}$; some galaxies near a particular $z_{spec}$ map to a range of 
-$z_{phot}$ if the training set galaxies at that $z_{spec}$ have inconsistent 
-$z_{phot}$, as might occur if their SED's features fall between photometric 
-filters.
-
-\begin{figure}
-	\begin{center}
-
-\includegraphics[width=0.5\textwidth]{fig/training_outliers/scatter.png}\\
-
-\includegraphics[width=0.5\textwidth]{fig/training_outliers/estimators.png}	
-		\caption{[This has proven to be the most challenging test case 
-to implement, and there's clearly still a bug in the function that makes the 
-likelihoods.]}
-		\label{fig:traincatout}
-	\end{center}
-\end{figure}
-
-\subsection{Emulated Interim Prior Effects}
-\label{sec:priors}
-
-The interim prior encapsulates the the relationship between observed photometry 
-and redshift information upon which a photo-$z$ estimate is based.  Interim 
-priors are in general not identical to the true $n(z)$ we wish to estimate; if 
-they were, we would not need any data!  For template fitting photo-$z$ methods, 
-the interim prior is usually an input chosen by the researcher.  However, for 
-machine learning methods, the interim prior is some function of the training 
-set data that in many cases may be influenced by random numbers and is rarely 
-output with the redshift estimates.  Interim priors for template fitting 
-methods tend to have incomplete coverage in the space of true photometry, 
-because they are limited by the choice of the library of SEDs.  Interim priors 
-for machine learning methods tend to have incomplete coverage in the space of 
-redshifts, because there are fewer galaxies with spectroscopically confirmed 
-redshifts at high redshifts than low redshifts.
-
-Existing $n(z)$ estimation routines will always produce a biased estimator when 
-the interim prior is not equal to the true $n(z)$.  We demonstrate here that 
-regardless of the appropriateness of the interim prior as an approximation to 
-the true $n(z)$, \chippr is not affected by the choice of the interim prior so 
-long as it has nontrivial coverage in the space of redshift.  These tests 
-modify Step \ref{itm:interim} of Sec. \ref{sec:mockdata}.
-
-\subsubsection{Template-like Interim Prior}
-\label{sec:tempintpr}
-
-An interim prior based on a template library may be a sum of smooth functions 
-representing $n(z)$ for each SED type in the library.  Template libraries do 
-not include every possible galaxy SED, and the $n(z)$ used for each SED type 
-may not be accurate.  The interim prior shown in Fig. \ref{fig:tempintpr} is an 
-emulation of an interim prior corresponding to a template library of this type.
-
-\begin{figure}
-	\begin{center}
-
-\includegraphics[width=0.5\textwidth]{fig/template_prior/scatter.png}\\
-
-\includegraphics[width=0.5\textwidth]{fig/template_prior/estimators.png}	
-		\caption{[The case of a multimodal interim prior was a very 
-compelling test in the previous version but somehow isn't anymore.]}
-		\label{fig:tempintpr}
-	\end{center}
-\end{figure}
-
-\subsubsection{Training-like Interim Prior}
-\label{sec:trainintpr}
-
-An interim prior based on a training set may be biased toward low redshifts due 
-to the dearth of distant galaxies with spectroscopic redshifts.  The interim 
-prior shown in Fig. \ref{fig:trainintpr} is an emulation of an interim prior 
-corresponding to a training set biased in this way.  We chose an interpolation 
-of the interim prior used for the SDSS DR8 photo-$z$ PDFs.
-
-\begin{figure}
-	\begin{center}
-
-\includegraphics[width=0.5\textwidth]{fig/training_prior/scatter.png}\\
-
-\includegraphics[width=0.5\textwidth]{fig/training_prior/estimators.png}	
-		\caption{[The case of a low-$z$ favoring interim prior was a 
-very compelling test in the previous version but somehow isn't anymore.]}
-		\label{fig:trainintpr}
-	\end{center}
-\end{figure}
-
-\section{Application to Realistic Mock Data}
-\label{sec:application}
-
-To show how the choice of $n(z)$ estimator propagates to cosmological 
-constraints, we apply \chippr to a data from a realistic cosmological 
-simulation (probably \texttt{Buzzard}) with photo-$z$ PDFs produced by a 
-popular method (probably \texttt{BPZ}).
-
-\subsection{Mock Data \& Metrics}
-\label{sec:appintro}
-
-As in Sec. \ref{sec:validintro}, the mock data takes the form of photo-$z$ 
-interim posteriors, but that is where the similarity ends.  These photo-$z$ 
-interim posteriors are derived from the photometry resulting from the 
-\texttt{Buzzard} simulation by way of \texttt{BPZ}.  Because the simulation 
-begins with setting true values of the cosmological parameters, we can 
-propagate the different estimators of $n(z)$ through a forecasting code (which 
-one?) to generate error ellipses on the cosmological parameters.
-
-\subsubsection{Mock Data}
-\label{sec:buzzard}
-
-We summarize the details of the \texttt{Buzzard} simulation here.
-
-\subsubsection{Cosmological Constraint Metric}
-\label{sec:cosmo}
-
-Because the realistic mock data of Sec. \ref{sec:buzzard} is associated with 
-true values of the cosmological parameters, we may compare the quality of 
-cosmological constraints under different estimators of $n(z)$, creating a 
-figure like Fig. \ref{fig:scgsr}.  We perform this analysis using a forecasting 
-code that takes $n(z)$ and produces projected error ellipses in the space of 
-cosmological parameters.
-
-\begin{figure}
-	\begin{center}
-		Not my figure! (N. MacCrann via J. DeRose)
-		\includegraphics[width=0.5\textwidth]{cosmo_constraints.png}
-		\caption{[I'd like to be able to make a plot like this with the 
-different $n(z)$ estimators produced by \texttt{chippr} from the same set of 
-photo-$z$ PDFs.  Then I can calculate some metrics of the accuracy and 
-precision of the error distributions relative to the true values of the 
-cosmological parameters that produced the mock data.]}
-		\label{fig:scgsr}
-	\end{center}
-\end{figure}
-
-\subsection{Results}
-\label{sec:results}
-
-\begin{figure}
-	\begin{center}
-		\caption{[moneyplot of error ellipses resulting from different 
-$n(z)$ estimators]}
-		\label{fig:money}
-	\end{center}
-\end{figure}
+\end{figure*}
 
@@
 
-The results of Fig \ref{fig:money} have significant implications for the 
-developing data analysis pipelines of next-generation telescope surveys.  
-However, the method presented here has its own limitations, which are 
-reiterated to discourage the community from applying this work inappropriately.
-
-We intend to pursue a number of extensions of the work presented in this paper 
-in future work.
-
 \section{Conclusion}
@@
 
-We now summarize answers to the questions posed in the introduction:
+\section*{Appendices}
 
-\begin{itemize}
-	\item Existing $n(z)$ estimation methods produce biased estimators that 
-propagate to inaccuracies in characterizing the cosmological parameters.
-	\item Photo-$z$ PDFs are probabilistic data products so must be handled 
-in a mathematically consistent manner such as the probabilistic graphical model 
-outlined in this paper.
-	\item In addition to coming with its own error distribution, the $n(z)$ 
-estimator produced by \chippr is quantifiably more accurate than established 
-estimators.
-	\item Propagation of the \chippr result leads to a quantifiable 
-improvement in the constraints on cosmological parameters.
-\end{itemize}
+\begin{appendices}
 
-In conclusion, we discourage the community from continuing to use the stacked 
-estimator and reductions of photo-$z$ PDFs to redshift point estimates in 
-obtaining estimators of $n(z)$.  \chippr is freely available to the community 
-for incorporation into evolving data analysis pipelines.
+\section{Mathematical derivation}
+\label{app:math}
 
-\appendix{Catalog Production}
-\label{app:catalog}
+\section{Mock data generation}
+\label{app:data}
 
-Sec. \ref{sec:mockdata} outlines one way to make mock data, but it's not 
-necessarily a logically consistent way to do it, a problem identified in trying 
-to use it for the test cases in Sec. \ref{sec:likelihoods}.  If the goal is to 
-emulate the traditional $z_{spec}$ vs. $z_{phot}$ scatterplots, we need to 
-start from probability distributions in that space.  This means there are two 
-parameters $z_{spec}$ and $z_{phot}$ drawn from a two-dimensional distribution 
-$p(z_{spec}, z_{phot} | \textul{\sigma}, \vec{\phi})$, where $\textul{\sigma}$ 
-contains parameters concerning the relationship between $z_{spec}$ and 
-$z_{phot}$, and $\vec{\phi}$ contains hyperparameters concerning $n(z)$.  We 
-know from basic probability that
-\begin{align}
-\label{eq:trivial}
-p(z_{spec}, z_{phot} | \textul{\sigma}, \vec{\phi}) =& p(z_{phot} | z_{spec}, 
-\textul{\sigma}, \vec{\phi})\ p(z_{sped} | \textul{\sigma}, \vec{\phi}).
-\end{align}
-Because $z_{spec}$ depends only on $\vec{\phi}$ and $z_{phot}$ is independent 
-of $\vec{\phi}$, this becomes
-\begin{align}
-\label{eq:simpler}
-p(z_{spec}, z_{phot} | \textul{\sigma}, \vec{\phi}) =& p(z_{phot} | z_{spec}, 
-\textul{\sigma})\ p(z_{spec} | \vec{\phi}).
-\end{align}
-We know how to make reasonable choices for both terms on the righthand side of 
-Eq. \ref{eq:simpler}.
+\textul{Plots: flow chart of forward model}
 
-In the $z_{spec}$ vs. $z_{phot}$ scatterplot, $p(z_{phot} | z_{spec}, 
-\textul{\sigma}, \vec{\phi})$ would be represented by vertical slices, but we 
-are interested in the horizontal slices that represent $p(z_{spec} | z_{phot}, 
-\textul{\sigma}, \vec{\phi})$.  Now we must distinguish the true 
-hyperparameters $\vec{\phi}_{true}$ from the interim hyperparameters 
-$\vec{\phi}_{int}$.  If the horizontal slices carry information about 
-$\vec{\phi}$, how can we get a catalog of $p(z_{spec} | z_{phot}, 
-\textul{\sigma}, \vec{\phi}_{int})$ that's consistent with $p(z_{spec} | 
-z_{phot}, \textul{\sigma}, \vec{\phi}_{true})$?  Somehow that information needs 
-to be propagated through in order to recover it.
+\begin{figure*}
+	\begin{center}
+		\includegraphics[width=0.75\textwidth]{fig/flowchart.png}
+		\caption{In appendix?}
+		\label{fig:flowchart}
+	\end{center}
+\end{figure*}
 
-We achieve this goal by dividing out the true $n(z)$ and multiplying by the 
-interim $n(z)$ as in
-\begin{align}
-\label{eq:transformation}
-p(z_{spec} | z_{phot}, \textul{\sigma}, \vec{\phi}_{int}) =& p(z_{spec} | 
-z_{phot}, \textul{\sigma}, \vec{\phi}_{true})\ \frac{p(z_{spec} | 
-\vec{\phi}_{int})}{p(z_{spec} | \vec{\phi}_{true})},
-\end{align}
-since $p(z_{spec} | z_{phot}, \textul{\sigma}, \vec{\phi})$ is separable and 
-the term $p(z_{spec} | z_{phot}, \textul{\sigma})$ changes independently of the 
-$p(z_{spec} | \vec{\phi})$ term.  The information about $\vec{\phi}_{true}$ is 
-retained by the elements of the catalog: only $(z_{spec}, z_{phot})$ pairs 
-drawn from the true distribution will be represented in the catalog!
-
-It appears as if this is the only fully self-consistent way to ensure that a 
-draw from the posterior distribution $p(z_{spec} | z_{phot}, \textul{\sigma}, 
-\vec{\phi})$ does in fact follow that distribution.  In the cases with fiducial 
-posteriors $p(z_{spec} | z_{phot}, \sigma, \vec{\phi})$, the method of Sec. 
-\ref{sec:mockdata} was equivalent to this, but Sec. \ref{sec:likelihoods} 
-requires a more sophisticated approach.
+\end{appendices}
 
@@
 preparation of this paper.
+This work was completed under the generous nutritional support of the Center 
+for Computational Astrophysics.
 \end{acknowledgements}
 
+\bibliographystyle{apj}
+\bibliography{references}
+
 \end{document}

### 2018-06-28 research/paper/cleaner.py

@@ -1,31 +1,0 @@
-import os, re
-#import subprocess
-#import argparse
-def clean(fileName):
-	remove_keys = ['abstract', 'annote', 'isbn', 'mendeley-groups', 'keywords', 'file', 'issn']
-	blind_keys = ['url', 'language']
-	bibf = open(fileName, 'r')
-	cleanf = open(fileName + '.tmp', 'w')
-	lines = bibf.readlines()
-	for line in lines:
-		if re.search('^'+'|'.join(blind_keys),line):
-			cleanf.write(line)
-		elif not re.search('^'+'|'.join(remove_keys),line):
-			#print line
-			cleanf.write(line)
-	cleanf.close()
-	os.remove(fileName)
-	os.rename(fileName + '.tmp', fileName)
-def readable(infile, outfile):
-	intfile = 'temp.tex'
-	os.system('sed \'/^%/ d\' < ' + infile + ' > ' + intfile)
-	os.system('fold -w80 -s ' + intfile + ' > ' + outfile)
-
-clean('references.bib')
-
-cwd = os.getcwd()
-infile = os.path.join(cwd, 'draft-editing.tex')
-outfile = os.path.join(cwd, 'draft.tex')
-print(infile, outfile)
-readable(infile, outfile)
-

### 2018-06-28 research/paper/cleaner.sh

@@ -1,0 +1,7 @@
+infile='draft-descope.tex'
+outfile='draft.tex'
+
+sed -e "s/{\\textbackslash}'\\\\{i\\\\}/\\'{i}/g" references.bib | grep -E -v '^\W*(urldate|abstract|file|keywords)' > references.bib.tmp
+mv references.bib.tmp references.bib
+
+grep -E -v '^(%|\\COMMENT)' "$infile" | fold -w80 -s > "$outfile"

### 2018-07-03 research/paper/draft.tex

@@
 \newcommand{\myemail}{aimalz@nyu.edu}
-\newcommand{\mathul}{\underline}
+\newcommand{\mathul}[1]{\underline{#1}}
 \newcommand{\chippr}{CHIPPR}
@@
 \title{How to obtain the redshift distribution from probabilistic redshift 
-estimates}%:\\
+estimates}
 
@@
 \altaffiltext{1}{Center for Cosmology and Particle Physics, Department of 
-Physics,
-  New York University, 726 Broadway, 9th floor, New York, NY 10003, USA}
+Physics, New York University, 726 Broadway, 9th floor, New York, NY 10003, USA}
 \altaffiltext{2}{Simons Center for Computational Astrophysics, 162 Fifth 
@@
 \pzpdf s in inference?  How does the result of \chippr\ compare to established 
-estimators in terms of the accuracy of \nz ?  \textcolor{blue}{Reach goal: How 
-significant is the effect of the discrepancy between \nz\ estimators on 
-cosmological constraints?}
+estimators in terms of the accuracy of \nz ?
+\textcolor{blue}{Reach goal: How significant is the effect of the discrepancy 
+between \nz\ estimators on cosmological constraints?}
 
@@
 A: Hierarchical inference is the only self-consistent way.}
-
-This paper presents the Cosmological Hierarchical Inference with Probabilistic 
-Photometric Redshifts (\chippr ) model as a mathematically consistent method 
-for constraining \nz s using a catalog of \pzpdf s.
-In Section~\ref{sec:model}, we introduce the \chippr\ formalism; in 
-Section~\ref{sec:application}, we outline how to use \chippr\ to constrain \nz\ 
-from a catalog of \pzpdf s; in Section~\ref{sec:limitations} we document the 
-assumptions that must hold for \chippr\ to be valid.
 
@@
 probability density over redshift; this fact can also be confirmed from 
-dimensional analysis of Equation \ref{eqn:nz}.
+dimensional analysis of Equation~\ref{eqn:nz}.
 Therefore, it can be rewritten as
@@
 
-\includegraphics[width=0.3\textwidth]{fig/forward_model_example.png}
-    \includegraphics[width=0.3\textwidth]{fig/forward_model_example.png}
-    \includegraphics[width=0.3\textwidth]{fig/forward_model_example.png}
-		\caption{Show this for each individually.}
+\includegraphics[width=0.3\textwidth]{fig/hivarsigmas_scatter.png}
+    \includegraphics[width=0.3\textwidth]{fig/uouthi_scatter.png}
+    \includegraphics[width=0.3\textwidth]{fig/neghivarbias_scatter.png}
+		\caption{intrinsic scatter (left), uniform outliers (center), 
+bias (right);
+    Not sure how best to broach the subject of nonuniform outliers and bias as 
+model misspecification
+    }
 		\label{fig:mega_scatter}
@@
 \end{figure*}
-Tests including all three effects at the tolerance levels of LSST and at twice 
-the tolerance levels of LSST are presented in Section~\ref{sec:results}.
+Tests including all three effects at the tolerance levels of LSST are presented 
+in Section~\ref{sec:results}.
 
@@
   \item implicit prior $\phi^{*}$ known
-	\item \pzpdf s are accurate posteriors %draws from photo-$z$ PDFs do 
-not match marginal distributions in $z_true$ vs. $z_phot$ space, i.e. toy model 
-with $z_phot$ drawn from an uglier space but pdfs evaluated in lsst-limits space
+	\item \pzpdf s are accurate posteriors
 	\end{itemize}
@@
 
-\textcolor{blue}{Q: How does the result of \chippr compare to established 
+\textcolor{blue}{Q: How does the result of \chippr\ compare to established 
 estimators in terms of the accuracy of $n(z)$?\\
-A: \chippr yields the best possible $n(z)$, conditional on the accuracy of the 
+A: \chippr\ yields the best possible $n(z)$, conditional on the accuracy of the 
 photo-$z$ PDFs used.}
 
-\subsection{Idealized example}
-\label{sec:idealized}
+\subsection{LSST Requirements}
+\label{sec:lsstdemo}
 
-\textcolor{blue}{LSST specs: unbiased, uniform outliers, medium z-dependent 
-scatter}
+\textcolor{blue}{LSST specs: biased to 0.003, 10\% uniform outliers, 
+z-dependent scatter of 0.05}
 
@@
 	\begin{center}
+		\includegraphics[width=0.45\textwidth]{fig/lsst_scatter.png}
 
-\includegraphics[width=0.45\textwidth]{fig/forward_model_example.png}
-		\includegraphics[width=0.45\textwidth]{fig/results_example.png}
-		\caption{Show this for idealized case.}
-		\label{fig:idealized}
-	\end{center}
-\end{figure*}
-
-\subsection{Realistic example}
-\label{sec:realistic}
-
-\textcolor{blue}{DES(?) specs: biased to enhance true $n(z)$, systematic 
-outliers, high z-dependent scatter}
-
-\begin{figure*}
-	\begin{center}
-
-\includegraphics[width=0.45\textwidth]{fig/forward_model_example.png}
-		\includegraphics[width=0.45\textwidth]{fig/results_example.png}
-		\caption{Show this for realistic case.}
-		\label{fig:realistic}
+\includegraphics[width=0.45\textwidth]{fig/lsst_log_estimators.png}
+		\caption{Using LSST numbers}
+		\label{fig:lsstdemo}
 	\end{center}
@@
 	\begin{center}
+		\includegraphics[width=0.45\textwidth]{fig/wrong_scatter.png}
 
-\includegraphics[width=0.45\textwidth]{fig/forward_model_example.png}
-		\includegraphics[width=0.45\textwidth]{fig/results_example.png}
-		\caption{Show this for mischaracterized case.}
+\includegraphics[width=0.45\textwidth]{fig/wrong_log_estimators.png}
+		\caption{this is for wrong implicit prior, not bias, haven't 
+done that one yet}
 		\label{fig:mischaracterized}
@@
 \bibliographystyle{apj}
-\bibliography{references}
 
